---
title: "clinical_var"
author: "Ruwen Zhou"
date: "12/2/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(plyr)
library(tidyverse)
library(dplyr)
library(readxl)
library(ggplot2)
library(phyloseq)
library(vegan)
library(DESeq2)
library(pheatmap)
library(RColorBrewer)
library(ggraph)
library(igraph)
library(Hmisc)
library(writexl)

library(tidygraph)
library(ppcor)

library(reshape2)
library(ComplexHeatmap)
```


```{r}
food_data = read_xlsx("data/FFQA862.xlsx")
nutrition_data = read_xlsx("data/FFQB800.xlsx")
species1 = read_xlsx("data/species_866.xlsx")
metadata = read_xlsx("data/helios500_clinical.xlsx")

food_data <- food_data %>%
  mutate(across(c("Mushrooms", "Papaya", "Orange", "Chapati"), as.numeric))
food_data
nutrition_data
species1
metadata
```

```{r}
colnames(species1)[1] = "Participant_ID"
colnames(metadata)[2] = "Participant_ID"
threshold <- nrow(species1) * 0.5
species_filtered1 <- species1 %>% select_if(function(x) sum(x > 0) > threshold)
species_filtered2 <- species1 %>% select_if(function(x) sum(x > 0) > nrow(species1) * 0.1)
# species_filtered2 <- species %>% select_if(function(x) sum(x > 0) > threshold)
# 
# #abundance_threshold <- 0.0001
# #species_filtered2 <- species_filtered %>% select_if(function(x) sum(x) > (abundance_threshold * sum(species)))
# #species_filtered2
# species_filtered2 <- species_filtered2[, !names(species_filtered2) %in% c("unclassified", "others", "Homo sapiens")]
# #species_filtered2$`Homo sapiens`
# #species_filtered2
# food_data[is.na(food_data)] <- 0
# #food_data <- data.frame(lapply(food_data, function(x) as.numeric(as.character(x))))
# nutrition_data[is.na(nutrition_data)] <- 0
```


```{r}
unique_species <- read.csv("figure/unique_significant_species.csv")
unique_species$species
colnames(species1)[1] = "Participant_ID"
sig_species <- species1 %>%
  dplyr::select("Participant_ID", all_of(unique_species$species))

colnames(metadata)[2] = "Participant_ID"
```


```{r}
sig_species
metadata

sig_species <- sig_species %>%
  mutate(Participant_ID = as.character(Participant_ID))

metadata <- metadata %>%
  mutate(Participant_ID = as.character(Participant_ID))
```

```{r}
#species_filtered2
merged_data <- sig_species %>%
  inner_join(metadata, by = "Participant_ID")
merged_data_all = merged_data %>%
  inner_join(food_data, by = "Participant_ID")
merged_data
metadata
```
# four groups based on bifido and thosai
```{r}
# # Create binary categories using median splits
# merged_data_all$bifido_cat <- ifelse(
#   merged_data_all$`Bifidobacterium catenulatum` > 
#     median(merged_data_all$`Bifidobacterium catenulatum`), 
#   "High", "Low"
# )
# 
# merged_data_all$thosai_cat <- ifelse(
#   merged_data_all$Idli_Thosai > median(merged_data_all$Idli_Thosai), 
#   "High", "Low"
# )
# 
# # Create combined groups
# merged_data_all$combined_group <- paste(
#   merged_data_all$bifido_cat, 
#   merged_data_all$thosai_cat
# )
# 
# # Create a contingency table
# contingency_table <- table(
#   merged_data_all$combined_group, 
#   merged_data_all$diabetes_status
# )
# 
# # View the contingency table
# print(contingency_table)
```


```{r}
# Perform chi-square test
# chi_test <- chisq.test(contingency_table)
# print(chi_test)
```


```{r}
# Create a visual representation using ggplot2
# library(ggplot2)
# 
# # Convert to percentages
# prop_table <- prop.table(contingency_table, margin = 1) * 100
# 
# # Convert to data frame for plotting
# plot_data <- as.data.frame(prop_table)
# colnames(plot_data) <- c("Group", "Diabetes_Status", "Percentage")
# 
# # Create stacked bar plot
# ggplot(plot_data, aes(x = Group, y = Percentage, fill = Diabetes_Status)) +
#   geom_bar(stat = "identity") +
#   theme_minimal() +
#   labs(
#     title = "Distribution of Diabetes Status by Bifido-Thosai Groups",
#     x = "Bifido-Thosai Group",
#     y = "Percentage (%)"
#   ) +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
# Get summary statistics for each group
# library(dplyr)
# 
# group_summary <- merged_data_all %>%
#   group_by(combined_group, diabetes_status) %>%
#   summarise(
#     n = n(),
#     percent = n() / nrow(merged_data_all) * 100
#   ) %>%
#   arrange(combined_group, diabetes_status)
# 
# print(group_summary)
```
```{r}
# # First, create a specific group for high-high vs others
# merged_data_all$special_group <- ifelse(
#   merged_data_all$bifido_cat == "High" & merged_data_all$thosai_cat == "High",
#   "High Bifido & High Thosai",
#   "Other Groups"
# )
# 
# # Create contingency table for this specific comparison
# high_high_table <- table(
#   merged_data_all$special_group,
#   merged_data_all$diabetes_status
# )
# 
# # Chi-square test
# high_high_chi <- chisq.test(high_high_table)
# print(high_high_chi)
# 
# # Calculate proportions within each group
# prop_table <- prop.table(high_high_table, margin = 1) * 100
# 
# # Print percentages
# print("Percentage of diabetes status in each group:")
# print(prop_table)
# 
# # Visualize using ggplot2
# library(ggplot2)
# 
# # Convert to data frame for plotting
# plot_data <- as.data.frame(prop_table)
# colnames(plot_data) <- c("Group", "Diabetes_Status", "Percentage")
# 
# # Create bar plot
# ggplot(plot_data, aes(x = Group, y = Percentage, fill = Diabetes_Status)) +
#   geom_bar(stat = "identity") +
#   theme_minimal() +
#   labs(
#     title = "Diabetes Status Distribution",
#     subtitle = "High Bifido & High Thosai vs Other Groups",
#     x = "Group",
#     y = "Percentage (%)"
#   ) +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
# 
# # You might also want to perform a logistic regression
# # This will give you odds ratios
# library(stats)
# 
# # Create binary outcome for diabetes (1 = diabetes, 0 = normal/pre-diabetes)
# merged_data_all$diabetes_binary <- ifelse(
#   merged_data_all$diabetes_status == "diabetes", 1, 0
# )
# 
# # Fit logistic regression
# logistic_model <- glm(
#   diabetes_binary ~ special_group,
#   family = binomial(link = "logit"),
#   data = merged_data_all
# )
# 
# # Print odds ratios
# odds_ratios <- exp(coef(logistic_model))
# print("Odds Ratios:")
# print(odds_ratios)
# 
# # Calculate risk ratios
# risk_ratio <- function(table) {
#   # Risk in high-high group
#   risk_high <- table["High Bifido & High Thosai", "diabetes"] / 
#     sum(table["High Bifido & High Thosai", ])
#   
#   # Risk in other groups
#   risk_other <- table["Other Groups", "diabetes"] / 
#     sum(table["Other Groups", ])
#   
#   return(risk_high / risk_other)
# }
# 
# rr <- risk_ratio(high_high_table)
# print("Risk Ratio (High-High vs Others):")
# print(rr)
```
# only in indian
```{r}
# only_i = subset(merged_data_all, FREG5_Ethnic_Group == "I")
# only_i$`Bifidobacterium catenulatum`
# ```
# ```{r}
# # Create binary categories using median splits
# only_i$bifido_cat <- ifelse(
#   only_i$`Bifidobacterium catenulatum` > 
#     median(only_i$`Bifidobacterium catenulatum`), 
#   "High", "Low"
# )
# 
# only_i$thosai_cat <- ifelse(
#   only_i$Idli_Thosai > median(only_i$Idli_Thosai), 
#   "High", "Low"
# )
# 
# # Create combined groups
# only_i$combined_group <- paste(
#   only_i$bifido_cat, 
#   only_i$thosai_cat
# )
# 
# # Create a contingency table
# contingency_table_i <- table(
#   only_i$combined_group, 
#   only_i$diabetes_status
# )
# 
# # View the contingency table
# print(contingency_table_i)
```


```{r}
# # Perform chi-square test
# chi_test <- chisq.test(contingency_table_i)
# print(chi_test)
```


```{r}
# Create a visual representation using ggplot2
# library(ggplot2)
# 
# # Convert to percentages
# prop_table <- prop.table(contingency_table_i, margin = 1) * 100
# 
# # Convert to data frame for plotting
# plot_data <- as.data.frame(prop_table)
# colnames(plot_data) <- c("Group", "Diabetes_Status", "Percentage")
# 
# # Create stacked bar plot
# ggplot(plot_data, aes(x = Group, y = Percentage, fill = Diabetes_Status)) +
#   geom_bar(stat = "identity") +
#   theme_minimal() +
#   labs(
#     title = "Distribution of Diabetes Status by Bifido-Thosai Groups",
#     x = "Bifido-Thosai Group",
#     y = "Percentage (%)"
#   ) +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}

# First, create a specific group for high-high vs others
# only_i$special_group <- ifelse(
#   only_i$bifido_cat == "High" & only_i$thosai_cat == "High",
#   "High Bifido & High Thosai",
#   "Other Groups"
# )
# 
# # Create contingency table for this specific comparison
# high_high_table <- table(
#   only_i$special_group,
#   only_i$diabetes_status
# )
# 
# # Chi-square test
# high_high_chi <- chisq.test(high_high_table)
# print(high_high_chi)
# 
# # Calculate proportions within each group
# prop_table <- prop.table(high_high_table, margin = 1) * 100
# 
# # Print percentages
# print("Percentage of diabetes status in each group:")
# print(prop_table)
# 
# # Visualize using ggplot2
# library(ggplot2)
# 
# # Convert to data frame for plotting
# plot_data <- as.data.frame(prop_table)
# colnames(plot_data) <- c("Group", "Diabetes_Status", "Percentage")
# 
# # Create bar plot
# ggplot(plot_data, aes(x = Group, y = Percentage, fill = Diabetes_Status)) +
#   geom_bar(stat = "identity") +
#   theme_minimal() +
#   labs(
#     title = "Diabetes Status Distribution",
#     subtitle = "High Bifido & High Thosai vs Other Groups",
#     x = "Group",
#     y = "Percentage (%)"
#   ) +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
# 
# # You might also want to perform a logistic regression
# # This will give you odds ratios
# library(stats)
# 
# # Create binary outcome for diabetes (1 = diabetes, 0 = normal/pre-diabetes)
# only_i$diabetes_binary <- ifelse(
#   only_i$diabetes_status == "diabetes", 1, 0
# )
# 
# # Fit logistic regression
# logistic_model <- glm(
#   diabetes_binary ~ special_group,
#   family = binomial(link = "logit"),
#   data = only_i
# )
# 
# # Print odds ratios
# odds_ratios <- exp(coef(logistic_model))
# print("Odds Ratios:")
# print(odds_ratios)
# 
# # Calculate risk ratios
# risk_ratio <- function(table) {
#   # Risk in high-high group
#   risk_high <- table["High Bifido & High Thosai", "diabetes"] / 
#     sum(table["High Bifido & High Thosai", ])
#   
#   # Risk in other groups
#   risk_other <- table["Other Groups", "diabetes"] / 
#     sum(table["Other Groups", ])
#   
#   return(risk_high / risk_other)
# }
# 
# rr <- risk_ratio(high_high_table)
# print("Risk Ratio (High-High vs Others):")
# print(rr)
```






# heatmaps
```{r}
# Define columns
species_cols <- colnames(sig_species)[colnames(sig_species) != "Participant_ID"]
clinical_vars <- c("DLAB27_Hba1C_Percent", "DLAB83_CRP", "BMI", "HOMA_IR", "HOMA_B")

# Initialize matrices for correlations and p-values
cor_matrix <- matrix(NA, nrow = length(species_cols), ncol = length(clinical_vars))
p_matrix <- matrix(NA, nrow = length(species_cols), ncol = length(clinical_vars))

# Calculate partial correlations
for(i in 1:length(species_cols)) {
  for(j in 1:length(clinical_vars)) {
    # Create temporary dataset with ranked variables
    temp_data <- data.frame(
      species = rank(merged_data[[species_cols[i]]]),
      clinical = rank(merged_data[[clinical_vars[j]]]),
      gender = as.numeric(merged_data$FREG7_Gender == "M"),
      age = rank(merged_data$FREG8_Age),
      ethnic_C = as.numeric(merged_data$FREG5_Ethnic_Group == "C"),
      ethnic_I = as.numeric(merged_data$FREG5_Ethnic_Group == "I"),
      ethnic_M = as.numeric(merged_data$FREG5_Ethnic_Group == "M")
    )
    
    # Calculate partial correlation
    pcor_result <- pcor.test(temp_data$species, temp_data$clinical,
                            temp_data[,c("gender", "age", 
                                       "ethnic_C", "ethnic_I", "ethnic_M")])
    
    # Cap correlation values at ±0.2
    cor_matrix[i,j] <- pmin(pmax(pcor_result$estimate, -0.15), 0.15)
    p_matrix[i,j] <- pcor_result$p.value
  }
}

# Add row and column names
rownames(cor_matrix) <- species_cols
colnames(cor_matrix) <- clinical_vars
rownames(p_matrix) <- species_cols
colnames(p_matrix) <- clinical_vars

# Perform clustering
species_dist <- dist(cor_matrix)
species_clust <- hclust(species_dist)
species_order <- species_clust$labels[species_clust$order]

clinical_dist <- dist(t(cor_matrix))
clinical_clust <- hclust(clinical_dist)
clinical_order <- clinical_clust$labels[clinical_clust$order]

# Convert to long format
melted_cor <- melt(cor_matrix)
melted_p <- melt(p_matrix)
colnames(melted_cor) <- c("Species", "Clinical", "Correlation")
melted_cor$Significant <- melted_p$value < 0.05

# Set factor levels based on clustering
melted_cor$Species <- factor(melted_cor$Species, levels = species_order)
melted_cor$Clinical <- factor(melted_cor$Clinical, levels = clinical_order)
#melted_cor$Clinical
# Create the plot
clinical_labels <- c(
  "BMI" = "BMI",
  "DLAB27_Hba1C_Percent" = "HbaA1C",
  "DLAB83_CRP" = "CRP",
  "HOMA_IR" = "HOMA-IR",
  "HOMA_B" = "HOMA-B"
)

# Plot with updated x-axis labels
ggplot(melted_cor, aes(x = Clinical, y = Species, fill = Correlation)) +
  geom_tile() +
  scale_fill_gradientn(colors = rev(colorRampPalette(brewer.pal(11, "RdBu"))(100)), 
                      limits = c(-0.15, 0.15)) +
  geom_text(aes(label = ifelse(Significant, "*", "")), 
            color = "black", size = 5) +
  scale_x_discrete(labels = clinical_labels) +  # Rename x-axis labels
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(face = "italic"),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold")
  ) +
  labs(
    title = "Correlations of Species and Clinical Variable",
    subtitle = "Adjusted for Gender, Age, and Ethnicity\n* indicates p < 0.05",
    x = "Clinical Variables",
    y = "Species"
  )

# Save the plot if needed
ggsave("figure/fig6_clinical_species_correlations.pdf", width = 6, height = 11)
```
```{r}
melted_cor
p_matrix
```





```{r}
library(gridExtra)
merged_data <- sig_species %>%
  inner_join(metadata, by = "Participant_ID")

# Define columns
species_cols <- colnames(sig_species)[colnames(sig_species) != "Participant_ID"]
continuous_vars <- c("DLAB27_Hba1C_Percent", "DLAB83_CRP", "BMI", "HOMA_IR", "HOMA_B")

# Initialize matrices for continuous variables
cont_cor_matrix <- matrix(NA, nrow = length(species_cols), ncol = length(continuous_vars))
cont_p_matrix <- matrix(NA, nrow = length(species_cols), ncol = length(continuous_vars))

# Initialize vectors for diabetes status
diab_effect <- numeric(length(species_cols))
diab_p <- numeric(length(species_cols))

# Calculate correlations for continuous variables
for(i in 1:length(species_cols)) {
  for(j in 1:length(continuous_vars)) {
    temp_data <- data.frame(
      species = rank(merged_data[[species_cols[i]]]),
      clinical = rank(merged_data[[continuous_vars[j]]]),
      gender = as.numeric(merged_data$FREG7_Gender == "M"),
      age = rank(merged_data$FREG8_Age),
      ethnic_C = as.numeric(merged_data$FREG5_Ethnic_Group == "C"),
      ethnic_I = as.numeric(merged_data$FREG5_Ethnic_Group == "I"),
      ethnic_M = as.numeric(merged_data$FREG5_Ethnic_Group == "M")
    )
    
    pcor_result <- pcor.test(temp_data$species, temp_data$clinical,
                            temp_data[,c("gender", "age", 
                                       "ethnic_C", "ethnic_I", "ethnic_M")])
    
    cont_cor_matrix[i,j] <- pmin(pmax(pcor_result$estimate, -0.15), 0.15)
    cont_p_matrix[i,j] <- pcor_result$p.value
  }
  
  # Calculate diabetes association
  temp_data <- data.frame(
    species_abundance = merged_data[[species_cols[i]]],
    gender = merged_data$FREG7_Gender,
    age = merged_data$FREG8_Age,
    ethnicity = merged_data$FREG5_Ethnic_Group,
    diabetes = merged_data$diabetes_status
  )
  
  model <- lm(species_abundance ~ gender + age + ethnicity, data = temp_data)
  species_residuals <- residuals(model)
  
  kw_test <- kruskal.test(species_residuals ~ temp_data$diabetes)
  
  groups <- split(species_residuals, temp_data$diabetes)
  means <- sapply(groups, mean)
  overall_mean <- mean(species_residuals)
  ss_between <- sum(sapply(groups, length) * (means - overall_mean)^2)
  ss_total <- sum((species_residuals - overall_mean)^2)
  eta_squared <- ss_between / ss_total
  
  if("diabetes" %in% names(means) && "normal" %in% names(means)) {
    direction <- sign(means["diabetes"] - means["normal"])
  } else {
    direction <- 1
  }
  
  diab_effect[i] <- direction * eta_squared
  diab_p[i] <- kw_test$p.value
}

# Prepare continuous variables plot
rownames(cont_cor_matrix) <- species_cols
colnames(cont_cor_matrix) <- continuous_vars
rownames(cont_p_matrix) <- species_cols
colnames(cont_p_matrix) <- continuous_vars

# Clustering for continuous variables
species_dist <- dist(cont_cor_matrix)
species_clust <- hclust(species_dist)
species_order <- species_clust$labels[species_clust$order]

clinical_dist <- dist(t(cont_cor_matrix))
clinical_clust <- hclust(clinical_dist)
clinical_order <- clinical_clust$labels[clinical_clust$order]

# Prepare data for plots
# Continuous variables
melted_cont <- melt(cont_cor_matrix)
melted_cont_p <- melt(cont_p_matrix)
colnames(melted_cont) <- c("Species", "Clinical", "Correlation")
melted_cont$Significant <- melted_cont_p$value < 0.05
melted_cont$Species <- factor(melted_cont$Species, levels = species_order)
melted_cont$Clinical <- factor(melted_cont$Clinical, levels = clinical_order)

# Diabetes status
diab_df <- data.frame(
  Species = species_cols,
  Effect = diab_effect,
  P_value = diab_p
)
diab_df$Species <- factor(diab_df$Species, levels = species_order)
diab_df$Significant <- diab_df$P_value < 0.05

# Create continuous variables plot
p1 <- ggplot(melted_cont, aes(x = Clinical, y = Species, fill = Correlation)) +
  geom_tile() +
  scale_fill_gradientn(colors = rev(colorRampPalette(brewer.pal(11, "RdBu"))(100)), 
                      limits = c(-0.15, 0.15)) +
  geom_text(aes(label = ifelse(Significant, "*", "")), 
            color = "black", size = 5) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(face = "italic"),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold")
  ) +
  labs(
    title = "Species Correlations with Clinical Variables",
    subtitle = "Adjusted for Gender, Age, and Ethnicity\n* indicates p < 0.05",
    x = "Clinical Variables",
    y = "Species",
    fill = "Correlation"
  )

# Create diabetes status plot
p2 <- ggplot(diab_df, aes(y = Species, x = Effect)) +
  geom_bar(stat = "identity", aes(fill = abs(Effect))) +
  #scale_fill_gradientn(colors = rev(brewer.pal(9, "PuBu"))) +
  scale_fill_gradient2(low = "white", high = "plum4") +
  geom_text(aes(label = ifelse(Significant, "*", ""), x = 0), 
            color = "black", size = 5) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(face = "italic"),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold")
  ) +
  labs(
    title = "Species Associations with Diabetes Status",
    subtitle = "Adjusted for Gender, Age, and Ethnicity\n* indicates p < 0.05",
    x = "Effect Size (Eta Squared)",
    y = NULL,
    fill = "Absolute\nEffect Size"
  )

# Arrange plots side by side
grid.arrange(p1, p2, ncol = 2, widths = c(1.5, 1))

# Optionally save the combined plot
#ggsave("figure/clinical_diabetes_associations2.pdf", arrangeGrob(p1, p2, ncol = 2, widths = c(1.5, 1)), width = 15, height = 10)
```
## separate ethnicity
```{r}
library(gridExtra)

# Function to perform analysis for one ethnic group
analyze_ethnic_group <- function(merged_data, ethnic_group) {
  # Filter data for the specific ethnic group
  ethnic_data <- merged_data %>% 
    filter(FREG5_Ethnic_Group == ethnic_group)
  
  # Get columns
  species_cols <- colnames(sig_species)[colnames(sig_species) != "Participant_ID"]
  continuous_vars <- c("DLAB27_Hba1C_Percent", "DLAB83_CRP", "BMI", "HOMA_IR", "HOMA_B")
  
  # Initialize matrices
  cont_cor_matrix <- matrix(NA, nrow = length(species_cols), ncol = length(continuous_vars))
  cont_p_matrix <- matrix(NA, nrow = length(species_cols), ncol = length(continuous_vars))
  diab_effect <- numeric(length(species_cols))
  diab_p <- numeric(length(species_cols))
  
  # Calculate correlations for continuous variables
  for(i in 1:length(species_cols)) {
    for(j in 1:length(continuous_vars)) {
      temp_data <- data.frame(
        species = rank(ethnic_data[[species_cols[i]]]),
        clinical = rank(ethnic_data[[continuous_vars[j]]]),
        gender = as.numeric(ethnic_data$FREG7_Gender == "M"),
        age = rank(ethnic_data$FREG8_Age)
      )
      
      # Now adjusting only for gender and age
      pcor_result <- pcor.test(temp_data$species, temp_data$clinical,
                              temp_data[,c("gender", "age")])
      
      cont_cor_matrix[i,j] <- pmin(pmax(pcor_result$estimate, -0.15), 0.15)
      cont_p_matrix[i,j] <- pcor_result$p.value
    }
    
    # Calculate diabetes association
    temp_data <- data.frame(
      species_abundance = ethnic_data[[species_cols[i]]],
      gender = ethnic_data$FREG7_Gender,
      age = ethnic_data$FREG8_Age,
      diabetes = ethnic_data$diabetes_status
    )
    
    # Adjust for gender and age only
    model <- lm(species_abundance ~ gender + age, data = temp_data)
    species_residuals <- residuals(model)
    
    kw_test <- kruskal.test(species_residuals ~ temp_data$diabetes)
    
    groups <- split(species_residuals, temp_data$diabetes)
    means <- sapply(groups, mean)
    overall_mean <- mean(species_residuals)
    ss_between <- sum(sapply(groups, length) * (means - overall_mean)^2)
    ss_total <- sum((species_residuals - overall_mean)^2)
    eta_squared <- ss_between / ss_total
    
    if("diabetes" %in% names(means) && "normal" %in% names(means)) {
      direction <- sign(means["diabetes"] - means["normal"])
    } else {
      direction <- 1
    }
    
    diab_effect[i] <- direction * eta_squared
    diab_p[i] <- kw_test$p.value
  }
  
  # Prepare matrices
  rownames(cont_cor_matrix) <- species_cols
  colnames(cont_cor_matrix) <- continuous_vars
  rownames(cont_p_matrix) <- species_cols
  colnames(cont_p_matrix) <- continuous_vars
  
  # Return results
  list(
    cont_cor = cont_cor_matrix,
    cont_p = cont_p_matrix,
    diab_effect = diab_effect,
    diab_p = diab_p
  )
}

# Analyze each ethnic group
results_C <- analyze_ethnic_group(merged_data, "C")
results_I <- analyze_ethnic_group(merged_data, "I")
results_M <- analyze_ethnic_group(merged_data, "M")

# Function to create plots for one ethnic group
create_plots <- function(results, species_cols, continuous_vars, group_name) {
  # Prepare continuous variables plot
  melted_cont <- melt(results$cont_cor)
  melted_cont_p <- melt(results$cont_p)
  colnames(melted_cont) <- c("Species", "Clinical", "Correlation")
  melted_cont$Significant <- melted_cont_p$value < 0.05
  
  # Create continuous variables plot
  p1 <- ggplot(melted_cont, aes(x = Clinical, y = Species, fill = Correlation)) +
    geom_tile() +
    scale_fill_gradientn(colors = rev(colorRampPalette(brewer.pal(11, "RdBu"))(100)), 
                        limits = c(-0.15, 0.15)) +
    geom_text(aes(label = ifelse(Significant, "*", "")), 
              color = "black", size = 5) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.text.y = element_text(face = "italic"),
      axis.title = element_text(face = "bold"),
      legend.title = element_text(face = "bold")
    ) +
    labs(
      title = paste("Species Correlations with Clinical Variables -", group_name),
      subtitle = "Adjusted for Gender and Age\n* indicates p < 0.05",
      x = "Clinical Variables",
      y = "Species",
      fill = "Correlation"
    )
  
  # Prepare diabetes plot data
  diab_df <- data.frame(
    Species = species_cols,
    Effect = results$diab_effect,
    P_value = results$diab_p
  )
  diab_df$Significant <- diab_df$P_value < 0.05
  
  # Create diabetes plot
  p2 <- ggplot(diab_df, aes(y = Species, x = Effect)) +
    geom_bar(stat = "identity", aes(fill = abs(Effect))) +
    scale_fill_gradient2(low = "white", high = "plum4") +
    geom_text(aes(label = ifelse(Significant, "*", ""), x = 0), 
              color = "black", size = 5) +
    theme_minimal() +
    theme(
      axis.text.y = element_text(face = "italic"),
      axis.title = element_text(face = "bold"),
      legend.title = element_text(face = "bold")
    ) +
    labs(
      title = paste("Species Associations with Diabetes Status -", group_name),
      subtitle = "Adjusted for Gender and Age\n* indicates p < 0.05",
      x = "Effect Size (Eta Squared)",
      y = NULL,
      fill = "Absolute\nEffect Size"
    )
  
  list(continuous = p1, diabetes = p2)
}
```


```{r}
# Create plots for each ethnic group
plots_C <- create_plots(results_C, species_cols, continuous_vars, "Chinese")
plots_I <- create_plots(results_I, species_cols, continuous_vars, "Indian")
plots_M <- create_plots(results_M, species_cols, continuous_vars, "Malay")

# plots_C
# plots_I 
# plots_M
```


```{r}
# Arrange all plots
# Continuous variables plots on top row, diabetes plots on bottom row
grid.arrange(
  plots_C$continuous, plots_I$continuous, plots_M$continuous,
  plots_C$diabetes, plots_I$diabetes, plots_M$diabetes,
  ncol = 3, nrow = 2
)


# ggsave("figure/ethnic_stratified_analysis.pdf",
#        arrangeGrob(plots_C$continuous, plots_I$continuous, plots_M$continuous,
#                   plots_C$diabetes, plots_I$diabetes, plots_M$diabetes,
#                   ncol = 3, nrow = 2),
#        width = 24, height = 16)
```



```{r}
# Create summary table

diabetes_dist <- merged_data %>%
  group_by(FREG5_Ethnic_Group, diabetes_status) %>%
  summarise(count = n(), .groups = 'drop') %>%
  # Calculate percentages within each ethnic group
  group_by(FREG5_Ethnic_Group) %>%
  mutate(percentage = count/sum(count) * 100)

diabetes_dist <- diabetes_dist %>%
  mutate(diabetes_status = factor(diabetes_status, levels = c("diabetes", "pre-diabetes", "normal")))
# Create a plot
ggplot(diabetes_dist, aes(x = FREG5_Ethnic_Group, y = count, fill = diabetes_status)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal() +
  labs(
    title = "Distribution of Diabetes Status (based on Hba1C) by Ethnic Group",
    x = "Ethnic Group",
    y = "Count",
    fill = "Diabetes Status"
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 14, face = "bold")
  )

# Print numerical summary
print(diabetes_dist)
```
```{r}
merged_data$glucose
diabetes_dist <- merged_data %>%
  group_by(FREG5_Ethnic_Group, glucose) %>%
  summarise(count = n(), .groups = 'drop') %>%
  # Calculate percentages within each ethnic group
  group_by(FREG5_Ethnic_Group) %>%
  mutate(percentage = count/sum(count) * 100)

diabetes_dist <- diabetes_dist %>%
  mutate(glucose = factor(glucose, levels = c("diabetes", "pre-diabetes", "normal", "hypoglycemia")))

# Create a plot
ggplot(diabetes_dist, aes(x = FREG5_Ethnic_Group, y = count, fill = glucose)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal() +
  labs(
    title = "Distribution of Diabetes Status (based on glucose) by Ethnic Group",
    x = "Ethnic Group",
    y = "Count",
    fill = "Diabetes Status"
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 14, face = "bold")
  )

# Print numerical summary
print(diabetes_dist)
```
```{r}
species_filtered2
```


## overall analysis for clinical variables
```{r}
species_filtered2
# Match samples and create merged dataset
merged_data_a <- species_filtered2 %>%
  inner_join(metadata, by = "Participant_ID")
species_cols2 = colnames(species_filtered2[,-1])
species_cols2
```


```{r}
# Assuming merged_data is already loaded and contains species and metadata columns
# Create OTU Table
merged_data_a$BMI_group <- ifelse(merged_data_a$BMI > 25, "Obese", "Non-Obese")
OTU <- otu_table(as.matrix(merged_data_a[, species_cols2]), taxa_are_rows = FALSE)
merged_data_a
# Make sure the row names of the OTU table are the Participant IDs
rownames(OTU) <- merged_data_a$Participant_ID

# Create Sample Data
# Ensure that the sample_data dataframe has Participant_ID as row names
samples_data <- sample_data(merged_data_a[, c("diabetes_status", "FREG5_Ethnic_Group", "DLAB27_Hba1C_Percent", "DLAB83_CRP", "DLAB81_Insulin", "DLAB25_Gluf_Mmol_L", "BMI","BMI_group")])
rownames(samples_data) <- merged_data_a$Participant_ID

# Create a phyloseq object
physeq <- phyloseq(OTU, samples_data)
OTU
```
##obese alpha beta
```{r}
comparisons <- list(c("Obese", "Non-Obese"))
a1 = plot_richness(physeq, x = "BMI_group", color = "BMI_group", measures = c("Shannon", "Simpson", "Chao1", "Fisher")) + geom_boxplot() + theme_bw()+
    stat_compare_means(comparisons = comparisons, method = "t.test", label = "p.signif", symnum.args = list(cutpoints = c(0, 0.001, 0.01, 0.05, 1), symbols = c("***", "**", "*", "ns")))
ggsave(
  filename = "figure/figure6/obese_alpha.pdf",
  plot = a1,
  height = 6, 
  width = 9,
  device = cairo_pdf)

```
```{r}
# Compute alpha diversity measures for the physeq object
# alpha_div <- estimate_richness(physeq, measures = c("Shannon", "Simpson", "Chao1", "Fisher"))
# 
# # Add BMI group information to the alpha diversity data
# alpha_div$BMI_group <- sam_data$BMI_group
# 
# # Run statistical tests for each diversity measure
# p_values <- data.frame(
#   Measure = c("Shannon", "Simpson", "Chao1", "Fisher"),
#   p_value = c(
#     t.test(alpha_div$Shannon ~ alpha_div$BMI_group)$p.value,
#     t.test(alpha_div$Simpson ~ alpha_div$BMI_group)$p.value,
#     t.test(alpha_div$Chao1 ~ alpha_div$BMI_group)$p.value,
#     t.test(alpha_div$Fisher ~ alpha_div$BMI_group)$p.value
#   )
# )
# 
# # Optionally, adjust p-values for multiple comparisons (e.g., using FDR correction)
# p_values$p_adj <- p.adjust(p_values$p_value, method = "fdr")
# print(p_values)
```



```{r}
dist = phyloseq::distance(physeq, method="bray")
phylo.beta = ordinate(physeq, method="PCoA", distance=dist)

b1 = plot_ordination(physeq, phylo.beta, type = "samples", color = "BMI_group") + theme_bw() + stat_ellipse(aes(group=BMI_group))
ggsave(
  filename = "figure/figure6/obese_beta.pdf",
  plot = b1,
  height = 6, 
  width = 9,
  device = cairo_pdf)
```
```{r}
library(vegan)
sample_data(physeq)$BMI_group <- factor(sample_data(physeq)$BMI_group)

physeq <- subset_samples(physeq, !is.na(sample_data(physeq)$BMI_group))

# Calculate the distance matrix using the ape package
dist <- phyloseq::distance(physeq, method = "bray")
dist_matrix <- as.matrix(dist)

# Perform PERMANOVA using adonis2
permanova <- adonis2(dist_matrix ~ sample_data(physeq)$BMI_group)

# Print the PERMANOVA results
print(permanova)
```

##diabete alpha beta
```{r}
#comparisons <- list(c("Obese", "Non-Obese"))
physeq_f <- subset_samples(physeq, !is.na(sample_data(physeq)$diabetes_status))
a2 = plot_richness(physeq_f, x = "diabetes_status", color = "diabetes_status", measures = c("Shannon", "Simpson", "Chao1", "Fisher")) + geom_boxplot() + theme_bw()
ggsave(
  filename = "figure/figure6/diabete_alpha.pdf",
  plot = a2,
  height = 6, 
  width = 9,
  device = cairo_pdf)

```

```{r}
dist = phyloseq::distance(physeq_f, method="bray")
phylo.beta = ordinate(physeq_f, method="PCoA", distance=dist)

b2 = plot_ordination(physeq_f, phylo.beta, type = "samples", color = "diabetes_status") + theme_bw() + stat_ellipse(aes(group=diabetes_status))
ggsave(
  filename = "figure/figure6/diabetes_beta.pdf",
  plot = b2,
  height = 6, 
  width = 9,
  device = cairo_pdf)
```
```{r}
library(vegan)
sample_data(physeq_f)$diabetes_status <- factor(sample_data(physeq_f)$diabetes_status)

#physeq <- subset_samples(physeq, !is.na(sample_data(physeq)$BMI_group))

# Calculate the distance matrix using the ape package
dist <- phyloseq::distance(physeq_f, method = "bray")
dist_matrix <- as.matrix(dist)

# Perform PERMANOVA using adonis2
permanova <- adonis2(dist_matrix ~ sample_data(physeq_f)$diabetes_status)

# Print the PERMANOVA results
print(permanova)
```


```{r}
library(phyloseq)
library(dplyr)

alpha_diversity <- estimate_richness(physeq, measures = "Shannon")
# Convert rownames to a column for merging
alpha_diversity$Participant_ID <- rownames(alpha_diversity)

# Extract sample metadata from the phyloseq object and make it a regular data frame
sample_metadata <- as.data.frame(sample_data(physeq))
# Ensure 'Participant_ID' is a column for merging
sample_metadata$Participant_ID <- rownames(sample_metadata)

# Merge alpha diversity data with sample metadata
alpha_diversity_full <- merge(alpha_diversity, sample_metadata, by = "Participant_ID")


# Now you can plot or analyze your merged data
# For example, to visualize alpha diversity across diabetes status:
library(ggplot2)
ggplot(alpha_diversity_full, aes(x = diabetes_status, y = Shannon)) +
  geom_boxplot() +
  labs(title = "Alpha Diversity across Diabetes Status", y = "Shannon Diversity Index", x = "Diabetes Status")

```

## food and clinical variables
```{r}
food_data
merged_data_f = food_data %>%
  inner_join(metadata, by = "Participant_ID")
```

```{r}
food_cols <- colnames(food_data)[colnames(food_data) != "Participant_ID"]
clinical_vars <- c("DLAB27_Hba1C_Percent", "DLAB83_CRP", "BMI", "HOMA_IR", "HOMA_B")

# Initialize matrices for correlations and p-values
cor_matrix <- matrix(NA, nrow = length(food_cols), ncol = length(clinical_vars))
p_matrix <- matrix(NA, nrow = length(food_cols), ncol = length(clinical_vars))

# Calculate partial correlations
for(i in 1:length(food_cols)) {
  for(j in 1:length(clinical_vars)) {
    # Create temporary dataset with ranked variables
    temp_data <- data.frame(
      food = rank(merged_data_f[[food_cols[i]]]),
      clinical = rank(merged_data_f[[clinical_vars[j]]]),
      gender = as.numeric(merged_data_f$FREG7_Gender == "M"),
      age = rank(merged_data_f$FREG8_Age),
      ethnic_C = as.numeric(merged_data_f$FREG5_Ethnic_Group == "C"),
      ethnic_I = as.numeric(merged_data_f$FREG5_Ethnic_Group == "I"),
      ethnic_M = as.numeric(merged_data_f$FREG5_Ethnic_Group == "M")
    )
    
    # Calculate partial correlation
    pcor_result <- pcor.test(temp_data$food, temp_data$clinical,
                            temp_data[,c("gender", "age", 
                                       "ethnic_C", "ethnic_I", "ethnic_M")])
    
    # Cap correlation values at ±0.2
    cor_matrix[i,j] <- pmin(pmax(pcor_result$estimate, -0.15), 0.15)
    p_matrix[i,j] <- pcor_result$p.value
  }
}

# Add row and column names
rownames(cor_matrix) <- food_cols
colnames(cor_matrix) <- clinical_vars
rownames(p_matrix) <- food_cols
colnames(p_matrix) <- clinical_vars

# Perform clustering
food_dist <- dist(cor_matrix)
food_clust <- hclust(food_dist)
food_order <- food_clust$labels[food_clust$order]

clinical_dist <- dist(t(cor_matrix))
clinical_clust <- hclust(clinical_dist)
clinical_order <- clinical_clust$labels[clinical_clust$order]

# Convert to long format
melted_cor <- melt(cor_matrix)
melted_p <- melt(p_matrix)
colnames(melted_cor) <- c("Food", "Clinical", "Correlation")
melted_cor$Significant <- melted_p$value < 0.05

# Set factor levels based on clustering
melted_cor$Food <- factor(melted_cor$Food, levels = food_order)
melted_cor$Clinical <- factor(melted_cor$Clinical, levels = clinical_order)

# Create the plot
ggplot(melted_cor, aes(x = Clinical, y = Food, fill = Correlation)) +
  geom_tile() +
  scale_fill_gradientn(colors = rev(colorRampPalette(brewer.pal(11, "RdBu"))(100)), 
                      limits = c(-0.15, 0.15)) +
  geom_text(aes(label = ifelse(Significant, "*", "")), 
            color = "black", size = 5) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(face = "italic"),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold")
  ) +
  labs(
    title = "Spearman Correlations between Food and Clinical Variables",
    subtitle = "Adjusted for Gender, Age, and Ethnicity\n* indicates p < 0.05",
    x = "Clinical Variables",
    y = "Food"
  )

```



```{r}


# Define columns
food_cols <- colnames(food_data)[colnames(food_data) != "Participant_ID"]
continuous_vars <- c("DLAB27_Hba1C_Percent", "DLAB83_CRP", "BMI", "HOMA_IR", "HOMA_B")

# Initialize matrices for continuous variables
cont_cor_matrix <- matrix(NA, nrow = length(food_cols), ncol = length(continuous_vars))
cont_p_matrix <- matrix(NA, nrow = length(food_cols), ncol = length(continuous_vars))

# Initialize vectors for diabetes status
diab_effect <- numeric(length(food_cols))
diab_p <- numeric(length(food_cols))

# Calculate correlations for continuous variables
for(i in 1:length(food_cols)) {
  for(j in 1:length(continuous_vars)) {
    temp_data <- data.frame(
      food = rank(merged_data_f[[food_cols[i]]]),
      clinical = rank(merged_data_f[[continuous_vars[j]]]),
      gender = as.numeric(merged_data_f$FREG7_Gender == "M"),
      age = rank(merged_data_f$FREG8_Age),
      ethnic_C = as.numeric(merged_data_f$FREG5_Ethnic_Group == "C"),
      ethnic_I = as.numeric(merged_data_f$FREG5_Ethnic_Group == "I"),
      ethnic_M = as.numeric(merged_data_f$FREG5_Ethnic_Group == "M")
    )
    
    pcor_result <- pcor.test(temp_data$food, temp_data$clinical,
                            temp_data[,c("gender", "age", 
                                       "ethnic_C", "ethnic_I", "ethnic_M")])
    
    cont_cor_matrix[i,j] <- pmin(pmax(pcor_result$estimate, -0.15), 0.15)
    cont_p_matrix[i,j] <- pcor_result$p.value
  }
  
  # Calculate diabetes association
  temp_data <- data.frame(
    food_abundance = merged_data_f[[food_cols[i]]],
    gender = merged_data_f$FREG7_Gender,
    age = merged_data_f$FREG8_Age,
    ethnicity = merged_data_f$FREG5_Ethnic_Group,
    diabetes = merged_data_f$diabetes_status
  )
  
  model <- lm(food_abundance ~ gender + age + ethnicity, data = temp_data)
  food_residuals <- residuals(model)
  
  kw_test <- kruskal.test(food_residuals ~ temp_data$diabetes)
  
  groups <- split(food_residuals, temp_data$diabetes)
  means <- sapply(groups, mean)
  overall_mean <- mean(food_residuals)
  ss_between <- sum(sapply(groups, length) * (means - overall_mean)^2)
  ss_total <- sum((food_residuals - overall_mean)^2)
  eta_squared <- ss_between / ss_total
  
  if("diabetes" %in% names(means) && "normal" %in% names(means)) {
    direction <- sign(means["diabetes"] - means["normal"])
  } else {
    direction <- 1
  }
  
  diab_effect[i] <- direction * eta_squared
  diab_p[i] <- kw_test$p.value
}

# Prepare continuous variables plot
rownames(cont_cor_matrix) <- food_cols
colnames(cont_cor_matrix) <- continuous_vars
rownames(cont_p_matrix) <- food_cols
colnames(cont_p_matrix) <- continuous_vars

# Clustering for continuous variables
food_dist <- dist(cont_cor_matrix)
food_clust <- hclust(food_dist)
food_order <- food_clust$labels[food_clust$order]

clinical_dist <- dist(t(cont_cor_matrix))
clinical_clust <- hclust(clinical_dist)
clinical_order <- clinical_clust$labels[clinical_clust$order]

# Prepare data for plots
# Continuous variables
melted_cont <- melt(cont_cor_matrix)
melted_cont_p <- melt(cont_p_matrix)
colnames(melted_cont) <- c("Food", "Clinical", "Correlation")
melted_cont$Significant <- melted_cont_p$value < 0.05
melted_cont$Food <- factor(melted_cont$Food, levels = food_order)
melted_cont$Clinical <- factor(melted_cont$Clinical, levels = clinical_order)

# Diabetes status
diab_df <- data.frame(
  Food = food_cols,
  Effect = diab_effect,
  P_value = diab_p
)
diab_df$Food <- factor(diab_df$Food, levels = food_order)
diab_df$Significant <- diab_df$P_value < 0.05

# Create continuous variables plot
p1 <- ggplot(melted_cont, aes(x = Clinical, y = Food, fill = Correlation)) +
  geom_tile() +
  scale_fill_gradientn(colors = rev(colorRampPalette(brewer.pal(11, "RdBu"))(100)), 
                      limits = c(-0.15, 0.15)) +
  geom_text(aes(label = ifelse(Significant, "*", "")), 
            color = "black", size = 5) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(face = "italic"),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold")
  ) +
  labs(
    title = "Food Correlations with Clinical Variables",
    subtitle = "Adjusted for Gender, Age, and Ethnicity\n* indicates p < 0.05",
    x = "Clinical Variables",
    y = "Food",
    fill = "Correlation"
  )

# Create diabetes status plot
p2 <- ggplot(diab_df, aes(y = Food, x = Effect)) +
  geom_bar(stat = "identity", aes(fill = abs(Effect))) +
  #scale_fill_gradientn(colors = rev(brewer.pal(9, "PuBu"))) +
  scale_fill_gradient2(low = "white", high = "darkmagenta") +
  geom_text(aes(label = ifelse(Significant, "*", ""), x = 0), 
            color = "black", size = 5) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(face = "italic"),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold")
  ) +
  labs(
    title = "Food Associations with Diabetes Status",
    subtitle = "Adjusted for Gender, Age, and Ethnicity\n* indicates p < 0.05",
    x = "Effect Size (Eta Squared)",
    y = NULL,
    fill = "Absolute\nEffect Size"
  )

# Arrange plots side by side
grid.arrange(p1, p2, ncol = 2, widths = c(1.5, 1))

# Optionally save the combined plot
ggsave("figure/clinical_food_associations3.pdf", arrangeGrob(p1, p2, ncol = 2, widths = c(1.5, 1)), width = 15, height = 10)
```

## separate ethnic groups
```{r}

# Function to perform analysis for one ethnic group
analyze_ethnic_group <- function(merged_data_f, ethnic_group) {
  # Filter data for the specific ethnic group
  ethnic_data <- merged_data_f %>% 
    filter(FREG5_Ethnic_Group == ethnic_group)
  
  # Get columns
  food_cols <- colnames(food_data)[colnames(food_data) != "Participant_ID"]
  continuous_vars <- c("DLAB27_Hba1C_Percent", "DLAB83_CRP", 
                      "DLAB81_Insulin", "DLAB25_Gluf_Mmol_L", "BMI", "HOMA")
  
  # Initialize matrices
  cont_cor_matrix <- matrix(NA, nrow = length(food_cols), ncol = length(continuous_vars))
  cont_p_matrix <- matrix(NA, nrow = length(food_cols), ncol = length(continuous_vars))
  diab_effect <- numeric(length(food_cols))
  diab_p <- numeric(length(food_cols))
  
  # Calculate correlations for continuous variables
  for(i in 1:length(food_cols)) {
    for(j in 1:length(continuous_vars)) {
      temp_data <- data.frame(
        food = rank(ethnic_data[[food_cols[i]]]),
        clinical = rank(ethnic_data[[continuous_vars[j]]]),
        gender = as.numeric(ethnic_data$FREG7_Gender == "M"),
        age = rank(ethnic_data$FREG8_Age)
      )
      
      # Now adjusting only for gender and age
      pcor_result <- pcor.test(temp_data$food, temp_data$clinical,
                              temp_data[,c("gender", "age")])
      
      cont_cor_matrix[i,j] <- pmin(pmax(pcor_result$estimate, -0.15), 0.15)
      cont_p_matrix[i,j] <- pcor_result$p.value
    }
    
    # Calculate diabetes association
    temp_data <- data.frame(
      food_abundance = ethnic_data[[food_cols[i]]],
      gender = ethnic_data$FREG7_Gender,
      age = ethnic_data$FREG8_Age,
      diabetes = ethnic_data$diabetes_status
    )
    
    # Adjust for gender and age only
    model <- lm(food_abundance ~ gender + age, data = temp_data)
    food_residuals <- residuals(model)
    
    kw_test <- kruskal.test(food_residuals ~ temp_data$diabetes)
    
    groups <- split(food_residuals, temp_data$diabetes)
    means <- sapply(groups, mean)
    overall_mean <- mean(food_residuals)
    ss_between <- sum(sapply(groups, length) * (means - overall_mean)^2)
    ss_total <- sum((food_residuals - overall_mean)^2)
    eta_squared <- ss_between / ss_total
    
    if("diabetes" %in% names(means) && "normal" %in% names(means)) {
      direction <- sign(means["diabetes"] - means["normal"])
    } else {
      direction <- 1
    }
    
    diab_effect[i] <- direction * eta_squared
    diab_p[i] <- kw_test$p.value
  }
  
  # Prepare matrices
  rownames(cont_cor_matrix) <- food_cols
  colnames(cont_cor_matrix) <- continuous_vars
  rownames(cont_p_matrix) <- food_cols
  colnames(cont_p_matrix) <- continuous_vars
  
  # Return results
  list(
    cont_cor = cont_cor_matrix,
    cont_p = cont_p_matrix,
    diab_effect = diab_effect,
    diab_p = diab_p
  )
}

# Analyze each ethnic group
results_C <- analyze_ethnic_group(merged_data_f, "C")
results_I <- analyze_ethnic_group(merged_data_f, "I")
results_M <- analyze_ethnic_group(merged_data_f, "M")

# Function to create plots for one ethnic group
create_plots <- function(results, food_cols, continuous_vars, group_name) {
  # Prepare continuous variables plot
  melted_cont <- melt(results$cont_cor)
  melted_cont_p <- melt(results$cont_p)
  colnames(melted_cont) <- c("Food", "Clinical", "Correlation")
  melted_cont$Significant <- melted_cont_p$value < 0.05
  
  # Create continuous variables plot
  p1 <- ggplot(melted_cont, aes(x = Clinical, y = Food, fill = Correlation)) +
    geom_tile() +
    scale_fill_gradientn(colors = rev(colorRampPalette(brewer.pal(11, "RdBu"))(100)), 
                        limits = c(-0.15, 0.15)) +
    geom_text(aes(label = ifelse(Significant, "*", "")), 
              color = "black", size = 5) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.text.y = element_text(face = "italic"),
      axis.title = element_text(face = "bold"),
      legend.title = element_text(face = "bold")
    ) +
    labs(
      title = paste("Food Correlations with Clinical Variables -", group_name),
      subtitle = "Adjusted for Gender and Age\n* indicates p < 0.05",
      x = "Clinical Variables",
      y = "Food",
      fill = "Correlation"
    )
  
  # Prepare diabetes plot data
  diab_df <- data.frame(
    Food = food_cols,
    Effect = results$diab_effect,
    P_value = results$diab_p
  )
  diab_df$Significant <- diab_df$P_value < 0.05
  
  # Create diabetes plot
  p2 <- ggplot(diab_df, aes(y = Food, x = Effect)) +
    geom_bar(stat = "identity", aes(fill = abs(Effect))) +
    scale_fill_gradient2(low = "white", high = "plum4") +
    geom_text(aes(label = ifelse(Significant, "*", ""), x = 0), 
              color = "black", size = 5) +
    theme_minimal() +
    theme(
      axis.text.y = element_text(face = "italic"),
      axis.title = element_text(face = "bold"),
      legend.title = element_text(face = "bold")
    ) +
    labs(
      title = paste("Food Associations with Diabetes Status -", group_name),
      subtitle = "Adjusted for Gender and Age\n* indicates p < 0.05",
      x = "Effect Size (Eta Squared)",
      y = NULL,
      fill = "Absolute\nEffect Size"
    )
  
  list(continuous = p1, diabetes = p2)
}
```


```{r}
# Create plots for each ethnic group
plots_C <- create_plots(results_C, food_cols, continuous_vars, "Chinese")
plots_I <- create_plots(results_I, food_cols, continuous_vars, "Indian")
plots_M <- create_plots(results_M, food_cols, continuous_vars, "Malay")

plots_C
plots_I
plots_M
```


```{r}
# Arrange all plots
# Continuous variables plots on top row, diabetes plots on bottom row
grid.arrange(
  plots_C$continuous, plots_I$continuous, plots_M$continuous,
  plots_C$diabetes, plots_I$diabetes, plots_M$diabetes,
  ncol = 3, nrow = 2
)


ggsave("figure/ethnic_food_clinical_analysis.pdf",
       arrangeGrob(plots_C$continuous, plots_I$continuous, plots_M$continuous,
                  plots_C$diabetes, plots_I$diabetes, plots_M$diabetes,
                  ncol = 3, nrow = 2),
       width = 24, height = 16)
```



## volcano diabete vs non-diabete
```{r}
threshold <- nrow(species1) * 0.05
species_filtered2 <- species1 %>% select_if(function(x) sum(x > 0) > threshold)
species_filtered2

merged_data_filter = species_filtered2 %>%
  inner_join(metadata, by = "Participant_ID")
merged_data_filter
```

```{r}
# Create a new diabetes status grouping
merged_data_filter <- merged_data_filter %>%
  mutate(diabetes_group = case_when(
    diabetes_status == "diabetes" ~ "diabetes",
    diabetes_status %in% c("pre-diabetes", "normal") ~ "non_diabetes"
  ))

merged_data_filter <- merged_data_filter %>%
  mutate(obesity_status = case_when(
    BMI < 18.5 ~ "Underweight",
    BMI >= 18.5 & BMI < 23 ~ "Normal",
    BMI >= 23 & BMI < 25 ~ "Overweight",
    BMI >= 25 & BMI < 30 ~ "Obese Class I",
    BMI >= 30 ~ "Obese Class II",
    TRUE ~ NA_character_  # For any missing or invalid BMI values
  ))

# If you want a simpler classification, you can use this instead:
merged_data_filter <- merged_data_filter %>%
  mutate(obesity_binary = case_when(
    BMI < 25 ~ "Non-obese",
    BMI >= 25 ~ "Obese",
    TRUE ~ NA_character_
  ))
```


```{r}
# Get species columns (excluding Participant_ID)
# species_cols <- colnames(species_filtered2)[-1]  # Remove first column
# species_cols
# # Function to perform t-test for each species
# calculate_stats <- function(species_name) {
#   t_test <- t.test(
#     get(species_name) ~ diabetes_group, 
#     data = diabetes_data
#   )
#   
#   # Calculate mean values for each group
#   means <- diabetes_data %>%
#     group_by(diabetes_group) %>%
#     summarise(mean_val = mean(get(species_name))) %>%
#     pivot_wider(names_from = diabetes_group, values_from = mean_val)
#   
#   # Calculate log2 fold change
#   log2FC <- log2(means$diabetes / means$non_diabetes)
#   
#   return(data.frame(
#     species = species_name,
#     log2FoldChange = log2FC,
#     pvalue = t_test$p.value
#   ))
# }
# 
# # Apply statistical testing to all species
# volcano_data <- map_dfr(species_cols, calculate_stats)
# 
# # Add adjusted p-values
# volcano_data$padj <- p.adjust(volcano_data$pvalue, method = "BH")
# 
# # Create volcano plot using ggplot2
# library(ggplot2)
# library(ggrepel)  # For label repulsion
# 
# ggplot(volcano_data, aes(x = log2FoldChange, y = -log10(padj))) +
#   geom_point(aes(color = abs(log2FoldChange) > 0.5 & padj < 0.1)) +
#   geom_text_repel(
#     data = subset(volcano_data, abs(log2FoldChange) > 0.5 & padj < 0.1),
#     aes(label = species),
#     max.overlaps = 15
#   ) +
#   geom_hline(yintercept = -log10(0.1), linetype = "dashed") +
#   geom_vline(xintercept = c(-0.5, 0.5), linetype = "dashed") +
#   scale_color_manual(values = c("grey", "red")) +
#   labs(
#     x = "log2 Fold Change (Diabetes vs Non-diabetes)",
#     y = "-log10 adjusted p-value",
#     title = "Volcano Plot of Species Abundance Changes in Diabetes vs Non-diabetes"
#   ) +
#   theme_minimal() +
#   theme(legend.position = "none")
```


## literature species
```{r}
diabetes_enriched <- c(
  "Lactobacillus gasseri",
  "Streptococcus mutans",
  "Clostridium citroniae",
  "Clostridium bolteae",
  "Escherichia coli",
  "Bacteroides fragilis",
  "Streptococcus salivarius",  # Example Streptococcus species, you may need to adjust based on your data
  "Streptococcus parasanguinis"
  # Add more species from literature
)

diabetes_depleted <- c(
  "Roseburia intestinalis",
  "Faecalibacterium prausnitzii",
  "Butyrivibrio crossotus",
  "Coprococcus eutactus", 
  "Turicibacter sanguinis",
  "Bacteroides plebeius",
  "Ruminococcus lactaris"
  # Add more species from literature
)

obesity_enriched <- c(
  "Ruminococcus gnavus",
  "Lactobacillus reuteri",
  "Lactobacillus gasseri"
  # Add more species from literature
)

obesity_depleted <- c(
  "Akkermansia muciniphila",
  "Lactobacillus paracasei",
  "Methanobacteriales smithii",
  "Bifidobacterium animalis"
  # Add more species from literature
)
# Function to analyze trends based on literature species
analyze_literature_species_trends <- function(species_list, condition_data, condition_column, expected_direction) {
  
  # Filter for species that exist in your dataset
  existing_species <- intersect(species_list, colnames(condition_data))
  
  if (length(existing_species) == 0) {
    message("No matching species found in the dataset for this analysis.")
    return(NULL)
  }
  
  # Analyze each species
  results <- map_dfr(existing_species, function(species) {
    # Calculate mean values for each group
    means <- condition_data %>%
      group_by(get(condition_column)) %>%
      summarise(mean_val = mean(.data[[species]], na.rm = TRUE)) %>%
      pivot_wider(names_from = `get(condition_column)`, values_from = mean_val)
    
    # Calculate log2 fold change
    log2FC <- log2(means[[2]] / means[[1]])
    
    # Check if the trend matches the expected direction
    trend_matches <- if (expected_direction == "enriched") {
      log2FC > 0
    } else {
      log2FC < 0
    }
    
    data.frame(
      species = species,
      log2FoldChange = log2FC,
      matches_literature = trend_matches
    )
  })
  
  return(results)
}

# Analyze diabetes-associated species trends
diabetes_enriched_trends <- analyze_literature_species_trends(
  diabetes_enriched,
  merged_data_filter,
  "diabetes_group",
  "enriched"
)

diabetes_depleted_trends <- analyze_literature_species_trends(
  diabetes_depleted,
  merged_data_filter,
  "diabetes_group",
  "depleted"
)

# Analyze obesity-associated species trends
obesity_enriched_trends <- analyze_literature_species_trends(
  obesity_enriched,
  merged_data_filter,
  "obesity_binary",
  "enriched"
)

obesity_depleted_trends <- analyze_literature_species_trends(
  obesity_depleted,
  merged_data_filter,
  "obesity_binary",
  "depleted"
)

# Combine and summarize results
create_summary_table <- function(trends, condition) {
  if (is.null(trends) || nrow(trends) == 0) return(NULL)
  
  trends %>%
    mutate(condition = condition) %>%
    dplyr::select(species, log2FoldChange, matches_literature, condition)
}

diabetes_enriched_summary <- create_summary_table(diabetes_enriched_trends, "Diabetes Enriched")
diabetes_depleted_summary <- create_summary_table(diabetes_depleted_trends, "Diabetes Depleted")
obesity_enriched_summary <- create_summary_table(obesity_enriched_trends, "Obesity Enriched")
obesity_depleted_summary <- create_summary_table(obesity_depleted_trends, "Obesity Depleted")

# Combine summaries into a single table
all_summaries <- bind_rows(
  diabetes_enriched_summary,
  diabetes_depleted_summary,
  obesity_enriched_summary,
  obesity_depleted_summary
)

# View the summary table
library(knitr)
kable(all_summaries, caption = "Summary of Literature Trends in Microbiome Data")

```

```{r}
# Function to create violin plots for species with log transformation
plot_violin_species_log <- function(species_list, condition_data, condition_column, title) {
  
  # Filter for species that exist in your dataset
  existing_species <- intersect(species_list, colnames(condition_data))
  
  if (length(existing_species) == 0) {
    message("No matching species found in the dataset for this plot.")
    return(NULL)
  }
  
  # Reshape the data for plotting
  plot_data <- condition_data %>%
    dplyr::select(all_of(existing_species), condition_column) %>%
    pivot_longer(cols = -all_of(condition_column), names_to = "species", values_to = "abundance") %>%
    mutate(abundance_log = log10(abundance + 1))  # Log transform (add 1 to avoid log(0))
  
  # Create violin plot
  ggplot(plot_data, aes(x = .data[[condition_column]], y = abundance_log, fill = .data[[condition_column]])) +
    geom_violin(alpha = 0.5) +
    geom_jitter(width = 0.2, size = 0.8, alpha = 0.6) +
    facet_wrap(~ species, scales = "free", ncol = 3) +  # One plot per species
    theme_minimal() +
    labs(
      title = title,
      x = condition_column,
      y = "Log10 Abundance",
      fill = condition_column
    ) +
    scale_fill_manual(values = c("Non-obese" = "blue", "Obese" = "orange", "Non-diabetic" = "blue", "Diabetic" = "orange"))
}

# Create separate plots for obesity and diabetes
# Obesity Enriched
obesity_enriched_plot <- plot_violin_species_log(
  obesity_enriched, 
  merged_data_filter, 
  "obesity_binary", 
  "Obesity Enriched Species (Log Transformed)"
)

# Obesity Depleted
obesity_depleted_plot <- plot_violin_species_log(
  obesity_depleted, 
  merged_data_filter, 
  "obesity_binary", 
  "Obesity Depleted Species (Log Transformed)"
)

# Diabetes Enriched
diabetes_enriched_plot <- plot_violin_species_log(
  diabetes_enriched, 
  merged_data_filter, 
  "diabetes_group", 
  "Diabetes Enriched Species (Log Transformed)"
)

# Diabetes Depleted
diabetes_depleted_plot <- plot_violin_species_log(
  diabetes_depleted, 
  merged_data_filter, 
  "diabetes_group", 
  "Diabetes Depleted Species (Log Transformed)"
)

# Save plots
ggsave("figure/obesity_enriched_species_log.png", obesity_enriched_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/obesity_depleted_species_log.png", obesity_depleted_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/diabetes_enriched_species_log.png", diabetes_enriched_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/diabetes_depleted_species_log.png", diabetes_depleted_plot, width = 12, height = 8, dpi = 300)

```
## separate ethnicity
```{r}
merged_data_filter
# Function to create violin plots with log transformation, stratification, and consistent color mapping
plot_violin_species_log_ethnic <- function(species_list, condition_data, condition_column, title) {
  
  # Filter for species that exist in your dataset
  existing_species <- intersect(species_list, colnames(condition_data))
  
  if (length(existing_species) == 0) {
    message("No matching species found in the dataset for this plot.")
    return(NULL)
  }
  
  # Standardize the values in the condition column to match the expected levels
  condition_data <- condition_data %>%
    filter(!is.na(.data[[condition_column]]) & FREG5_Ethnic_Group %in% c("C", "I", "M")) %>%  # Remove NA and exclude "O"
    mutate(!!condition_column := case_when(
      .data[[condition_column]] == "non_diabetes" ~ "Non-diabetic",
      .data[[condition_column]] == "diabetes" ~ "Diabetic",
      .data[[condition_column]] == "non_obese" ~ "Non-obese",
      .data[[condition_column]] == "obese" ~ "Obese",
      TRUE ~ as.character(.data[[condition_column]])
    )) %>%
    mutate(!!condition_column := factor(
      .data[[condition_column]], 
      levels = c("Non-diabetic", "Diabetic", "Non-obese", "Obese")
    ))
  
  # Reshape data for plotting
  plot_data <- condition_data %>%
    dplyr::select(all_of(existing_species), condition_column, FREG5_Ethnic_Group) %>%
    pivot_longer(cols = -c(all_of(condition_column), FREG5_Ethnic_Group), 
                 names_to = "species", values_to = "abundance") %>%
    mutate(abundance_log = log10(abundance + 1))  # Log transform (add 1 to avoid log(0))
  
  # Debugging: Check levels
  print(levels(plot_data[[condition_column]]))
  
  # Create violin plot
  ggplot(plot_data, aes(x = FREG5_Ethnic_Group, y = abundance_log, fill = .data[[condition_column]])) +
    geom_violin(alpha = 0.5) +
    geom_jitter(aes(color = .data[[condition_column]]), 
                width = 0.2, size = 0.8, alpha = 0.6) +  # Adjust jitter without dodge
    facet_wrap(~ species, scales = "free", ncol = 3) +  # One plot per species
    theme_minimal() +
    labs(
      title = title,
      x = "Ethnic Group",
      y = "Log10 Abundance",
      fill = "Condition",
      color = "Condition"
    ) +
    scale_fill_manual(values = c("Non-diabetic" = "#4E6691", "Diabetic" = "#B8474D",
                                 "Non-obese" = "#4E6691", "Obese" = "#B8474D")) +
    scale_color_manual(values = c("Non-diabetic" = "#4E6691", "Diabetic" = "#B8474D",
                                  "Non-obese" = "#4E6691", "Obese" = "#B8474D")) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.title = element_text(size = 12, face = "bold"),
      plot.title = element_text(size = 14, face = "bold")
    )
}

# Create separate plots for diabetes and obesity
# Diabetes Enriched
diabetes_enriched_plot <- plot_violin_species_log_ethnic(
  diabetes_enriched, 
  merged_data_filter, 
  "diabetes_group", 
  "Diabetes Enriched Species (Log Transformed, Stratified by Ethnic Group)"
)

# Diabetes Depleted
diabetes_depleted_plot <- plot_violin_species_log_ethnic(
  diabetes_depleted, 
  merged_data_filter, 
  "diabetes_group", 
  "Diabetes Depleted Species (Log Transformed, Stratified by Ethnic Group)"
)

# Obesity Enriched
obesity_enriched_plot <- plot_violin_species_log_ethnic(
  obesity_enriched, 
  merged_data_filter, 
  "obesity_binary", 
  "Obesity Enriched Species (Log Transformed, Stratified by Ethnic Group)"
)

# Obesity Depleted
obesity_depleted_plot <- plot_violin_species_log_ethnic(
  obesity_depleted, 
  merged_data_filter, 
  "obesity_binary", 
  "Obesity Depleted Species (Log Transformed, Stratified by Ethnic Group)"
)

# Save plots
ggsave("figure/diabetes_enriched_species_log_ethnic.png", diabetes_enriched_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/diabetes_depleted_species_log_ethnic.png", diabetes_depleted_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/obesity_enriched_species_log_ethnic.png", obesity_enriched_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/obesity_depleted_species_log_ethnic.png", obesity_depleted_plot, width = 12, height = 8, dpi = 300)

```
## separate ethnic box plot
```{r}
plot_box_species_log_ethnic <- function(species_list, condition_data, condition_column, title) {
  
  # Filter for species that exist in your dataset
  existing_species <- intersect(species_list, colnames(condition_data))
  
  if (length(existing_species) == 0) {
    message("No matching species found in the dataset for this plot.")
    return(NULL)
  }
  
  # Standardize the values in the condition column to match the expected levels
  condition_data <- condition_data %>%
    filter(!is.na(.data[[condition_column]]) & FREG5_Ethnic_Group %in% c("C", "I", "M")) %>%  # Remove NA and exclude "O"
    mutate(!!condition_column := case_when(
      .data[[condition_column]] == "non_diabetes" ~ "Non-diabetic",
      .data[[condition_column]] == "diabetes" ~ "Diabetic",
      .data[[condition_column]] == "non_obese" ~ "Non-obese",
      .data[[condition_column]] == "obese" ~ "Obese",
      TRUE ~ as.character(.data[[condition_column]])
    )) %>%
    mutate(!!condition_column := factor(
      .data[[condition_column]], 
      levels = c("Non-diabetic", "Diabetic", "Non-obese", "Obese")
    ))
  
  # Reshape data for plotting
  plot_data <- condition_data %>%
    dplyr::select(all_of(existing_species), condition_column, FREG5_Ethnic_Group) %>%
    pivot_longer(cols = -c(all_of(condition_column), FREG5_Ethnic_Group), 
                 names_to = "species", values_to = "abundance") %>%
    mutate(abundance_log = log10(abundance + 1))  # Log transform (add 1 to avoid log(0))
  
  # Create box plot with dots
  ggplot(plot_data, aes(x = FREG5_Ethnic_Group, y = abundance_log, fill = .data[[condition_column]])) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA) +  # Remove default outliers
    geom_jitter(aes(color = .data[[condition_column]]), 
                position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2),
                size = 0.8, alpha = 0.6) +
    facet_wrap(~ species, scales = "free", ncol = 3) +  # One plot per species
    theme_minimal() +
    labs(
      title = title,
      x = "Ethnic Group",
      y = "Log10 Abundance",
      fill = "Condition",
      color = "Condition"
    ) +
    scale_fill_manual(values = c("Non-diabetic" = "#4E6691", "Diabetic" = "#B8474D",
                                "Non-obese" = "#4E6691", "Obese" = "#B8474D")) +
    scale_color_manual(values = c("Non-diabetic" = "#4E6691", "Diabetic" = "#B8474D",
                                 "Non-obese" = "#4E6691", "Obese" = "#B8474D")) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.title = element_text(size = 12, face = "bold"),
      plot.title = element_text(size = 14, face = "bold")
    )
}

# Create separate plots for diabetes and obesity
# Diabetes Enriched
diabetes_enriched_plot <- plot_box_species_log_ethnic(
  diabetes_enriched, 
  merged_data_filter, 
  "diabetes_group", 
  "Diabetes Enriched Species (Log Transformed, Stratified by Ethnic Group)"
)

# Diabetes Depleted
diabetes_depleted_plot <- plot_box_species_log_ethnic(
  diabetes_depleted, 
  merged_data_filter, 
  "diabetes_group", 
  "Diabetes Depleted Species (Log Transformed, Stratified by Ethnic Group)"
)

# Obesity Enriched
obesity_enriched_plot <- plot_box_species_log_ethnic(
  obesity_enriched, 
  merged_data_filter, 
  "obesity_binary", 
  "Obesity Enriched Species (Log Transformed, Stratified by Ethnic Group)"
)

# Obesity Depleted
obesity_depleted_plot <- plot_box_species_log_ethnic(
  obesity_depleted, 
  merged_data_filter, 
  "obesity_binary", 
  "Obesity Depleted Species (Log Transformed, Stratified by Ethnic Group)"
)

# Save plots
ggsave("figure/diabetes_enriched_species_log_ethnic1.png", diabetes_enriched_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/diabetes_depleted_species_log_ethnic1.png", diabetes_depleted_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/obesity_enriched_species_log_ethnic1.png", obesity_enriched_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/obesity_depleted_species_log_ethnic1.png", obesity_depleted_plot, width = 12, height = 8, dpi = 300)
```

# with p value
```{r}
# Function to create box plots with pre-calculated p-values
plot_box_species_log_ethnic <- function(species_list, condition_data, condition_column, 
                                      stratified_results, title) {
  
  # Filter for species that exist in your dataset
  existing_species <- intersect(species_list, colnames(condition_data))
  
  if (length(existing_species) == 0) {
    message("No matching species found in the dataset for this plot.")
    return(NULL)
  }
  
  # Standardize the values in the condition column
  condition_data <- condition_data %>%
    filter(!is.na(.data[[condition_column]]) & FREG5_Ethnic_Group %in% c("C", "I", "M")) %>%
    mutate(condition = case_when(
      .data[[condition_column]] == "non_diabetes" ~ "Non-diabetic",
      .data[[condition_column]] == "diabetes" ~ "Diabetic",
      .data[[condition_column]] == "non_obese" ~ "Non-obese",
      .data[[condition_column]] == "obese" ~ "Obese",
      TRUE ~ as.character(.data[[condition_column]])
    )) %>%
    mutate(condition = factor(
      condition, 
      levels = c("Non-diabetic", "Diabetic", "Non-obese", "Obese")
    ))
  
  # Reshape data for plotting
  plot_data <- condition_data %>%
    dplyr::select(all_of(existing_species), condition, FREG5_Ethnic_Group) %>%
    pivot_longer(cols = -c(condition, FREG5_Ethnic_Group), 
                 names_to = "species", values_to = "abundance") %>%
    mutate(abundance_log = log10(abundance + 1))
  
  # Prepare p-values for plotting
  p_values <- stratified_results %>%
    group_by(species, FREG5_Ethnic_Group) %>%
    mutate(
      p_label = sprintf("p = %.3f", pvalue),
      # Calculate y position
      y_pos = max(plot_data$abundance_log[plot_data$species == species]) + 0.2
    )
  
  # Create box plot with dots and p-values
  ggplot(plot_data, aes(x = FREG5_Ethnic_Group, y = abundance_log, fill = condition)) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA) +
    geom_jitter(aes(color = condition), 
                position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2),
                size = 0.8, alpha = 0.6) +
    # Add p-value annotations
    geom_text(data = p_values,
              aes(x = FREG5_Ethnic_Group, y = y_pos, label = p_label),
              position = position_dodge(width = 0.75),
              size = 3) +
    facet_wrap(~ species, scales = "free", ncol = 3) +
    theme_minimal() +
    labs(
      title = title,
      x = "Ethnic Group",
      y = "Log10 Abundance",
      fill = "Condition",
      color = "Condition"
    ) +
    scale_fill_manual(values = c("Non-diabetic" = "#4E6691", "Diabetic" = "#B8474D",
                                "Non-obese" = "#4E6691", "Obese" = "#B8474D")) +
    scale_color_manual(values = c("Non-diabetic" = "#4E6691", "Diabetic" = "#B8474D",
                                 "Non-obese" = "#4E6691", "Obese" = "#B8474D")) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.title = element_text(size = 12, face = "bold"),
      plot.title = element_text(size = 14, face = "bold")
    )
}

# Create separate plots for diabetes and obesity using pre-calculated p-values
# Diabetes Enriched
diabetes_enriched_plot <- plot_box_species_log_ethnic(
  diabetes_enriched, 
  merged_data_filter, 
  "diabetes_group",
  diabetes_enriched_stratified,
  "Diabetes Enriched Species (Log Transformed, Stratified by Ethnic Group)"
)

# Diabetes Depleted
diabetes_depleted_plot <- plot_box_species_log_ethnic(
  diabetes_depleted, 
  merged_data_filter, 
  "diabetes_group",
  diabetes_depleted_stratified,
  "Diabetes Depleted Species (Log Transformed, Stratified by Ethnic Group)"
)

# Obesity Enriched
obesity_enriched_plot <- plot_box_species_log_ethnic(
  obesity_enriched, 
  merged_data_filter, 
  "obesity_binary",
  obesity_enriched_stratified,
  "Obesity Enriched Species (Log Transformed, Stratified by Ethnic Group)"
)

# Obesity Depleted
obesity_depleted_plot <- plot_box_species_log_ethnic(
  obesity_depleted, 
  merged_data_filter, 
  "obesity_binary",
  obesity_depleted_stratified,
  "Obesity Depleted Species (Log Transformed, Stratified by Ethnic Group)"
)

# Save plots
ggsave("figure/diabetes_enriched_species_log_ethnic_box.png", diabetes_enriched_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/diabetes_depleted_species_log_ethnic_box.png", diabetes_depleted_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/obesity_enriched_species_log_ethnic_box.png", obesity_enriched_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/obesity_depleted_species_log_ethnic_box.png", obesity_depleted_plot, width = 12, height = 8, dpi = 300)
```


```{r}
library(ggrepel)
library(dplyr)
diabetes_enriched <- c(
  "Lactobacillus gasseri",
  "Streptococcus mutans",
  "Clostridium citroniae",
  "Clostridium bolteae",
  "Escherichia coli",
  "Bacteroides fragilis",
  "Streptococcus salivarius",  # Example Streptococcus species, you may need to adjust based on your data
  "Streptococcus parasanguinis"
  # Add more species from literature
)

diabetes_depleted <- c(
  "Roseburia intestinalis",
  "Faecalibacterium prausnitzii",
  "Butyrivibrio crossotus",
  "Coprococcus eutactus", 
  "Turicibacter sanguinis",
  "Bacteroides plebeius",
  "Ruminococcus lactaris"
  # Add more species from literature
)

obesity_enriched <- c(
  "Ruminococcus gnavus",
  "Lactobacillus reuteri",
  "Lactobacillus gasseri"
  # Add more species from literature
)

obesity_depleted <- c(
  "Akkermansia muciniphila",
  "Lactobacillus paracasei",
  "Methanobacteriales smithii",
  "Bifidobacterium animalis"
  # Add more species from literature
)

# Function to analyze literature-based species
analyze_literature_species <- function(species_list, condition_data, condition_column, 
                                     expected_direction = "enriched") {
  
  # Filter for species that exist in your dataset
  existing_species <- intersect(species_list, colnames(condition_data))
  
  if(length(existing_species) == 0) {
    return(NULL)
  }
  
  # Analyze each species
  results <- map_dfr(existing_species, function(species) {
    # Perform Wilcoxon test
    test_result <- wilcox.test(
      get(species) ~ get(condition_column), 
      data = condition_data
    )
    
    # Calculate fold change
    means <- condition_data %>%
      group_by(get(condition_column)) %>%
      summarise(mean_val = mean(get(species))) %>%
      pivot_wider(names_from = `get(condition_column)`, values_from = mean_val)
    
    # Assuming binary columns like diabetes/non-diabetes or obese/non-obese
    if(expected_direction == "enriched") {
      log2FC <- log2(means[[2]] / means[[1]])  # condition/control
    } else {
      log2FC <- -log2(means[[2]] / means[[1]])  # negative of condition/control
    }
    
    data.frame(
      species = species,
      pvalue = test_result$p.value,
      log2FoldChange = log2FC,
      matches_literature = sign(log2FC) == if(expected_direction == "enriched") 1 else -1
    )
  })
  
  results$padj <- p.adjust(results$pvalue, method = "BH")
  return(results)
}

# Analyze diabetes-associated species
diabetes_enriched_results <- analyze_literature_species(
  diabetes_enriched, 
  merged_data_filter,
  "diabetes_group",
  "enriched"
)

diabetes_depleted_results <- analyze_literature_species(
  diabetes_depleted, 
  merged_data_filter,
  "diabetes_group",
  "depleted"
)

# Analyze obesity-associated species
obesity_enriched_results <- analyze_literature_species(
  obesity_enriched, 
  merged_data_filter,
  "obesity_binary",
  "enriched"
)

obesity_depleted_results <- analyze_literature_species(
  obesity_depleted, 
  merged_data_filter,
  "obesity_binary",
  "depleted"
)

# Create summary visualization
library(gridExtra)

plot_literature_comparison <- function(results, title) {
  if(is.null(results) || nrow(results) == 0) return(NULL)
  
  ggplot(results, aes(x = log2FoldChange, y = -log10(padj))) +
    geom_point(aes(color = matches_literature)) +
    geom_text_repel(aes(label = species)) +
    scale_color_manual(values = c("red", "blue")) +
    labs(title = title) +
    theme_minimal()
}

# Create plots
p1 <- plot_literature_comparison(diabetes_enriched_results, "Diabetes Enriched Species")
p2 <- plot_literature_comparison(diabetes_depleted_results, "Diabetes Depleted Species")
p3 <- plot_literature_comparison(obesity_enriched_results, "Obesity Enriched Species")
p4 <- plot_literature_comparison(obesity_depleted_results, "Obesity Depleted Species")

# Arrange plots
g = grid.arrange(p1, p2, p3, p4, ncol = 2)
ggsave("figure/microbiota_diabete_obesity2.png", g, width = 12, height = 10, dpi = 300)
# Create summary table
create_summary_table <- function(results, condition) {
  if(is.null(results) || nrow(results) == 0) return(NULL)
  
  results %>%
    mutate(
      significant = padj < 0.1,
      agrees_with_literature = matches_literature & significant,
      condition = condition
    ) %>%
    dplyr::select(species, significant, agrees_with_literature, condition)
}

summary_table <- bind_rows(
  create_summary_table(diabetes_enriched_results, "Diabetes Enriched"),
  create_summary_table(diabetes_depleted_results, "Diabetes Depleted"),
  create_summary_table(obesity_enriched_results, "Obesity Enriched"),
  create_summary_table(obesity_depleted_results, "Obesity Depleted")
)

print(summary_table)
```


```{r}
analyze_literature_species_stratified <- function(species_list, condition_data, condition_column,
                                                ethnic_groups = c("C", "I", "M"),
                                                expected_direction = "enriched") {
  
  # Filter for species that exist in your dataset
  existing_species <- intersect(species_list, colnames(condition_data))
  
  if(length(existing_species) == 0) {
    return(NULL)
  }
  
  # Create a list to store results for each ethnic group
  results_list <- map(ethnic_groups, function(ethnic_group) {
    # Filter data for current ethnic group
    ethnic_data <- condition_data %>%
      filter(FREG5_Ethnic_Group == ethnic_group)
    
    # Skip if no data for this ethnic group
    if(nrow(ethnic_data) == 0) {
      return(NULL)
    }
    
    # Analyze each species within this ethnic group
    results <- map_dfr(existing_species, function(species) {
      # Check if there's enough data in each condition group
      condition_counts <- table(ethnic_data[[condition_column]])
      if(any(condition_counts < 2)) {
        return(data.frame(
          species = species,
          pvalue = NA,
          log2FoldChange = NA,
          matches_literature = NA
        ))
      }
      
      # Perform Wilcoxon test
      tryCatch({
        test_result <- wilcox.test(
          get(species) ~ get(condition_column), 
          data = ethnic_data
        )
        
        # Calculate fold change
        means <- ethnic_data %>%
          group_by(get(condition_column)) %>%
          summarise(mean_val = mean(get(species))) %>%
          pivot_wider(names_from = `get(condition_column)`, values_from = mean_val)
        
        if(expected_direction == "enriched") {
          log2FC <- log2(means[[2]] / means[[1]])  # condition/control
        } else {
          log2FC <- -log2(means[[2]] / means[[1]])  # negative of condition/control
        }
        
        data.frame(
          species = species,
          pvalue = test_result$p.value,
          log2FoldChange = log2FC,
          matches_literature = sign(log2FC) == if(expected_direction == "enriched") 1 else -1
        )
      }, error = function(e) {
        data.frame(
          species = species,
          pvalue = NA,
          log2FoldChange = NA,
          matches_literature = NA
        )
      })
    })
    
    # Add FDR correction within each ethnic group
    results$padj <- p.adjust(results$pvalue, method = "BH")
    results$FREG5_Ethnic_Group <- ethnic_group
    
    return(results)
  })
  
  # Combine results from all ethnic groups
  final_results <- bind_rows(results_list)
  return(final_results)
}

# Usage example for diabetes analysis
diabetes_enriched_stratified <- analyze_literature_species_stratified(
  diabetes_enriched, 
  merged_data_filter,
  "diabetes_group",
  c("C", "I", "M"),
  "enriched"
)

diabetes_depleted_stratified <- analyze_literature_species_stratified(
  diabetes_depleted, 
  merged_data_filter,
  "diabetes_group",
  c("C", "I", "M"),
  "depleted"
)

# Usage example for obesity analysis
obesity_enriched_stratified <- analyze_literature_species_stratified(
  obesity_enriched, 
  merged_data_filter,
  "obesity_binary",
  c("C", "I", "M"),
  "enriched"
)

obesity_depleted_stratified <- analyze_literature_species_stratified(
  obesity_depleted, 
  merged_data_filter,
  "obesity_binary",
  c("C", "I", "M"),
  "depleted"
)

# Optional: Create summary tables for each condition
# create_summary_table <- function(results_data) {
#   results_data %>%
#     filter(!is.na(pvalue)) %>%
#     group_by(FREG5_Ethnic_Group) %>%
#     summarise(
#       total_species = n(),
#       significant_species = sum(padj < 0.05),
#       matching_direction = sum(matches_literature & padj < 0.05)
#     )
# }
# 
# # Example usage of summary table
# diabetes_enriched_summary <- create_summary_table(diabetes_enriched_stratified)
# diabetes_depleted_summary <- create_summary_table(diabetes_depleted_stratified)
# obesity_enriched_summary <- create_summary_table(obesity_enriched_stratified)
# obesity_depleted_summary <- create_summary_table(obesity_depleted_stratified)
```


```{r}
diabetes_enriched_stratified
diabetes_depleted_stratified
obesity_enriched_stratified
obesity_depleted_stratified
```
##FB ratio
```{r}
firmicutes_species <- grep("^(Clostridium|Ruminococcus|Lactobacillus|Bacillus|Enterococcus|Streptococcus|Staphylococcus|Eubacterium|Roseburia|Faecalibacterium|Butyrivibrio)", 
                         colnames(merged_data_filter), value = TRUE)
bacteroidetes_species <- grep("^(Bacteroides|Prevotella|Porphyromonas|Alistipes|Parabacteroides)", 
                            colnames(merged_data_filter), value = TRUE)
# Calculate sum of relative abundances for each phylum
merged_data_filter <- merged_data_filter %>%
  mutate(
    Firmicutes_total = rowSums(select(., all_of(firmicutes_species))),
    Bacteroidetes_total = rowSums(select(., all_of(bacteroidetes_species))),
    FB_ratio = Firmicutes_total / Bacteroidetes_total
  )
```

## regression
```{r}

# firmicutes_species <- grep("^(Clostridium|Ruminococcus|Lactobacillus|Bacillus|Enterococcus|Streptococcus|Staphylococcus|Eubacterium|Roseburia|Faecalibacterium|Butyrivibrio)", 
#                          colnames(merged_data_filter), value = TRUE)
# 
# bacteroidetes_species <- grep("^(Bacteroides|Prevotella|Porphyromonas|Alistipes|Parabacteroides)", 
#                             colnames(merged_data_filter), value = TRUE)
```




```{r}
# First normalize the counts to relative abundances
# Get species columns (all except the first column from species_filtered2)
merged_data_fb <- species1 %>%
  inner_join(metadata, by = "Participant_ID")

species_cols <- colnames(species1)[-1]

# Create normalized abundances
merged_data_fb <- merged_data_fb %>%
  mutate(across(all_of(species_cols), 
                ~./rowSums(select(merged_data_fb, all_of(species_cols))), 
                .names = "{.col}_norm"))

# Now get the normalized species for each phylum
firmicutes_species_norm <- grep("^(Clostridium|Ruminococcus|Lactobacillus|Bacillus|Enterococcus|Streptococcus|Staphylococcus|Eubacterium|Roseburia|Faecalibacterium|Butyrivibrio).*_norm$", 
                         colnames(merged_data_fb), value = TRUE)

bacteroidetes_species_norm <- grep("^(Bacteroides|Prevotella|Porphyromonas|Alistipes|Parabacteroides).*_norm$", 
                            colnames(merged_data_fb), value = TRUE)

# Calculate FB ratio using normalized abundances
merged_data_fb <- merged_data_fb %>%
  mutate(
    Firmicutes_total = rowSums(select(., all_of(firmicutes_species_norm))),
    Bacteroidetes_total = rowSums(select(., all_of(bacteroidetes_species_norm))),
    FB_ratio = Firmicutes_total / Bacteroidetes_total
  )
```


```{r}
merged_data_fb <- merged_data_fb[merged_data_fb$FB_ratio <= 5, ]
# Create the plot with normalized data
ggplot(merged_data_fb, aes(x = BMI, y = FB_ratio)) +
  geom_point(color = "blue", alpha = 0.5) +
  geom_smooth(method = "lm", color = "red") +
  theme_minimal() +
  labs(
    title = "Relationship between BMI and FB Ratio (Using Normalized Abundances)",
    x = "BMI",
    y = "FB Ratio"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold")
  )

# Check the regression
model <- lm(FB_ratio ~ BMI, data = merged_data_fb)
summary(model)
```
```{r}
# Create a BMI category variable
merged_data_filter <- merged_data_filter %>%
  mutate(BMI_category = ifelse(BMI < 25, "Normal Weight (BMI < 25)", "Obese (BMI ≥ 25)"))

# Create the plot with two regression lines
ggplot(merged_data_filter, aes(x = BMI, y = FB_ratio, color = BMI_category)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE) +
  theme_minimal() +
  scale_color_manual(values = c("Normal Weight (BMI < 25)" = "blue", 
                               "Overweight/Obese (BMI ≥ 25)" = "red")) +
  labs(
    title = "FB Ratio vs BMI by Weight Category",
    x = "BMI",
    y = "FB Ratio",
    color = "BMI Category"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom"
  )

# Perform separate regression analyses
# For BMI < 25
model_normal <- lm(FB_ratio ~ BMI, 
                  data = merged_data_filter[merged_data_filter$BMI < 25,])
print("Regression results for BMI < 25:")
summary(model_normal)

# For BMI ≥ 25
model_overweight <- lm(FB_ratio ~ BMI, 
                      data = merged_data_filter[merged_data_filter$BMI >= 25,])
print("Regression results for BMI ≥ 25:")
summary(model_overweight)
```


##box plot
```{r}
# Log-transform FB_ratio to reduce the influence of extreme values
merged_data_filter <- merged_data_filter %>%
  mutate(FB_ratio_log = log10(FB_ratio + 0.001))  # Add small constant to handle zeros

# Re-run Wilcoxon test with log-transformed values
wilcox_test_result <- wilcox.test(FB_ratio_log ~ obesity_binary, data = merged_data_filter)
print(wilcox_test_result)

# Create a boxplot with log-transformed data
ggplot(merged_data_filter, aes(x = obesity_binary, y = FB_ratio_log)) +
  geom_boxplot(outlier.shape = NA) +  # Exclude outliers from boxplot display
  geom_jitter(width = 0.2, alpha = 0.3, color = "blue") +
  theme_minimal() +
  labs(
    title = "Log-Transformed Firmicutes/Bacteroidetes Ratio by Obesity Status",
    x = "Obesity Status",
    y = "Log10(F/B Ratio)"
  )

# Optional: Visualize the untransformed data but highlight the extreme value
ggplot(merged_data_filter, aes(x = obesity_binary, y = FB_ratio)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.3, color = "blue") +
  theme_minimal() +
  labs(
    title = "Firmicutes/Bacteroidetes Ratio by Obesity Status",
    x = "Obesity Status",
    y = "F/B Ratio"
  ) +
  scale_y_continuous(trans = "log10", breaks = c(1, 10, 100, 1000, 8000))  # Log scale for clarity

```






##stratify by ethnic
```{r}
# Filter out the "O" ethnic group
ethnic_group_data_filtered <- ethnic_group_data %>%
  filter(FREG5_Ethnic_Group != "O")

# Re-plot with properly aligned jitter points
ggplot(ethnic_group_data_filtered, aes(x = obesity_binary, y = FB_ratio, color = FREG5_Ethnic_Group)) +
  geom_boxplot(aes(fill = FREG5_Ethnic_Group), alpha = 0.3, outlier.shape = NA) +  # Boxplot without outlier symbols
  geom_jitter(aes(x = obesity_binary), width = 0.2, alpha = 0.5) +  # Ensure jitter aligns with box groups
  geom_point(data = subset(ethnic_group_data_filtered, FB_ratio > 1000), 
             aes(x = obesity_binary, y = FB_ratio), 
             color = "red", size = 3, shape = 8) +  # Highlight extreme value
  theme_minimal() +
  labs(
    title = "Firmicutes/Bacteroidetes Ratio by Obesity Status and Ethnic Group",
    x = "Obesity Status",
    y = "F/B Ratio (Untransformed)"
  ) +
  scale_y_continuous(trans = "log10", breaks = c(1, 10, 100, 1000, 8000)) +
  scale_fill_manual(values = c("C" = "blue", "I" = "green", "M" = "orange")) +
  scale_color_manual(values = c("C" = "blue", "I" = "green", "M" = "orange"))

```
```{r}
# Filter out the "O" ethnic group
ethnic_group_data_filtered <- ethnic_group_data %>%
  filter(FREG5_Ethnic_Group != "O")

# Re-plot: Group by ethnic group, compare obese vs non-obese
ggplot(ethnic_group_data_filtered, aes(x = obesity_binary, y = FB_ratio, color = obesity_binary)) +
  geom_boxplot(aes(fill = obesity_binary), alpha = 0.3, outlier.shape = NA) +  # Boxplot without outlier symbols
  geom_jitter(width = 0.2, alpha = 0.5) +  # Jitter points for visualization
  geom_point(data = subset(ethnic_group_data_filtered, FB_ratio > 1000), 
             aes(x = obesity_binary, y = FB_ratio), 
             color = "red", size = 3, shape = 8) +  # Highlight extreme value
  facet_wrap(~ FREG5_Ethnic_Group, scales = "free") +  # Facet by ethnic group
  theme_minimal() +
  labs(
    title = "Firmicutes/Bacteroidetes Ratio by Ethnic Group and Obesity Status",
    x = "Obesity Status",
    y = "F/B Ratio (Untransformed)"
  ) +
  scale_y_continuous(trans = "log10", breaks = c(1, 10, 100, 1000, 8000)) +
  scale_fill_manual(values = c("Non-obese" = "blue", "Obese" = "orange")) +
  scale_color_manual(values = c("Non-obese" = "blue", "Obese" = "orange"))

```

```{r}
# Calculate the distribution of obesity_binary by ethnic group
obesity_dist <- merged_data_filter %>%
  group_by(FREG5_Ethnic_Group, obesity_binary) %>%
  summarise(count = n(), .groups = 'drop') %>%
  # Calculate percentages within each ethnic group
  group_by(FREG5_Ethnic_Group) %>%
  mutate(percentage = count / sum(count) * 100)


obesity_dist <- obesity_dist %>%
  mutate(obesity_binary = factor(obesity_binary, levels = c("Obese", "Non-obese")))

# Create the bar plot with updated stacking order
ggplot(obesity_dist, aes(x = FREG5_Ethnic_Group, y = count, fill = obesity_binary)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_stack(vjust = 0.5), size = 4) +
  scale_fill_manual(values = c("Obese" = "#B8474D", "Non-obese" = "4E6691")) +
  theme_minimal() +
  labs(
    title = "Distribution of Obesity Status by Ethnic Group",
    x = "Ethnic Group",
    y = "Count",
    fill = "Obesity Status"
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 14, face = "bold")
  )


```


