---
title: "clinical_var"
author: "Ruwen Zhou"
date: "12/2/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(plyr)
library(tidyverse)
library(dplyr)
library(readxl)
library(ggplot2)
library(phyloseq)
library(vegan)
library(DESeq2)
library(pheatmap)
library(RColorBrewer)
library(ggraph)
library(igraph)
library(Hmisc)
library(writexl)

library(tidygraph)
library(ppcor)

library(reshape2)
library(ComplexHeatmap)
```


```{r}
food_data = read_xlsx("data/FFQA862.xlsx")
nutrition_data = read_xlsx("data/FFQB800.xlsx")
species1 = read_xlsx("data/species_866.xlsx")
metadata = read_xlsx("data/helios500_clinical.xlsx")

food_data <- food_data %>%
  mutate(across(c("Mushrooms", "Papaya", "Orange", "Chapati"), as.numeric))
food_data
nutrition_data
species1
metadata
```

```{r}
colnames(species1)[1] = "Participant_ID"
colnames(metadata)[2] = "Participant_ID"
threshold <- nrow(species1) * 0.5
species_filtered1 <- species1 %>% select_if(function(x) sum(x > 0) > threshold)
species_filtered2 <- species1 %>% select_if(function(x) sum(x > 0) > nrow(species1) * 0.1)
# species_filtered2 <- species %>% select_if(function(x) sum(x > 0) > threshold)
# 
# #abundance_threshold <- 0.0001
# #species_filtered2 <- species_filtered %>% select_if(function(x) sum(x) > (abundance_threshold * sum(species)))
# #species_filtered2
# species_filtered2 <- species_filtered2[, !names(species_filtered2) %in% c("unclassified", "others", "Homo sapiens")]
# #species_filtered2$`Homo sapiens`
# #species_filtered2
# food_data[is.na(food_data)] <- 0
# #food_data <- data.frame(lapply(food_data, function(x) as.numeric(as.character(x))))
# nutrition_data[is.na(nutrition_data)] <- 0
```


```{r}
unique_species <- read.csv("figure/unique_significant_species.csv")
unique_species$species
colnames(species1)[1] = "Participant_ID"
sig_species <- species1 %>%
  dplyr::select("Participant_ID", all_of(unique_species$species))

colnames(metadata)[2] = "Participant_ID"
```


```{r}
sig_species
metadata

sig_species <- sig_species %>%
  mutate(Participant_ID = as.character(Participant_ID))

metadata <- metadata %>%
  mutate(Participant_ID = as.character(Participant_ID))
```

```{r}
#species_filtered2
merged_data <- sig_species %>%
  inner_join(metadata, by = "Participant_ID")
merged_data_all = merged_data %>%
  inner_join(food_data, by = "Participant_ID")
merged_data
metadata
```
# four groups based on bifido and thosai
```{r}
# # Create binary categories using median splits
# merged_data_all$bifido_cat <- ifelse(
#   merged_data_all$`Bifidobacterium catenulatum` > 
#     median(merged_data_all$`Bifidobacterium catenulatum`), 
#   "High", "Low"
# )
# 
# merged_data_all$thosai_cat <- ifelse(
#   merged_data_all$Idli_Thosai > median(merged_data_all$Idli_Thosai), 
#   "High", "Low"
# )
# 
# # Create combined groups
# merged_data_all$combined_group <- paste(
#   merged_data_all$bifido_cat, 
#   merged_data_all$thosai_cat
# )
# 
# # Create a contingency table
# contingency_table <- table(
#   merged_data_all$combined_group, 
#   merged_data_all$diabetes_status
# )
# 
# # View the contingency table
# print(contingency_table)
```


```{r}
# Perform chi-square test
# chi_test <- chisq.test(contingency_table)
# print(chi_test)
```


```{r}
# Create a visual representation using ggplot2
# library(ggplot2)
# 
# # Convert to percentages
# prop_table <- prop.table(contingency_table, margin = 1) * 100
# 
# # Convert to data frame for plotting
# plot_data <- as.data.frame(prop_table)
# colnames(plot_data) <- c("Group", "Diabetes_Status", "Percentage")
# 
# # Create stacked bar plot
# ggplot(plot_data, aes(x = Group, y = Percentage, fill = Diabetes_Status)) +
#   geom_bar(stat = "identity") +
#   theme_minimal() +
#   labs(
#     title = "Distribution of Diabetes Status by Bifido-Thosai Groups",
#     x = "Bifido-Thosai Group",
#     y = "Percentage (%)"
#   ) +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
# Get summary statistics for each group
# library(dplyr)
# 
# group_summary <- merged_data_all %>%
#   group_by(combined_group, diabetes_status) %>%
#   summarise(
#     n = n(),
#     percent = n() / nrow(merged_data_all) * 100
#   ) %>%
#   arrange(combined_group, diabetes_status)
# 
# print(group_summary)
```
```{r}
# # First, create a specific group for high-high vs others
# merged_data_all$special_group <- ifelse(
#   merged_data_all$bifido_cat == "High" & merged_data_all$thosai_cat == "High",
#   "High Bifido & High Thosai",
#   "Other Groups"
# )
# 
# # Create contingency table for this specific comparison
# high_high_table <- table(
#   merged_data_all$special_group,
#   merged_data_all$diabetes_status
# )
# 
# # Chi-square test
# high_high_chi <- chisq.test(high_high_table)
# print(high_high_chi)
# 
# # Calculate proportions within each group
# prop_table <- prop.table(high_high_table, margin = 1) * 100
# 
# # Print percentages
# print("Percentage of diabetes status in each group:")
# print(prop_table)
# 
# # Visualize using ggplot2
# library(ggplot2)
# 
# # Convert to data frame for plotting
# plot_data <- as.data.frame(prop_table)
# colnames(plot_data) <- c("Group", "Diabetes_Status", "Percentage")
# 
# # Create bar plot
# ggplot(plot_data, aes(x = Group, y = Percentage, fill = Diabetes_Status)) +
#   geom_bar(stat = "identity") +
#   theme_minimal() +
#   labs(
#     title = "Diabetes Status Distribution",
#     subtitle = "High Bifido & High Thosai vs Other Groups",
#     x = "Group",
#     y = "Percentage (%)"
#   ) +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
# 
# # You might also want to perform a logistic regression
# # This will give you odds ratios
# library(stats)
# 
# # Create binary outcome for diabetes (1 = diabetes, 0 = normal/pre-diabetes)
# merged_data_all$diabetes_binary <- ifelse(
#   merged_data_all$diabetes_status == "diabetes", 1, 0
# )
# 
# # Fit logistic regression
# logistic_model <- glm(
#   diabetes_binary ~ special_group,
#   family = binomial(link = "logit"),
#   data = merged_data_all
# )
# 
# # Print odds ratios
# odds_ratios <- exp(coef(logistic_model))
# print("Odds Ratios:")
# print(odds_ratios)
# 
# # Calculate risk ratios
# risk_ratio <- function(table) {
#   # Risk in high-high group
#   risk_high <- table["High Bifido & High Thosai", "diabetes"] / 
#     sum(table["High Bifido & High Thosai", ])
#   
#   # Risk in other groups
#   risk_other <- table["Other Groups", "diabetes"] / 
#     sum(table["Other Groups", ])
#   
#   return(risk_high / risk_other)
# }
# 
# rr <- risk_ratio(high_high_table)
# print("Risk Ratio (High-High vs Others):")
# print(rr)
```
# only in indian
```{r}
# only_i = subset(merged_data_all, FREG5_Ethnic_Group == "I")
# only_i$`Bifidobacterium catenulatum`
# ```
# ```{r}
# # Create binary categories using median splits
# only_i$bifido_cat <- ifelse(
#   only_i$`Bifidobacterium catenulatum` > 
#     median(only_i$`Bifidobacterium catenulatum`), 
#   "High", "Low"
# )
# 
# only_i$thosai_cat <- ifelse(
#   only_i$Idli_Thosai > median(only_i$Idli_Thosai), 
#   "High", "Low"
# )
# 
# # Create combined groups
# only_i$combined_group <- paste(
#   only_i$bifido_cat, 
#   only_i$thosai_cat
# )
# 
# # Create a contingency table
# contingency_table_i <- table(
#   only_i$combined_group, 
#   only_i$diabetes_status
# )
# 
# # View the contingency table
# print(contingency_table_i)
```


```{r}
# # Perform chi-square test
# chi_test <- chisq.test(contingency_table_i)
# print(chi_test)
```


```{r}
# Create a visual representation using ggplot2
# library(ggplot2)
# 
# # Convert to percentages
# prop_table <- prop.table(contingency_table_i, margin = 1) * 100
# 
# # Convert to data frame for plotting
# plot_data <- as.data.frame(prop_table)
# colnames(plot_data) <- c("Group", "Diabetes_Status", "Percentage")
# 
# # Create stacked bar plot
# ggplot(plot_data, aes(x = Group, y = Percentage, fill = Diabetes_Status)) +
#   geom_bar(stat = "identity") +
#   theme_minimal() +
#   labs(
#     title = "Distribution of Diabetes Status by Bifido-Thosai Groups",
#     x = "Bifido-Thosai Group",
#     y = "Percentage (%)"
#   ) +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}

# First, create a specific group for high-high vs others
# only_i$special_group <- ifelse(
#   only_i$bifido_cat == "High" & only_i$thosai_cat == "High",
#   "High Bifido & High Thosai",
#   "Other Groups"
# )
# 
# # Create contingency table for this specific comparison
# high_high_table <- table(
#   only_i$special_group,
#   only_i$diabetes_status
# )
# 
# # Chi-square test
# high_high_chi <- chisq.test(high_high_table)
# print(high_high_chi)
# 
# # Calculate proportions within each group
# prop_table <- prop.table(high_high_table, margin = 1) * 100
# 
# # Print percentages
# print("Percentage of diabetes status in each group:")
# print(prop_table)
# 
# # Visualize using ggplot2
# library(ggplot2)
# 
# # Convert to data frame for plotting
# plot_data <- as.data.frame(prop_table)
# colnames(plot_data) <- c("Group", "Diabetes_Status", "Percentage")
# 
# # Create bar plot
# ggplot(plot_data, aes(x = Group, y = Percentage, fill = Diabetes_Status)) +
#   geom_bar(stat = "identity") +
#   theme_minimal() +
#   labs(
#     title = "Diabetes Status Distribution",
#     subtitle = "High Bifido & High Thosai vs Other Groups",
#     x = "Group",
#     y = "Percentage (%)"
#   ) +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
# 
# # You might also want to perform a logistic regression
# # This will give you odds ratios
# library(stats)
# 
# # Create binary outcome for diabetes (1 = diabetes, 0 = normal/pre-diabetes)
# only_i$diabetes_binary <- ifelse(
#   only_i$diabetes_status == "diabetes", 1, 0
# )
# 
# # Fit logistic regression
# logistic_model <- glm(
#   diabetes_binary ~ special_group,
#   family = binomial(link = "logit"),
#   data = only_i
# )
# 
# # Print odds ratios
# odds_ratios <- exp(coef(logistic_model))
# print("Odds Ratios:")
# print(odds_ratios)
# 
# # Calculate risk ratios
# risk_ratio <- function(table) {
#   # Risk in high-high group
#   risk_high <- table["High Bifido & High Thosai", "diabetes"] / 
#     sum(table["High Bifido & High Thosai", ])
#   
#   # Risk in other groups
#   risk_other <- table["Other Groups", "diabetes"] / 
#     sum(table["Other Groups", ])
#   
#   return(risk_high / risk_other)
# }
# 
# rr <- risk_ratio(high_high_table)
# print("Risk Ratio (High-High vs Others):")
# print(rr)
```






# heatmaps
```{r}
# Define columns
species_cols <- colnames(sig_species)[colnames(sig_species) != "Participant_ID"]
clinical_vars <- c("DLAB27_Hba1C_Percent", "DLAB83_CRP", "BMI", "HOMA_IR", "HOMA_B")

# Initialize matrices for correlations and p-values
cor_matrix <- matrix(NA, nrow = length(species_cols), ncol = length(clinical_vars))
p_matrix <- matrix(NA, nrow = length(species_cols), ncol = length(clinical_vars))

# Calculate partial correlations
for(i in 1:length(species_cols)) {
  for(j in 1:length(clinical_vars)) {
    # Create temporary dataset with ranked variables
    temp_data <- data.frame(
      species = rank(merged_data[[species_cols[i]]]),
      clinical = rank(merged_data[[clinical_vars[j]]]),
      gender = as.numeric(merged_data$FREG7_Gender == "M"),
      age = rank(merged_data$FREG8_Age),
      ethnic_C = as.numeric(merged_data$FREG5_Ethnic_Group == "C"),
      ethnic_I = as.numeric(merged_data$FREG5_Ethnic_Group == "I"),
      ethnic_M = as.numeric(merged_data$FREG5_Ethnic_Group == "M")
    )
    
    # Calculate partial correlation
    pcor_result <- pcor.test(temp_data$species, temp_data$clinical,
                            temp_data[,c("gender", "age", 
                                       "ethnic_C", "ethnic_I", "ethnic_M")])
    
    # Cap correlation values at ±0.2
    cor_matrix[i,j] <- pmin(pmax(pcor_result$estimate, -0.15), 0.15)
    p_matrix[i,j] <- pcor_result$p.value
  }
}

# Add row and column names
rownames(cor_matrix) <- species_cols
colnames(cor_matrix) <- clinical_vars
rownames(p_matrix) <- species_cols
colnames(p_matrix) <- clinical_vars

# Perform clustering
species_dist <- dist(cor_matrix)
species_clust <- hclust(species_dist)
species_order <- species_clust$labels[species_clust$order]

clinical_dist <- dist(t(cor_matrix))
clinical_clust <- hclust(clinical_dist)
clinical_order <- clinical_clust$labels[clinical_clust$order]

# Convert to long format
melted_cor <- melt(cor_matrix)
melted_p <- melt(p_matrix)
colnames(melted_cor) <- c("Species", "Clinical", "Correlation")
melted_cor$Significant <- melted_p$value < 0.05

# Set factor levels based on clustering
melted_cor$Species <- factor(melted_cor$Species, levels = species_order)
melted_cor$Clinical <- factor(melted_cor$Clinical, levels = clinical_order)
#melted_cor$Clinical
# Create the plot
clinical_labels <- c(
  "BMI" = "BMI",
  "DLAB27_Hba1C_Percent" = "HbaA1C",
  "DLAB83_CRP" = "CRP",
  "HOMA_IR" = "HOMA-IR",
  "HOMA_B" = "HOMA-B"
)

# Plot with updated x-axis labels
ggplot(melted_cor, aes(x = Clinical, y = Species, fill = Correlation)) +
  geom_tile() +
  scale_fill_gradientn(colors = rev(colorRampPalette(brewer.pal(11, "RdBu"))(100)), 
                      limits = c(-0.15, 0.15)) +
  geom_text(aes(label = ifelse(Significant, "*", "")), 
            color = "black", size = 5) +
  scale_x_discrete(labels = clinical_labels) +  # Rename x-axis labels
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(face = "italic"),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold")
  ) +
  labs(
    title = "Correlations of Species and Clinical Variable",
    subtitle = "Adjusted for Gender, Age, and Ethnicity\n* indicates p < 0.05",
    x = "Clinical Variables",
    y = "Species"
  )

# Save the plot if needed
#ggsave("figure/fig6_clinical_species_correlations.pdf", width = 6, height = 11)
```
```{r}
melted_cor
p_matrix
```





```{r}
library(gridExtra)
merged_data <- sig_species %>%
  inner_join(metadata, by = "Participant_ID")

# Define columns
species_cols <- colnames(sig_species)[colnames(sig_species) != "Participant_ID"]
continuous_vars <- c("DLAB27_Hba1C_Percent", "DLAB83_CRP", "BMI", "HOMA_IR", "HOMA_B")

# Initialize matrices for continuous variables
cont_cor_matrix <- matrix(NA, nrow = length(species_cols), ncol = length(continuous_vars))
cont_p_matrix <- matrix(NA, nrow = length(species_cols), ncol = length(continuous_vars))

# Initialize vectors for diabetes status
diab_effect <- numeric(length(species_cols))
diab_p <- numeric(length(species_cols))

# Calculate correlations for continuous variables
for(i in 1:length(species_cols)) {
  for(j in 1:length(continuous_vars)) {
    temp_data <- data.frame(
      species = rank(merged_data[[species_cols[i]]]),
      clinical = rank(merged_data[[continuous_vars[j]]]),
      gender = as.numeric(merged_data$FREG7_Gender == "M"),
      age = rank(merged_data$FREG8_Age),
      ethnic_C = as.numeric(merged_data$FREG5_Ethnic_Group == "C"),
      ethnic_I = as.numeric(merged_data$FREG5_Ethnic_Group == "I"),
      ethnic_M = as.numeric(merged_data$FREG5_Ethnic_Group == "M")
    )
    
    pcor_result <- pcor.test(temp_data$species, temp_data$clinical,
                            temp_data[,c("gender", "age", 
                                       "ethnic_C", "ethnic_I", "ethnic_M")])
    
    cont_cor_matrix[i,j] <- pmin(pmax(pcor_result$estimate, -0.15), 0.15)
    cont_p_matrix[i,j] <- pcor_result$p.value
  }
  
  # Calculate diabetes association
  temp_data <- data.frame(
    species_abundance = merged_data[[species_cols[i]]],
    gender = merged_data$FREG7_Gender,
    age = merged_data$FREG8_Age,
    ethnicity = merged_data$FREG5_Ethnic_Group,
    diabetes = merged_data$diabetes_status
  )
  
  model <- lm(species_abundance ~ gender + age + ethnicity, data = temp_data)
  species_residuals <- residuals(model)
  
  kw_test <- kruskal.test(species_residuals ~ temp_data$diabetes)
  
  groups <- split(species_residuals, temp_data$diabetes)
  means <- sapply(groups, mean)
  overall_mean <- mean(species_residuals)
  ss_between <- sum(sapply(groups, length) * (means - overall_mean)^2)
  ss_total <- sum((species_residuals - overall_mean)^2)
  eta_squared <- ss_between / ss_total
  
  if("diabetes" %in% names(means) && "normal" %in% names(means)) {
    direction <- sign(means["diabetes"] - means["normal"])
  } else {
    direction <- 1
  }
  
  diab_effect[i] <- direction * eta_squared
  diab_p[i] <- kw_test$p.value
}

# Prepare continuous variables plot
rownames(cont_cor_matrix) <- species_cols
colnames(cont_cor_matrix) <- continuous_vars
rownames(cont_p_matrix) <- species_cols
colnames(cont_p_matrix) <- continuous_vars

# Clustering for continuous variables
species_dist <- dist(cont_cor_matrix)
species_clust <- hclust(species_dist)
species_order <- species_clust$labels[species_clust$order]

clinical_dist <- dist(t(cont_cor_matrix))
clinical_clust <- hclust(clinical_dist)
clinical_order <- clinical_clust$labels[clinical_clust$order]

# Prepare data for plots
# Continuous variables
melted_cont <- melt(cont_cor_matrix)
melted_cont_p <- melt(cont_p_matrix)
colnames(melted_cont) <- c("Species", "Clinical", "Correlation")
melted_cont$Significant <- melted_cont_p$value < 0.05
melted_cont$Species <- factor(melted_cont$Species, levels = species_order)
melted_cont$Clinical <- factor(melted_cont$Clinical, levels = clinical_order)

# Diabetes status
diab_df <- data.frame(
  Species = species_cols,
  Effect = diab_effect,
  P_value = diab_p
)
diab_df$Species <- factor(diab_df$Species, levels = species_order)
diab_df$Significant <- diab_df$P_value < 0.05

# Create continuous variables plot
p1 <- ggplot(melted_cont, aes(x = Clinical, y = Species, fill = Correlation)) +
  geom_tile() +
  scale_fill_gradientn(colors = rev(colorRampPalette(brewer.pal(11, "RdBu"))(100)), 
                      limits = c(-0.15, 0.15)) +
  geom_text(aes(label = ifelse(Significant, "*", "")), 
            color = "black", size = 5) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(face = "italic"),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold")
  ) +
  labs(
    title = "Species Correlations with Clinical Variables",
    subtitle = "Adjusted for Gender, Age, and Ethnicity\n* indicates p < 0.05",
    x = "Clinical Variables",
    y = "Species",
    fill = "Correlation"
  )

# Create diabetes status plot
p2 <- ggplot(diab_df, aes(y = Species, x = Effect)) +
  geom_bar(stat = "identity", aes(fill = abs(Effect))) +
  #scale_fill_gradientn(colors = rev(brewer.pal(9, "PuBu"))) +
  scale_fill_gradient2(low = "white", high = "plum4") +
  geom_text(aes(label = ifelse(Significant, "*", ""), x = 0), 
            color = "black", size = 5) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(face = "italic"),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold")
  ) +
  labs(
    title = "Species Associations with Diabetes Status",
    subtitle = "Adjusted for Gender, Age, and Ethnicity\n* indicates p < 0.05",
    x = "Effect Size (Eta Squared)",
    y = NULL,
    fill = "Absolute\nEffect Size"
  )

# Arrange plots side by side
grid.arrange(p1, p2, ncol = 2, widths = c(1.5, 1))

# Optionally save the combined plot
#ggsave("figure/clinical_diabetes_associations2.pdf", arrangeGrob(p1, p2, ncol = 2, widths = c(1.5, 1)), width = 15, height = 10)
```
## separate ethnicity
```{r}
library(gridExtra)

# Function to perform analysis for one ethnic group
analyze_ethnic_group <- function(merged_data, ethnic_group) {
  # Filter data for the specific ethnic group
  ethnic_data <- merged_data %>% 
    filter(FREG5_Ethnic_Group == ethnic_group)
  
  # Get columns
  species_cols <- colnames(sig_species)[colnames(sig_species) != "Participant_ID"]
  continuous_vars <- c("DLAB27_Hba1C_Percent", "DLAB83_CRP", "BMI", "HOMA_IR", "HOMA_B")
  
  # Initialize matrices
  cont_cor_matrix <- matrix(NA, nrow = length(species_cols), ncol = length(continuous_vars))
  cont_p_matrix <- matrix(NA, nrow = length(species_cols), ncol = length(continuous_vars))
  diab_effect <- numeric(length(species_cols))
  diab_p <- numeric(length(species_cols))
  
  # Calculate correlations for continuous variables
  for(i in 1:length(species_cols)) {
    for(j in 1:length(continuous_vars)) {
      temp_data <- data.frame(
        species = rank(ethnic_data[[species_cols[i]]]),
        clinical = rank(ethnic_data[[continuous_vars[j]]]),
        gender = as.numeric(ethnic_data$FREG7_Gender == "M"),
        age = rank(ethnic_data$FREG8_Age)
      )
      
      # Now adjusting only for gender and age
      pcor_result <- pcor.test(temp_data$species, temp_data$clinical,
                              temp_data[,c("gender", "age")])
      
      cont_cor_matrix[i,j] <- pmin(pmax(pcor_result$estimate, -0.15), 0.15)
      cont_p_matrix[i,j] <- pcor_result$p.value
    }
    
    # Calculate diabetes association
    temp_data <- data.frame(
      species_abundance = ethnic_data[[species_cols[i]]],
      gender = ethnic_data$FREG7_Gender,
      age = ethnic_data$FREG8_Age,
      diabetes = ethnic_data$diabetes_status
    )
    
    # Adjust for gender and age only
    model <- lm(species_abundance ~ gender + age, data = temp_data)
    species_residuals <- residuals(model)
    
    kw_test <- kruskal.test(species_residuals ~ temp_data$diabetes)
    
    groups <- split(species_residuals, temp_data$diabetes)
    means <- sapply(groups, mean)
    overall_mean <- mean(species_residuals)
    ss_between <- sum(sapply(groups, length) * (means - overall_mean)^2)
    ss_total <- sum((species_residuals - overall_mean)^2)
    eta_squared <- ss_between / ss_total
    
    if("diabetes" %in% names(means) && "normal" %in% names(means)) {
      direction <- sign(means["diabetes"] - means["normal"])
    } else {
      direction <- 1
    }
    
    diab_effect[i] <- direction * eta_squared
    diab_p[i] <- kw_test$p.value
  }
  
  # Prepare matrices
  rownames(cont_cor_matrix) <- species_cols
  colnames(cont_cor_matrix) <- continuous_vars
  rownames(cont_p_matrix) <- species_cols
  colnames(cont_p_matrix) <- continuous_vars
  
  # Return results
  list(
    cont_cor = cont_cor_matrix,
    cont_p = cont_p_matrix,
    diab_effect = diab_effect,
    diab_p = diab_p
  )
}

# Analyze each ethnic group
results_C <- analyze_ethnic_group(merged_data, "C")
results_I <- analyze_ethnic_group(merged_data, "I")
results_M <- analyze_ethnic_group(merged_data, "M")

# Function to create plots for one ethnic group
create_plots <- function(results, species_cols, continuous_vars, group_name) {
  # Prepare continuous variables plot
  melted_cont <- melt(results$cont_cor)
  melted_cont_p <- melt(results$cont_p)
  colnames(melted_cont) <- c("Species", "Clinical", "Correlation")
  melted_cont$Significant <- melted_cont_p$value < 0.05
  
  # Create continuous variables plot
  p1 <- ggplot(melted_cont, aes(x = Clinical, y = Species, fill = Correlation)) +
    geom_tile() +
    scale_fill_gradientn(colors = rev(colorRampPalette(brewer.pal(11, "RdBu"))(100)), 
                        limits = c(-0.15, 0.15)) +
    geom_text(aes(label = ifelse(Significant, "*", "")), 
              color = "black", size = 5) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.text.y = element_text(face = "italic"),
      axis.title = element_text(face = "bold"),
      legend.title = element_text(face = "bold")
    ) +
    labs(
      title = paste("Species Correlations with Clinical Variables -", group_name),
      subtitle = "Adjusted for Gender and Age\n* indicates p < 0.05",
      x = "Clinical Variables",
      y = "Species",
      fill = "Correlation"
    )
  
  # Prepare diabetes plot data
  diab_df <- data.frame(
    Species = species_cols,
    Effect = results$diab_effect,
    P_value = results$diab_p
  )
  diab_df$Significant <- diab_df$P_value < 0.05
  
  # Create diabetes plot
  p2 <- ggplot(diab_df, aes(y = Species, x = Effect)) +
    geom_bar(stat = "identity", aes(fill = abs(Effect))) +
    scale_fill_gradient2(low = "white", high = "plum4") +
    geom_text(aes(label = ifelse(Significant, "*", ""), x = 0), 
              color = "black", size = 5) +
    theme_minimal() +
    theme(
      axis.text.y = element_text(face = "italic"),
      axis.title = element_text(face = "bold"),
      legend.title = element_text(face = "bold")
    ) +
    labs(
      title = paste("Species Associations with Diabetes Status -", group_name),
      subtitle = "Adjusted for Gender and Age\n* indicates p < 0.05",
      x = "Effect Size (Eta Squared)",
      y = NULL,
      fill = "Absolute\nEffect Size"
    )
  
  list(continuous = p1, diabetes = p2)
}
```


```{r}
# Create plots for each ethnic group
plots_C <- create_plots(results_C, species_cols, continuous_vars, "Chinese")
plots_I <- create_plots(results_I, species_cols, continuous_vars, "Indian")
plots_M <- create_plots(results_M, species_cols, continuous_vars, "Malay")

# plots_C
# plots_I 
# plots_M
```


```{r}
# Arrange all plots
# Continuous variables plots on top row, diabetes plots on bottom row
grid.arrange(
  plots_C$continuous, plots_I$continuous, plots_M$continuous,
  plots_C$diabetes, plots_I$diabetes, plots_M$diabetes,
  ncol = 3, nrow = 2
)


# ggsave("figure/ethnic_stratified_analysis.pdf",
#        arrangeGrob(plots_C$continuous, plots_I$continuous, plots_M$continuous,
#                   plots_C$diabetes, plots_I$diabetes, plots_M$diabetes,
#                   ncol = 3, nrow = 2),
#        width = 24, height = 16)
```



```{r}
# Create summary table

diabetes_dist <- merged_data %>%
  group_by(FREG5_Ethnic_Group, diabetes_status) %>%
  summarise(count = n(), .groups = 'drop') %>%
  # Calculate percentages within each ethnic group
  group_by(FREG5_Ethnic_Group) %>%
  mutate(percentage = count/sum(count) * 100)

diabetes_dist <- diabetes_dist %>%
  mutate(diabetes_status = factor(diabetes_status, levels = c("diabetes", "pre-diabetes", "normal")))
# Create a plot
ggplot(diabetes_dist, aes(x = FREG5_Ethnic_Group, y = count, fill = diabetes_status)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal() +
  labs(
    title = "Distribution of Diabetes Status (based on Hba1C) by Ethnic Group",
    x = "Ethnic Group",
    y = "Count",
    fill = "Diabetes Status"
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 14, face = "bold")
  )

# Print numerical summary
print(diabetes_dist)
```
```{r}
merged_data$glucose
diabetes_dist <- merged_data %>%
  group_by(FREG5_Ethnic_Group, glucose) %>%
  summarise(count = n(), .groups = 'drop') %>%
  # Calculate percentages within each ethnic group
  group_by(FREG5_Ethnic_Group) %>%
  mutate(percentage = count/sum(count) * 100)

diabetes_dist <- diabetes_dist %>%
  mutate(glucose = factor(glucose, levels = c("diabetes", "pre-diabetes", "normal", "hypoglycemia")))

# Create a plot
ggplot(diabetes_dist, aes(x = FREG5_Ethnic_Group, y = count, fill = glucose)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal() +
  labs(
    title = "Distribution of Diabetes Status (based on glucose) by Ethnic Group",
    x = "Ethnic Group",
    y = "Count",
    fill = "Diabetes Status"
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 14, face = "bold")
  )

# Print numerical summary
print(diabetes_dist)
```
```{r}
species_filtered2
```


## overall analysis for clinical variables
```{r}
species_filtered2
# Match samples and create merged dataset
merged_data_a <- species_filtered2 %>%
  inner_join(metadata, by = "Participant_ID")
species_cols2 = colnames(species_filtered2[,-1])
species_cols2
```


```{r}
# Assuming merged_data is already loaded and contains species and metadata columns
# Create OTU Table
merged_data_a$BMI_group <- ifelse(merged_data_a$BMI > 25, "Obese", "Non-Obese")
OTU <- otu_table(as.matrix(merged_data_a[, species_cols2]), taxa_are_rows = FALSE)
merged_data_a
# Make sure the row names of the OTU table are the Participant IDs
rownames(OTU) <- merged_data_a$Participant_ID

# Create Sample Data
# Ensure that the sample_data dataframe has Participant_ID as row names
samples_data <- sample_data(merged_data_a[, c("diabetes_status", "FREG5_Ethnic_Group", "DLAB27_Hba1C_Percent", "DLAB83_CRP", "DLAB81_Insulin", "DLAB25_Gluf_Mmol_L", "BMI","BMI_group")])
rownames(samples_data) <- merged_data_a$Participant_ID

# Create a phyloseq object
physeq <- phyloseq(OTU, samples_data)
OTU
```
##obese alpha beta
```{r}
comparisons <- list(c("Obese", "Non-Obese"))
a1 = plot_richness(physeq, x = "BMI_group", color = "BMI_group", measures = c("Shannon", "Simpson", "Chao1", "Fisher")) + geom_boxplot() + theme_bw()+
    stat_compare_means(comparisons = comparisons, method = "t.test", label = "p.signif", symnum.args = list(cutpoints = c(0, 0.001, 0.01, 0.05, 1), symbols = c("***", "**", "*", "ns")))
ggsave(
  filename = "figure/figure6/obese_alpha.pdf",
  plot = a1,
  height = 6, 
  width = 9,
  device = cairo_pdf)

```
```{r}
# Compute alpha diversity measures for the physeq object
# alpha_div <- estimate_richness(physeq, measures = c("Shannon", "Simpson", "Chao1", "Fisher"))
# 
# # Add BMI group information to the alpha diversity data
# alpha_div$BMI_group <- sam_data$BMI_group
# 
# # Run statistical tests for each diversity measure
# p_values <- data.frame(
#   Measure = c("Shannon", "Simpson", "Chao1", "Fisher"),
#   p_value = c(
#     t.test(alpha_div$Shannon ~ alpha_div$BMI_group)$p.value,
#     t.test(alpha_div$Simpson ~ alpha_div$BMI_group)$p.value,
#     t.test(alpha_div$Chao1 ~ alpha_div$BMI_group)$p.value,
#     t.test(alpha_div$Fisher ~ alpha_div$BMI_group)$p.value
#   )
# )
# 
# # Optionally, adjust p-values for multiple comparisons (e.g., using FDR correction)
# p_values$p_adj <- p.adjust(p_values$p_value, method = "fdr")
# print(p_values)
```



```{r}
dist = phyloseq::distance(physeq, method="bray")
phylo.beta = ordinate(physeq, method="PCoA", distance=dist)

b1 = plot_ordination(physeq, phylo.beta, type = "samples", color = "BMI_group") + theme_bw() + stat_ellipse(aes(group=BMI_group))
ggsave(
  filename = "figure/figure6/obese_beta.pdf",
  plot = b1,
  height = 6, 
  width = 9,
  device = cairo_pdf)
```
```{r}
library(vegan)
sample_data(physeq)$BMI_group <- factor(sample_data(physeq)$BMI_group)

physeq <- subset_samples(physeq, !is.na(sample_data(physeq)$BMI_group))

# Calculate the distance matrix using the ape package
dist <- phyloseq::distance(physeq, method = "bray")
dist_matrix <- as.matrix(dist)

# Perform PERMANOVA using adonis2
permanova <- adonis2(dist_matrix ~ sample_data(physeq)$BMI_group)

# Print the PERMANOVA results
print(permanova)
```

##diabete alpha beta
```{r}
#comparisons <- list(c("Obese", "Non-Obese"))
physeq_f <- subset_samples(physeq, !is.na(sample_data(physeq)$diabetes_status))
a2 = plot_richness(physeq_f, x = "diabetes_status", color = "diabetes_status", measures = c("Shannon", "Simpson", "Chao1", "Fisher")) + geom_boxplot() + theme_bw()
ggsave(
  filename = "figure/figure6/diabete_alpha.pdf",
  plot = a2,
  height = 6, 
  width = 9,
  device = cairo_pdf)

```

```{r}
dist = phyloseq::distance(physeq_f, method="bray")
phylo.beta = ordinate(physeq_f, method="PCoA", distance=dist)

b2 = plot_ordination(physeq_f, phylo.beta, type = "samples", color = "diabetes_status") + theme_bw() + stat_ellipse(aes(group=diabetes_status))
ggsave(
  filename = "figure/figure6/diabetes_beta.pdf",
  plot = b2,
  height = 6, 
  width = 9,
  device = cairo_pdf)
```
```{r}
library(vegan)
sample_data(physeq_f)$diabetes_status <- factor(sample_data(physeq_f)$diabetes_status)

#physeq <- subset_samples(physeq, !is.na(sample_data(physeq)$BMI_group))

# Calculate the distance matrix using the ape package
dist <- phyloseq::distance(physeq_f, method = "bray")
dist_matrix <- as.matrix(dist)

# Perform PERMANOVA using adonis2
permanova <- adonis2(dist_matrix ~ sample_data(physeq_f)$diabetes_status)

# Print the PERMANOVA results
print(permanova)
```


```{r}
library(phyloseq)
library(dplyr)

alpha_diversity <- estimate_richness(physeq, measures = "Shannon")
# Convert rownames to a column for merging
alpha_diversity$Participant_ID <- rownames(alpha_diversity)

# Extract sample metadata from the phyloseq object and make it a regular data frame
sample_metadata <- as.data.frame(sample_data(physeq))
# Ensure 'Participant_ID' is a column for merging
sample_metadata$Participant_ID <- rownames(sample_metadata)

# Merge alpha diversity data with sample metadata
alpha_diversity_full <- merge(alpha_diversity, sample_metadata, by = "Participant_ID")


# Now you can plot or analyze your merged data
# For example, to visualize alpha diversity across diabetes status:
library(ggplot2)
ggplot(alpha_diversity_full, aes(x = diabetes_status, y = Shannon)) +
  geom_boxplot() +
  labs(title = "Alpha Diversity across Diabetes Status", y = "Shannon Diversity Index", x = "Diabetes Status")

```

## food and clinical variables
```{r}
food_data
merged_data_f = food_data %>%
  inner_join(metadata, by = "Participant_ID")
```

```{r}
food_cols <- colnames(food_data)[colnames(food_data) != "Participant_ID"]
clinical_vars <- c("DLAB27_Hba1C_Percent", "DLAB83_CRP", "BMI", "HOMA_IR", "HOMA_B")

# Initialize matrices for correlations and p-values
cor_matrix <- matrix(NA, nrow = length(food_cols), ncol = length(clinical_vars))
p_matrix <- matrix(NA, nrow = length(food_cols), ncol = length(clinical_vars))

# Calculate partial correlations
for(i in 1:length(food_cols)) {
  for(j in 1:length(clinical_vars)) {
    # Create temporary dataset with ranked variables
    temp_data <- data.frame(
      food = rank(merged_data_f[[food_cols[i]]]),
      clinical = rank(merged_data_f[[clinical_vars[j]]]),
      gender = as.numeric(merged_data_f$FREG7_Gender == "M"),
      age = rank(merged_data_f$FREG8_Age),
      ethnic_C = as.numeric(merged_data_f$FREG5_Ethnic_Group == "C"),
      ethnic_I = as.numeric(merged_data_f$FREG5_Ethnic_Group == "I"),
      ethnic_M = as.numeric(merged_data_f$FREG5_Ethnic_Group == "M")
    )
    
    # Calculate partial correlation
    pcor_result <- pcor.test(temp_data$food, temp_data$clinical,
                            temp_data[,c("gender", "age", 
                                       "ethnic_C", "ethnic_I", "ethnic_M")])
    
    # Cap correlation values at ±0.2
    cor_matrix[i,j] <- pmin(pmax(pcor_result$estimate, -0.15), 0.15)
    p_matrix[i,j] <- pcor_result$p.value
  }
}

# Add row and column names
rownames(cor_matrix) <- food_cols
colnames(cor_matrix) <- clinical_vars
rownames(p_matrix) <- food_cols
colnames(p_matrix) <- clinical_vars

# Perform clustering
food_dist <- dist(cor_matrix)
food_clust <- hclust(food_dist)
food_order <- food_clust$labels[food_clust$order]

clinical_dist <- dist(t(cor_matrix))
clinical_clust <- hclust(clinical_dist)
clinical_order <- clinical_clust$labels[clinical_clust$order]

# Convert to long format
melted_cor <- melt(cor_matrix)
melted_p <- melt(p_matrix)
colnames(melted_cor) <- c("Food", "Clinical", "Correlation")
melted_cor$Significant <- melted_p$value < 0.05

# Set factor levels based on clustering
melted_cor$Food <- factor(melted_cor$Food, levels = food_order)
melted_cor$Clinical <- factor(melted_cor$Clinical, levels = clinical_order)

# Create the plot
ggplot(melted_cor, aes(x = Clinical, y = Food, fill = Correlation)) +
  geom_tile() +
  scale_fill_gradientn(colors = rev(colorRampPalette(brewer.pal(11, "RdBu"))(100)), 
                      limits = c(-0.15, 0.15)) +
  geom_text(aes(label = ifelse(Significant, "*", "")), 
            color = "black", size = 5) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(face = "italic"),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold")
  ) +
  labs(
    title = "Spearman Correlations between Food and Clinical Variables",
    subtitle = "Adjusted for Gender, Age, and Ethnicity\n* indicates p < 0.05",
    x = "Clinical Variables",
    y = "Food"
  )

```



```{r}


# Define columns
food_cols <- colnames(food_data)[colnames(food_data) != "Participant_ID"]
continuous_vars <- c("DLAB27_Hba1C_Percent", "DLAB83_CRP", "BMI", "HOMA_IR", "HOMA_B")

# Initialize matrices for continuous variables
cont_cor_matrix <- matrix(NA, nrow = length(food_cols), ncol = length(continuous_vars))
cont_p_matrix <- matrix(NA, nrow = length(food_cols), ncol = length(continuous_vars))

# Initialize vectors for diabetes status
diab_effect <- numeric(length(food_cols))
diab_p <- numeric(length(food_cols))

# Calculate correlations for continuous variables
for(i in 1:length(food_cols)) {
  for(j in 1:length(continuous_vars)) {
    temp_data <- data.frame(
      food = rank(merged_data_f[[food_cols[i]]]),
      clinical = rank(merged_data_f[[continuous_vars[j]]]),
      gender = as.numeric(merged_data_f$FREG7_Gender == "M"),
      age = rank(merged_data_f$FREG8_Age),
      ethnic_C = as.numeric(merged_data_f$FREG5_Ethnic_Group == "C"),
      ethnic_I = as.numeric(merged_data_f$FREG5_Ethnic_Group == "I"),
      ethnic_M = as.numeric(merged_data_f$FREG5_Ethnic_Group == "M")
    )
    
    pcor_result <- pcor.test(temp_data$food, temp_data$clinical,
                            temp_data[,c("gender", "age", 
                                       "ethnic_C", "ethnic_I", "ethnic_M")])
    
    cont_cor_matrix[i,j] <- pmin(pmax(pcor_result$estimate, -0.15), 0.15)
    cont_p_matrix[i,j] <- pcor_result$p.value
  }
  
  # Calculate diabetes association
  temp_data <- data.frame(
    food_abundance = merged_data_f[[food_cols[i]]],
    gender = merged_data_f$FREG7_Gender,
    age = merged_data_f$FREG8_Age,
    ethnicity = merged_data_f$FREG5_Ethnic_Group,
    diabetes = merged_data_f$diabetes_status
  )
  
  model <- lm(food_abundance ~ gender + age + ethnicity, data = temp_data)
  food_residuals <- residuals(model)
  
  kw_test <- kruskal.test(food_residuals ~ temp_data$diabetes)
  
  groups <- split(food_residuals, temp_data$diabetes)
  means <- sapply(groups, mean)
  overall_mean <- mean(food_residuals)
  ss_between <- sum(sapply(groups, length) * (means - overall_mean)^2)
  ss_total <- sum((food_residuals - overall_mean)^2)
  eta_squared <- ss_between / ss_total
  
  if("diabetes" %in% names(means) && "normal" %in% names(means)) {
    direction <- sign(means["diabetes"] - means["normal"])
  } else {
    direction <- 1
  }
  
  diab_effect[i] <- direction * eta_squared
  diab_p[i] <- kw_test$p.value
}

# Prepare continuous variables plot
rownames(cont_cor_matrix) <- food_cols
colnames(cont_cor_matrix) <- continuous_vars
rownames(cont_p_matrix) <- food_cols
colnames(cont_p_matrix) <- continuous_vars

# Clustering for continuous variables
food_dist <- dist(cont_cor_matrix)
food_clust <- hclust(food_dist)
food_order <- food_clust$labels[food_clust$order]

clinical_dist <- dist(t(cont_cor_matrix))
clinical_clust <- hclust(clinical_dist)
clinical_order <- clinical_clust$labels[clinical_clust$order]

# Prepare data for plots
# Continuous variables
melted_cont <- melt(cont_cor_matrix)
melted_cont_p <- melt(cont_p_matrix)
colnames(melted_cont) <- c("Food", "Clinical", "Correlation")
melted_cont$Significant <- melted_cont_p$value < 0.05
melted_cont$Food <- factor(melted_cont$Food, levels = food_order)
melted_cont$Clinical <- factor(melted_cont$Clinical, levels = clinical_order)

# Diabetes status
diab_df <- data.frame(
  Food = food_cols,
  Effect = diab_effect,
  P_value = diab_p
)
diab_df$Food <- factor(diab_df$Food, levels = food_order)
diab_df$Significant <- diab_df$P_value < 0.05

# Create continuous variables plot
p1 <- ggplot(melted_cont, aes(x = Clinical, y = Food, fill = Correlation)) +
  geom_tile() +
  scale_fill_gradientn(colors = rev(colorRampPalette(brewer.pal(11, "RdBu"))(100)), 
                      limits = c(-0.15, 0.15)) +
  geom_text(aes(label = ifelse(Significant, "*", "")), 
            color = "black", size = 5) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(face = "italic"),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold")
  ) +
  labs(
    title = "Food Correlations with Clinical Variables",
    subtitle = "Adjusted for Gender, Age, and Ethnicity\n* indicates p < 0.05",
    x = "Clinical Variables",
    y = "Food",
    fill = "Correlation"
  )

# Create diabetes status plot
p2 <- ggplot(diab_df, aes(y = Food, x = Effect)) +
  geom_bar(stat = "identity", aes(fill = abs(Effect))) +
  #scale_fill_gradientn(colors = rev(brewer.pal(9, "PuBu"))) +
  scale_fill_gradient2(low = "white", high = "darkmagenta") +
  geom_text(aes(label = ifelse(Significant, "*", ""), x = 0), 
            color = "black", size = 5) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(face = "italic"),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold")
  ) +
  labs(
    title = "Food Associations with Diabetes Status",
    subtitle = "Adjusted for Gender, Age, and Ethnicity\n* indicates p < 0.05",
    x = "Effect Size (Eta Squared)",
    y = NULL,
    fill = "Absolute\nEffect Size"
  )

# Arrange plots side by side
grid.arrange(p1, p2, ncol = 2, widths = c(1.5, 1))

# Optionally save the combined plot
ggsave("figure/clinical_food_associations3.pdf", arrangeGrob(p1, p2, ncol = 2, widths = c(1.5, 1)), width = 15, height = 10)
```

## separate ethnic groups
```{r}

# Function to perform analysis for one ethnic group
analyze_ethnic_group <- function(merged_data_f, ethnic_group) {
  # Filter data for the specific ethnic group
  ethnic_data <- merged_data_f %>% 
    filter(FREG5_Ethnic_Group == ethnic_group)
  
  # Get columns
  food_cols <- colnames(food_data)[colnames(food_data) != "Participant_ID"]
  continuous_vars <- c("DLAB27_Hba1C_Percent", "DLAB83_CRP", 
                      "DLAB81_Insulin", "DLAB25_Gluf_Mmol_L", "BMI", "HOMA")
  
  # Initialize matrices
  cont_cor_matrix <- matrix(NA, nrow = length(food_cols), ncol = length(continuous_vars))
  cont_p_matrix <- matrix(NA, nrow = length(food_cols), ncol = length(continuous_vars))
  diab_effect <- numeric(length(food_cols))
  diab_p <- numeric(length(food_cols))
  
  # Calculate correlations for continuous variables
  for(i in 1:length(food_cols)) {
    for(j in 1:length(continuous_vars)) {
      temp_data <- data.frame(
        food = rank(ethnic_data[[food_cols[i]]]),
        clinical = rank(ethnic_data[[continuous_vars[j]]]),
        gender = as.numeric(ethnic_data$FREG7_Gender == "M"),
        age = rank(ethnic_data$FREG8_Age)
      )
      
      # Now adjusting only for gender and age
      pcor_result <- pcor.test(temp_data$food, temp_data$clinical,
                              temp_data[,c("gender", "age")])
      
      cont_cor_matrix[i,j] <- pmin(pmax(pcor_result$estimate, -0.15), 0.15)
      cont_p_matrix[i,j] <- pcor_result$p.value
    }
    
    # Calculate diabetes association
    temp_data <- data.frame(
      food_abundance = ethnic_data[[food_cols[i]]],
      gender = ethnic_data$FREG7_Gender,
      age = ethnic_data$FREG8_Age,
      diabetes = ethnic_data$diabetes_status
    )
    
    # Adjust for gender and age only
    model <- lm(food_abundance ~ gender + age, data = temp_data)
    food_residuals <- residuals(model)
    
    kw_test <- kruskal.test(food_residuals ~ temp_data$diabetes)
    
    groups <- split(food_residuals, temp_data$diabetes)
    means <- sapply(groups, mean)
    overall_mean <- mean(food_residuals)
    ss_between <- sum(sapply(groups, length) * (means - overall_mean)^2)
    ss_total <- sum((food_residuals - overall_mean)^2)
    eta_squared <- ss_between / ss_total
    
    if("diabetes" %in% names(means) && "normal" %in% names(means)) {
      direction <- sign(means["diabetes"] - means["normal"])
    } else {
      direction <- 1
    }
    
    diab_effect[i] <- direction * eta_squared
    diab_p[i] <- kw_test$p.value
  }
  
  # Prepare matrices
  rownames(cont_cor_matrix) <- food_cols
  colnames(cont_cor_matrix) <- continuous_vars
  rownames(cont_p_matrix) <- food_cols
  colnames(cont_p_matrix) <- continuous_vars
  
  # Return results
  list(
    cont_cor = cont_cor_matrix,
    cont_p = cont_p_matrix,
    diab_effect = diab_effect,
    diab_p = diab_p
  )
}

# Analyze each ethnic group
results_C <- analyze_ethnic_group(merged_data_f, "C")
results_I <- analyze_ethnic_group(merged_data_f, "I")
results_M <- analyze_ethnic_group(merged_data_f, "M")

# Function to create plots for one ethnic group
create_plots <- function(results, food_cols, continuous_vars, group_name) {
  # Prepare continuous variables plot
  melted_cont <- melt(results$cont_cor)
  melted_cont_p <- melt(results$cont_p)
  colnames(melted_cont) <- c("Food", "Clinical", "Correlation")
  melted_cont$Significant <- melted_cont_p$value < 0.05
  
  # Create continuous variables plot
  p1 <- ggplot(melted_cont, aes(x = Clinical, y = Food, fill = Correlation)) +
    geom_tile() +
    scale_fill_gradientn(colors = rev(colorRampPalette(brewer.pal(11, "RdBu"))(100)), 
                        limits = c(-0.15, 0.15)) +
    geom_text(aes(label = ifelse(Significant, "*", "")), 
              color = "black", size = 5) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.text.y = element_text(face = "italic"),
      axis.title = element_text(face = "bold"),
      legend.title = element_text(face = "bold")
    ) +
    labs(
      title = paste("Food Correlations with Clinical Variables -", group_name),
      subtitle = "Adjusted for Gender and Age\n* indicates p < 0.05",
      x = "Clinical Variables",
      y = "Food",
      fill = "Correlation"
    )
  
  # Prepare diabetes plot data
  diab_df <- data.frame(
    Food = food_cols,
    Effect = results$diab_effect,
    P_value = results$diab_p
  )
  diab_df$Significant <- diab_df$P_value < 0.05
  
  # Create diabetes plot
  p2 <- ggplot(diab_df, aes(y = Food, x = Effect)) +
    geom_bar(stat = "identity", aes(fill = abs(Effect))) +
    scale_fill_gradient2(low = "white", high = "plum4") +
    geom_text(aes(label = ifelse(Significant, "*", ""), x = 0), 
              color = "black", size = 5) +
    theme_minimal() +
    theme(
      axis.text.y = element_text(face = "italic"),
      axis.title = element_text(face = "bold"),
      legend.title = element_text(face = "bold")
    ) +
    labs(
      title = paste("Food Associations with Diabetes Status -", group_name),
      subtitle = "Adjusted for Gender and Age\n* indicates p < 0.05",
      x = "Effect Size (Eta Squared)",
      y = NULL,
      fill = "Absolute\nEffect Size"
    )
  
  list(continuous = p1, diabetes = p2)
}
```


```{r}
# Create plots for each ethnic group
plots_C <- create_plots(results_C, food_cols, continuous_vars, "Chinese")
plots_I <- create_plots(results_I, food_cols, continuous_vars, "Indian")
plots_M <- create_plots(results_M, food_cols, continuous_vars, "Malay")

plots_C
plots_I
plots_M
```


```{r}
# Arrange all plots
# Continuous variables plots on top row, diabetes plots on bottom row
grid.arrange(
  plots_C$continuous, plots_I$continuous, plots_M$continuous,
  plots_C$diabetes, plots_I$diabetes, plots_M$diabetes,
  ncol = 3, nrow = 2
)


ggsave("figure/ethnic_food_clinical_analysis.pdf",
       arrangeGrob(plots_C$continuous, plots_I$continuous, plots_M$continuous,
                  plots_C$diabetes, plots_I$diabetes, plots_M$diabetes,
                  ncol = 3, nrow = 2),
       width = 24, height = 16)
```



## volcano diabete vs non-diabete
```{r}
threshold <- nrow(species1) * 0.05
species_filtered2 <- species1 %>% select_if(function(x) sum(x > 0) > threshold)
species_filtered2

merged_data_filter = species_filtered2 %>%
  inner_join(metadata, by = "Participant_ID")

species_filtered2
```

```{r}
# Create a new diabetes status grouping
merged_data_filter <- merged_data_filter %>%
  mutate(diabetes_group = case_when(
    diabetes_status == "diabetes" ~ "diabetes",
    diabetes_status %in% c("pre-diabetes", "normal") ~ "non_diabetes"
  ))

merged_data_filter <- merged_data_filter %>%
  mutate(obesity_status = case_when(
    BMI < 18.5 ~ "Underweight",
    BMI >= 18.5 & BMI < 23 ~ "Normal",
    BMI >= 23 & BMI < 25 ~ "Overweight",
    BMI >= 25 & BMI < 30 ~ "Obese Class I",
    BMI >= 30 ~ "Obese Class II",
    TRUE ~ NA_character_  # For any missing or invalid BMI values
  ))

# If you want a simpler classification, you can use this instead:
merged_data_filter <- merged_data_filter %>%
  mutate(obesity_binary = case_when(
    BMI < 25 ~ "Non-obese",
    BMI >= 25 ~ "Obese",
    TRUE ~ NA_character_
  ))
```


```{r}
# Get species columns (excluding Participant_ID)
# species_cols <- colnames(species_filtered2)[-1]  # Remove first column
# species_cols
# # Function to perform t-test for each species
# calculate_stats <- function(species_name) {
#   t_test <- t.test(
#     get(species_name) ~ diabetes_group, 
#     data = diabetes_data
#   )
#   
#   # Calculate mean values for each group
#   means <- diabetes_data %>%
#     group_by(diabetes_group) %>%
#     summarise(mean_val = mean(get(species_name))) %>%
#     pivot_wider(names_from = diabetes_group, values_from = mean_val)
#   
#   # Calculate log2 fold change
#   log2FC <- log2(means$diabetes / means$non_diabetes)
#   
#   return(data.frame(
#     species = species_name,
#     log2FoldChange = log2FC,
#     pvalue = t_test$p.value
#   ))
# }
# 
# # Apply statistical testing to all species
# volcano_data <- map_dfr(species_cols, calculate_stats)
# 
# # Add adjusted p-values
# volcano_data$padj <- p.adjust(volcano_data$pvalue, method = "BH")
# 
# # Create volcano plot using ggplot2
# library(ggplot2)
# library(ggrepel)  # For label repulsion
# 
# ggplot(volcano_data, aes(x = log2FoldChange, y = -log10(padj))) +
#   geom_point(aes(color = abs(log2FoldChange) > 0.5 & padj < 0.1)) +
#   geom_text_repel(
#     data = subset(volcano_data, abs(log2FoldChange) > 0.5 & padj < 0.1),
#     aes(label = species),
#     max.overlaps = 15
#   ) +
#   geom_hline(yintercept = -log10(0.1), linetype = "dashed") +
#   geom_vline(xintercept = c(-0.5, 0.5), linetype = "dashed") +
#   scale_color_manual(values = c("grey", "red")) +
#   labs(
#     x = "log2 Fold Change (Diabetes vs Non-diabetes)",
#     y = "-log10 adjusted p-value",
#     title = "Volcano Plot of Species Abundance Changes in Diabetes vs Non-diabetes"
#   ) +
#   theme_minimal() +
#   theme(legend.position = "none")
```
##GMHI
```{r}
# GMHI Calculation adapted for your data format
# Your data: species_filtered2 with Participant_ID in first column, absolute abundances

library(tidyverse)

# First, let's create the species lists adapted to your naming format
# (converting from s__Genus_species to "Genus species")

# Health-prevalent species (MH) - 7 species
MH_species_adapted <- c(
  "Alistipes senegalensis",
  "Bacteroidales bacterium ph8", 
  "Bifidobacterium adolescentis",
  "Bifidobacterium angulatum", 
  "Bifidobacterium catenulatum",
  "Lachnospiraceae bacterium 8 1 57FAA",  # Note: multiple underscores become spaces
  "Sutterella wadsworthensis"
)

# Health-scarce species (MN) - 43 species  
MN_species_adapted <- c(
  "Anaerotruncus colihominis",
  "Atopobium parvulum",
  "Bifidobacterium dentium",
  "Blautia producta", 
  "candidate division TM7 single cell isolate TM7c",
  "Clostridiales bacterium 1 7 47FAA",
  "Clostridium asparagiforme",
  "Clostridium bolteae",
  "Clostridium citroniae", 
  "Clostridium clostridioforme",
  "Clostridium hathewayi",
  "Clostridium nexile",
  "Clostridium ramosum",
  "Clostridium symbiosum",
  "Eggerthella lenta",
  "Erysipelotrichaceae bacterium 2 2 44A",
  "Flavonifractor plautii",
  "Fusobacterium nucleatum", 
  "Gemella morbillorum",
  "Gemella sanguinis",
  "Granulicatella adiacens",
  "Holdemania filiformis",
  "Klebsiella pneumoniae",
  "Lachnospiraceae bacterium 1 4 56FAA",
  "Lachnospiraceae bacterium 2 1 58FAA", 
  "Lachnospiraceae bacterium 3 1 57FAA CT1",
  "Lachnospiraceae bacterium 5 1 57FAA",
  "Lachnospiraceae bacterium 9 1 43BFAA",
  "Lactobacillus salivarius",
  "Peptostreptococcus stomatis",
  "Ruminococcaceae bacterium D16",
  "Ruminococcus gnavus", 
  "Solobacterium moorei",
  "Streptococcus anginosus",
  "Streptococcus australis", 
  "Streptococcus gordonii",
  "Streptococcus infantis",
  "Streptococcus mitis oralis pneumoniae", 
  "Streptococcus sanguinis",
  "Streptococcus vestibularis",
  "Subdoligranulum sp 4 3 54A2FAA",
  "Subdoligranulum variabile",
  "Veillonella atypica"
)

# Function to calculate GMHI
calculate_GMHI <- function(species_data) {
  
  # Step 0: Debug - check data structure
  print("Data structure:")
  print(str(species_data))
  print("First few columns:")
  print(head(species_data[, 1:min(5, ncol(species_data))]))
  
  # Step 1: Prepare data - set Participant_ID as rownames
  rownames(species_data) <- species_data$Participant_ID
  species_data <- species_data[, -1]  # Remove Participant_ID column
  
  # Step 1.5: Convert all columns to numeric (handle any character columns)
  # Check which columns are not numeric
  non_numeric_cols <- sapply(species_data, function(x) !is.numeric(x))
  if(any(non_numeric_cols)) {
    print("Non-numeric columns found:")
    print(names(species_data)[non_numeric_cols])
    print("Converting to numeric...")
    
    # Convert to numeric, coercing characters to NA, then NA to 0
    species_data[non_numeric_cols] <- lapply(species_data[non_numeric_cols], function(x) {
      numeric_x <- as.numeric(as.character(x))
      numeric_x[is.na(numeric_x)] <- 0
      return(numeric_x)
    })
  }
  
  # Check for any remaining issues
  print("Data summary after conversion:")
  print(summary(species_data[, 1:min(3, ncol(species_data))]))
  
  # Step 2: Convert to relative abundances (normalize by row sums)
  row_sums <- rowSums(species_data, na.rm = TRUE)
  print(paste("Row sums range:", min(row_sums), "to", max(row_sums)))
  
  # Avoid division by zero
  row_sums[row_sums == 0] <- 1
  species_profile <- sweep(species_data, 1, row_sums, `/`)
  
  # Step 3: Remove very low abundances and transpose for processing
  species_profile[species_profile < 0.00001] <- 0
  species_profile_t <- data.frame(t(species_profile), check.names = FALSE)
  
  # Step 4: Extract MH and MN species present in your data
  # Check which species names match (you may need to adjust these)
  available_species <- rownames(species_profile_t)
  
  # Find matching MH species
  MH_matches <- available_species[available_species %in% MH_species_adapted]
  MN_matches <- available_species[available_species %in% MN_species_adapted]
  
  print(paste("Found", length(MH_matches), "MH species out of 7"))
  print(paste("Found", length(MN_matches), "MN species out of 43"))
  if(length(MH_matches) > 0) {
    print(paste("MH species found:", paste(MH_matches, collapse = ", ")))
  } else {
    print("No MH species found in data")
  }
  if(length(MN_matches) > 0) {
    print(paste("MN species found:", paste(MN_matches, collapse = ", ")))
  } else {
    print("No MN species found in data")
  }
  
  # Extract the matching species data
  if(length(MH_matches) > 0) {
    MH_species_data <- species_profile_t[MH_matches, , drop = FALSE]
  } else {
    MH_species_data <- data.frame(matrix(0, nrow = 1, ncol = ncol(species_profile_t)))
    colnames(MH_species_data) <- colnames(species_profile_t)
  }
  
  if(length(MN_matches) > 0) {
    MN_species_data <- species_profile_t[MN_matches, , drop = FALSE]  
  } else {
    MN_species_data <- data.frame(matrix(0, nrow = 1, ncol = ncol(species_profile_t)))
    colnames(MN_species_data) <- colnames(species_profile_t)
  }
  
  # Step 5: Calculate Shannon diversity
  alpha <- function(x) {
    x_pos <- x[x > 0]
    if(length(x_pos) == 0) return(0)
    sum(log(x_pos) * x_pos) * (-1)
  }
  
  MH_shannon <- apply(MH_species_data, 2, alpha)
  MN_shannon <- apply(MN_species_data, 2, alpha)
  
  # Step 6: Calculate richness
  R_MH <- apply(MH_species_data, 2, function(x) sum(x > 0))
  R_MN <- apply(MN_species_data, 2, function(x) sum(x > 0))
  
  # Step 7: Calculate collective abundances
  MH_prime <- 7   # Total possible MH species
  MN_prime <- 31  # Reference value from original study
  
  psi_MH <- (R_MH / MH_prime) * MH_shannon
  psi_MN <- (R_MN / MN_prime) * MN_shannon
  
  # Step 8: Calculate GMHI
  GMHI <- log10((psi_MH + 0.00001) / (psi_MN + 0.00001))
  
  # Return results with original participant IDs
  results <- data.frame(
    Participant_ID = rownames(species_data),  # This should preserve your original IDs
    GMHI = as.numeric(GMHI),
    R_MH = R_MH,
    R_MN = R_MN,
    MH_shannon = MH_shannon,
    MN_shannon = MN_shannon,
    stringsAsFactors = FALSE
  )
  
  return(results)
}

```


```{r}

# Calculate GMHI for your data
GMHI_results <- calculate_GMHI(species_filtered2)
GMHI_results$Participant_ID <- species_filtered2$Participant_ID
# View results
print(GMHI_results)

# Save results
#write.csv(GMHI_results, "GMHI_results.csv", row.names = FALSE)
```
```{r}
GMHI_results
```

```{r}
merged_data_filter_GMHI <- inner_join(GMHI_results, metadata, by = "Participant_ID")

# Step 2: Create diabetes status grouping
merged_data_filter_GMHI <- merged_data_filter_GMHI %>%
  mutate(diabetes_group = case_when(
    diabetes_status == "diabetes" ~ "diabetes",
    diabetes_status %in% c("pre-diabetes", "normal") ~ "non_diabetes",
    TRUE ~ NA_character_  # Handle any other values
  ))

# Step 3: Create detailed obesity status classification
merged_data_filter_GMHI <- merged_data_filter_GMHI %>%
  mutate(obesity_status = case_when(
    BMI < 18.5 ~ "Underweight",
    BMI >= 18.5 & BMI < 23 ~ "Normal",
    BMI >= 23 & BMI < 25 ~ "Overweight", 
    BMI >= 25 & BMI < 30 ~ "Obese Class I",
    BMI >= 30 ~ "Obese Class II",
    TRUE ~ NA_character_  # For any missing or invalid BMI values
  ))

# Step 4: Create binary obesity classification
merged_data_filter_GMHI <- merged_data_filter_GMHI %>%
  mutate(obesity_binary = case_when(
    BMI < 25 ~ "Non-obese",
    BMI >= 25 ~ "Obese", 
    TRUE ~ NA_character_
  ))
merged_data_filter_GMHI
```
### test gmhi with obesity
```{r}
obesity_summary <- merged_data_filter_GMHI %>%
  filter(!is.na(obesity_binary) & !is.na(GMHI)) %>%
  group_by(obesity_binary) %>%
  summarise(
    n = n(),
    mean_GMHI = round(mean(GMHI), 3),
    sd_GMHI = round(sd(GMHI), 3),
    median_GMHI = round(median(GMHI), 3),
    min_GMHI = round(min(GMHI), 3),
    max_GMHI = round(max(GMHI), 3),
    mean_BMI = round(mean(BMI, na.rm = TRUE), 1),
    .groups = 'drop'
  )
print(obesity_summary)

# Detailed obesity categories
print("\nGMHI by Detailed Obesity Categories:")
detailed_summary <- merged_data_filter_GMHI %>%
  filter(!is.na(obesity_status) & !is.na(GMHI)) %>%
  group_by(obesity_status) %>%
  summarise(
    n = n(),
    mean_GMHI = round(mean(GMHI), 3),
    sd_GMHI = round(sd(GMHI), 3),
    median_GMHI = round(median(GMHI), 3),
    mean_BMI = round(mean(BMI, na.rm = TRUE), 1),
    .groups = 'drop'
  ) %>%
  arrange(mean_BMI)
print(detailed_summary)
```
### t test
```{r}
valid_data <- merged_data_filter_GMHI %>% 
  filter(!is.na(obesity_binary) & !is.na(GMHI))

if(nrow(valid_data) > 0 && length(unique(valid_data$obesity_binary)) == 2) {
  ttest_result <- t.test(GMHI ~ obesity_binary, data = valid_data)
  print("T-test (Non-obese vs Obese):")
  print(paste("Mean difference:", round(ttest_result$estimate[1] - ttest_result$estimate[2], 4)))
  print(paste("P-value:", round(ttest_result$p.value, 4)))
  print(paste("95% CI for difference:", round(ttest_result$conf.int[1], 4), "to", round(ttest_result$conf.int[2], 4)))
  
  # Effect size (Cohen's d)
  pooled_sd <- sqrt(((sum(!is.na(valid_data$GMHI[valid_data$obesity_binary == unique(valid_data$obesity_binary)[1]])) - 1) * 
                     var(valid_data$GMHI[valid_data$obesity_binary == unique(valid_data$obesity_binary)[1]], na.rm = TRUE) +
                     (sum(!is.na(valid_data$GMHI[valid_data$obesity_binary == unique(valid_data$obesity_binary)[2]])) - 1) * 
                     var(valid_data$GMHI[valid_data$obesity_binary == unique(valid_data$obesity_binary)[2]], na.rm = TRUE)) /
                    (sum(!is.na(valid_data$GMHI)) - 2))
  cohens_d <- (ttest_result$estimate[1] - ttest_result$estimate[2]) / pooled_sd
  print(paste("Cohen's d (effect size):", round(cohens_d, 3)))
}

# ANOVA for detailed obesity categories
valid_detailed <- merged_data_filter_GMHI %>% 
  filter(!is.na(obesity_status) & !is.na(GMHI))

if(nrow(valid_detailed) > 0 && length(unique(valid_detailed$obesity_status)) > 2) {
  anova_result <- aov(GMHI ~ obesity_status, data = valid_detailed)
  print("\nANOVA (Detailed Obesity Categories):")
  print(summary(anova_result))
  
  # Post-hoc test if ANOVA is significant
  anova_p <- summary(anova_result)[[1]][["Pr(>F)"]][1]
  if(anova_p < 0.05) {
    print("\nPost-hoc Tukey HSD test:")
    tukey_result <- TukeyHSD(anova_result)
    print(tukey_result)
  }
}
```
```{r}
p2 <- ggplot(merged_data_filter_GMHI %>% filter(!is.na(obesity_binary)), 
             aes(x = obesity_binary, y = GMHI, fill = obesity_binary)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 3, fill = "white") +
  labs(title = "GMHI by Obesity Status (Binary)",
       x = "Obesity Status",
       y = "Gut Microbiome Health Index (GMHI)") +
  theme_minimal() +
  theme(legend.position = "none",
        plot.title = element_text(size = 14, face = "bold")) +
  scale_fill_manual(values = c("Non-obese" = "lightblue", "Obese" = "salmon"))
p2

```
```{r}
p3 <- ggplot(merged_data_filter_GMHI %>% filter(!is.na(obesity_status)), 
             aes(x = reorder(obesity_status, BMI, FUN = median, na.rm = TRUE), 
                 y = GMHI, fill = obesity_status)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 3, fill = "white") +
  labs(title = "GMHI by Detailed Obesity Categories",
       x = "Obesity Category",
       y = "Gut Microbiome Health Index (GMHI)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none",
        plot.title = element_text(size = 14, face = "bold")) +
  scale_fill_brewer(type = "qual", palette = "Set3")
p3
```
### linear regression
```{r}
lm_model <- lm(GMHI ~ BMI, data = merged_data_filter_GMHI)
print("Linear Regression Summary (GMHI ~ BMI):")
print(summary(lm_model))

# Extract key metrics
r_squared <- summary(lm_model)$r.squared
adj_r_squared <- summary(lm_model)$adj.r.squared
print(paste("R-squared:", round(r_squared, 4)))
print(paste("Adjusted R-squared:", round(adj_r_squared, 4)))

```

## bifido in obese and in indian
```{r}
species_names <- colnames(species_filtered2)[-1]  # excluding Participant_ID

# Find Bifidobacterium species
bifido_species <- species_names[grepl("Bifidobacterium", species_names, ignore.case = TRUE)]
print(paste("Found", length(bifido_species), "Bifidobacterium species:"))
print(bifido_species)

if(length(bifido_species) == 0) {
  # Try alternative patterns
  bifido_species <- species_names[grepl("Bifido", species_names, ignore.case = TRUE)]
  print("Trying 'Bifido' pattern:")
  print(bifido_species)
}

# Step 2: Calculate total Bifidobacterium abundance for each participant
print("\n2. CALCULATING BIFIDOBACTERIUM ABUNDANCES")
print("----------------------------------------")

if(length(bifido_species) > 0) {
  # Extract Bifidobacterium data
  bifido_data <- species_filtered2 %>%
    dplyr::select(Participant_ID, all_of(bifido_species))
  
  # Calculate total Bifidobacterium abundance
  bifido_data$Total_Bifidobacterium <- rowSums(bifido_data[, -1], na.rm = TRUE)
  
  # Calculate total abundance across all species for relative abundance
  total_abundance <- rowSums(species_filtered2[, -1], na.rm = TRUE)
  bifido_data$Relative_Bifidobacterium <- (bifido_data$Total_Bifidobacterium / total_abundance) * 100
  
  # Add individual species relative abundances
  for(species in bifido_species) {
    rel_col_name <- paste0("Rel_", gsub(" ", "_", species))
    bifido_data[[rel_col_name]] <- (species_filtered2[[species]] / total_abundance) * 100
  }
  
  print("Bifidobacterium abundance summary:")
  print(summary(bifido_data$Total_Bifidobacterium))
  print("Relative abundance (%) summary:")
  print(summary(bifido_data$Relative_Bifidobacterium))
  
} else {
  print("No Bifidobacterium species found! Please check species names.")
  print("First 20 species in your dataset:")
  print(head(species_names, 20))
  stop("Cannot proceed without Bifidobacterium data")
}
```
```{r}
if(length(bifido_species) > 0) {
  # Extract Bifidobacterium data
  bifido_data <- species_filtered2 %>%
    dplyr::select(Participant_ID, all_of(bifido_species))
  
  # Calculate total Bifidobacterium abundance
  bifido_data$Total_Bifidobacterium <- rowSums(bifido_data[, -1], na.rm = TRUE)
  
  # Calculate total abundance across all species for relative abundance
  total_abundance <- rowSums(species_filtered2[, -1], na.rm = TRUE)
  bifido_data$Relative_Bifidobacterium <- (bifido_data$Total_Bifidobacterium / total_abundance) * 100
  
  # Add individual species relative abundances
  for(species in bifido_species) {
    rel_col_name <- paste0("Rel_", gsub(" ", "_", species))
    bifido_data[[rel_col_name]] <- (species_filtered2[[species]] / total_abundance) * 100
  }
  
  print("Bifidobacterium abundance summary:")
  print(summary(bifido_data$Total_Bifidobacterium))
  print("Relative abundance (%) summary:")
  print(summary(bifido_data$Relative_Bifidobacterium))
  
} else {
  print("No Bifidobacterium species found! Please check species names.")
  print("First 20 species in your dataset:")
  print(head(species_names, 20))
  stop("Cannot proceed without Bifidobacterium data")
}
bifido_data
```
```{r}
merged_data_filter_GMHI
```

```{r}
library(ggpubr)
bifido_merged <- inner_join(bifido_data, merged_data_filter_GMHI, by = "Participant_ID")
valid_data <- bifido_merged %>% 
  filter(!is.na(obesity_binary) & !is.na(Relative_Bifidobacterium))

print(paste("Valid data for analysis:", nrow(valid_data), "participants"))

# T-test for relative abundance
if(nrow(valid_data) > 0 && length(unique(valid_data$obesity_binary)) == 2) {
  ttest_rel <- t.test(Relative_Bifidobacterium ~ obesity_binary, data = valid_data)
  
  print("T-test Results:")
  print(paste("Non-obese mean:", round(ttest_rel$estimate[1], 3), "%"))
  print(paste("Obese mean:", round(ttest_rel$estimate[2], 3), "%"))
  print(paste("Difference:", round(ttest_rel$estimate[1] - ttest_rel$estimate[2], 3), "%"))
  print(paste("P-value:", round(ttest_rel$p.value, 6)))
  
  # Determine significance level for stars
  p_value <- ttest_rel$p.value
  if(p_value < 0.001) {
    sig_star <- "***"
    sig_text <- "p < 0.001"
  } else if(p_value < 0.01) {
    sig_star <- "**"
    sig_text <- paste("p =", round(p_value, 3))
  } else if(p_value < 0.05) {
    sig_star <- "*"
    sig_text <- paste("p =", round(p_value, 3))
  } else {
    sig_star <- "ns"
    sig_text <- paste("p =", round(p_value, 3))
  }
  
  print(paste("Significance:", sig_star, "-", sig_text))
} else {
  print("Cannot perform t-test - insufficient data or groups")
  sig_star <- "ns"
  sig_text <- "Cannot test"
  p_value <- 1
}

# Step 4: Create summary statistics for the plot
summary_stats <- valid_data %>%
  group_by(obesity_binary) %>%
  summarise(
    n = n(),
    mean = round(mean(Relative_Bifidobacterium, na.rm = TRUE), 3),
    sd = round(sd(Relative_Bifidobacterium, na.rm = TRUE), 3),
    median = round(median(Relative_Bifidobacterium, na.rm = TRUE), 3),
    q25 = quantile(Relative_Bifidobacterium, 0.25, na.rm = TRUE),
    q75 = quantile(Relative_Bifidobacterium, 0.75, na.rm = TRUE),
    .groups = 'drop'
  )

print("\nSummary Statistics:")
print(summary_stats)

# Step 5: Create the box plot with significance stars
print("\nCreating Box Plot...")

# Calculate position for significance annotation
y_max <- max(valid_data$Relative_Bifidobacterium, na.rm = TRUE)
y_position <- y_max * 1.1

# Create the main box plot
p1 <- ggplot(valid_data, aes(x = obesity_binary, y = Relative_Bifidobacterium, fill = obesity_binary)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.6) +
  geom_jitter(width = 0.2, alpha = 0.5, size = 1.5) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 3, fill = "white", stroke = 1) +
  
  # Add significance annotation
  annotate("segment", x = 1, xend = 2, y = y_position, yend = y_position, color = "black") +
  annotate("segment", x = 1, xend = 1, y = y_position, yend = y_position * 0.95, color = "black") +
  annotate("segment", x = 2, xend = 2, y = y_position, yend = y_position * 0.95, color = "black") +
  annotate("text", x = 1.5, y = y_position * 1.05, label = sig_star, size = 6, fontface = "bold") +
  
  # Labels and styling
  labs(
    title = "Bifidobacterium Relative Abundance by Obesity Status",
    subtitle = paste("T-test:", sig_text, "| n =", nrow(valid_data)),
    x = "Obesity Status",
    y = "Bifidobacterium Relative Abundance (%)",
    caption = paste("* p < 0.05, ** p < 0.01, *** p < 0.001, ns = not significant")
  ) +
  
  # Theme and colors
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 11),
    legend.position = "none",
    panel.grid.minor = element_blank()
  ) +
  scale_fill_manual(values = c("Non-obese" = "lightblue", "Obese" = "salmon")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))  # Extra space on top for annotation

# Display the plot
print(p1)

# Step 6: Alternative plot with ggpubr for automatic significance
print("\nCreating Alternative Plot with ggpubr...")

# Using ggpubr for automatic significance testing
p2 <- ggboxplot(valid_data, 
                x = "obesity_binary", 
                y = "Relative_Bifidobacterium",
                fill = "obesity_binary",
                palette = c("lightblue", "salmon"),
                add = "jitter",
                add.params = list(alpha = 0.5)) +
  
  stat_compare_means(method = "t.test", 
                     label = "p.signif",
                     label.y = y_position,
                     size = 6) +
  
  labs(
    title = "Bifidobacterium Relative Abundance by Obesity Status",
    subtitle = paste("Automatic significance testing | n =", nrow(valid_data)),
    x = "Obesity Status",
    y = "Bifidobacterium Relative Abundance (%)"
  ) +
  
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 11),
    legend.position = "none"
  )

print(p2)
```
```{r}
valid_diabetes_data <- bifido_merged %>% 
  filter(!is.na(diabetes_status) & !is.na(Relative_Bifidobacterium) & 
         diabetes_status %in% c("pre-diabetes", "diabetes", "normal"))

print(paste("Valid data for diabetes analysis:", nrow(valid_diabetes_data), "participants"))
print("Distribution by diabetes status:")
print(table(valid_diabetes_data$diabetes_status))

# Step 3: Statistical tests for diabetes groups
print("\nStatistical Tests:")
print("==================")

# ANOVA test for overall differences
if(nrow(valid_diabetes_data) > 0 && length(unique(valid_diabetes_data$diabetes_status)) >= 2) {
  anova_result <- aov(Relative_Bifidobacterium ~ diabetes_status, data = valid_diabetes_data)
  anova_summary <- summary(anova_result)
  anova_p <- anova_summary[[1]][["Pr(>F)"]][1]
  
  print("ANOVA Results:")
  print(anova_summary)
  print(paste("Overall ANOVA p-value:", round(anova_p, 6)))
  
  # Determine overall significance
  if(anova_p < 0.001) {
    overall_sig <- "***"
  } else if(anova_p < 0.01) {
    overall_sig <- "**"
  } else if(anova_p < 0.05) {
    overall_sig <- "*"
  } else {
    overall_sig <- "ns"
  }
  
  print(paste("Overall significance:", overall_sig))
  
  # Post-hoc pairwise comparisons if ANOVA is significant
  if(anova_p < 0.05) {
    print("\nPost-hoc Pairwise Comparisons (Tukey HSD):")
    tukey_result <- TukeyHSD(anova_result)
    print(tukey_result)
    
    # Extract pairwise p-values for plotting
    pairwise_p <- tukey_result$diabetes_status[, "p adj"]
    pairwise_comparisons <- data.frame(
      comparison = names(pairwise_p),
      p_value = pairwise_p,
      stringsAsFactors = FALSE
    )
    
    # Add significance stars for each comparison
    pairwise_comparisons$sig_star <- sapply(pairwise_comparisons$p_value, function(p) {
      if(p < 0.001) "***"
      else if(p < 0.01) "**" 
      else if(p < 0.05) "*"
      else "ns"
    })
    
    print("Pairwise significance:")
    print(pairwise_comparisons)
  } else {
    pairwise_comparisons <- NULL
  }
} else {
  print("Cannot perform ANOVA - insufficient data or groups")
  overall_sig <- "ns"
  anova_p <- 1
  pairwise_comparisons <- NULL
}

# Step 4: Summary statistics by diabetes status
summary_diabetes <- valid_diabetes_data %>%
  group_by(diabetes_status) %>%
  summarise(
    n = n(),
    mean = round(mean(Relative_Bifidobacterium, na.rm = TRUE), 3),
    sd = round(sd(Relative_Bifidobacterium, na.rm = TRUE), 3),
    median = round(median(Relative_Bifidobacterium, na.rm = TRUE), 3),
    q25 = round(quantile(Relative_Bifidobacterium, 0.25, na.rm = TRUE), 3),
    q75 = round(quantile(Relative_Bifidobacterium, 0.75, na.rm = TRUE), 3),
    .groups = 'drop'
  ) %>%
  arrange(match(diabetes_status, c("normal", "pre-diabetes", "diabetes")))

print("\nSummary Statistics by Diabetes Status:")
print(summary_diabetes)

# Step 5: Create the main box plot
print("\nCreating Box Plot...")

# Order the diabetes status for logical progression
valid_diabetes_data$diabetes_status <- factor(valid_diabetes_data$diabetes_status, 
                                             levels = c("normal", "pre-diabetes", "diabetes"))

# Calculate y-position for significance annotations
y_max <- max(valid_diabetes_data$Relative_Bifidobacterium, na.rm = TRUE)
y_positions <- c(y_max * 1.1, y_max * 1.2, y_max * 1.3)

# Create the main box plot
p1 <- ggplot(valid_diabetes_data, aes(x = diabetes_status, y = Relative_Bifidobacterium, fill = diabetes_status)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.6) +
  geom_jitter(width = 0.2, alpha = 0.5, size = 1.5) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 3, fill = "white", stroke = 1) +
  
  # Labels and styling
  labs(
    title = "Bifidobacterium Relative Abundance by Diabetes Status",
    subtitle = paste("ANOVA:", overall_sig, "| p =", round(anova_p, 4), "| n =", nrow(valid_diabetes_data)),
    x = "Diabetes Status",
    y = "Bifidobacterium Relative Abundance (%)",
    caption = "* p < 0.05, ** p < 0.01, *** p < 0.001, ns = not significant"
  ) +
  
  # Theme and colors
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 11),
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    legend.position = "none",
    panel.grid.minor = element_blank()
  ) +
  scale_fill_manual(values = c("normal" = "lightgreen", "pre-diabetes" = "orange", "diabetes" = "red")) +
  scale_x_discrete(labels = c("normal" = "Normal", "pre-diabetes" = "Pre-diabetes", "diabetes" = "Diabetes")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.2)))  # Extra space on top for annotations

# Add pairwise significance annotations if available
if(!is.null(pairwise_comparisons) && nrow(pairwise_comparisons) > 0) {
  # Add significance bars for significant comparisons
  sig_comparisons <- pairwise_comparisons[pairwise_comparisons$sig_star != "ns", ]
  
  if(nrow(sig_comparisons) > 0) {
    for(i in 1:nrow(sig_comparisons)) {
      comp <- sig_comparisons$comparison[i]
      sig <- sig_comparisons$sig_star[i]
      
      # Parse comparison (e.g., "diabetes-normal")
      groups <- unlist(strsplit(comp, "-"))
      group1 <- groups[1]
      group2 <- groups[2]
      
      # Get x positions
      x1 <- which(levels(valid_diabetes_data$diabetes_status) == group1)
      x2 <- which(levels(valid_diabetes_data$diabetes_status) == group2)
      
      y_pos <- y_positions[i]
      
      # Add significance annotation
      p1 <- p1 + 
        annotate("segment", x = x1, xend = x2, y = y_pos, yend = y_pos, color = "black") +
        annotate("segment", x = x1, xend = x1, y = y_pos, yend = y_pos * 0.98, color = "black") +
        annotate("segment", x = x2, xend = x2, y = y_pos, yend = y_pos * 0.98, color = "black") +
        annotate("text", x = (x1 + x2)/2, y = y_pos * 1.02, label = sig, size = 5, fontface = "bold")
    }
  }
}

print(p1)
```


### combine
```{r}
library(gridExtra)  # For combining plots
library(grid)
valid_ethnic_data <- bifido_merged %>%
  filter(!is.na(FREG5_Ethnic_Group) & 
         !is.na(Relative_Bifidobacterium) &
         FREG5_Ethnic_Group %in% c("C", "I", "M"))

print(paste("Participants with valid ethnic and Bifidobacterium data:", nrow(valid_ethnic_data)))

# Check sample sizes by ethnic group
ethnic_counts <- table(valid_ethnic_data$FREG5_Ethnic_Group)
print("Sample sizes by ethnic group:")
print(ethnic_counts)

# Step 2: Create ethnic group labels
ethnic_labels <- c("C" = "Chinese", "I" = "Indian", "M" = "Malay")
valid_ethnic_data$Ethnic_Group_Label <- ethnic_labels[valid_ethnic_data$FREG5_Ethnic_Group]

# ===============================================
# DIABETES ANALYSIS BY ETHNICITY
# ===============================================

# Filter for diabetes analysis
diabetes_ethnic_data <- valid_ethnic_data %>%
  filter(!is.na(diabetes_status) & 
         diabetes_status %in% c("normal", "pre-diabetes", "diabetes"))

print(paste("Participants for diabetes analysis:", nrow(diabetes_ethnic_data)))

# Sample sizes by ethnic group and diabetes status
print("Sample sizes by ethnic group and diabetes status:")
diabetes_ethnic_table <- table(diabetes_ethnic_data$FREG5_Ethnic_Group, 
                              diabetes_ethnic_data$diabetes_status)
print(diabetes_ethnic_table)

# Summary statistics by ethnicity and diabetes
diabetes_ethnic_summary <- diabetes_ethnic_data %>%
  group_by(FREG5_Ethnic_Group, Ethnic_Group_Label, diabetes_status) %>%
  summarise(
    n = n(),
    mean = round(mean(Relative_Bifidobacterium, na.rm = TRUE), 3),
    sd = round(sd(Relative_Bifidobacterium, na.rm = TRUE), 3),
    median = round(median(Relative_Bifidobacterium, na.rm = TRUE), 3),
    .groups = 'drop'
  ) %>%
  arrange(FREG5_Ethnic_Group, diabetes_status)

print("Summary statistics:")
print(diabetes_ethnic_summary)

# Statistical tests within each ethnic group
print("\nStatistical tests within each ethnic group (ANOVA):")
for(ethnic in c("C", "I", "M")) {
  ethnic_subset <- diabetes_ethnic_data %>% 
    filter(FREG5_Ethnic_Group == ethnic)
  
  if(nrow(ethnic_subset) > 0 && length(unique(ethnic_subset$diabetes_status)) >= 2) {
    anova_ethnic <- aov(Relative_Bifidobacterium ~ diabetes_status, data = ethnic_subset)
    anova_p <- summary(anova_ethnic)[[1]][["Pr(>F)"]][1]
    
    print(paste(ethnic_labels[ethnic], "- ANOVA p-value:", round(anova_p, 4)))
    
    if(anova_p < 0.05 && length(unique(ethnic_subset$diabetes_status)) > 2) {
      tukey_ethnic <- TukeyHSD(anova_ethnic)
      print(paste("  Significant differences found - Tukey HSD:"))
      print(tukey_ethnic$diabetes_status)
    }
  } else {
    print(paste(ethnic_labels[ethnic], "- Insufficient data for ANOVA"))
  }
}

# Create diabetes box plots by ethnicity with significance lines
diabetes_ethnic_data$diabetes_status <- factor(diabetes_ethnic_data$diabetes_status,
                                              levels = c("normal", "pre-diabetes", "diabetes"))

# Create separate plots for each ethnicity to add significance properly
diabetes_plots <- list()

for(ethnic in c("C", "I", "M")) {
  ethnic_subset <- diabetes_ethnic_data %>% 
    filter(FREG5_Ethnic_Group == ethnic)
  
  if(nrow(ethnic_subset) > 0) {
    # Perform statistical tests for this ethnicity
    comparisons_list <- list(c("normal", "pre-diabetes"), 
                           c("normal", "diabetes"), 
                           c("pre-diabetes", "diabetes"))
    
    # Filter comparisons to only include groups that exist
    available_groups <- unique(ethnic_subset$diabetes_status)
    valid_comparisons <- comparisons_list[sapply(comparisons_list, function(x) all(x %in% available_groups))]
    
    if(length(valid_comparisons) > 0 && length(available_groups) >= 2) {
      p_ethnic <- ggboxplot(ethnic_subset,
                           x = "diabetes_status", 
                           y = "Relative_Bifidobacterium",
                           fill = "diabetes_status",
                           palette = c("lightgreen", "orange", "red"),
                           add = "jitter",
                           add.params = list(alpha = 0.5, size = 1)) +
        
        stat_compare_means(comparisons = valid_comparisons,
                          method = "t.test",
                          label = "p.signif",
                          step.increase = 0.1,
                          tip.length = 0.02) +
        
        labs(
          title = paste(ethnic_labels[ethnic]),
          x = "Diabetes Status",
          y = if(ethnic == "C") "Bifidobacterium Relative Abundance (%)" else ""
        ) +
        
        theme_minimal() +
        theme(
          plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
          axis.title = element_text(size = 10),
          axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
          axis.text.y = element_text(size = 9),
          legend.position = "none"
        ) +
        scale_x_discrete(labels = c("normal" = "Normal", "pre-diabetes" = "Pre-diabetes", "diabetes" = "Diabetes"))
      
    } else {
      # Create basic plot without significance if insufficient groups
      p_ethnic <- ggplot(ethnic_subset, aes(x = diabetes_status, y = Relative_Bifidobacterium, fill = diabetes_status)) +
        geom_boxplot(alpha = 0.7) +
        geom_jitter(width = 0.2, alpha = 0.5, size = 1) +
        labs(
          title = paste(ethnic_labels[ethnic]),
          x = "Diabetes Status",
          y = if(ethnic == "C") "Bifidobacterium Relative Abundance (%)" else ""
        ) +
        theme_minimal() +
        theme(
          plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
          axis.title = element_text(size = 10),
          axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
          axis.text.y = element_text(size = 9),
          legend.position = "none"
        ) +
        scale_fill_manual(values = c("normal" = "lightgreen", "pre-diabetes" = "orange", "diabetes" = "red")) +
        scale_x_discrete(labels = c("normal" = "Normal", "pre-diabetes" = "Pre-diabetes", "diabetes" = "Diabetes"))
    }
    
    diabetes_plots[[ethnic]] <- p_ethnic
  }
}

# Combine plots
if(length(diabetes_plots) > 0) {
  p_diabetes_ethnic <- grid.arrange(grobs = diabetes_plots, ncol = 3,
                                   top = textGrob("Bifidobacterium by Diabetes Status - Stratified by Ethnicity", 
                                                 gp = gpar(fontsize = 14, fontface = "bold")))
} else {
  # Fallback to original plot style
  p_diabetes_ethnic <- ggplot(diabetes_ethnic_data, 
                             aes(x = diabetes_status, y = Relative_Bifidobacterium, fill = diabetes_status)) +
    geom_boxplot(alpha = 0.7) +
    geom_jitter(width = 0.2, alpha = 0.5, size = 1) +
    facet_wrap(~Ethnic_Group_Label, scales = "free_y") +
    labs(
      title = "Bifidobacterium by Diabetes Status - Stratified by Ethnicity",
      x = "Diabetes Status",
      y = "Bifidobacterium Relative Abundance (%)"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      axis.title = element_text(size = 11),
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none",
      strip.text = element_text(size = 11, face = "bold")
    ) +
    scale_fill_manual(values = c("normal" = "lightgreen", "pre-diabetes" = "orange", "diabetes" = "red")) +
    scale_x_discrete(labels = c("normal" = "Normal", "pre-diabetes" = "Pre-diabetes", "diabetes" = "Diabetes"))
}

print(p_diabetes_ethnic)

# ===============================================
# OBESITY ANALYSIS BY ETHNICITY
# ===============================================



# Filter for obesity analysis
obesity_ethnic_data <- valid_ethnic_data %>%
  filter(!is.na(obesity_binary))

print(paste("Participants for obesity analysis:", nrow(obesity_ethnic_data)))

# Sample sizes by ethnic group and obesity status
print("Sample sizes by ethnic group and obesity status:")
obesity_ethnic_table <- table(obesity_ethnic_data$FREG5_Ethnic_Group, 
                             obesity_ethnic_data$obesity_binary)
print(obesity_ethnic_table)

# Summary statistics by ethnicity and obesity
obesity_ethnic_summary <- obesity_ethnic_data %>%
  group_by(FREG5_Ethnic_Group, Ethnic_Group_Label, obesity_binary) %>%
  summarise(
    n = n(),
    mean = round(mean(Relative_Bifidobacterium, na.rm = TRUE), 3),
    sd = round(sd(Relative_Bifidobacterium, na.rm = TRUE), 3),
    median = round(median(Relative_Bifidobacterium, na.rm = TRUE), 3),
    .groups = 'drop'
  ) %>%
  arrange(FREG5_Ethnic_Group, obesity_binary)

print("Summary statistics:")
print(obesity_ethnic_summary)

# Statistical tests within each ethnic group (t-tests)
print("\nStatistical tests within each ethnic group (t-tests):")
ttest_results <- data.frame()

for(ethnic in c("C", "I", "M")) {
  ethnic_subset <- obesity_ethnic_data %>% 
    filter(FREG5_Ethnic_Group == ethnic)
  
  if(nrow(ethnic_subset) > 0 && length(unique(ethnic_subset$obesity_binary)) == 2) {
    ttest_ethnic <- t.test(Relative_Bifidobacterium ~ obesity_binary, data = ethnic_subset)
    
    # Determine significance
    p_val <- ttest_ethnic$p.value
    if(p_val < 0.001) sig_star <- "***"
    else if(p_val < 0.01) sig_star <- "**"
    else if(p_val < 0.05) sig_star <- "*"
    else sig_star <- "ns"
    
    print(paste(ethnic_labels[ethnic], ":"))
    print(paste("  Non-obese mean:", round(ttest_ethnic$estimate[1], 3)))
    print(paste("  Obese mean:", round(ttest_ethnic$estimate[2], 3)))
    print(paste("  Difference:", round(ttest_ethnic$estimate[1] - ttest_ethnic$estimate[2], 3)))
    print(paste("  P-value:", round(p_val, 4), "(", sig_star, ")"))
    
    # Store results for plotting
    ttest_results <- rbind(ttest_results, data.frame(
      Ethnic_Group = ethnic,
      Ethnic_Label = ethnic_labels[ethnic],
      p_value = p_val,
      sig_star = sig_star,
      stringsAsFactors = FALSE
    ))
  } else {
    print(paste(ethnic_labels[ethnic], "- Insufficient data for t-test"))
  }
}

# Create obesity box plots by ethnicity with significance lines
obesity_plots <- list()

for(ethnic in c("C", "I", "M")) {
  ethnic_subset <- obesity_ethnic_data %>% 
    filter(FREG5_Ethnic_Group == ethnic)
  
  if(nrow(ethnic_subset) > 0 && length(unique(ethnic_subset$obesity_binary)) == 2) {
    # Get significance result for this ethnicity
    ttest_result <- ttest_results %>% filter(Ethnic_Group == ethnic)
    
    if(nrow(ttest_result) > 0) {
      p_ethnic <- ggboxplot(ethnic_subset,
                           x = "obesity_binary", 
                           y = "Relative_Bifidobacterium",
                           fill = "obesity_binary",
                           palette = c("lightblue", "salmon"),
                           add = "jitter",
                           add.params = list(alpha = 0.5, size = 1)) +
        
        stat_compare_means(method = "t.test",
                          label = "p.signif",
                          comparisons = list(c("Non-obese", "Obese")),
                          tip.length = 0.02) +
        
        labs(
          title = paste(ethnic_labels[ethnic]),
          x = "Obesity Status",
          y = if(ethnic == "C") "Bifidobacterium Relative Abundance (%)" else ""
        ) +
        
        theme_minimal() +
        theme(
          plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
          axis.title = element_text(size = 10),
          axis.text = element_text(size = 9),
          legend.position = "none"
        )
      
    } else {
      # Create basic plot without significance
      p_ethnic <- ggplot(ethnic_subset, aes(x = obesity_binary, y = Relative_Bifidobacterium, fill = obesity_binary)) +
        geom_boxplot(alpha = 0.7) +
        geom_jitter(width = 0.2, alpha = 0.5, size = 1) +
        labs(
          title = paste(ethnic_labels[ethnic]),
          x = "Obesity Status",
          y = if(ethnic == "C") "Bifidobacterium Relative Abundance (%)" else ""
        ) +
        theme_minimal() +
        theme(
          plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
          axis.title = element_text(size = 10),
          axis.text = element_text(size = 9),
          legend.position = "none"
        ) +
        scale_fill_manual(values = c("Non-obese" = "lightblue", "Obese" = "salmon"))
    }
    
    obesity_plots[[ethnic]] <- p_ethnic
  }
}

# Combine obesity plots
if(length(obesity_plots) > 0) {
  p_obesity_ethnic <- grid.arrange(grobs = obesity_plots, ncol = 3,
                                  top = textGrob("Bifidobacterium by Obesity Status - Stratified by Ethnicity", 
                                                gp = gpar(fontsize = 14, fontface = "bold")))
} else {
  # Fallback plot
  p_obesity_ethnic <- ggplot(obesity_ethnic_data, 
                            aes(x = obesity_binary, y = Relative_Bifidobacterium, fill = obesity_binary)) +
    geom_boxplot(alpha = 0.7) +
    geom_jitter(width = 0.2, alpha = 0.5, size = 1) +
    facet_wrap(~Ethnic_Group_Label, scales = "free_y") +
    labs(
      title = "Bifidobacterium by Obesity Status - Stratified by Ethnicity",
      x = "Obesity Status", 
      y = "Bifidobacterium Relative Abundance (%)"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      axis.title = element_text(size = 11),
      axis.text = element_text(size = 10),
      legend.position = "none",
      strip.text = element_text(size = 11, face = "bold")
    ) +
    scale_fill_manual(values = c("Non-obese" = "lightblue", "Obese" = "salmon"))
}

print(p_obesity_ethnic)

# ===============================================
# COMBINED ANALYSIS
# ===============================================



# Create combined dataset with both diabetes and obesity info
combined_data <- valid_ethnic_data %>%
  filter(!is.na(diabetes_status) & diabetes_status %in% c("normal", "pre-diabetes", "diabetes") &
         !is.na(obesity_binary))

print(paste("Participants with both diabetes and obesity data:", nrow(combined_data)))

# Create interaction plots
print("\nCreating interaction plots...")

# 1. Diabetes by ethnicity with obesity as color
p_diabetes_obesity <- ggplot(combined_data, 
                            aes(x = diabetes_status, y = Relative_Bifidobacterium, 
                                fill = obesity_binary)) +
  geom_boxplot(alpha = 0.7, position = position_dodge(0.8)) +
  facet_wrap(~Ethnic_Group_Label) +
  
  labs(
    title = "Bifidobacterium by Diabetes Status and Obesity - Stratified by Ethnicity",
    x = "Diabetes Status",
    y = "Bifidobacterium Relative Abundance (%)",
    fill = "Obesity Status"
  ) +
  
  theme_minimal() +
  theme(
    plot.title = element_text(size = 13, face = "bold", hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 11, face = "bold")
  ) +
  scale_fill_manual(values = c("Non-obese" = "lightblue", "Obese" = "salmon")) +
  scale_x_discrete(labels = c("normal" = "Normal", "pre-diabetes" = "Pre-diabetes", "diabetes" = "Diabetes"))

print(p_diabetes_obesity)

# 2. Statistical summary table for all combinations
print("\nComprehensive summary table:")
comprehensive_summary <- combined_data %>%
  group_by(Ethnic_Group_Label, diabetes_status, obesity_binary) %>%
  summarise(
    n = n(),
    mean = round(mean(Relative_Bifidobacterium, na.rm = TRUE), 3),
    sd = round(sd(Relative_Bifidobacterium, na.rm = TRUE), 3),
    .groups = 'drop'
  ) %>%
  arrange(Ethnic_Group_Label, diabetes_status, obesity_binary)

print(comprehensive_summary)
```
## each bifido
```{r}
bifido_species_cols <- colnames(species_filtered2)[grepl("Bifidobacterium", colnames(species_filtered2), ignore.case = TRUE)]
bifido_species_cols
```
```{r}
species_long <- valid_ethnic_data %>%
  dplyr::select(Participant_ID, obesity_binary, all_of(bifido_species_cols)) %>%
  pivot_longer(cols = all_of(bifido_species_cols),
               names_to = "Species",
               values_to = "Abundance") %>%
  mutate(
    Species = gsub("_", " ", Species),  # Replace underscores with spaces
    Species = str_to_title(Species)     # Capitalize first letter of each word
  ) %>%
  filter(!is.na(obesity_binary))  # Remove any missing obesity data




species_test_results <- data.frame()

for(species in unique(species_long$Species)) {
  species_data <- species_long %>% filter(Species == species)
  
  if(nrow(species_data) > 0 && length(unique(species_data$obesity_binary)) == 2) {
    # Perform t-test
    ttest_result <- t.test(Abundance ~ obesity_binary, data = species_data)
    
    # Calculate summary statistics
    summary_stats <- species_data %>%
      group_by(obesity_binary) %>%
      summarise(
        mean = mean(Abundance, na.rm = TRUE),
        median = median(Abundance, na.rm = TRUE),
        n = n(),
        .groups = 'drop'
      )
    
    # Determine significance
    p_val <- ttest_result$p.value
    if(p_val < 0.001) sig_star <- "***"
    else if(p_val < 0.01) sig_star <- "**"
    else if(p_val < 0.05) sig_star <- "*"
    else sig_star <- "ns"
    
    # Store results
    non_obese_mean <- ifelse(nrow(summary_stats %>% filter(obesity_binary == "Non-obese")) > 0,
                            summary_stats$mean[summary_stats$obesity_binary == "Non-obese"], NA)
    obese_mean <- ifelse(nrow(summary_stats %>% filter(obesity_binary == "Obese")) > 0,
                        summary_stats$mean[summary_stats$obesity_binary == "Obese"], NA)
    
    species_test_results <- rbind(species_test_results, data.frame(
      Species = species,
      Non_obese_mean = non_obese_mean,
      Obese_mean = obese_mean,
      Difference = ttest_result$estimate[1] - ttest_result$estimate[2],
      p_value = p_val,
      sig_star = sig_star,
      stringsAsFactors = FALSE
    ))
    
    print(paste(species, ":"))
    print(paste("  Non-obese mean:", round(ttest_result$estimate[1], 2)))
    print(paste("  Obese mean:", round(ttest_result$estimate[2], 2)))
    print(paste("  Difference:", round(ttest_result$estimate[1] - ttest_result$estimate[2], 2)))
    print(paste("  P-value:", round(p_val, 6), "(", sig_star, ")"))
  }
}

print("\nSpecies test results summary:")
print(species_test_results)
write.csv(species_test_results, "figure/bifido.csv")
# Step 5: Create stacked box plot
print("\nCreating stacked box plot...")

# Order species by mean abundance for better visualization
species_order <- species_long %>%
  group_by(Species) %>%
  summarise(mean_abundance = mean(Abundance, na.rm = TRUE), .groups = 'drop') %>%
  arrange(desc(mean_abundance)) %>%
  pull(Species)

species_long$Species <- factor(species_long$Species, levels = species_order)

# Create the main stacked box plot
p_species_stacked <- ggplot(species_long, aes(x = Species, y = Abundance, fill = obesity_binary)) +
  geom_boxplot(alpha = 0.7, position = position_dodge(0.8), outlier.alpha = 0.6) +
  
  # Add individual points
  geom_point(position = position_jitterdodge(dodge.width = 0.8, jitter.width = 0.2), 
             alpha = 0.4, size = 0.8) +
  
  labs(
    title = "Individual Bifidobacterium Species Abundance by Obesity Status",
    x = "Bifidobacterium Species",
    y = "Absolute Abundance (counts)",
    fill = "Obesity Status"
  ) +
  
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.text.y = element_text(size = 10),
    legend.position = "bottom",
    legend.title = element_text(size = 11),
    legend.text = element_text(size = 10),
    panel.grid.minor = element_blank()
  ) +
  
  scale_fill_manual(values = c("Non-obese" = "lightblue", "Obese" = "salmon")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))  # Extra space for significance annotations

# Add significance annotations
if(nrow(species_test_results) > 0) {
  for(i in 1:nrow(species_test_results)) {
    species_name <- species_test_results$Species[i]
    sig_star <- species_test_results$sig_star[i]
    
    if(sig_star != "ns") {
      # Get max y value for this species
      species_max_y <- max(species_long$Abundance[species_long$Species == species_name], na.rm = TRUE)
      y_position <- species_max_y * 1.1
      
      # Get x position (species index)
      x_position <- which(levels(species_long$Species) == species_name)
      
      # Add significance annotation
      p_species_stacked <- p_species_stacked +
        annotate("segment", x = x_position - 0.2, xend = x_position + 0.2, 
                y = y_position, yend = y_position, color = "black", size = 0.5) +
        annotate("segment", x = x_position - 0.2, xend = x_position - 0.2, 
                y = y_position, yend = y_position * 0.98, color = "black", size = 0.5) +
        annotate("segment", x = x_position + 0.2, xend = x_position + 0.2, 
                y = y_position, yend = y_position * 0.98, color = "black", size = 0.5) +
        annotate("text", x = x_position, y = y_position * 1.03, 
                label = sig_star, size = 4, fontface = "bold")
    }
  }
}

print(p_species_stacked)

```



## literature species
```{r}
diabetes_enriched <- c(
  "Lactobacillus gasseri",
  "Streptococcus mutans",
  "Clostridium citroniae",
  "Clostridium bolteae",
  "Escherichia coli",
  "Bacteroides fragilis",
  "Streptococcus salivarius",  # Example Streptococcus species, you may need to adjust based on your data
  "Streptococcus parasanguinis"
  # Add more species from literature
)

diabetes_depleted <- c(
  "Roseburia intestinalis",
  "Faecalibacterium prausnitzii",
  "Butyrivibrio crossotus",
  "Coprococcus eutactus", 
  "Turicibacter sanguinis",
  "Bacteroides plebeius",
  "Ruminococcus lactaris"
  # Add more species from literature
)

obesity_enriched <- c(
  "Ruminococcus gnavus",
  "Lactobacillus reuteri",
  "Lactobacillus gasseri"
  # Add more species from literature
)

obesity_depleted <- c(
  "Akkermansia muciniphila",
  "Lactobacillus paracasei",
  "Methanobacteriales smithii",
  "Bifidobacterium animalis"
  # Add more species from literature
)
# Function to analyze trends based on literature species
analyze_literature_species_trends <- function(species_list, condition_data, condition_column, expected_direction) {
  
  # Filter for species that exist in your dataset
  existing_species <- intersect(species_list, colnames(condition_data))
  
  if (length(existing_species) == 0) {
    message("No matching species found in the dataset for this analysis.")
    return(NULL)
  }
  
  # Analyze each species
  results <- map_dfr(existing_species, function(species) {
    # Calculate mean values for each group
    means <- condition_data %>%
      group_by(get(condition_column)) %>%
      summarise(mean_val = mean(.data[[species]], na.rm = TRUE)) %>%
      pivot_wider(names_from = `get(condition_column)`, values_from = mean_val)
    
    # Calculate log2 fold change
    log2FC <- log2(means[[2]] / means[[1]])
    
    # Check if the trend matches the expected direction
    trend_matches <- if (expected_direction == "enriched") {
      log2FC > 0
    } else {
      log2FC < 0
    }
    
    data.frame(
      species = species,
      log2FoldChange = log2FC,
      matches_literature = trend_matches
    )
  })
  
  return(results)
}

# Analyze diabetes-associated species trends
diabetes_enriched_trends <- analyze_literature_species_trends(
  diabetes_enriched,
  merged_data_filter,
  "diabetes_group",
  "enriched"
)

diabetes_depleted_trends <- analyze_literature_species_trends(
  diabetes_depleted,
  merged_data_filter,
  "diabetes_group",
  "depleted"
)

# Analyze obesity-associated species trends
obesity_enriched_trends <- analyze_literature_species_trends(
  obesity_enriched,
  merged_data_filter,
  "obesity_binary",
  "enriched"
)

obesity_depleted_trends <- analyze_literature_species_trends(
  obesity_depleted,
  merged_data_filter,
  "obesity_binary",
  "depleted"
)

# Combine and summarize results
create_summary_table <- function(trends, condition) {
  if (is.null(trends) || nrow(trends) == 0) return(NULL)
  
  trends %>%
    mutate(condition = condition) %>%
    dplyr::select(species, log2FoldChange, matches_literature, condition)
}

diabetes_enriched_summary <- create_summary_table(diabetes_enriched_trends, "Diabetes Enriched")
diabetes_depleted_summary <- create_summary_table(diabetes_depleted_trends, "Diabetes Depleted")
obesity_enriched_summary <- create_summary_table(obesity_enriched_trends, "Obesity Enriched")
obesity_depleted_summary <- create_summary_table(obesity_depleted_trends, "Obesity Depleted")

# Combine summaries into a single table
all_summaries <- bind_rows(
  diabetes_enriched_summary,
  diabetes_depleted_summary,
  obesity_enriched_summary,
  obesity_depleted_summary
)

# View the summary table
library(knitr)
kable(all_summaries, caption = "Summary of Literature Trends in Microbiome Data")

```

## microbiome score of obesity
## dysbiosis score
```{r}
# Select species present in your dataset
obesity_enriched_present <- intersect(obesity_enriched, colnames(merged_data_filter))
obesity_depleted_present <- intersect(obesity_depleted, colnames(merged_data_filter))

# Compute row-wise sums of enriched and depleted species
enriched_sum <- rowSums(merged_data_filter[, obesity_enriched_present], na.rm = TRUE)
depleted_sum <- rowSums(merged_data_filter[, obesity_depleted_present], na.rm = TRUE)

# Calculate dysbiosis score with pseudocount to avoid division by 0
merged_data_filter$dysbiosis_score <- log((enriched_sum + 1e-6) / (depleted_sum + 1e-6))

```
## imflamatrory score
```{r}
# Calculate inflammatory score for each participant

# Define inflammatory weights for each food category
# Positive values = pro-inflammatory
# Negative values = anti-inflammatory
inflammatory_weights <- c(
  Fish = -0.5,             # Anti-inflammatory (omega-3)
  Crustacean = 0.1,        # Slightly pro-inflammatory
  Mollusk = 0.1,           # Slightly pro-inflammatory
  Poultry = 0.2,           # Slightly pro-inflammatory
  Red_meat = 0.8,          # Pro-inflammatory
  Processed_meat = 1.0,    # Strongly pro-inflammatory
  Dairy = 0.3,             # Moderately pro-inflammatory for some people
  Cruciferous_veg = -0.7,  # Strongly anti-inflammatory
  Legumes = -0.6,          # Anti-inflammatory
  Mushrooms = -0.4,        # Anti-inflammatory
  Fruits = -0.6,           # Anti-inflammatory
  Papaya = -0.8,           # Strongly anti-inflammatory
  Orange = -0.7,           # Anti-inflammatory (vitamin C)
  Wholegrain = -0.5,       # Anti-inflammatory
  Chapati = -0.3,          # Slightly anti-inflammatory (depends on grain)
  Idli_Thosai = -0.2,      # Slightly anti-inflammatory (fermented)
  Noodle_dishes = 0.4,     # Moderately pro-inflammatory (refined carbs)
  Deep_fried = 0.9,        # Strongly pro-inflammatory
  Coffee = -0.3,           # Slightly anti-inflammatory
  Tea = -0.4,              # Anti-inflammatory (polyphenols)
  Lontong = 0.1,           # Slightly pro-inflammatory
  Pork = 0.7,              # Pro-inflammatory
  Rice = 0.2               # Slightly pro-inflammatory (refined carbs)
)

# Calculate inflammatory score for each participant
food_data$inflammatory_score <- 0

for (food in names(inflammatory_weights)) {
  if (food %in% colnames(food_data)) {
    food_data$inflammatory_score <- food_data$inflammatory_score + 
      food_data[[food]] * inflammatory_weights[food]
  }
}


# Summary statistics
summary(food_data$inflammatory_score)
#merged_data_filter
```
```{r}
metadata
```

## combine 3 scores
```{r}
combined_data <- merge(food_data, 
                       merged_data_filter[, c("Participant_ID", "dysbiosis_score", "obesity_binary", "FREG5_Ethnic_Group", "FREG7_Gender","diabetes_status")],
                       by = "Participant_ID", 
                       all.x = FALSE,  # Keep only matching participants
                       all.y = FALSE)

# Check the structure of the combined data
str(combined_data)

# Basic summary statistics
summary(combined_data[, c("inflammatory_score", "dysbiosis_score", "obesity_binary")])

# Examine correlations between inflammatory score and dysbiosis score
correlation <- cor.test(combined_data$inflammatory_score, combined_data$dysbiosis_score)
print(correlation)

# Compare inflammatory scores between obese and non-obese individuals
# Convert obesity_binary to factor if it isn't already
combined_data$obesity_binary <- as.factor(combined_data$obesity_binary)

# T-test to compare inflammatory scores by obesity status
t_test_result <- t.test(inflammatory_score ~ obesity_binary, data = combined_data)
print(t_test_result)

# Visualize the relationships
# Load ggplot2 for better visualizations
library(ggplot2)

# 1. Scatterplot of inflammatory score vs dysbiosis score, colored by obesity status
ggplot(combined_data, aes(x = inflammatory_score, y = dysbiosis_score, color = FREG5_Ethnic_Group)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Relationship between Inflammatory Score and Dysbiosis Score",
       x = "Dietary Inflammatory Score",
       y = "Gut Microbiome Dysbiosis Score",
       color = "Obesity Status") +
  theme_minimal()

# 2. Boxplots comparing scores by obesity status
# For inflammatory score
p1 <- ggplot(combined_data, aes(x = obesity_binary, y = inflammatory_score, fill = obesity_binary)) +
  geom_boxplot() +
  labs(title = "Inflammatory Score by Obesity Status",
       x = "Obesity Status",
       y = "Dietary Inflammatory Score") +
  theme_minimal() +
  theme(legend.position = "none")

# For dysbiosis score
p2 <- ggplot(combined_data, aes(x = obesity_binary, y = dysbiosis_score, fill = obesity_binary)) +
  geom_boxplot() +
  labs(title = "Dysbiosis Score by Obesity Status",
       x = "Obesity Status",
       y = "Gut Microbiome Dysbiosis Score") +
  theme_minimal() +
  theme(legend.position = "none")

```
```{r}
# Load required libraries
library(ggplot2)
library(ggrepel)  # For non-overlapping text labels
library(viridis)  # For colorblind-friendly palettes

# Create an integrated visualization showing relationships between all three variables
# This creates a scatter plot with inflammatory score vs dysbiosis score
# with point size representing obesity status and color representing the combined risk

# First, let's calculate a combined score to use for coloring
# Standardize both scores to put them on the same scale
combined_data$inflammatory_z <- scale(combined_data$inflammatory_score)
combined_data$dysbiosis_z <- scale(combined_data$dysbiosis_score)

# Calculate a combined risk score (sum of z-scores)
combined_data$combined_risk <- combined_data$inflammatory_z + combined_data$dysbiosis_z

# Create a factor for obesity (for better labeling)
combined_data$obesity_status <- ifelse(combined_data$obesity_binary == 1, "Obese", "Non-obese")

# Create the integrated plot
integrated_plot <- ggplot(combined_data, 
                         aes(x = inflammatory_score, 
                             y = dysbiosis_score, 
                             size = obesity_status,
                             color = combined_risk)) +
  # Add points
  geom_point(alpha = 0.7) +
  
  # Add contour lines to show density (optional)
  stat_density_2d(linewidth = 0.5, color = "gray30", alpha = 0.5) +
  
  # Add correlation line
  geom_smooth(method = "lm", se = TRUE, color = "black", linewidth = 0.5) +
  
  # Use a color scale that transitions from blue (low risk) to red (high risk)
  scale_color_viridis(option = "plasma", name = "Combined Risk\n(Inflammatory + Dysbiosis)") +
  
  # Set point sizes for obesity status
  scale_size_manual(values = c("Non-obese" = 2, "Obese" = 4), name = "Obesity Status") +
  
  # Add quadrant lines at median values
  geom_hline(yintercept = median(combined_data$dysbiosis_score), 
             linetype = "dashed", color = "gray50", alpha = 0.7) +
  geom_vline(xintercept = median(combined_data$inflammatory_score), 
             linetype = "dashed", color = "gray50", alpha = 0.7) +
  
  # Add quadrant labels
  annotate("text", 
           x = c(min(combined_data$inflammatory_score) + 0.1 * diff(range(combined_data$inflammatory_score)),
                 max(combined_data$inflammatory_score) - 0.1 * diff(range(combined_data$inflammatory_score)),
                 min(combined_data$inflammatory_score) + 0.1 * diff(range(combined_data$inflammatory_score)),
                 max(combined_data$inflammatory_score) - 0.1 * diff(range(combined_data$inflammatory_score))),
           y = c(max(combined_data$dysbiosis_score) - 0.1 * diff(range(combined_data$dysbiosis_score)),
                 max(combined_data$dysbiosis_score) - 0.1 * diff(range(combined_data$dysbiosis_score)),
                 min(combined_data$dysbiosis_score) + 0.1 * diff(range(combined_data$dysbiosis_score)),
                 min(combined_data$dysbiosis_score) + 0.1 * diff(range(combined_data$dysbiosis_score))),
           label = c("Low Inflammatory\nHigh Dysbiosis", 
                     "High Inflammatory\nHigh Dysbiosis", 
                     "Low Inflammatory\nLow Dysbiosis", 
                     "High Inflammatory\nLow Dysbiosis"),
           size = 3, fontface = "bold", color = "gray20") +
  
  # Add labels and title
  labs(title = "Relationship between Dietary Inflammation, Gut Dysbiosis, and Obesity",
       subtitle = paste("Correlation: r =", round(cor(combined_data$inflammatory_score, 
                                                     combined_data$dysbiosis_score, 
                                                     use = "complete.obs"), 2)),
       x = "Dietary Inflammatory Score",
       y = "Gut Microbiome Dysbiosis Score",
       caption = "Dashed lines represent median values. Color indicates combined risk score.") +
  
  # Use a clean theme
  theme_minimal() +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(color = "gray40"),
        panel.grid.minor = element_blank())

# Add marginal plots to show distributions by obesity status (optional)
# This requires the 'ggExtra' package
# library(ggExtra)
# integrated_plot_with_marginals <- ggMarginal(integrated_plot, 
#                                            type = "density", 
#                                            groupFill = TRUE,
#                                            groupColour = TRUE)

# Display the plot
print(integrated_plot)

# Save the plot
ggsave("inflammatory_dysbiosis_obesity_relationship.png", integrated_plot, 
       width = 10, height = 8, dpi = 300)

# Optional: Generate summary statistics for the figure caption
obesity_summary <- aggregate(cbind(inflammatory_score, dysbiosis_score) ~ obesity_status, 
                            data = combined_data, 
                            FUN = function(x) c(mean = mean(x), sd = sd(x)))
print(obesity_summary)

# Calculate percentage of obese participants in each quadrant
combined_data$quadrant <- with(combined_data, 
                              paste0(ifelse(inflammatory_score > median(inflammatory_score), 
                                           "High", "Low"), 
                                    " Inflammatory / ",
                                    ifelse(dysbiosis_score > median(dysbiosis_score), 
                                           "High", "Low"), 
                                    " Dysbiosis"))
quadrant_summary <- table(combined_data$quadrant, combined_data$obesity_status)
quadrant_percentages <- prop.table(quadrant_summary, 1) * 100
print(quadrant_percentages)
```
```{r}
combined_data$obesity_status <- factor(ifelse(combined_data$obesity_binary == 1, "Obese", "Non-obese"))

# Create the integrated plot with clearer obesity differentiation
improved_plot <- ggplot(combined_data, 
                       aes(x = inflammatory_score, 
                           y = dysbiosis_score, 
                           shape = obesity_status,  # Use shape instead of just size
                           size = obesity_status,   # Keep size difference too
                           color = combined_risk)) +
  # Add points with higher alpha for better visibility
  geom_point(alpha = 0.8) +
  
  # Add contour lines to show density
  stat_density_2d(linewidth = 0.5, color = "gray30", alpha = 0.4) +
  
  # Add correlation line
  geom_smooth(method = "lm", se = TRUE, color = "black", linewidth = 0.5) +
  
  # Color scale
  scale_color_viridis(option = "plasma", name = "Combined Risk\n(Inflammatory + Dysbiosis)") +
  
  # Set point sizes AND shapes for obesity status (this is the key improvement)
  scale_size_manual(values = c("Non-obese" = 2, "Obese" = 4), name = "Obesity Status") +
  scale_shape_manual(values = c("Non-obese" = 16, "Obese" = 18), name = "Obesity Status") +
  
  # Quadrant lines and other elements as before
  geom_hline(yintercept = median(combined_data$dysbiosis_score), 
             linetype = "dashed", color = "gray50", alpha = 0.7) +
  geom_vline(xintercept = median(combined_data$inflammatory_score), 
             linetype = "dashed", color = "gray50", alpha = 0.7) +
  
  # Other elements from original plot...
  labs(title = "Relationship between Dietary Inflammation, Gut Dysbiosis, and Obesity",
       subtitle = paste("Correlation: r =", 
                        round(cor(combined_data$inflammatory_score, combined_data$dysbiosis_score, 
                                 use = "complete.obs"), 2)),
       x = "Dietary Inflammatory Score",
       y = "Gut Microbiome Dysbiosis Score",
       caption = "Dashed lines represent median values. Color indicates combined risk score.\nObese participants shown as larger diamond symbols.") +
  theme_minimal() +
  # Make legend more prominent
  theme(legend.position = "right",
        legend.key.size = unit(1, "cm"),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 11, face = "bold"),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(color = "gray40"))

# Display the improved plot
print(improved_plot)

# Save the improved plot
ggsave("improved_inflammatory_dysbiosis_obesity_relationship.png", improved_plot, 
       width = 10, height = 8, dpi = 300)

# Additional analysis to help interpretation
# Calculate mean values by obesity status
obesity_means <- aggregate(cbind(inflammatory_score, dysbiosis_score) ~ obesity_status, 
                          data = combined_data, 
                          FUN = mean)
print(obesity_means)

# Count participants in each quadrant by obesity status
combined_data$quadrant <- paste0(
  ifelse(combined_data$inflammatory_score > median(combined_data$inflammatory_score), 
         "High", "Low"), 
  " Inflammatory / ",
  ifelse(combined_data$dysbiosis_score > median(combined_data$dysbiosis_score), 
         "High", "Low"), 
  " Dysbiosis")

# Create a contingency table
quadrant_obesity_table <- table(combined_data$quadrant, combined_data$obesity_status)
print(quadrant_obesity_table)

# Calculate percentages
quadrant_percentages <- prop.table(quadrant_obesity_table, 1) * 100
print(quadrant_percentages)
```



```{r}
# Boxplot of dysbiosis score
library(ggplot2)
ggplot(merged_data_filter, aes(x = factor(obesity_binary), y = dysbiosis_score)) +
  geom_boxplot() +
  labs(x = "Obesity Status", y = "Dysbiosis Score", title = "Literature-Based Dysbiosis Score")

# Wilcoxon test
wilcox.test(dysbiosis_score ~ obesity_binary, data = merged_data_filter)

```
```{r}
ggplot(merged_data_filter, aes(x = factor(obesity_binary), y = dysbiosis_score)) +
  geom_violin(fill = "lightgray", alpha = 0.4) +
  geom_jitter(width = 0.1, alpha = 0.6, color = "steelblue") +
  geom_boxplot(width = 0.1, outlier.shape = NA) +
  labs(x = "Obesity Status", y = "Dysbiosis Score", title = "Dysbiosis Score by Obesity Status")

```


```{r}
library(effsize)
cliff.delta(dysbiosis_score ~ obesity_binary, data = merged_data_filter)
```

```{r}
library(pROC)
roc_obj <- roc(merged_data_filter$obesity_binary, merged_data_filter$dysbiosis_score)
plot(roc_obj, col = "darkred", print.auc = TRUE)

```
```{r}
library(pROC)

roc_obj <- roc(merged_data_filter$obesity_binary, merged_data_filter$dysbiosis_score)
plot(roc_obj, col = "firebrick", main = "ROC Curve for Dysbiosis Score")
auc(roc_obj)  # print AUC value

```

```{r}
merged_data_filter$obesity_binary_numeric <- ifelse(merged_data_filter$obesity_binary == "Obese", 1, 0)

# Now re-run logistic regression
model <- glm(obesity_binary_numeric ~ dysbiosis_score, data = merged_data_filter, family = "binomial")
summary(model)

```

```{r}
# Load needed packages
library(dplyr)
library(pROC)
library(ggplot2)

### 1. FILTER FOR MATCHING SPECIES
# These are your trend results from earlier
refined_enriched <- obesity_enriched_trends %>%
  filter(matches_literature == TRUE) %>%
  pull(species)

refined_depleted <- obesity_depleted_trends %>%
  filter(matches_literature == TRUE) %>%
  pull(species)

# Check which ones exist in your dataset
refined_enriched_present <- intersect(refined_enriched, colnames(merged_data_filter))
refined_depleted_present <- intersect(refined_depleted, colnames(merged_data_filter))

### 2. COMPUTE REFINED DYSBIOSIS SCORE
# Compute row-wise sums of enriched and depleted species
refined_enriched_sum <- rowSums(merged_data_filter[, refined_enriched_present], na.rm = TRUE)
refined_depleted_sum <- rowSums(merged_data_filter[, refined_depleted_present], na.rm = TRUE)

# Calculate log-ratio score with pseudocount
merged_data_filter$refined_dysbiosis_score <- log((refined_enriched_sum + 1e-6) / (refined_depleted_sum + 1e-6))

### 3. COMPARE BETWEEN GROUPS
# Optional boxplot
ggplot(merged_data_filter, aes(x = factor(obesity_binary), y = refined_dysbiosis_score)) +
  geom_boxplot() +
  geom_jitter(width = 0.1, alpha = 0.6) +
  labs(x = "Obesity Status", y = "Refined Dysbiosis Score", title = "Refined Score by Obesity")

# Wilcoxon test
wilcox.test(refined_dysbiosis_score ~ obesity_binary, data = merged_data_filter)

### 4. LOGISTIC REGRESSION
# Make sure outcome is numeric (0/1)
merged_data_filter$obesity_binary_numeric <- ifelse(merged_data_filter$obesity_binary == "Obese", 1, 0)

# Logistic model
model_refined <- glm(obesity_binary_numeric ~ refined_dysbiosis_score, data = merged_data_filter, family = "binomial")
summary(model_refined)

# Odds ratio
exp(coef(model_refined))

### 5. ROC & AUC
roc_refined <- roc(merged_data_filter$obesity_binary_numeric, merged_data_filter$refined_dysbiosis_score)
plot(roc_refined, col = "darkblue", main = "ROC - Refined Dysbiosis Score", print.auc = TRUE)
auc(roc_refined)

# Optional: get best cutoff and confusion matrix
best_cutoff <- coords(roc_refined, "best", "threshold")
merged_data_filter$refined_predicted_class <- ifelse(merged_data_filter$refined_dysbiosis_score > best_cutoff, 1, 0)

table(Predicted = merged_data_filter$refined_predicted_class,
      Actual = merged_data_filter$obesity_binary_numeric)

```
#machine learning score
```{r}
# Load libraries
library(randomForest)
library(pROC)
library(dplyr)
library(ggplot2)
library(tibble)

### 1. Prepare Data

# Ensure Participant_ID is row identifier
X <- species_filtered2 %>%
  column_to_rownames("Participant_ID")

# Match sample order and extract obesity label
y_df <- merged_data_filter %>%
  dplyr::select(Participant_ID, obesity_binary_numeric) %>%
  filter(Participant_ID %in% rownames(X))  # ensure only matching samples

# Align rows between X and y
X <- X[match(y_df$Participant_ID, rownames(X)), ]
y <- as.factor(y_df$obesity_binary_numeric)

### 2. Train Random Forest
set.seed(123)
rf_model <- randomForest(x = X, y = y, ntree = 500, importance = TRUE)

### 3. Predict Probabilities (Dysbiosis Score)
rf_scores <- predict(rf_model, X, type = "prob")[, "1"]  # probability of being obese

# Add scores back to merged_data_filter
merged_data_filter <- merged_data_filter %>%
  left_join(tibble(Participant_ID = y_df$Participant_ID, rf_dysbiosis_score = rf_scores),
            by = "Participant_ID")

### 4. Evaluate ROC & AUC
roc_rf <- roc(y, rf_scores)
plot(roc_rf, col = "tomato", print.auc = TRUE, main = "Random Forest Dysbiosis Score")
auc_rf <- auc(roc_rf)
print(paste("AUC:", round(auc_rf, 3)))

### 5. View Top Important Features
importance_df <- as.data.frame(importance(rf_model)) %>%
  rownames_to_column("Species") %>%
  arrange(desc(MeanDecreaseGini))

head(importance_df, 10)  # Top 10 important species

```
```{r}
set.seed(123)
sample_idx <- sample(1:nrow(X), size = 0.8 * nrow(X))
X_train <- X[sample_idx, ]
X_test <- X[-sample_idx, ]
y_train <- y[sample_idx]
y_test <- y[-sample_idx]

# Train
rf_model <- randomForest(x = X_train, y = y_train, ntree = 500, importance = TRUE)

# Predict on test set
rf_scores_test <- predict(rf_model, X_test, type = "prob")[, "1"]

# Evaluate
roc_test <- roc(y_test, rf_scores_test)
plot(roc_test, col = "steelblue", print.auc = TRUE, main = "Random Forest - Test Set")
auc(roc_test)

```
```{r}
library(caret)

# Create a combined dataframe
rf_input <- data.frame(y = y, X)

# Train with cross-validation
set.seed(123)
ctrl <- trainControl(method = "cv", number = 5, classProbs = TRUE, summaryFunction = twoClassSummary)

# y must be labeled with strings, not numeric
rf_input$y <- ifelse(rf_input$y == 1, "Obese", "NonObese")

rf_cv <- train(y ~ ., data = rf_input,
               method = "rf",
               metric = "ROC",
               trControl = ctrl)

print(rf_cv)

```






```{r}
# Function to create violin plots for species with log transformation
plot_violin_species_log <- function(species_list, condition_data, condition_column, title) {
  
  # Filter for species that exist in your dataset
  existing_species <- intersect(species_list, colnames(condition_data))
  
  if (length(existing_species) == 0) {
    message("No matching species found in the dataset for this plot.")
    return(NULL)
  }
  
  # Reshape the data for plotting
  plot_data <- condition_data %>%
    dplyr::select(all_of(existing_species), condition_column) %>%
    pivot_longer(cols = -all_of(condition_column), names_to = "species", values_to = "abundance") %>%
    mutate(abundance_log = log10(abundance + 1))  # Log transform (add 1 to avoid log(0))
  
  # Create violin plot
  ggplot(plot_data, aes(x = .data[[condition_column]], y = abundance_log, fill = .data[[condition_column]])) +
    geom_violin(alpha = 0.5) +
    geom_jitter(width = 0.2, size = 0.8, alpha = 0.6) +
    facet_wrap(~ species, scales = "free", ncol = 3) +  # One plot per species
    theme_minimal() +
    labs(
      title = title,
      x = condition_column,
      y = "Log10 Abundance",
      fill = condition_column
    ) +
    scale_fill_manual(values = c("Non-obese" = "blue", "Obese" = "orange", "Non-diabetic" = "blue", "Diabetic" = "orange"))
}

# Create separate plots for obesity and diabetes
# Obesity Enriched
obesity_enriched_plot <- plot_violin_species_log(
  obesity_enriched, 
  merged_data_filter, 
  "obesity_binary", 
  "Obesity Enriched Species (Log Transformed)"
)

# Obesity Depleted
obesity_depleted_plot <- plot_violin_species_log(
  obesity_depleted, 
  merged_data_filter, 
  "obesity_binary", 
  "Obesity Depleted Species (Log Transformed)"
)

# Diabetes Enriched
diabetes_enriched_plot <- plot_violin_species_log(
  diabetes_enriched, 
  merged_data_filter, 
  "diabetes_group", 
  "Diabetes Enriched Species (Log Transformed)"
)

# Diabetes Depleted
diabetes_depleted_plot <- plot_violin_species_log(
  diabetes_depleted, 
  merged_data_filter, 
  "diabetes_group", 
  "Diabetes Depleted Species (Log Transformed)"
)

# Save plots
ggsave("figure/obesity_enriched_species_log.png", obesity_enriched_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/obesity_depleted_species_log.png", obesity_depleted_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/diabetes_enriched_species_log.png", diabetes_enriched_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/diabetes_depleted_species_log.png", diabetes_depleted_plot, width = 12, height = 8, dpi = 300)

```
## separate ethnicity
```{r}
merged_data_filter
# Function to create violin plots with log transformation, stratification, and consistent color mapping
plot_violin_species_log_ethnic <- function(species_list, condition_data, condition_column, title) {
  
  # Filter for species that exist in your dataset
  existing_species <- intersect(species_list, colnames(condition_data))
  
  if (length(existing_species) == 0) {
    message("No matching species found in the dataset for this plot.")
    return(NULL)
  }
  
  # Standardize the values in the condition column to match the expected levels
  condition_data <- condition_data %>%
    filter(!is.na(.data[[condition_column]]) & FREG5_Ethnic_Group %in% c("C", "I", "M")) %>%  # Remove NA and exclude "O"
    mutate(!!condition_column := case_when(
      .data[[condition_column]] == "non_diabetes" ~ "Non-diabetic",
      .data[[condition_column]] == "diabetes" ~ "Diabetic",
      .data[[condition_column]] == "non_obese" ~ "Non-obese",
      .data[[condition_column]] == "obese" ~ "Obese",
      TRUE ~ as.character(.data[[condition_column]])
    )) %>%
    mutate(!!condition_column := factor(
      .data[[condition_column]], 
      levels = c("Non-diabetic", "Diabetic", "Non-obese", "Obese")
    ))
  
  # Reshape data for plotting
  plot_data <- condition_data %>%
    dplyr::select(all_of(existing_species), condition_column, FREG5_Ethnic_Group) %>%
    pivot_longer(cols = -c(all_of(condition_column), FREG5_Ethnic_Group), 
                 names_to = "species", values_to = "abundance") %>%
    mutate(abundance_log = log10(abundance + 1))  # Log transform (add 1 to avoid log(0))
  
  # Debugging: Check levels
  print(levels(plot_data[[condition_column]]))
  
  # Create violin plot
  ggplot(plot_data, aes(x = FREG5_Ethnic_Group, y = abundance_log, fill = .data[[condition_column]])) +
    geom_violin(alpha = 0.5) +
    geom_jitter(aes(color = .data[[condition_column]]), 
                width = 0.2, size = 0.8, alpha = 0.6) +  # Adjust jitter without dodge
    facet_wrap(~ species, scales = "free", ncol = 3) +  # One plot per species
    theme_minimal() +
    labs(
      title = title,
      x = "Ethnic Group",
      y = "Log10 Abundance",
      fill = "Condition",
      color = "Condition"
    ) +
    scale_fill_manual(values = c("Non-diabetic" = "#4E6691", "Diabetic" = "#B8474D",
                                 "Non-obese" = "#4E6691", "Obese" = "#B8474D")) +
    scale_color_manual(values = c("Non-diabetic" = "#4E6691", "Diabetic" = "#B8474D",
                                  "Non-obese" = "#4E6691", "Obese" = "#B8474D")) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.title = element_text(size = 12, face = "bold"),
      plot.title = element_text(size = 14, face = "bold")
    )
}

# Create separate plots for diabetes and obesity
# Diabetes Enriched
diabetes_enriched_plot <- plot_violin_species_log_ethnic(
  diabetes_enriched, 
  merged_data_filter, 
  "diabetes_group", 
  "Diabetes Enriched Species (Log Transformed, Stratified by Ethnic Group)"
)

# Diabetes Depleted
diabetes_depleted_plot <- plot_violin_species_log_ethnic(
  diabetes_depleted, 
  merged_data_filter, 
  "diabetes_group", 
  "Diabetes Depleted Species (Log Transformed, Stratified by Ethnic Group)"
)

# Obesity Enriched
obesity_enriched_plot <- plot_violin_species_log_ethnic(
  obesity_enriched, 
  merged_data_filter, 
  "obesity_binary", 
  "Obesity Enriched Species (Log Transformed, Stratified by Ethnic Group)"
)

# Obesity Depleted
obesity_depleted_plot <- plot_violin_species_log_ethnic(
  obesity_depleted, 
  merged_data_filter, 
  "obesity_binary", 
  "Obesity Depleted Species (Log Transformed, Stratified by Ethnic Group)"
)

# Save plots
ggsave("figure/diabetes_enriched_species_log_ethnic.png", diabetes_enriched_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/diabetes_depleted_species_log_ethnic.png", diabetes_depleted_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/obesity_enriched_species_log_ethnic.png", obesity_enriched_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/obesity_depleted_species_log_ethnic.png", obesity_depleted_plot, width = 12, height = 8, dpi = 300)

```
## separate ethnic box plot
```{r}
plot_box_species_log_ethnic <- function(species_list, condition_data, condition_column, title) {
  
  # Filter for species that exist in your dataset
  existing_species <- intersect(species_list, colnames(condition_data))
  
  if (length(existing_species) == 0) {
    message("No matching species found in the dataset for this plot.")
    return(NULL)
  }
  
  # Standardize the values in the condition column to match the expected levels
  condition_data <- condition_data %>%
    filter(!is.na(.data[[condition_column]]) & FREG5_Ethnic_Group %in% c("C", "I", "M")) %>%  # Remove NA and exclude "O"
    mutate(!!condition_column := case_when(
      .data[[condition_column]] == "non_diabetes" ~ "Non-diabetic",
      .data[[condition_column]] == "diabetes" ~ "Diabetic",
      .data[[condition_column]] == "non_obese" ~ "Non-obese",
      .data[[condition_column]] == "obese" ~ "Obese",
      TRUE ~ as.character(.data[[condition_column]])
    )) %>%
    mutate(!!condition_column := factor(
      .data[[condition_column]], 
      levels = c("Non-diabetic", "Diabetic", "Non-obese", "Obese")
    ))
  
  # Reshape data for plotting
  plot_data <- condition_data %>%
    dplyr::select(all_of(existing_species), condition_column, FREG5_Ethnic_Group) %>%
    pivot_longer(cols = -c(all_of(condition_column), FREG5_Ethnic_Group), 
                 names_to = "species", values_to = "abundance") %>%
    mutate(abundance_log = log10(abundance + 1))  # Log transform (add 1 to avoid log(0))
  
  # Create box plot with dots
  ggplot(plot_data, aes(x = FREG5_Ethnic_Group, y = abundance_log, fill = .data[[condition_column]])) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA) +  # Remove default outliers
    geom_jitter(aes(color = .data[[condition_column]]), 
                position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2),
                size = 0.8, alpha = 0.6) +
    facet_wrap(~ species, scales = "free", ncol = 3) +  # One plot per species
    theme_minimal() +
    labs(
      title = title,
      x = "Ethnic Group",
      y = "Log10 Abundance",
      fill = "Condition",
      color = "Condition"
    ) +
    scale_fill_manual(values = c("Non-diabetic" = "#4E6691", "Diabetic" = "#B8474D",
                                "Non-obese" = "#4E6691", "Obese" = "#B8474D")) +
    scale_color_manual(values = c("Non-diabetic" = "#4E6691", "Diabetic" = "#B8474D",
                                 "Non-obese" = "#4E6691", "Obese" = "#B8474D")) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.title = element_text(size = 12, face = "bold"),
      plot.title = element_text(size = 14, face = "bold")
    )
}

# Create separate plots for diabetes and obesity
# Diabetes Enriched
diabetes_enriched_plot <- plot_box_species_log_ethnic(
  diabetes_enriched, 
  merged_data_filter, 
  "diabetes_group", 
  "Diabetes Enriched Species (Log Transformed, Stratified by Ethnic Group)"
)

# Diabetes Depleted
diabetes_depleted_plot <- plot_box_species_log_ethnic(
  diabetes_depleted, 
  merged_data_filter, 
  "diabetes_group", 
  "Diabetes Depleted Species (Log Transformed, Stratified by Ethnic Group)"
)

# Obesity Enriched
obesity_enriched_plot <- plot_box_species_log_ethnic(
  obesity_enriched, 
  merged_data_filter, 
  "obesity_binary", 
  "Obesity Enriched Species (Log Transformed, Stratified by Ethnic Group)"
)

# Obesity Depleted
obesity_depleted_plot <- plot_box_species_log_ethnic(
  obesity_depleted, 
  merged_data_filter, 
  "obesity_binary", 
  "Obesity Depleted Species (Log Transformed, Stratified by Ethnic Group)"
)

# Save plots
ggsave("figure/diabetes_enriched_species_log_ethnic1.png", diabetes_enriched_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/diabetes_depleted_species_log_ethnic1.png", diabetes_depleted_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/obesity_enriched_species_log_ethnic1.png", obesity_enriched_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/obesity_depleted_species_log_ethnic1.png", obesity_depleted_plot, width = 12, height = 8, dpi = 300)
```

# with p value
```{r}
# Function to create box plots with pre-calculated p-values
plot_box_species_log_ethnic <- function(species_list, condition_data, condition_column, 
                                      stratified_results, title) {
  
  # Filter for species that exist in your dataset
  existing_species <- intersect(species_list, colnames(condition_data))
  
  if (length(existing_species) == 0) {
    message("No matching species found in the dataset for this plot.")
    return(NULL)
  }
  
  # Standardize the values in the condition column
  condition_data <- condition_data %>%
    filter(!is.na(.data[[condition_column]]) & FREG5_Ethnic_Group %in% c("C", "I", "M")) %>%
    mutate(condition = case_when(
      .data[[condition_column]] == "non_diabetes" ~ "Non-diabetic",
      .data[[condition_column]] == "diabetes" ~ "Diabetic",
      .data[[condition_column]] == "non_obese" ~ "Non-obese",
      .data[[condition_column]] == "obese" ~ "Obese",
      TRUE ~ as.character(.data[[condition_column]])
    )) %>%
    mutate(condition = factor(
      condition, 
      levels = c("Non-diabetic", "Diabetic", "Non-obese", "Obese")
    ))
  
  # Reshape data for plotting
  plot_data <- condition_data %>%
    dplyr::select(all_of(existing_species), condition, FREG5_Ethnic_Group) %>%
    pivot_longer(cols = -c(condition, FREG5_Ethnic_Group), 
                 names_to = "species", values_to = "abundance") %>%
    mutate(abundance_log = log10(abundance + 1))
  
  # Prepare p-values for plotting
  p_values <- stratified_results %>%
    group_by(species, FREG5_Ethnic_Group) %>%
    mutate(
      p_label = sprintf("p = %.3f", pvalue),
      # Calculate y position
      y_pos = max(plot_data$abundance_log[plot_data$species == species]) + 0.2
    )
  
  # Create box plot with dots and p-values
  ggplot(plot_data, aes(x = FREG5_Ethnic_Group, y = abundance_log, fill = condition)) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA) +
    geom_jitter(aes(color = condition), 
                position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2),
                size = 0.8, alpha = 0.6) +
    # Add p-value annotations
    geom_text(data = p_values,
              aes(x = FREG5_Ethnic_Group, y = y_pos, label = p_label),
              position = position_dodge(width = 0.75),
              size = 3) +
    facet_wrap(~ species, scales = "free", ncol = 3) +
    theme_minimal() +
    labs(
      title = title,
      x = "Ethnic Group",
      y = "Log10 Abundance",
      fill = "Condition",
      color = "Condition"
    ) +
    scale_fill_manual(values = c("Non-diabetic" = "#4E6691", "Diabetic" = "#B8474D",
                                "Non-obese" = "#4E6691", "Obese" = "#B8474D")) +
    scale_color_manual(values = c("Non-diabetic" = "#4E6691", "Diabetic" = "#B8474D",
                                 "Non-obese" = "#4E6691", "Obese" = "#B8474D")) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.title = element_text(size = 12, face = "bold"),
      plot.title = element_text(size = 14, face = "bold")
    )
}

# Create separate plots for diabetes and obesity using pre-calculated p-values
# Diabetes Enriched
diabetes_enriched_plot <- plot_box_species_log_ethnic(
  diabetes_enriched, 
  merged_data_filter, 
  "diabetes_group",
  diabetes_enriched_stratified,
  "Diabetes Enriched Species (Log Transformed, Stratified by Ethnic Group)"
)

# Diabetes Depleted
diabetes_depleted_plot <- plot_box_species_log_ethnic(
  diabetes_depleted, 
  merged_data_filter, 
  "diabetes_group",
  diabetes_depleted_stratified,
  "Diabetes Depleted Species (Log Transformed, Stratified by Ethnic Group)"
)

# Obesity Enriched
obesity_enriched_plot <- plot_box_species_log_ethnic(
  obesity_enriched, 
  merged_data_filter, 
  "obesity_binary",
  obesity_enriched_stratified,
  "Obesity Enriched Species (Log Transformed, Stratified by Ethnic Group)"
)

# Obesity Depleted
obesity_depleted_plot <- plot_box_species_log_ethnic(
  obesity_depleted, 
  merged_data_filter, 
  "obesity_binary",
  obesity_depleted_stratified,
  "Obesity Depleted Species (Log Transformed, Stratified by Ethnic Group)"
)

# Save plots
ggsave("figure/diabetes_enriched_species_log_ethnic_box.png", diabetes_enriched_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/diabetes_depleted_species_log_ethnic_box.png", diabetes_depleted_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/obesity_enriched_species_log_ethnic_box.png", obesity_enriched_plot, width = 12, height = 8, dpi = 300)
ggsave("figure/obesity_depleted_species_log_ethnic_box.png", obesity_depleted_plot, width = 12, height = 8, dpi = 300)
```


```{r}
library(ggrepel)
library(dplyr)
diabetes_enriched <- c(
  "Lactobacillus gasseri",
  "Streptococcus mutans",
  "Clostridium citroniae",
  "Clostridium bolteae",
  "Escherichia coli",
  "Bacteroides fragilis",
  "Streptococcus salivarius",  # Example Streptococcus species, you may need to adjust based on your data
  "Streptococcus parasanguinis"
  # Add more species from literature
)

diabetes_depleted <- c(
  "Roseburia intestinalis",
  "Faecalibacterium prausnitzii",
  "Butyrivibrio crossotus",
  "Coprococcus eutactus", 
  "Turicibacter sanguinis",
  "Bacteroides plebeius",
  "Ruminococcus lactaris"
  # Add more species from literature
)

obesity_enriched <- c(
  "Ruminococcus gnavus",
  "Lactobacillus reuteri",
  "Lactobacillus gasseri"
  # Add more species from literature
)

obesity_depleted <- c(
  "Akkermansia muciniphila",
  "Lactobacillus paracasei",
  "Methanobacteriales smithii",
  "Bifidobacterium animalis"
  # Add more species from literature
)

# Function to analyze literature-based species
analyze_literature_species <- function(species_list, condition_data, condition_column, 
                                     expected_direction = "enriched") {
  
  # Filter for species that exist in your dataset
  existing_species <- intersect(species_list, colnames(condition_data))
  
  if(length(existing_species) == 0) {
    return(NULL)
  }
  
  # Analyze each species
  results <- map_dfr(existing_species, function(species) {
    # Perform Wilcoxon test
    test_result <- wilcox.test(
      get(species) ~ get(condition_column), 
      data = condition_data
    )
    
    # Calculate fold change
    means <- condition_data %>%
      group_by(get(condition_column)) %>%
      summarise(mean_val = mean(get(species))) %>%
      pivot_wider(names_from = `get(condition_column)`, values_from = mean_val)
    
    # Assuming binary columns like diabetes/non-diabetes or obese/non-obese
    if(expected_direction == "enriched") {
      log2FC <- log2(means[[2]] / means[[1]])  # condition/control
    } else {
      log2FC <- -log2(means[[2]] / means[[1]])  # negative of condition/control
    }
    
    data.frame(
      species = species,
      pvalue = test_result$p.value,
      log2FoldChange = log2FC,
      matches_literature = sign(log2FC) == if(expected_direction == "enriched") 1 else -1
    )
  })
  
  results$padj <- p.adjust(results$pvalue, method = "BH")
  return(results)
}

# Analyze diabetes-associated species
diabetes_enriched_results <- analyze_literature_species(
  diabetes_enriched, 
  merged_data_filter,
  "diabetes_group",
  "enriched"
)

diabetes_depleted_results <- analyze_literature_species(
  diabetes_depleted, 
  merged_data_filter,
  "diabetes_group",
  "depleted"
)

# Analyze obesity-associated species
obesity_enriched_results <- analyze_literature_species(
  obesity_enriched, 
  merged_data_filter,
  "obesity_binary",
  "enriched"
)

obesity_depleted_results <- analyze_literature_species(
  obesity_depleted, 
  merged_data_filter,
  "obesity_binary",
  "depleted"
)

# Create summary visualization
library(gridExtra)

plot_literature_comparison <- function(results, title) {
  if(is.null(results) || nrow(results) == 0) return(NULL)
  
  ggplot(results, aes(x = log2FoldChange, y = -log10(padj))) +
    geom_point(aes(color = matches_literature)) +
    geom_text_repel(aes(label = species)) +
    scale_color_manual(values = c("red", "blue")) +
    labs(title = title) +
    theme_minimal()
}

# Create plots
p1 <- plot_literature_comparison(diabetes_enriched_results, "Diabetes Enriched Species")
p2 <- plot_literature_comparison(diabetes_depleted_results, "Diabetes Depleted Species")
p3 <- plot_literature_comparison(obesity_enriched_results, "Obesity Enriched Species")
p4 <- plot_literature_comparison(obesity_depleted_results, "Obesity Depleted Species")

# Arrange plots
g = grid.arrange(p1, p2, p3, p4, ncol = 2)
ggsave("figure/microbiota_diabete_obesity2.png", g, width = 12, height = 10, dpi = 300)
# Create summary table
create_summary_table <- function(results, condition) {
  if(is.null(results) || nrow(results) == 0) return(NULL)
  
  results %>%
    mutate(
      significant = padj < 0.1,
      agrees_with_literature = matches_literature & significant,
      condition = condition
    ) %>%
    dplyr::select(species, significant, agrees_with_literature, condition)
}

summary_table <- bind_rows(
  create_summary_table(diabetes_enriched_results, "Diabetes Enriched"),
  create_summary_table(diabetes_depleted_results, "Diabetes Depleted"),
  create_summary_table(obesity_enriched_results, "Obesity Enriched"),
  create_summary_table(obesity_depleted_results, "Obesity Depleted")
)

print(summary_table)
```


```{r}
analyze_literature_species_stratified <- function(species_list, condition_data, condition_column,
                                                ethnic_groups = c("C", "I", "M"),
                                                expected_direction = "enriched") {
  
  # Filter for species that exist in your dataset
  existing_species <- intersect(species_list, colnames(condition_data))
  
  if(length(existing_species) == 0) {
    return(NULL)
  }
  
  # Create a list to store results for each ethnic group
  results_list <- map(ethnic_groups, function(ethnic_group) {
    # Filter data for current ethnic group
    ethnic_data <- condition_data %>%
      filter(FREG5_Ethnic_Group == ethnic_group)
    
    # Skip if no data for this ethnic group
    if(nrow(ethnic_data) == 0) {
      return(NULL)
    }
    
    # Analyze each species within this ethnic group
    results <- map_dfr(existing_species, function(species) {
      # Check if there's enough data in each condition group
      condition_counts <- table(ethnic_data[[condition_column]])
      if(any(condition_counts < 2)) {
        return(data.frame(
          species = species,
          pvalue = NA,
          log2FoldChange = NA,
          matches_literature = NA
        ))
      }
      
      # Perform Wilcoxon test
      tryCatch({
        test_result <- wilcox.test(
          get(species) ~ get(condition_column), 
          data = ethnic_data
        )
        
        # Calculate fold change
        means <- ethnic_data %>%
          group_by(get(condition_column)) %>%
          summarise(mean_val = mean(get(species))) %>%
          pivot_wider(names_from = `get(condition_column)`, values_from = mean_val)
        
        if(expected_direction == "enriched") {
          log2FC <- log2(means[[2]] / means[[1]])  # condition/control
        } else {
          log2FC <- -log2(means[[2]] / means[[1]])  # negative of condition/control
        }
        
        data.frame(
          species = species,
          pvalue = test_result$p.value,
          log2FoldChange = log2FC,
          matches_literature = sign(log2FC) == if(expected_direction == "enriched") 1 else -1
        )
      }, error = function(e) {
        data.frame(
          species = species,
          pvalue = NA,
          log2FoldChange = NA,
          matches_literature = NA
        )
      })
    })
    
    # Add FDR correction within each ethnic group
    results$padj <- p.adjust(results$pvalue, method = "BH")
    results$FREG5_Ethnic_Group <- ethnic_group
    
    return(results)
  })
  
  # Combine results from all ethnic groups
  final_results <- bind_rows(results_list)
  return(final_results)
}

# Usage example for diabetes analysis
diabetes_enriched_stratified <- analyze_literature_species_stratified(
  diabetes_enriched, 
  merged_data_filter,
  "diabetes_group",
  c("C", "I", "M"),
  "enriched"
)

diabetes_depleted_stratified <- analyze_literature_species_stratified(
  diabetes_depleted, 
  merged_data_filter,
  "diabetes_group",
  c("C", "I", "M"),
  "depleted"
)

# Usage example for obesity analysis
obesity_enriched_stratified <- analyze_literature_species_stratified(
  obesity_enriched, 
  merged_data_filter,
  "obesity_binary",
  c("C", "I", "M"),
  "enriched"
)

obesity_depleted_stratified <- analyze_literature_species_stratified(
  obesity_depleted, 
  merged_data_filter,
  "obesity_binary",
  c("C", "I", "M"),
  "depleted"
)

# Optional: Create summary tables for each condition
# create_summary_table <- function(results_data) {
#   results_data %>%
#     filter(!is.na(pvalue)) %>%
#     group_by(FREG5_Ethnic_Group) %>%
#     summarise(
#       total_species = n(),
#       significant_species = sum(padj < 0.05),
#       matching_direction = sum(matches_literature & padj < 0.05)
#     )
# }
# 
# # Example usage of summary table
# diabetes_enriched_summary <- create_summary_table(diabetes_enriched_stratified)
# diabetes_depleted_summary <- create_summary_table(diabetes_depleted_stratified)
# obesity_enriched_summary <- create_summary_table(obesity_enriched_stratified)
# obesity_depleted_summary <- create_summary_table(obesity_depleted_stratified)
```


```{r}
diabetes_enriched_stratified
diabetes_depleted_stratified
obesity_enriched_stratified
obesity_depleted_stratified
```
##FB ratio
```{r}
firmicutes_species <- grep("^(Clostridium|Ruminococcus|Lactobacillus|Bacillus|Enterococcus|Streptococcus|Staphylococcus|Eubacterium|Roseburia|Faecalibacterium|Butyrivibrio)", 
                         colnames(merged_data_filter), value = TRUE)
bacteroidetes_species <- grep("^(Bacteroides|Prevotella|Porphyromonas|Alistipes|Parabacteroides)", 
                            colnames(merged_data_filter), value = TRUE)
# Calculate sum of relative abundances for each phylum
merged_data_filter <- merged_data_filter %>%
  mutate(
    Firmicutes_total = rowSums(select(., all_of(firmicutes_species))),
    Bacteroidetes_total = rowSums(select(., all_of(bacteroidetes_species))),
    FB_ratio = Firmicutes_total / Bacteroidetes_total
  )
```

## regression
```{r}

# firmicutes_species <- grep("^(Clostridium|Ruminococcus|Lactobacillus|Bacillus|Enterococcus|Streptococcus|Staphylococcus|Eubacterium|Roseburia|Faecalibacterium|Butyrivibrio)", 
#                          colnames(merged_data_filter), value = TRUE)
# 
# bacteroidetes_species <- grep("^(Bacteroides|Prevotella|Porphyromonas|Alistipes|Parabacteroides)", 
#                             colnames(merged_data_filter), value = TRUE)
```




```{r}
# First normalize the counts to relative abundances
# Get species columns (all except the first column from species_filtered2)
merged_data_fb <- species1 %>%
  inner_join(metadata, by = "Participant_ID")

species_cols <- colnames(species1)[-1]

# Create normalized abundances
merged_data_fb <- merged_data_fb %>%
  mutate(across(all_of(species_cols), 
                ~./rowSums(select(merged_data_fb, all_of(species_cols))), 
                .names = "{.col}_norm"))

# Now get the normalized species for each phylum
firmicutes_species_norm <- grep("^(Clostridium|Ruminococcus|Lactobacillus|Bacillus|Enterococcus|Streptococcus|Staphylococcus|Eubacterium|Roseburia|Faecalibacterium|Butyrivibrio).*_norm$", 
                         colnames(merged_data_fb), value = TRUE)

bacteroidetes_species_norm <- grep("^(Bacteroides|Prevotella|Porphyromonas|Alistipes|Parabacteroides).*_norm$", 
                            colnames(merged_data_fb), value = TRUE)

# Calculate FB ratio using normalized abundances
merged_data_fb <- merged_data_fb %>%
  mutate(
    Firmicutes_total = rowSums(select(., all_of(firmicutes_species_norm))),
    Bacteroidetes_total = rowSums(select(., all_of(bacteroidetes_species_norm))),
    FB_ratio = Firmicutes_total / Bacteroidetes_total
  )
```


```{r}
merged_data_fb <- merged_data_fb[merged_data_fb$FB_ratio <= 5, ]
# Create the plot with normalized data
ggplot(merged_data_fb, aes(x = BMI, y = FB_ratio)) +
  geom_point(color = "blue", alpha = 0.5) +
  geom_smooth(method = "lm", color = "red") +
  theme_minimal() +
  labs(
    title = "Relationship between BMI and FB Ratio (Using Normalized Abundances)",
    x = "BMI",
    y = "FB Ratio"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold")
  )

# Check the regression
model <- lm(FB_ratio ~ BMI, data = merged_data_fb)
summary(model)
```
```{r}
# Create a BMI category variable
merged_data_filter <- merged_data_filter %>%
  mutate(BMI_category = ifelse(BMI < 25, "Normal Weight (BMI < 25)", "Obese (BMI ≥ 25)"))

# Create the plot with two regression lines
ggplot(merged_data_filter, aes(x = BMI, y = FB_ratio, color = BMI_category)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE) +
  theme_minimal() +
  scale_color_manual(values = c("Normal Weight (BMI < 25)" = "blue", 
                               "Overweight/Obese (BMI ≥ 25)" = "red")) +
  labs(
    title = "FB Ratio vs BMI by Weight Category",
    x = "BMI",
    y = "FB Ratio",
    color = "BMI Category"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom"
  )

# Perform separate regression analyses
# For BMI < 25
model_normal <- lm(FB_ratio ~ BMI, 
                  data = merged_data_filter[merged_data_filter$BMI < 25,])
print("Regression results for BMI < 25:")
summary(model_normal)

# For BMI ≥ 25
model_overweight <- lm(FB_ratio ~ BMI, 
                      data = merged_data_filter[merged_data_filter$BMI >= 25,])
print("Regression results for BMI ≥ 25:")
summary(model_overweight)
```


##box plot
```{r}
# Log-transform FB_ratio to reduce the influence of extreme values
merged_data_filter <- merged_data_filter %>%
  mutate(FB_ratio_log = log10(FB_ratio + 0.001))  # Add small constant to handle zeros

# Re-run Wilcoxon test with log-transformed values
wilcox_test_result <- wilcox.test(FB_ratio_log ~ obesity_binary, data = merged_data_filter)
print(wilcox_test_result)

# Create a boxplot with log-transformed data
ggplot(merged_data_filter, aes(x = obesity_binary, y = FB_ratio_log)) +
  geom_boxplot(outlier.shape = NA) +  # Exclude outliers from boxplot display
  geom_jitter(width = 0.2, alpha = 0.3, color = "blue") +
  theme_minimal() +
  labs(
    title = "Log-Transformed Firmicutes/Bacteroidetes Ratio by Obesity Status",
    x = "Obesity Status",
    y = "Log10(F/B Ratio)"
  )

# Optional: Visualize the untransformed data but highlight the extreme value
ggplot(merged_data_filter, aes(x = obesity_binary, y = FB_ratio)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.3, color = "blue") +
  theme_minimal() +
  labs(
    title = "Firmicutes/Bacteroidetes Ratio by Obesity Status",
    x = "Obesity Status",
    y = "F/B Ratio"
  ) +
  scale_y_continuous(trans = "log10", breaks = c(1, 10, 100, 1000, 8000))  # Log scale for clarity

```






##stratify by ethnic
```{r}
# Filter out the "O" ethnic group
ethnic_group_data_filtered <- ethnic_group_data %>%
  filter(FREG5_Ethnic_Group != "O")

# Re-plot with properly aligned jitter points
ggplot(ethnic_group_data_filtered, aes(x = obesity_binary, y = FB_ratio, color = FREG5_Ethnic_Group)) +
  geom_boxplot(aes(fill = FREG5_Ethnic_Group), alpha = 0.3, outlier.shape = NA) +  # Boxplot without outlier symbols
  geom_jitter(aes(x = obesity_binary), width = 0.2, alpha = 0.5) +  # Ensure jitter aligns with box groups
  geom_point(data = subset(ethnic_group_data_filtered, FB_ratio > 1000), 
             aes(x = obesity_binary, y = FB_ratio), 
             color = "red", size = 3, shape = 8) +  # Highlight extreme value
  theme_minimal() +
  labs(
    title = "Firmicutes/Bacteroidetes Ratio by Obesity Status and Ethnic Group",
    x = "Obesity Status",
    y = "F/B Ratio (Untransformed)"
  ) +
  scale_y_continuous(trans = "log10", breaks = c(1, 10, 100, 1000, 8000)) +
  scale_fill_manual(values = c("C" = "blue", "I" = "green", "M" = "orange")) +
  scale_color_manual(values = c("C" = "blue", "I" = "green", "M" = "orange"))

```
```{r}
# Filter out the "O" ethnic group
ethnic_group_data_filtered <- ethnic_group_data %>%
  filter(FREG5_Ethnic_Group != "O")

# Re-plot: Group by ethnic group, compare obese vs non-obese
ggplot(ethnic_group_data_filtered, aes(x = obesity_binary, y = FB_ratio, color = obesity_binary)) +
  geom_boxplot(aes(fill = obesity_binary), alpha = 0.3, outlier.shape = NA) +  # Boxplot without outlier symbols
  geom_jitter(width = 0.2, alpha = 0.5) +  # Jitter points for visualization
  geom_point(data = subset(ethnic_group_data_filtered, FB_ratio > 1000), 
             aes(x = obesity_binary, y = FB_ratio), 
             color = "red", size = 3, shape = 8) +  # Highlight extreme value
  facet_wrap(~ FREG5_Ethnic_Group, scales = "free") +  # Facet by ethnic group
  theme_minimal() +
  labs(
    title = "Firmicutes/Bacteroidetes Ratio by Ethnic Group and Obesity Status",
    x = "Obesity Status",
    y = "F/B Ratio (Untransformed)"
  ) +
  scale_y_continuous(trans = "log10", breaks = c(1, 10, 100, 1000, 8000)) +
  scale_fill_manual(values = c("Non-obese" = "blue", "Obese" = "orange")) +
  scale_color_manual(values = c("Non-obese" = "blue", "Obese" = "orange"))

```

```{r}
# Calculate the distribution of obesity_binary by ethnic group
obesity_dist <- merged_data_filter %>%
  group_by(FREG5_Ethnic_Group, obesity_binary) %>%
  summarise(count = n(), .groups = 'drop') %>%
  # Calculate percentages within each ethnic group
  group_by(FREG5_Ethnic_Group) %>%
  mutate(percentage = count / sum(count) * 100)


obesity_dist <- obesity_dist %>%
  mutate(obesity_binary = factor(obesity_binary, levels = c("Obese", "Non-obese")))

# Create the bar plot with updated stacking order
ggplot(obesity_dist, aes(x = FREG5_Ethnic_Group, y = count, fill = obesity_binary)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_stack(vjust = 0.5), size = 4) +
  scale_fill_manual(values = c("Obese" = "#B8474D", "Non-obese" = "4E6691")) +
  theme_minimal() +
  labs(
    title = "Distribution of Obesity Status by Ethnic Group",
    x = "Ethnic Group",
    y = "Count",
    fill = "Obesity Status"
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 14, face = "bold")
  )


```


