---
title: "Untitled"
author: "Ruwen Zhou"
date: "7/19/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggalluvial)
library(readr)
library(plyr)
library(tidyverse)
library(dplyr)
library(readxl)
library(vegan)
library(phyloseq)
library(broom)
library(tidyr)
```

```{r}
meta_data <- read.csv("data/demo_535.csv")
food_raw <- read.csv("data/food_535.csv")
# species_abundance <- read.csv("data/butyrate_pick.csv")
# meta_data = meta_data[,c("Participant_ID","FREG5_Ethnic_Group")]
food_raw
meta_data

data_raw <- merge(meta_data[,c("Participant_ID", "FREG5_Ethnic_Group")], food_raw[-c(2:4)], by = "Participant_ID")  
data_raw <- subset(data_raw, FREG5_Ethnic_Group != "O")

rownames(data_raw) =data_raw[,1]
data_raw[1] = NULL
data_raw
```
#test Kruskal-Wallis
```{r}
data_raw <- data_raw %>% 
  mutate(FREG5_Ethnic_Group = as.factor(FREG5_Ethnic_Group)) %>% 
  mutate(across(-FREG5_Ethnic_Group, as.numeric))
data_raw


results <- data_raw %>%
  pivot_longer(-FREG5_Ethnic_Group, names_to = "food", values_to = "intake") %>%
  group_by(food) %>%
  do(tidy(kruskal.test(intake ~ FREG5_Ethnic_Group, data = .))) %>%
  ungroup() %>%
  mutate(p.adjusted = p.adjust(p.value, method = "BH")) %>%
  arrange(p.adjusted)

results
significant_results <- results %>%
  filter(p.value < 0.05)

print(significant_results)


```
not much food intake is related to ethnicity

#pca
```{r}
library(FactoMineR)
library(factoextra)

# Perform PCA
pca_result <- PCA(data_raw %>% select(-FREG5_Ethnic_Group), scale.unit = TRUE)

# Visualize PCA
fviz_pca_ind(pca_result, 
             geom.ind = "point", 
             col.ind = data_raw$FREG5_Ethnic_Group, 
             palette = "jco",
             addEllipses = TRUE,
             legend.title = "Ethnic Group") +
  labs(title = "PCA of Food Intake by Ethnic Group")

```


# random forest
```{r fig.height=10, fig.width=10}
library(randomForest)

data_imputed <- data_raw %>%
  mutate(across(-FREG5_Ethnic_Group, ~ ifelse(is.na(.), 0, .)))

rf_model <- randomForest(FREG5_Ethnic_Group ~ ., data = data_imputed, importance = TRUE)

# Plot variable importance
varImpPlot(rf_model, main = "Variable Importance in Random Forest")

```
MeanDecreaseAccuracy:

This column shows the overall importance of each variable for predicting the target variable (ethnic group) across all classes. It represents the mean decrease in accuracy of the model when the variable is permuted. Higher values indicate greater importance.

MeanDecreaseGini:

This column shows the importance of each variable based on the Gini impurity criterion. It represents the total decrease in node impurity that the variable contributes to, averaged over all the trees in the forest. Higher values indicate greater importance.





```{r}
library(dplyr)
library(ggplot2)

# Extract variable importance
importance <- importance(rf_model)
var_importance <- data.frame(Variable = rownames(importance), 
                             MeanDecreaseAccuracy = importance[, "MeanDecreaseAccuracy"], 
                             MeanDecreaseGini = importance[, "MeanDecreaseGini"])

# Set negative MeanDecreaseAccuracy values to zero
# var_importance <- var_importance %>%
#   mutate(MeanDecreaseAccuracy = ifelse(MeanDecreaseAccuracy < 0, 0, MeanDecreaseAccuracy))

# Select top 20 important variables based on MeanDecreaseAccuracy for better readability
top_vars <- var_importance %>% arrange(desc(MeanDecreaseAccuracy)) %>% head(20)

# Plot variable importance for top variables
ggplot(top_vars, aes(x = reorder(Variable, MeanDecreaseAccuracy), y = MeanDecreaseAccuracy)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_bw() +
  labs(title = "Top 20 Variable Importance in Random Forest",
       x = "Variable",
       y = "Mean Decrease in Accuracy")


```

```{r}
# Select top 10 important variables
top_vars <- head(var_importance[order(-var_importance$Importance), ], 12)
top_vars
# Create boxplots for top variables
data_long <- data_imputed %>%
  pivot_longer(-FREG5_Ethnic_Group, names_to = "food", values_to = "intake") %>%
  filter(food %in% top_vars$Variable)

ggplot(data_long, aes(x = FREG5_Ethnic_Group, y = intake, fill = FREG5_Ethnic_Group)) +
  geom_boxplot() +
  facet_wrap(~ food, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Top 10 Important Food Intake Variables by Ethnic Group",
       x = "Ethnic Group",
       y = "Food Intake")

```



