---
title: "genus_zscore"
author: "Ruwen Zhou"
date: "5/31/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(plyr)
library(tidyverse)
library(dplyr)
library(readxl)
library(ggplot2)
library(phyloseq)
library(vegan)
library(DESeq2)
library(pheatmap)
library(RColorBrewer)
```


#demographic
```{r}
helios_core = read_excel("data/HELIOS_Core_redacted.xlsx")
metadata_micro = read_csv("data/metadata_fil_01032024_redacted.csv")
sample_list = metadata_micro$`Original Sample name`
helios_new <- filter(helios_core, FREG1_Barcode %in% sample_list)
```
#food nutrition
```{r}
FFQ_food = read_xlsx("data/HELIOS_FFQA.xlsx")
FFQ_nutrition = read_xlsx("data/HELIOS_FFQB.xlsx")
pid_list = helios_new$FREG0_PID
food_new = filter(FFQ_food, FREG0_PID %in% pid_list)
nutrition_new = filter(FFQ_nutrition, FREG0_PID %in% pid_list)
```
#species
```{r}
species = read_csv("data/abundance_bracken_standard_16_20221209_taxa_G_unfiltered_reads_decont2.csv")
#species$`SAM id`
helios_new_list = helios_new$FREG1_Barcode
#sample_list = metadata_micro$`Original Sample name`
micro_list <- filter(metadata_micro, metadata_micro$`Original Sample name` %in% helios_new_list)
species_new = filter(species, species$`SAM id` %in% micro_list$`Library ID`)
species_new
```

```{r}
new_merge = merge(metadata_micro, species_new, by.x = "Library ID", by.y = "SAM id")
#new_merge
species_meta = merge(x = helios_new[,c("FREG1_Barcode", "DLAB27_Hba1C_Percent", "DLAB25_Gluf_Mmol_L", "FREG5_Ethnic_Group","FREG10_Fasting_Hours","FBP16_Caffeine")], y = new_merge, by.y = "Original Sample name", by.x = "FREG1_Barcode")
species_meta = species_meta %>% drop_na(DLAB27_Hba1C_Percent)
species_meta
```

```{r}
metadata = species_meta[, c("FREG1_Barcode", "DLAB27_Hba1C_Percent","DLAB25_Gluf_Mmol_L", "FREG5_Ethnic_Group","FREG10_Fasting_Hours", "FBP16_Caffeine")]
metadata
drop = c("DLAB27_Hba1C_Percent", "Library ID", "SampleType", "seq_batch", "DLAB25_Gluf_Mmol_L", "FREG5_Ethnic_Group","FREG10_Fasting_Hours", "FBP16_Caffeine")
species_data = species_meta[,!(names(species_meta) %in% drop)]
species_data
```

#diabetes
```{r}
metadata$diabetes_status <- ifelse(metadata$DLAB27_Hba1C_Percent > 6.4, "diabetes",
                                   ifelse(metadata$DLAB27_Hba1C_Percent >= 5.7 & metadata$DLAB27_Hba1C_Percent <= 6.4, "pre-diabetes",
                                          ifelse(metadata$DLAB27_Hba1C_Percent < 5.7, "normal", NA)))

#metadata$diabetes_status <- ifelse(metadata$DLAB27_Hba1C_Percent > 6.4, "diabetes",
                                   #ifelse(metadata$DLAB27_Hba1C_Percent <= 6.4, "non-diabetes", NA))



metadata$glucose <- ifelse(metadata$DLAB25_Gluf_Mmol_L > 7, "diabetes",
                                   ifelse(metadata$DLAB25_Gluf_Mmol_L >= 5.6 & metadata$DLAB25_Gluf_Mmol_L <= 7, "pre-diabetes",
                                          ifelse(metadata$DLAB25_Gluf_Mmol_L >= 3.9 & metadata$DLAB25_Gluf_Mmol_L <= 5.6, "normal",
                                                 ifelse(metadata$DLAB25_Gluf_Mmol_L < 3.9, "hypoglycemia", NA))))

                                         

metadata = metadata[!duplicated(metadata$FREG1_Barcode), ]
metadata
species_data = species_data[!duplicated(species_data$FREG1_Barcode), ]
species_data

```

```{r}
rownames(species_data) = species_data[,1]
species_data[1] = NULL
rownames(metadata) = metadata[,1]
species_data
metadata
```


```{r}
OTU = otu_table(as.matrix(species_data), taxa_are_rows = FALSE)              
SAM = sample_data(metadata, errorIfNULL = TRUE)

phylo <- phyloseq(OTU, SAM)
phylo
phylo = filter_taxa(phylo, function(x) sum(x > 2) > (0.11 * length(x)), TRUE)
```


## Drop species that are present in less than 5% of the samples
## Drop species with a total abundance less than 0.01% of the total reads
```{r}
prevalence_threshold <- 0.05
otu_table_filtered <- filter_taxa(phylo, 
                                  function(x) sum(x > 0) > (prevalence_threshold * nsamples(otu_table(phylo))),
                                  prune=TRUE)
otu_table_filtered

abundance_threshold <- 0.0001
phyloseq_filtered <- filter_taxa(otu_table_filtered, 
                                  function(x) sum(x) > (abundance_threshold * sum(otu_table(phylo))),
                                  prune=TRUE)

phyloseq_filtered

```

## alpha analysis
```{r}
#sample_data(phyloseq_filtered)$diabetes_status <- factor(sample_data(phyloseq_filtered)$diabetes_status,
                                                         #levels = c("non-diabetes", "diabetes"))

plot_richness(phyloseq_filtered, x = "diabetes_status", color = "diabetes_status", measures = c("Shannon", "Simpson", "Chao1", "Fisher")) + geom_boxplot() + theme_bw()
```
## Beta analysis
```{r}
dist = phyloseq::distance(phyloseq_filtered, method="bray")
phylo.beta = ordinate(phyloseq_filtered, method="PCoA", distance=dist)

plot_ordination(phyloseq_filtered, phylo.beta, type = "samples", color = "diabetes_status") + theme_bw() + stat_ellipse(aes(group=diabetes_status))
```
```{r}
# Convert glucose levels to a factor
sample_data(phyloseq_filtered)$diabetes_status <- factor(sample_data(phyloseq_filtered)$diabetes_status)

phyloseq_filtered <- subset_samples(phyloseq_filtered, !is.na(sample_data(phyloseq_filtered)$diabetes_status))

# Calculate the distance matrix using the ape package
dist <- phyloseq::distance(phyloseq_filtered, method = "bray")
dist_matrix <- as.matrix(dist)

# Perform PERMANOVA using adonis2
permanova <- adonis2(dist_matrix ~ sample_data(phyloseq_filtered)$diabetes_status)

# Print the PERMANOVA results
print(permanova)
```

## DA


```{r}
deseq_data <- phyloseq_to_deseq2(phyloseq_filtered, ~ diabetes_status)
deseq_data <- DESeq(deseq_data)
sample_data(phyloseq_filtered)
res <- results(deseq_data, contrast = c("diabetes_status", "diabetes", "normal"))
res <- res[order(res$padj, na.last=NA), ]
alpha <- 0.05
res <- res[which(res$padj < alpha), ]
res
```

```{r}
sig_taxa <- rownames(res)
sig_taxa <- sig_taxa[! sig_taxa %in% c('unclassified')]
phyloseq_filtered_subset = subset_samples(phyloseq_filtered, diabetes_status %in% c("normal", "diabetes"))

sig_otu_table <- otu_table(phyloseq_filtered_subset)[,sig_taxa]

sig_otu_table_df <- as.data.frame(sig_otu_table)

# Transpose the data frame to have taxa as rows and samples as columns


sig_otu_table_log <- log10(sig_otu_table_df + 1)

sig_otu_table_log_t <- t(sig_otu_table_log)
sig_otu_table_log_matrix <- as.matrix(sig_otu_table_log_t)
sig_otu_table_log_matrix
#pheatmap(sig_otu_table_log_matrix)

grouping_df = sample_data(phyloseq_filtered_subset)
grouping_df$diabetes_status <- factor(grouping_df$diabetes_status)
annotation_data <- as.data.frame(grouping_df$diabetes_status)
rownames(annotation_data) <- grouping_df$FREG1_Barcode
colnames(annotation_data) <- c('Group')

group_colors <- brewer.pal(n = length(levels(grouping_df$diabetes_status)), "Set1")
names(group_colors) <- levels(grouping_df$diabetes_status)

```

```{r fig.width=16, fig.height=12}
grouping_df

grouping_df <- grouping_df[order(grouping_df$diabetes_status),]
#sig_otu_table_log_matrix
new = as.character(grouping_df$FREG1_Barcode)
# Create an ordering vector that matches the order of the sample IDs in the grouped dataframe
order_matrix <- sig_otu_table_log_matrix[,new]

#order_matrix
# Reorder the matrix columns according to the order_vector
#sig_otu_table_log_matrix_ordered <- sig_otu_table_log_matrix[, order_vector]
#sig_otu_table_log_matrix_ordered
# Now you can create the heatmap with the ordered matrix

pheatmap(order_matrix, 
         annotation_col = annotation_data,
         cluster_cols = FALSE,
         cluster_rows = T,
         annotation_colors = list(Group = group_colors))

pheatmap(order_matrix, 
         annotation_col = annotation_data,
         cluster_cols = T,
         cluster_rows = T,
         annotation_colors = list(Group = group_colors))

```


```{r}

# Assuming phyloseq_filtered is already defined and contains the necessary data
phyloseq_filtered_subset = subset_samples(phyloseq_filtered, diabetes_status %in% c("normal", "diabetes"))

sig_otu_table <- otu_table(phyloseq_filtered_subset)[, sig_taxa]
sig_otu_table_df <- as.data.frame(sig_otu_table)
sample_data <- data.frame(sample_data(phyloseq_filtered_subset))
sample_data$SampleID <- rownames(sample_data)
combined_df = merge(x = sig_otu_table_df, y = sample_data, by.x = "row.names", by.y = "SampleID")
rownames(combined_df) <- combined_df$Row.names
combined_df$Row.names <- NULL
combined_df


metadata_columns <- c("FREG1_Barcode", "DLAB27_Hba1C_Percent", "DLAB25_Gluf_Mmol_L", "FREG10_Fasting_Hours", "FREG5_Ethnic_Group", "FBP16_Caffeine", "glucose")

# Ensure you're working with a data frame that excludes these metadata columns for calculations
abundance_data <- combined_df[, !(names(combined_df) %in% metadata_columns)]

# Filter to include only normal samples and calculate the mean abundances
normal_data <- abundance_data[abundance_data$diabetes_status == "normal", sapply(abundance_data, is.numeric)]

# Calculate the mean abundances for these filtered columns
normal_means <- rowMeans(normal_data, na.rm = TRUE)
numeric_columns <- sapply(abundance_data, is.numeric)

# Calculate log2 fold changes, ensuring to use only numeric columns
log2fc <- sweep(abundance_data[, numeric_columns], 1, normal_means, FUN = function(x, y) log2((x + 1)/(y + 1)))
log2fc_t = t(log2fc)
# Prepare matrix for heatmap (ensure this only contains the log2fc data)
log2fc_matrix <- as.matrix(log2fc_t)
log2fc_matrix

```

```{r}
grouping_df = sample_data(phyloseq_filtered_subset)
grouping_df$diabetes_status <- factor(grouping_df$diabetes_status)
annotation_data <- as.data.frame(grouping_df$diabetes_status)
rownames(annotation_data) <- grouping_df$FREG1_Barcode
colnames(annotation_data) <- c('Group')

group_colors <- brewer.pal(n = length(levels(grouping_df$diabetes_status)), "Set1")
names(group_colors) <- levels(grouping_df$diabetes_status)


grouping_df <- grouping_df[order(grouping_df$diabetes_status),]
#sig_otu_table_log_matrix
new = as.character(grouping_df$FREG1_Barcode)
# Create an ordering vector that matches the order of the sample IDs in the grouped dataframe
ordered_log2fc_matrix <- log2fc_matrix[,new]
ordered_log2fc_matrix
```


```{r fig.width=16, fig.height=12}
# Generate the heatmap with annotations
extended_palette <- colorRampPalette(brewer.pal(11, "RdBu"))(100)
pheatmap(ordered_log2fc_matrix, 
         cluster_rows = TRUE, 
         cluster_cols = FALSE, # Set to FALSE to maintain the custom order
         scale = "none",
         annotation_col = annotation_data,
         color = rev(extended_palette), 
         breaks = seq(-6, 6, length.out = 101),
         show_rownames = TRUE,
         show_colnames = FALSE)

pheatmap(ordered_log2fc_matrix, 
         cluster_rows = TRUE, 
         cluster_cols = TRUE, # Set to FALSE to maintain the custom order
         scale = "none",
         annotation_col = annotation_data,
         color = rev(extended_palette), 
         breaks = seq(-6, 6, length.out = 101),
         show_rownames = TRUE,
         show_colnames = FALSE)

```

##z score
mean of all samples 
```{r}
sig_taxa <- sig_taxa[! sig_taxa %in% c('unclassified', "Wujia", "Muribaculum")]
otu_table <- otu_table(phyloseq_filtered_subset)[, sig_taxa]

sample_ids <- rownames(otu_table)
otu_table
# Apply Z-score normalization across columns

z_score_matrix <- apply(otu_table, 2, scale)

# Reapply the row names to the scaled matrix
rownames(z_score_matrix) <- sample_ids
#z_score_matrix

grouping_df = sample_data(phyloseq_filtered_subset)
grouping_df$diabetes_status <- factor(grouping_df$diabetes_status)
annotation_data <- as.data.frame(grouping_df$diabetes_status)
rownames(annotation_data) <- grouping_df$FREG1_Barcode
colnames(annotation_data) <- c('Group')

group_colors <- brewer.pal(n = length(levels(grouping_df$diabetes_status)), "Set1")
names(group_colors) <- levels(grouping_df$diabetes_status)


grouping_df <- grouping_df[order(grouping_df$diabetes_status),]

new = as.character(grouping_df$FREG1_Barcode)
# Create an ordering vector that matches the order of the sample IDs in the grouped dataframe
z_score_matrix_t = t(z_score_matrix)

z_score_matrix_new <- z_score_matrix_t[,new]


```

```{r fig.width=8, fig.height=5}
extended_palette <- colorRampPalette(brewer.pal(11, "RdBu"))(100)
pheatmap(z_score_matrix_new, 
         cluster_rows = TRUE, 
         cluster_cols = FALSE, # Set to FALSE to maintain the custom order
         scale = "none",
         annotation_col = annotation_data,
         color = rev(extended_palette), 
         breaks = seq(-1, 1, length.out = 101),
         show_rownames = TRUE,
         show_colnames = FALSE,
         fontsize_row = 14)

pheatmap(z_score_matrix_new, 
         cluster_rows = TRUE, 
         cluster_cols = TRUE, # Set to FALSE to maintain the custom order
         scale = "none",
         annotation_col = annotation_data,
         color = rev(extended_palette), 
         breaks = seq(-1, 1, length.out = 101),
         show_rownames = TRUE,
         show_colnames = FALSE,
         fontsize_row = 14)
```


## ethinity

### turn into relative count
```{r}

phyloseq_relative <- transform_sample_counts(phyloseq_filtered, function(x) x / sum(x))

# Print the resulting phyloseq object
phyloseq_relative@otu_table

```

```{r}
phyloseq_filtered_race_raw = subset_samples(phyloseq_filtered, FREG5_Ethnic_Group %in% c("C", "I", "M"))

deseq_data1 <- phyloseq_to_deseq2(phyloseq_filtered_race_raw, ~ FREG5_Ethnic_Group)
deseq_data1 <- DESeq(deseq_data1)
res1 <- results(deseq_data1, contrast = c("FREG5_Ethnic_Group", "C", "I"))
res1 <- res1[order(res1$padj, na.last=NA), ]
alpha <- 0.05
res1 <- res1[which(res1$padj < alpha), ]
rownames(res1)
```

```{r}
sig_taxa1 <- rownames(res1)
sig_taxa1 <- sig_taxa1[! sig_taxa1 %in% c('unclassified')]
```

##z score
```{r}
phyloseq_filtered_race = subset_samples(phyloseq_relative, FREG5_Ethnic_Group %in% c("C", "I", "M"))
otu_table1 <- otu_table(phyloseq_filtered_race)[, sig_taxa1]

sample_ids <- rownames(otu_table1)

# Apply Z-score normalization across columns

z_score_matrix1 <- apply(otu_table1, 2, scale)

# Reapply the row names to the scaled matrix
rownames(z_score_matrix1) <- sample_ids
#z_score_matrix

grouping_df = sample_data(phyloseq_filtered_race)
grouping_df$FREG5_Ethnic_Group <- factor(grouping_df$FREG5_Ethnic_Group)
annotation_data <- as.data.frame(grouping_df$FREG5_Ethnic_Group)
rownames(annotation_data) <- grouping_df$FREG1_Barcode
colnames(annotation_data) <- c('Group')

group_colors <- brewer.pal(n = length(levels(grouping_df$FREG5_Ethnic_Group)), "Set1")
names(group_colors) <- levels(grouping_df$FREG5_Ethnic_Group)


grouping_df <- grouping_df[order(grouping_df$FREG5_Ethnic_Group),]

new = as.character(grouping_df$FREG1_Barcode)
# Create an ordering vector that matches the order of the sample IDs in the grouped dataframe
z_score_matrix1_t = t(z_score_matrix1)

z_score_matrix_new1 <- z_score_matrix1_t[,new]
#z_score_matrix_new1


```

```{r}
# Ensure the rownames of annotation_data match the order of columns in z_score_matrix_new1
annotation_data_ordered <- annotation_data[match(colnames(z_score_matrix_new1), rownames(annotation_data)), , drop = FALSE]

ethnicity_colors <- c("C" = "#9FDDCA", "I" = "#9EA3F1", "M" = "#E17887")

gene_names_italics <- lapply(rownames(z_score_matrix_new1), function(x) bquote(italic(.(x))))
col_annotation <- list(Group = ethnicity_colors)

pheatmap(z_score_matrix_new1, 
         cluster_rows = TRUE, 
         cluster_cols = FALSE, # Set to FALSE to maintain the custom order
         scale = "none",
         annotation_col = annotation_data_ordered,
         color = rev(extended_palette), 
         breaks = seq(-1, 1, length.out = 101),
         labels_row = as.expression(gene_names_italics),
         show_rownames = TRUE,
         show_colnames = FALSE,
         fontsize_row = 10,
         annotation_colors = col_annotation)


```
# cluster within each groups
```{r fig.width=15, fig.height=10}
library(dplyr)
library(pheatmap)
#extended_palette <- colorRampPalette(brewer.pal(11, "RdBu"))(100)
extended_palette <- coolwarm <- colorRampPalette(c("#3C4EC2", "#FFFFFF", "#CE4C3A"))(100)



# Convert numeric IDs to character if they aren't already
grouping_df$FREG1_Barcode <- as.character(grouping_df$FREG1_Barcode)

# Split the matrix according to the groups, cluster within each group, and then combine
groups <- unique(grouping_df$FREG5_Ethnic_Group)  # get unique groups
list_of_matrices <- list()

for(group in groups) {
  # Get column names for this group by matching the IDs in the dataframe to column names
  cols_for_group <- grouping_df$FREG1_Barcode[grouping_df$FREG5_Ethnic_Group == group]
  sub_matrix <- z_score_matrix_new1[, cols_for_group]
  # Perform clustering on columns of sub_matrix
  dist_mat <- dist(t(sub_matrix)) 
  hclust_mat <- hclust(dist_mat)

  sub_matrix <- sub_matrix[, hclust_mat$order]
  list_of_matrices[[group]] <- sub_matrix
}


# Combine matrices in the order of the groups
final_matrix <- do.call(cbind, list_of_matrices)

# Ensure annotation data is ordered correctly
final_annotation_data <- annotation_data[match(colnames(final_matrix), rownames(annotation_data)), , drop = FALSE]

# Create heatmap
pheatmap(final_matrix, 
         cluster_rows = TRUE, 
         cluster_cols = FALSE, # Clustering was manually done
         scale = "none",
         annotation_col = final_annotation_data,
         color = rev(extended_palette), 
         breaks = seq(-1, 1, length.out = 101),
         labels_row = as.expression(gene_names_italics),
         show_rownames = TRUE,
         show_colnames = FALSE,
         fontsize_row = 10,
         annotation_colors = col_annotation)


```
```{r fig.width=15, fig.height=10}
library(dplyr)
library(pheatmap)

# Assuming 'grouping_df' has the group information
groups <- unique(grouping_df$FREG5_Ethnic_Group)  # get unique groups
list_of_matrices <- list()

for(group in groups) {
  # Get column names for this group
  cols_for_group <- grouping_df$FREG1_Barcode[grouping_df$FREG5_Ethnic_Group == group]
  
  # Extract sub-matrix for this group
  sub_matrix <- z_score_matrix_new1[, cols_for_group, drop = FALSE]

  # Calculate sum of each column, focus on positive values
  col_sums <- colSums(sub_matrix * (sub_matrix > 0), na.rm = TRUE)

  # Order columns by descending sum
  ordered_cols <- colnames(sub_matrix)[order(-col_sums)]

  # Reorder sub_matrix according to the sums
  sub_matrix <- sub_matrix[, ordered_cols]

  list_of_matrices[[group]] <- sub_matrix
}

# Combine matrices in the order of the groups
final_matrix <- do.call(cbind, list_of_matrices)

# Assuming 'annotation_data' is correctly ordered to match 'grouping_df' ordering initially
# If the 'grouping_df' order changes, 'annotation_data' should be reordered to match 'final_matrix'
final_annotation_data <- annotation_data[match(colnames(final_matrix), rownames(annotation_data)), , drop = FALSE]

# Generate heatmap
pheatmap(final_matrix, 
         cluster_rows = TRUE, 
         cluster_cols = FALSE, # Custom order maintained
         scale = "none",
         annotation_col = final_annotation_data,
         color = rev(extended_palette), 
         breaks = seq(-1, 1, length.out = 101),
         labels_row = as.expression(gene_names_italics),
         show_rownames = TRUE,
         show_colnames = FALSE,
         fontsize_row = 10,
         annotation_colors = col_annotation)

```

```{r fig.width=6, fig.height=4}
library(dplyr)
library(pheatmap)

# Assuming 'grouping_df' has the group information
groups <- unique(grouping_df$FREG5_Ethnic_Group)  # get unique groups
list_of_matrices <- list()

for(group in groups) {
  # Get column names for this group
  cols_for_group <- grouping_df$FREG1_Barcode[grouping_df$FREG5_Ethnic_Group == group]
  
  # Extract sub-matrix for this group
  sub_matrix <- z_score_matrix_new1[, cols_for_group, drop = FALSE]

  # Calculate the number of positive entries for each column in the sub-matrix
  col_positives <- colSums(sub_matrix > 0, na.rm = TRUE)

  # Order columns by descending number of positives
  ordered_cols <- names(sort(col_positives, decreasing = TRUE))

  # Reorder sub-matrix according to the counts
  sub_matrix <- sub_matrix[, ordered_cols]

  list_of_matrices[[group]] <- sub_matrix
}

# Combine matrices in the order of the groups
final_matrix <- do.call(cbind, list_of_matrices)
# Calculate the number of positive entries for each row in the final matrix
row_positives <- rowSums(final_matrix > 0, na.rm = TRUE)

# Create a distance matrix for rows based on the count of positives
row_distance <- as.dist(outer(row_positives, row_positives, FUN = function(x, y) abs(x - y)))

# Perform hierarchical clustering of rows based on this distance
row_clustering <- hclust(row_distance, method = "complete")

# Reorder the final matrix rows according to the clustering
final_matrix_ordered <- final_matrix[row_clustering$order, ]
# Assuming 'annotation_data' is correctly ordered to match 'grouping_df' ordering initially
final_annotation_data <- annotation_data[match(colnames(final_matrix), rownames(annotation_data)), , drop = FALSE]

# Generate heatmap
pheatmap(final_matrix_ordered, 
         cluster_rows = FALSE,  # Custom clustering done
         cluster_cols = FALSE,  # Custom clustering done within groups
         scale = "none",
         annotation_col = final_annotation_data,
         color = rev(extended_palette), 
         breaks = seq(-1, 1, length.out = 101),
         labels_row = as.expression(gene_names_italics),
         show_rownames = TRUE,
         show_colnames = FALSE,
         fontsize_row = 10,
         annotation_colors = col_annotation)

```
```{r fig.width=12, fig.height=8}
library(dplyr)
library(pheatmap)

# First, proceed with the column clustering within each ethnic group as previously described

# Assuming 'grouping_df' has the group information
groups <- unique(grouping_df$FREG5_Ethnic_Group)  # get unique groups
list_of_matrices <- list()

for(group in groups) {
  # Get column names for this group
  cols_for_group <- grouping_df$FREG1_Barcode[grouping_df$FREG5_Ethnic_Group == group]
  
  # Extract sub-matrix for this group
  sub_matrix <- z_score_matrix_new1[, cols_for_group, drop = FALSE]

  # Calculate the number of positive entries for each column in the sub-matrix
  col_positives <- colSums(sub_matrix > 0, na.rm = TRUE)

  # Order columns by descending number of positives
  ordered_cols <- names(sort(col_positives, decreasing = TRUE))

  # Reorder sub-matrix according to the counts
  sub_matrix <- sub_matrix[, ordered_cols]

  list_of_matrices[[group]] <- sub_matrix
}

# Combine matrices in the order of the groups
final_matrix <- do.call(cbind, list_of_matrices)

# Calculate the number of positive entries for each row
row_positives <- rowSums(final_matrix > 0, na.rm = TRUE)

# Order rows by descending number of positives
ordered_rows <- names(sort(row_positives, decreasing = TRUE))

# Reorder the final matrix rows according to the sorted order
final_matrix_ordered <- final_matrix[ordered_rows, ]

# Assuming 'annotation_data' is correctly ordered to match 'final_matrix'
final_annotation_data <- annotation_data[match(colnames(final_matrix), rownames(annotation_data)), , drop = FALSE]

# Generate heatmap with sorted rows and custom clustered columns
pheatmap(final_matrix_ordered, 
         cluster_rows = FALSE,  # Rows are manually sorted
         cluster_cols = FALSE,  # Columns are manually clustered within groups
         scale = "none",
         annotation_col = final_annotation_data,
         color = rev(extended_palette), 
         breaks = seq(-1, 1, length.out = 101),
         labels_row = as.expression(gene_names_italics),
         show_rownames = TRUE,
         show_colnames = FALSE,
         fontsize_row = 10,
         annotation_colors = col_annotation)

```



```{r fig.width=10, fig.height=6}
extended_palette <- colorRampPalette(brewer.pal(11, "RdBu"))(100)
#png("ethinity_genus_zscore.png", width = 1400, height = 1000, res = 120)

pheatmap(z_score_matrix_new1, 
         cluster_rows = TRUE, 
         cluster_cols = FALSE, # Set to FALSE to maintain the custom order
         scale = "none",
         annotation_col = annotation_data,
         color = rev(extended_palette), 
         breaks = seq(-1, 1, length.out = 101),
         show_rownames = TRUE,
         show_colnames = FALSE,
         fontsize_row = 10)
#dev.off()
```



```{r fig.width=10, fig.height=6}
pheatmap(z_score_matrix_new1, 
         cluster_rows = FALSE, 
         cluster_cols = TRUE, # Set to FALSE to maintain the custom order
         scale = "none",
         annotation_col = annotation_data,
         color = rev(extended_palette), 
         breaks = seq(-1, 1, length.out = 101),
         show_rownames = TRUE,
         show_colnames = FALSE,
         fontsize_row = 14,
         gaps_row = 5)
```
```{r}
#write.csv(z_score_matrix_new1, "italic.csv")
#z_score_matrix_new1
```

```{r fig.width=10, fig.height=10}
ethnicity_colors <- c("C" = "#9FDDCA", "I" = "#9EA3F1", "M" = "#E17887")

gene_names_italics <- lapply(rownames(z_score_matrix_new1), function(x) bquote(italic(.(x))))
column_gap_indices <- c(395)
col_annotation <- list(Group = ethnicity_colors)
#png("figure/ethinity_genus_zscore_cut_auto.png", width = 1400, height = 1000, res = 120)
pheatmap(z_score_matrix_new1, 
         cluster_rows = FALSE,
         cluster_cols = FALSE,
         scale = "none", 
         annotation_col = annotation_data, 
         color = rev(extended_palette), 
         breaks = seq(-1, 1, length.out = 101), 
         show_rownames = TRUE, 
         show_colnames = FALSE, 
         fontsize_row = 10,
         labels_row = as.expression(gene_names_italics),
         cutree_rows = 4,  # Specify the number of clusters
         gaps_row = 1,
         gaps_col = column_gap_indices,
         annotation_colors = col_annotation)     


#dev.off()
```





# coffee
#FOOD AND MICROBIOME
```{r}
food_new

food_meta = merge(x = helios_new[,c("FREG0_PID", "FREG1_Barcode")], y = food_new, by.y = "FREG0_PID", by.x = "FREG0_PID")

food_meta = food_meta[!duplicated(food_meta$FREG1_Barcode),]
food_meta

#species_data$rowname <- rownames(species_data)
species_data_row = species_data
species_data_row$rowname <- rownames(species_data_row)
```
```{r}

#rownames(food_meta) = food_meta[,2]
#food_meta
#species_data
merged_data <- merge(species_data_row, food_meta, by.x = "rowname", by.y = "FREG1_Barcode", all = FALSE)
merged_data
```

```{r}
species_data2 = subset(species_data, rownames(species_data) %in% merged_data$rowname)
species_data2
food_meta2 = subset(food_meta, food_meta$FREG1_Barcode %in% merged_data$rowname)
food_meta2
```

DFFQA124_coffee_sugar_g_per_day

```{r}
c("DFFQA124_coffee_sugar_g_per_day", "DFFQA124a_Coffee_condensed_milk", "DFFQA124b_Coffee_evaporated_milk", "DFFQA124c_Coffee_whole_milk", "DFFQA124d_Coffee_lowfat_milk", "DFFQA124e_Coffee_skimmed_milk", "DFFQA124f_Coffee_soya_milk", "DFFQA124g_Coffee_creamer_no_sugar", "DFFQA124h_Coffee_3in1", "DFFQA124i_Coffee_2in1", "DFFQA124j_Coffee_no_milk", "DFFQA130_bottled_coffee_serv_day")
```

```{r}
coffee_cols <- c("DFFQA124_coffee_sugar_g_per_day", "DFFQA124a_Coffee_condensed_milk",
                 "DFFQA124b_Coffee_evaporated_milk", "DFFQA124c_Coffee_whole_milk",
                 "DFFQA124d_Coffee_lowfat_milk", "DFFQA124e_Coffee_skimmed_milk",
                 "DFFQA124f_Coffee_soya_milk", "DFFQA124g_Coffee_creamer_no_sugar",
                 "DFFQA124h_Coffee_3in1", "DFFQA124i_Coffee_2in1",
                 "DFFQA124j_Coffee_no_milk", "DFFQA130_bottled_coffee_serv_day")

food_meta2$coffee_drinker <- ifelse(rowSums(food_meta2[, coffee_cols] == 0) == length(coffee_cols),
                                    "No Coffee", "Drinks Coffee")
table(food_meta2$coffee_drinker)
```

```{r}

rownames(food_meta2) = food_meta2[,2]
food_meta3 = food_meta2[, c("FREG1_Barcode", "FREG0_PID", "coffee_drinker")]
food_meta3

OTU2 = otu_table(as.matrix(species_data2), taxa_are_rows = FALSE)              
SAM2 = sample_data(food_meta3, errorIfNULL = TRUE)

phylo2 <- phyloseq(OTU2, SAM2)
phylo2
#phylo = filter_taxa(phylo, function(x) sum(x > 2) > (0.11 * length(x)), TRUE)
```
## Drop species that are present in less than 5% of the samples
## Drop species with a total abundance less than 0.01% of the total reads
```{r}
prevalence_threshold <- 0.01
otu_table_filtered3 <- filter_taxa(phylo2, 
                                  function(x) sum(x > 0) > (prevalence_threshold * nsamples(otu_table(phylo2))),
                                  prune=TRUE)


abundance_threshold <- 0.0001
phyloseq_filtered3 <- filter_taxa(otu_table_filtered3, 
                                  function(x) sum(x) > (abundance_threshold * sum(otu_table(phylo2))),
                                  prune=TRUE)

phyloseq_filtered3

```

## alpha analysis
```{r}
plot_richness(phylo2, x = "coffee_drinker", color = "coffee_drinker", measures = c("Shannon", "Simpson", "Chao1", "Fisher")) + geom_boxplot() + theme_bw()
```
## Beta analysis
```{r}
dist = phyloseq::distance(phylo2, method="bray")
phylo.beta = ordinate(phylo2, method="PCoA", distance=dist)

plot_ordination(phylo2, phylo.beta, type = "samples", color = "coffee_drinker") +
  theme_bw() +
  stat_ellipse(aes(group = coffee_drinker)) +
  scale_color_discrete(name = "Coffee Drinker")
```

```{r}
dist_matrix <- as.matrix(dist)


permanova <- adonis2(dist_matrix ~ sample_data(phyloseq_filtered3)$coffee_drinker)


permanova
```

## da
```{r}

deseq_data2 <- phyloseq_to_deseq2(phyloseq_filtered3, ~ coffee_drinker)
deseq_data2 <- DESeq(deseq_data2)
res2 <- results(deseq_data2, contrast = c("coffee_drinker", "Drinks Coffee", "No Coffee"))
res2 <- res2[order(res2$padj, na.last=NA), ]
alpha <- 0.05
res2 <- res2[which(res2$padj < alpha), ]

rownames(res2)
```

```{r}
sig_taxa2 <- rownames(res2)
sig_taxa2 <- sig_taxa2[! sig_taxa2 %in% c('unclassified')]
```

##z score
mean of all samples 
```{r}
otu_table2 <- otu_table(phyloseq_filtered3)[, sig_taxa2]

sample_ids <- rownames(otu_table2)

# Apply Z-score normalization across columns

z_score_matrix2 <- apply(otu_table2, 2, scale)

# Reapply the row names to the scaled matrix
rownames(z_score_matrix2) <- sample_ids
#z_score_matrix

grouping_df = sample_data(phyloseq_filtered3)
grouping_df$coffee_drinker <- factor(grouping_df$coffee_drinker)
annotation_data <- as.data.frame(grouping_df$coffee_drinker)
rownames(annotation_data) <- grouping_df$FREG1_Barcode
colnames(annotation_data) <- c('Group')

group_colors <- brewer.pal(n = length(levels(grouping_df$coffee_drinker)), "Set1")
names(group_colors) <- levels(grouping_df$coffee_drinker)


grouping_df <- grouping_df[order(grouping_df$coffee_drinker),]

new = as.character(grouping_df$FREG1_Barcode)
# Create an ordering vector that matches the order of the sample IDs in the grouped dataframe
z_score_matrix2_t = t(z_score_matrix2)

z_score_matrix_new2 <- z_score_matrix2_t[,new]
#z_score_matrix_new1


```

```{r fig.width=10, fig.height=10}
extended_palette <- colorRampPalette(brewer.pal(11, "RdBu"))(100)
#png("coffee_genus_zscore.png", width = 1400, height = 1000, res = 120)
pheatmap(z_score_matrix_new2, 
         cluster_rows = TRUE, 
         cluster_cols = FALSE, # Set to FALSE to maintain the custom order
         scale = "none",
         annotation_col = annotation_data,
         color = rev(extended_palette), 
         breaks = seq(-1, 1, length.out = 101),
         show_rownames = TRUE,
         show_colnames = FALSE,
         fontsize_row = 10)
#dev.off()
pheatmap(z_score_matrix_new2, 
         cluster_rows = TRUE, 
         cluster_cols = TRUE, # Set to FALSE to maintain the custom order
         scale = "none",
         annotation_col = annotation_data,
         color = rev(extended_palette), 
         breaks = seq(-1, 1, length.out = 101),
         show_rownames = TRUE,
         show_colnames = FALSE,
         fontsize_row = 14)
```
