---
title: "helios_demog"
author: "Ruwen Zhou"
date: "4/3/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(plyr)
library(tidyverse)
library(dplyr)
library(readxl)
library(ggplot2)
library(phyloseq)
```

```{r}
helios_core = read_excel("data/HELIOS_Core_redacted.xlsx")
helios_core
table(helios_core$FREG14_Visit_number)
n_distinct(helios_core$FREG0_PID)
```
```{r}
gender_counts <- table(helios_core$FREG7_Gender)

# Convert the counts to a data frame
gender_df <- data.frame(
  gender = names(gender_counts),
  count = as.numeric(gender_counts)
)

# Calculate the percentage for each nationality
gender_df$percentage <- gender_df$count / sum(gender_df$count) * 100

# Create a pie chart using ggplot2
ggplot(gender_df, aes(x = "", y = count, fill = gender)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Gender Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(gender, "\n", 
                               count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5)) +
  guides(fill = guide_legend(title = "Nationality"))

```

```{r}
nationality_counts <- table(helios_core$FREG4_Nationality)

# Convert the counts to a data frame
nationality_df <- data.frame(
  nationality = names(nationality_counts),
  count = as.numeric(nationality_counts)
)

# Calculate the percentage for each nationality
nationality_df$percentage <- nationality_df$count / sum(nationality_df$count) * 100

# Create a pie chart using ggplot2
ggplot(nationality_df, aes(x = "", y = count, fill = nationality)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Nationality Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(nationality, "\n", 
                               count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5)) +
  guides(fill = guide_legend(title = "Nationality"))
```
```{r}
ggplot(nationality_df, aes(x = "", y = count, fill = nationality)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Patient Nationality Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(nationality, "\n", round(percentage, 1), "%")),
            position = position_stack(vjust = 0.5),
            size = 3, # Adjust the label size
            color = "white") + # Set label color for better contrast
  guides(fill = guide_legend(title = "Nationality"))

```

```{r}
age_groups <- cut(helios_core$FREG8_Age, breaks = seq(30, 90, by = 10),
                  labels = c("30-40", "40-50", "50-60", "60-70", "70-80", "80-90"),
                  include.lowest = TRUE, right = FALSE)

# Count the number of patients in each age group
age_counts <- table(age_groups)

# Convert the counts to a data frame
age_df <- data.frame(
  age_group = names(age_counts),
  count = as.numeric(age_counts)
)

# Calculate the percentage for each age group
age_df$percentage <- age_df$count / sum(age_df$count) * 100

# Create a pie chart using ggplot2
ggplot(age_df, aes(x = "", y = count, fill = age_group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Age Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(age_group, "\n", count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5),
            size = 4) +
  guides(fill = guide_legend(title = "Age Group"))
```
FREG5_Ethnic_Group
```{r}
ethnic_counts <- table(helios_core$FREG5_Ethnic_Group)

# Convert the counts to a data frame
ethnic_df <- data.frame(
  ethnic = names(ethnic_counts),
  count = as.numeric(ethnic_counts)
)

# Calculate the percentage for each nationality
ethnic_df$percentage <- ethnic_df$count / sum(ethnic_df$count) * 100

# Create a pie chart using ggplot2
ggplot(ethnic_df, aes(x = "", y = count, fill = ethnic)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Ethnic Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(ethnic, "\n", 
                               count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5)) +
  guides(fill = guide_legend(title = "Ethnic"))
```
## cancer screening
FREG14_Visit_number
```{r}
table(helios_core$FREG14_Visit_number)
table(helios_core$DHLQ490_Op4_4)
```


## blood pressure



## FFQA food

## FFQB nutrition 

## Blood test
### inflamatiion Hba1c

## combine to match microbiome
```{r}
metadata_micro = read_csv("data/metadata_fil_01032024_redacted.csv")
metadata_micro
sample_list = metadata_micro$`Original Sample name`

helios_new <- filter(helios_core, FREG1_Barcode %in% sample_list)
helios_new
```

```{r}
FFQ_food = read_xlsx("data/HELIOS_FFQA.xlsx")
FFQ_nutrition = read_xlsx("data/HELIOS_FFQB.xlsx")
FFQ_food
FFQ_nutrition
```
```{r}
pid_list = helios_new$FREG0_PID
food_new = filter(FFQ_food, FREG0_PID %in% pid_list)
#food_new$FREG0_PID[duplicated(food_new$FREG0_PID)]
nutrition_new = filter(FFQ_nutrition, FREG0_PID %in% pid_list)
nutrition_new
```

```{r}
food_new[food_new$FREG0_PID == "SHA-1468-6",]
```
## demographic of microbiome
```{r}
helios_new
```
## gender
```{r}
gender_counts <- table(helios_new$FREG7_Gender)

# Convert the counts to a data frame
gender_df <- data.frame(
  gender = names(gender_counts),
  count = as.numeric(gender_counts)
)


gender_df$percentage <- gender_df$count / sum(gender_df$count) * 100
gender_df
# Create a pie chart using ggplot2
ggplot(gender_df, aes(x = "", y = count, fill = gender)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Gender Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(gender, "\n", 
                               count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5)) +
  guides(fill = guide_legend(title = "Gender"))
```
## nationality
```{r}
nationality_counts <- table(helios_new$FREG4_Nationality)
nationality_counts
# Convert the counts to a data frame
nationality_df <- data.frame(
  nationality = names(nationality_counts),
  count = as.numeric(nationality_counts)
)

# Calculate the percentage for each nationality
nationality_df$percentage <- nationality_df$count / sum(nationality_df$count) * 100

# Create a pie chart using ggplot2
ggplot(nationality_df, aes(x = "", y = count, fill = nationality)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Nationality Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(nationality, "\n", 
                               count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5)) +
  guides(fill = guide_legend(title = "Nationality"))
```
## age
```{r}
age_groups <- cut(helios_new$FREG8_Age, breaks = seq(30, 90, by = 10),
                  labels = c("30-40", "40-50", "50-60", "60-70", "70-80", "80-90"),
                  include.lowest = TRUE, right = FALSE)

# Count the number of patients in each age group
age_counts <- table(age_groups)

# Convert the counts to a data frame
age_df <- data.frame(
  age_group = names(age_counts),
  count = as.numeric(age_counts)
)

# Calculate the percentage for each age group
age_df$percentage <- age_df$count / sum(age_df$count) * 100
age_df
# Create a pie chart using ggplot2
ggplot(age_df, aes(x = "", y = count, fill = age_group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Age Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(age_group, "\n", count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5),
            size = 4) +
  guides(fill = guide_legend(title = "Age Group"))
```
## ethinicity Chinese-Malay-Indian-Others
```{r}
ethnic_counts <- table(helios_new$FREG5_Ethnic_Group)

# Convert the counts to a data frame
ethnic_df <- data.frame(
  ethnic = names(ethnic_counts),
  count = as.numeric(ethnic_counts)
)

# Calculate the percentage for each nationality
ethnic_df$percentage <- ethnic_df$count / sum(ethnic_df$count) * 100
ethnic_df
# Create a pie chart using ggplot2
ggplot(ethnic_df, aes(x = "", y = count, fill = ethnic)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Ethnic Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(ethnic, "\n", 
                               count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5)) +
  guides(fill = guide_legend(title = "Ethnic"))
```

## BMI
```{r}
BMI = helios_new$DBI14_Weight/(helios_new$DBI13_Height*0.01)^2
BMI
BMI_groups <- cut(BMI, breaks = c(0,18.5,23,25,Inf),
                  labels = c("Underweight", "Normal", "Overweight", "Obese"),
                  include.lowest = TRUE, right = FALSE)

bmi_counts <- table(BMI_groups)

# Convert the counts to a data frame
bmi_df <- data.frame(
  bmi_group = names(bmi_counts),
  count = as.numeric(bmi_counts)
)

# Calculate the percentage for each age group
bmi_df$percentage <- bmi_df$count / sum(bmi_df$count) * 100

# Create a pie chart using ggplot2
ggplot(bmi_df, aes(x = "", y = count, fill = bmi_group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "BMI Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(bmi_group, "\n", count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5),
            size = 4) +
  guides(fill = guide_legend(title = "BMI Group"))
```
DLAB25_Gluf_Mmol_L

## Food and nutrition distribution 
```{r}
nutrition_new

# Select the columns to calculate mean and median
cols_to_summarize <- setdiff(names(nutrition_new), c("FREG0_PID", "FREG14_Visit_number"))

# Convert selected columns to numeric
nutrition_new[, cols_to_summarize] <- sapply(nutrition_new[, cols_to_summarize], as.numeric)

# Calculate mean for each selected column
column_means <- colMeans(nutrition_new[, cols_to_summarize], na.rm = TRUE)

# Calculate median for each selected column
column_medians <- apply(nutrition_new[, cols_to_summarize], 2, median, na.rm = TRUE)

# Create a summary dataframe
summary_df <- data.frame(
  Column = cols_to_summarize,
  Mean = column_means,
  Median = column_medians
)
summary_df

write.csv(summary_df, "data/nutrition_summary.csv", row.names=FALSE)
```

## blood test glucose （whether fasting)
```{r}
glucose = helios_new$DLAB25_Gluf_Mmol_L

glucose_groups <- cut(glucose, breaks = c(0,3.9,5.6,7,Inf),
                  labels = c("hypoglycemia", "Normal", "Pre-diabetes", "Diabetes"),
                  include.lowest = TRUE, right = FALSE)

glucose_counts <- table(glucose_groups)

# Convert the counts to a data frame
glucose_df <- data.frame(
  glucose_group = names(glucose_counts),
  count = as.numeric(glucose_counts)
)
glucose_df
# Calculate the percentage for each age group
glucose_df$percentage <- glucose_df$count / sum(glucose_df$count) * 100

# Create a pie chart using ggplot2
ggplot(glucose_df, aes(x = "", y = count, fill = glucose_group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Glucose Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(glucose_group, "\n", count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5),
            size = 4) +
  guides(fill = guide_legend(title = "Glucose Group"))
```

## DLAB27_Hba1C_Percent
```{r}
hba1c = helios_new$DLAB27_Hba1C_Percent

hba1c_groups <- cut(hba1c, breaks = c(0, 5.7, 6.4, Inf),
                  labels = c(),
                  include.lowest = TRUE, right = FALSE)

hba1c_counts <- table(hba1c_groups)

# Convert the counts to a data frame
hba1c_df <- data.frame(
  hba1c_group = names(hba1c_counts),
  count = as.numeric(hba1c_counts)
)
hba1c_df
# Calculate the percentage for each age group
hba1c_df$percentage <- hba1c_df$count / sum(hba1c_df$count) * 100

# Create a pie chart using ggplot2
ggplot(hba1c_df, aes(x = "", y = count, fill = hba1c_group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Hba1C Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(hba1c_group, "\n", count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5),
            size = 4) +
  guides(fill = guide_legend(title = "Hba1C Group"))
```
## CRP
```{r}
crp = helios_new$DLAB83_CRP

crp_groups <- cut(crp, breaks = c(0, 10, Inf),
                  labels = c("Normal", "inflammation"),
                  include.lowest = TRUE, right = FALSE)

crp_counts <- table(crp_groups)

# Convert the counts to a data frame
crp_df <- data.frame(
  crp_group = names(crp_counts),
  count = as.numeric(crp_counts)
)
crp_df
# Calculate the percentage for each age group
crp_df$percentage <- crp_df$count / sum(crp_df$count) * 100

# Create a pie chart using ggplot2
ggplot(crp_df, aes(x = "", y = count, fill = crp_group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "CRP Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(crp_group, "\n", count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5),
            size = 4) +
  guides(fill = guide_legend(title = "CRP Group"))
```

## DLAB15_Tc_Mmol_L Total Cholesterol (Mmol/L)
```{r}
tc = helios_new$DLAB15_Tc_Mmol_L

tc_groups <- cut(tc, breaks = c(0, 5.2, 6.1, Inf),
                  labels = c("Normal", "Borderline High", "High"),
                  include.lowest = TRUE, right = FALSE)
tc_groups
tc_counts <- table(tc_groups)

# Convert the counts to a data frame
tc_df <- data.frame(
  tc_group = names(tc_counts),
  count = as.numeric(tc_counts)
)
tc_df
# Calculate the percentage for each age group
tc_df$percentage <- tc_df$count / sum(tc_df$count) * 100

# Create a pie chart using ggplot2
ggplot(tc_df, aes(x = "", y = count, fill = tc_group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Total Cholesterol Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(tc_group, "\n", count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5),
            size = 4) +
  guides(fill = guide_legend(title = "Total Cholesterol Group"))
```

## DLAB17_Hdl_Mmol_L
```{r}
hdl = helios_new$DLAB17_Hdl_Mmol_L

hdl_groups <- cut(hdl, breaks = c(0, 1, 1.5, Inf),
                  labels = c("Low", "Normal", "High"),
                  include.lowest = TRUE, right = FALSE)

hdl_counts <- table(hdl_groups)

# Convert the counts to a data frame
hdl_df <- data.frame(
  hdl_group = names(hdl_counts),
  count = as.numeric(hdl_counts)
)
hdl_df
# Calculate the percentage for each age group
hdl_df$percentage <- hdl_df$count / sum(hdl_df$count) * 100

# Create a pie chart using ggplot2
ggplot(hdl_df, aes(x = "", y = count, fill = hdl_group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "HDL Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(hdl_group, "\n", count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5),
            size = 4) +
  guides(fill = guide_legend(title = "HDL Group"))
```

## DLAB19_Ldl_Mmol_L
```{r}
ldl = helios_new$DLAB19_Ldl_Mmol_L

ldl_groups <- cut(ldl, breaks = c(0, 3.3, 4, 4.9, Inf),
                  labels = c("Normal", "Borderline High", "High", "Very High"),
                  include.lowest = TRUE, right = FALSE)

ldl_counts <- table(ldl_groups)

# Convert the counts to a data frame
ldl_df <- data.frame(
  ldl_group = names(ldl_counts),
  count = as.numeric(ldl_counts)
)
ldl_df
# Calculate the percentage for each age group
ldl_df$percentage <- ldl_df$count / sum(ldl_df$count) * 100

# Create a pie chart using ggplot2
ggplot(ldl_df, aes(x = "", y = count, fill = ldl_group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "LDL Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(ldl_group, "\n", count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5),
            size = 4) +
  guides(fill = guide_legend(title = "LDL Group"))
```
## DLAB21_Trig_Mmol_L Triglyceride
```{r}
Triglyceride = helios_new$DLAB21_Trig_Mmol_L
Triglyceride
Triglyceride_groups <- cut(Triglyceride, breaks = c(0, 2.2, Inf),
                  labels = c("Normal", "High"),
                  include.lowest = TRUE, right = FALSE)

Triglyceride_counts <- table(Triglyceride_groups)

# Convert the counts to a data frame
Triglyceride_df <- data.frame(
  Triglyceride_group = names(Triglyceride_counts),
  count = as.numeric(Triglyceride_counts)
)
Triglyceride_df
# Calculate the percentage for each age group
Triglyceride_df$percentage <- Triglyceride_df$count / sum(Triglyceride_df$count) * 100

# Create a pie chart using ggplot2
ggplot(Triglyceride_df, aes(x = "", y = count, fill = Triglyceride_group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Triglyceride Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(Triglyceride_group, "\n", count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5),
            size = 4) +
  guides(fill = guide_legend(title = "Triglyceride Group"))
```


## DLAB23_Tc_Hdl_Ratio_Si
```{r}
ratio = helios_new$DLAB23_Tc_Hdl_Ratio_Si

ratio_groups <- cut(ratio, breaks = c(0, 3.5, 5.0, Inf),
                  labels = c("Normal", "Moderate risk", "High risk"),
                  include.lowest = TRUE, right = FALSE)

ratio_counts <- table(ratio_groups)

# Convert the counts to a data frame
ratio_df <- data.frame(
  ratio_group = names(ratio_counts),
  count = as.numeric(ratio_counts)
)
ratio_df
# Calculate the percentage for each age group
ratio_df$percentage <- ratio_df$count / sum(ratio_df$count) * 100

# Create a pie chart using ggplot2
ggplot(ratio_df, aes(x = "", y = count, fill = ratio_group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Total cholesterol to HDL ratio Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(ratio_group, "\n", count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5),
            size = 4) +
  guides(fill = guide_legend(title = "total cholesterol to HDL ratio Group"))
```

## DLAB28_Ua_Mmol_L
```{r}
helios_new2 <- helios_new

helios_new2$uric_acid_level <- ifelse(
  helios_new2$FREG7_Gender == "M" & helios_new2$DLAB28_Ua_Mmol_L >= 0.20 & helios_new2$DLAB28_Ua_Mmol_L <= 0.42,
  "Normal",
  ifelse(
    helios_new2$FREG7_Gender == "F" & helios_new2$DLAB28_Ua_Mmol_L >= 0.14 & helios_new2$DLAB28_Ua_Mmol_L <= 0.36,
    "Normal",
    ifelse(
      helios_new2$DLAB28_Ua_Mmol_L > 0.42,
      "High",
      "Low"
    )
  )
)

# Count the number of people with normal and high uric acid levels by gender
uric_acid_counts <- table(helios_new$FREG7_Gender, helios_new2$uric_acid_level)
uric_acid_counts

# Create a pie chart for males
male_counts <- uric_acid_counts["M", ]
male_df <- data.frame(
  male_group = names(male_counts),
  count = as.numeric(male_counts)
)

# Calculate the percentage for each age group
male_df$percentage <- male_df$count / sum(male_df$count) * 100

# Create a pie chart using ggplot2
ggplot(male_df, aes(x = "", y = count, fill = male_group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Male Uric Acid Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(male_group, "\n", count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5),
            size = 4) +
  guides(fill = guide_legend(title = "Male Uric Acid Group"))

# Create a pie chart for females
female_counts <- uric_acid_counts["F", ]
female_df <- data.frame(
  female_group = names(female_counts),
  count = as.numeric(female_counts)
)

# Calculate the percentage for each age group
female_df$percentage <- female_df$count / sum(female_df$count) * 100

# Create a pie chart using ggplot2
ggplot(female_df, aes(x = "", y = count, fill = female_group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Female Uric Acid Distribution") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(female_group, "\n", count, " (", round(percentage, 1), "%)")),
            position = position_stack(vjust = 0.5),
            size = 4) +
  guides(fill = guide_legend(title = "Female Uric Acid Group"))




```


## DLAB81_Insulin
```{r}
helios_new$DLAB81_Insulin
```

## microbiome and hba1c
### import data of microbiome
```{r}
species = read_csv("data/abundance_bracken_standard_16_20221209_taxa_S_unfiltered_reads_decont2.csv")
#species$`SAM id`
helios_new_list = helios_new$FREG1_Barcode
#sample_list = metadata_micro$`Original Sample name`
micro_list <- filter(metadata_micro, metadata_micro$`Original Sample name` %in% helios_new_list)
metadata_micro
species_new = filter(species, species$`SAM id` %in% micro_list$`Library ID`)

species_new
```
```{r}
new_merge = merge(metadata_micro, species_new, by.x = "Library ID", by.y = "SAM id")
#new_merge
species_meta = merge(x = helios_new[,c("FREG1_Barcode", "DLAB27_Hba1C_Percent", "DLAB25_Gluf_Mmol_L", "FREG5_Ethnic_Group","FREG10_Fasting_Hours","FBP16_Caffeine")], y = new_merge, by.y = "Original Sample name", by.x = "FREG1_Barcode")
species_meta = species_meta %>% drop_na(DLAB27_Hba1C_Percent)
species_meta
```

```{r}
metadata = species_meta[, c("FREG1_Barcode", "DLAB27_Hba1C_Percent","DLAB25_Gluf_Mmol_L", "FREG5_Ethnic_Group","FREG10_Fasting_Hours", "FBP16_Caffeine")]
metadata
drop = c("DLAB27_Hba1C_Percent", "Library ID", "SampleType", "seq_batch", "DLAB25_Gluf_Mmol_L", "FREG5_Ethnic_Group","FREG10_Fasting_Hours", "FBP16_Caffeine")
species_data = species_meta[,!(names(species_meta) %in% drop)]
species_data
```

```{r}
metadata$diabetes_status <- ifelse(metadata$DLAB27_Hba1C_Percent > 6.4, "diabetes",
                                   ifelse(metadata$DLAB27_Hba1C_Percent >= 5.7 & metadata$DLAB27_Hba1C_Percent <= 6.4, "pre-diabetes",
                                          ifelse(metadata$DLAB27_Hba1C_Percent < 5.7, "normal", NA)))

metadata$glucose <- ifelse(metadata$DLAB25_Gluf_Mmol_L > 7, "diabetes",
                                   ifelse(metadata$DLAB25_Gluf_Mmol_L >= 5.6 & metadata$DLAB25_Gluf_Mmol_L <= 7, "pre-diabetes",
                                          ifelse(metadata$DLAB25_Gluf_Mmol_L >= 3.9 & metadata$DLAB25_Gluf_Mmol_L <= 5.6, "normal",
                                                 ifelse(metadata$DLAB25_Gluf_Mmol_L < 3.9, "hypoglycemia", NA))))

                                         

metadata = metadata[!duplicated(metadata$FREG1_Barcode), ]
metadata
species_data = species_data[!duplicated(species_data$FREG1_Barcode), ]
species_data

```
```{r}
rownames(species_data) = species_data[,1]
species_data[1] = NULL
rownames(metadata) = metadata[,1]

species_data_t = t(species_data)
rownames(species_data_t) = colnames(species_data)
colnames(species_data_t) = rownames(species_data)
species_data_t
```


```{r}
OTU = otu_table(as.matrix(species_data), taxa_are_rows = FALSE)              
SAM = sample_data(metadata, errorIfNULL = TRUE)

phylo <- phyloseq(OTU, SAM)
phylo
phylo = filter_taxa(phylo, function(x) sum(x > 2) > (0.11 * length(x)), TRUE)
```
## Drop species that are present in less than 5% of the samples
## Drop species with a total abundance less than 0.01% of the total reads
```{r}
prevalence_threshold <- 0.05

otu_table_filtered <- filter_taxa(phylo, 
                                  function(x) sum(x > 0) > (prevalence_threshold * nsamples(otu_table(phylo))),
                                  prune=TRUE)
otu_table_filtered

abundance_threshold <- 0.0001
phyloseq_filtered <- filter_taxa(otu_table_filtered, 
                                  function(x) sum(x) > (abundance_threshold * sum(otu_table(phylo))),
                                  prune=TRUE)

phyloseq_filtered

```
```{r}
filtered_species = as.data.frame(otu_table(otu_table_filtered))
write.csv(filtered_species, "data/filtered_species.csv")
filtered_species
```

## alpha analysis
```{r}
sample_data(phyloseq_filtered)$diabetes_status <- factor(sample_data(phyloseq_filtered)$diabetes_status,
                                                         levels = c("normal", "pre-diabetes", "diabetes"))

plot_richness(phyloseq_filtered, x = "diabetes_status", color = "diabetes_status", measures = c("Shannon", "Simpson", "Chao1", "Fisher")) + geom_boxplot() + theme_bw()
```
## Beta analysis
```{r}
dist = phyloseq::distance(phyloseq_filtered, method="bray")
phylo.beta = ordinate(phyloseq_filtered, method="PCoA", distance=dist)

plot_ordination(phyloseq_filtered, phylo.beta, type = "samples", color = "diabetes_status") + theme_bw() + stat_ellipse(aes(group=diabetes_status))
```
```{r}
library(vegan)
sample_data(phyloseq_filtered)$diabetes_status <- factor(sample_data(phyloseq_filtered)$diabetes_status)

phyloseq_filtered <- subset_samples(phyloseq_filtered, !is.na(sample_data(phyloseq_filtered)$diabetes_status))

# Calculate the distance matrix using the ape package
dist <- phyloseq::distance(phyloseq_filtered, method = "bray")
dist_matrix <- as.matrix(dist)

# Perform PERMANOVA using adonis2
permanova <- adonis2(dist_matrix ~ sample_data(phyloseq_filtered)$diabetes_status)

# Print the PERMANOVA results
print(permanova)
```


## DA
```{r}
library(phyloseq)
library(DESeq2)
```

```{r}
deseq_data <- phyloseq_to_deseq2(phyloseq_filtered, ~ diabetes_status)
deseq_data <- DESeq(deseq_data)
res <- results(deseq_data, contrast = c("diabetes_status", "diabetes", "normal"))
res <- res[order(res$padj, na.last=NA), ]
alpha <- 0.05
res <- res[which(res$padj < alpha), ]
res
```

```{r}
plot_volcano <- function(res) {
  with(res, plot(log2FoldChange, -log10(padj), pch=20, main="Volcano plot", xlim=c(-5,5)))
  with(subset(res, padj<.05 ), points(log2FoldChange, -log10(padj), pch=20, col="red"))
  with(subset(res, abs(log2FoldChange)>1), points(log2FoldChange, -log10(padj), pch=20, col="orange"))
  with(subset(res, padj<.05 & abs(log2FoldChange)>1), points(log2FoldChange, -log10(padj), pch=20, col="green"))
}

plot_volcano(res)
```
The green dots mean taxons that are significant as well as having high fold change
The red dots mean taxons that ar significant but not having high fold change



```{r}
library(pheatmap)
library(RColorBrewer)

# Extract the significantly differentially abundant taxa
sig_taxa <- rownames(res)

sig_otu_table <- otu_table(phyloseq_filtered)[,sig_taxa]
sig_otu_table_df <- as.data.frame(sig_otu_table)

# Transpose the data frame to have taxa as rows and samples as columns


sig_otu_table_log <- log10(sig_otu_table_df + 1)

sig_otu_table_log_t <- t(sig_otu_table_log)
sig_otu_table_log_matrix <- as.matrix(sig_otu_table_log_t)
sig_otu_table_log_matrix
pheatmap(sig_otu_table_log_matrix)

grouping_df = sample_data(phyloseq_filtered)
grouping_df$diabetes_status <- factor(grouping_df$diabetes_status)
grouping_df
annotation_data <- as.data.frame(grouping_df$diabetes_status)
rownames(annotation_data) <- grouping_df$FREG1_Barcode
colnames(annotation_data) <- c('Group')
annotation_data

group_colors <- brewer.pal(n = length(levels(grouping_df$diabetes_status)), "Set1")
names(group_colors) <- levels(grouping_df$diabetes_status)

```
```{r}
grouping_df
# Assuming you have a dataframe 'grouping_df' with 'sample_id' and 'group' columns
# Make sure sample_id matches the column names of your sig_otu_table_log_matrix

# Order the grouping_df by the group column
grouping_df <- grouping_df[order(grouping_df$diabetes_status),]
sig_otu_table_log_matrix
new = as.character(grouping_df$FREG1_Barcode)
# Create an ordering vector that matches the order of the sample IDs in the grouped dataframe
order_matrix <- sig_otu_table_log_matrix[,new]

order_matrix
# Reorder the matrix columns according to the order_vector
#sig_otu_table_log_matrix_ordered <- sig_otu_table_log_matrix[, order_vector]
#sig_otu_table_log_matrix_ordered
# Now you can create the heatmap with the ordered matrix

pheatmap(order_matrix, 
         annotation_col = annotation_data,
         cluster_cols = FALSE,
         cluster_rows = T,
         annotation_colors = list(Group = group_colors))

pheatmap(order_matrix, 
         annotation_col = annotation_data,
         cluster_cols = T,
         cluster_rows = T,
         annotation_colors = list(Group = group_colors))

```

## ethinity
## alpha analysis
```{r}
library(ggpubr)

phyloseq_filtered_no_O <- subset_samples(phyloseq_filtered, FREG5_Ethnic_Group != "O")
```


```{r}
comparisons <- list(c("C", "I"), c("C", "M"), c("I", "M"))

# Plot alpha diversity and add pairwise statistical comparison
p <- plot_richness(phyloseq_filtered_no_O, x = "FREG5_Ethnic_Group", color = "FREG5_Ethnic_Group",
                  measures = c("Shannon", "Simpson", "Chao1", "Fisher")) +
    geom_boxplot() +
    theme_bw() +
    scale_color_manual(values = c("C" = "green4", "I" = "indianred1", "M" = "deepskyblue1")) +
    scale_fill_manual(values = c("C" = "green4", "I" = "indianred1", "M" = "deepskyblue1")) +
    stat_compare_means(comparisons = comparisons, method = "t.test", label = "p.signif", symnum.args = list(cutpoints = c(0, 0.001, 0.01, 0.05, 1), symbols = c("***", "**", "*", "ns")))
print(p)
#p + theme(text = element_text(family = "serif"))
```
## Beta analysis
```{r}
dist = phyloseq::distance(phyloseq_filtered_no_O, method="bray")
phylo.beta = ordinate(phyloseq_filtered_no_O, method="PCoA", distance=dist)
sample_data(phyloseq_filtered_no_O)$FREG5_Ethnic_Group <- factor(sample_data(phyloseq_filtered_no_O)$FREG5_Ethnic_Group,
                                                                 levels = c("C", "I", "M"))

plot_ordination(phyloseq_filtered_no_O, phylo.beta, type = "samples", color = "FREG5_Ethnic_Group") +
  theme_bw() +
  stat_ellipse(aes(group=FREG5_Ethnic_Group)) +
  scale_color_manual(values = c("C" = "green3", "I" = "indianred1", "M" = "deepskyblue1")) +
  scale_fill_manual(values = c("C" = "green3", "I" = "indianred1", "M" = "deepskyblue1"))

```


```{r}
library(ggplot2)

# Create the plot without adding the points
p <- plot_ordination(phyloseq_filtered_no_O, phylo.beta, type = "samples", color = "FREG5_Ethnic_Group") +
  theme_bw() +
  stat_ellipse(aes(group = FREG5_Ethnic_Group)) +
  scale_color_manual(values = c("C" = "green4", "I" = "indianred1", "M" = "deepskyblue1")) +
  scale_fill_manual(values = c("C" = "green4", "I" = "indianred1", "M" = "deepskyblue1"))

# Extract the data from the plot
plot_data <- p$data

# Add the points in the desired order using geom_point()
p +
  geom_point(data = subset(plot_data, FREG5_Ethnic_Group == "C"), color = "green4") +
  geom_point(data = subset(plot_data, FREG5_Ethnic_Group == "I"), color = "indianred1") +
  geom_point(data = subset(plot_data, FREG5_Ethnic_Group == "M"), color = "deepskyblue1")
  #theme(text = element_text(family = "serif"))
```


```{r}
# Convert glucose levels to a factor
sample_data(phyloseq_filtered)$FREG5_Ethnic_Group <- factor(sample_data(phyloseq_filtered)$FREG5_Ethnic_Group)


phyloseq_filtered2 <- subset_samples(phyloseq_filtered, !(sample_data(phyloseq_filtered)$FREG5_Ethnic_Group == "O"))
phyloseq_filtered2 <- subset_samples(phyloseq_filtered2, !is.na(sample_data(phyloseq_filtered2)$FREG5_Ethnic_Group))


dist <- phyloseq::distance(phyloseq_filtered2, method = "bray")
dist_matrix <- as.matrix(dist)

# Perform PERMANOVA using adonis2
permanova <- adonis2(dist_matrix ~ sample_data(phyloseq_filtered2)$FREG5_Ethnic_Group)

# Print the PERMANOVA results
permanova

plot_ordination(phyloseq_filtered2, phylo.beta, type = "samples", color = "FREG5_Ethnic_Group") + theme_bw() + stat_ellipse(aes(group=FREG5_Ethnic_Group))
```

1. The adonis2 model tested the effect of the FREG5_Ethnic_Group variable on the microbiome composition (dist_matrix). The model includes the Residual term, which represents the unexplained variation.
2. The FREG5_Ethnic_Group variable explains approximately 7% of the total variation in the microbiome composition (R2 = 0.0725). This means that the ethnic group factor accounts for a relatively small portion of the overall variability in the microbiome data.
3. The F-statistic (F = 3.8505) and the associated p-value (Pr(>F) = 0.001) indicate that the effect of FREG5_Ethnic_Group on the microbiome composition is highly significant. The F-statistic measures the ratio of the between-group variability to the within-group variability, and a higher F-value suggests a stronger effect of the grouping factor.
4. The low p-value (0.001) suggests that the observed differences in microbiome composition among the ethnic groups are unlikely to have occurred by chance alone. This provides strong evidence to reject the null hypothesis of no differences among the groups.
5. The Residual term explains the majority of the variation (approximately 93%) in the microbiome composition (R2 = 0.9275). This indicates that there are other factors or sources of variability not accounted for by the ethnic group factor that contribute to the overall variation in the microbiome data.
6. The total sum of squares (95.385) represents the total variation in the microbiome composition, while the sum of squares for the FREG5_Ethnic_Group (7.027) represents the variation explained by the ethnic group factor.

## DA
```{r}
library(phyloseq)
library(DESeq2)
```

```{r}
deseq_data1 <- phyloseq_to_deseq2(phyloseq_filtered2, ~ FREG5_Ethnic_Group)
deseq_data1 <- DESeq(deseq_data1)
res1 <- results(deseq_data1, contrast = c("FREG5_Ethnic_Group", "C", "I"))
res1 <- res1[order(res1$padj, na.last=NA), ]
alpha <- 0.05
res1 <- res1[which(res1$padj < alpha), ]
rownames(res1)
```

```{r}
c("Bifidobacterium angulatum", "Bifidobacterium catenulatum", "Bacteroides intestinalis", "Bifidobacterium longum", "Bifidobacterium breve", "Streptococcus thermophilus", "Streptococcus lutetiensis", "Bacteroides sp. D2", "Bacteroides sp. M10", "Bacteroides cellulosilyticus", "Bacteroides eggerthii", "Bacteroides sp. PHL 2737", "Bacteroides luhongzhouii", "Bifidobacterium dentium", "Bifidobacterium sp. KRGSERBCFTRI", "Bifidobacterium adolescentis", "Streptococcus parasanguinis", "Bacteroides sp. CBA7301", "Bifidobacterium pseudocatenulatum", "Streptococcus sp. LPB0220", "Akkermansia muciniphila", "Bacteroides salyersiae", "Bifidobacterium bifidum", "Bacteroides nordii", "Streptococcus sp. HSISM1", "Bacteroides ovatus", "Bacteroides thetaiotaomicron", "Bacteroides stercoris", "Bacteroides xylanisolvens", "Bacteroides zhangwenhongii", "Bacteroides faecis")
```


```{r}
plot_volcano <- function(res) {
  with(res, plot(log2FoldChange, -log10(padj), pch=20, main="Volcano plot", xlim=c(-5,5)))
  with(subset(res, padj<.05 ), points(log2FoldChange, -log10(padj), pch=20, col="red"))
  with(subset(res, abs(log2FoldChange)>1), points(log2FoldChange, -log10(padj), pch=20, col="orange"))
  with(subset(res, padj<.05 & abs(log2FoldChange)>1), points(log2FoldChange, -log10(padj), pch=20, col="green"))
}

plot_volcano(res1)
```


```{r}
deseq_data2 <- phyloseq_to_deseq2(phyloseq_filtered2, ~ FREG5_Ethnic_Group)
deseq_data2 <- DESeq(deseq_data2)
res2 <- results(deseq_data2, contrast = c("FREG5_Ethnic_Group", "C", "M"))
res2<- res2[order(res2$padj, na.last=NA), ]
alpha <- 0.05
res2 <- res2[which(res2$padj < alpha), ]
res2
```
```{r}
plot_volcano(res2)
```


```{r}
deseq_data3 <- phyloseq_to_deseq2(phyloseq_filtered2, ~ FREG5_Ethnic_Group)
deseq_data3 <- DESeq(deseq_data3)
res3 <- results(deseq_data3, contrast = c("FREG5_Ethnic_Group", "I", "M"))
res3<- res3[order(res3$padj, na.last=NA), ]
alpha <- 0.05
res3 <- res3[which(res3$padj < alpha), ]
res3
```
```{r}
plot_volcano(res3)
```


```{r}
library(pheatmap)
library(RColorBrewer)

# Extract the significantly differentially abundant taxa
sig_taxa <- rownames(res1)

sig_otu_table <- otu_table(phyloseq_filtered2)[,sig_taxa]
sig_otu_table_df <- as.data.frame(sig_otu_table)

# Transpose the data frame to have taxa as rows and samples as columns


sig_otu_table_log <- log10(sig_otu_table_df + 1)

sig_otu_table_log_t <- t(sig_otu_table_log)
sig_otu_table_log_matrix <- as.matrix(sig_otu_table_log_t)
sig_otu_table_log_matrix
pheatmap(sig_otu_table_log_matrix)

grouping_df = sample_data(phyloseq_filtered2)
grouping_df$FREG5_Ethnic_Group <- factor(grouping_df$FREG5_Ethnic_Group)
#grouping_df
annotation_data <- as.data.frame(grouping_df$FREG5_Ethnic_Group)
rownames(annotation_data) <- grouping_df$FREG1_Barcode
colnames(annotation_data) <- c('Group')
#annotation_data

group_colors <- brewer.pal(n = length(levels(grouping_df$FREG5_Ethnic_Group)), "Set1")
names(group_colors) <- levels(grouping_df$FREG5_Ethnic_Group)

```
```{r}
# Extract the significantly differentially abundant taxa
sig_taxa <- c("Bifidobacterium angulatum", "Bifidobacterium catenulatum", "Bacteroides intestinalis", "Bifidobacterium longum", "Bifidobacterium breve", "Streptococcus thermophilus", "Streptococcus lutetiensis", "Bacteroides sp. D2", "Bacteroides sp. M10", "Bacteroides cellulosilyticus", "Bacteroides eggerthii", "Bacteroides sp. PHL 2737", "Bacteroides luhongzhouii", "Bifidobacterium dentium", "Bifidobacterium sp. KRGSERBCFTRI", "Bifidobacterium adolescentis", "Streptococcus parasanguinis", "Bacteroides sp. CBA7301", "Bifidobacterium pseudocatenulatum", "Streptococcus sp. LPB0220", "Akkermansia muciniphila", "Bacteroides salyersiae", "Bifidobacterium bifidum", "Bacteroides nordii", "Streptococcus sp. HSISM1", "Bacteroides ovatus", "Bacteroides thetaiotaomicron", "Bacteroides stercoris", "Bacteroides xylanisolvens", "Bacteroides zhangwenhongii", "Bacteroides faecis")

sig_otu_table <- otu_table(phyloseq_filtered2)[,sig_taxa]
sig_otu_table_df <- as.data.frame(sig_otu_table)

# Transpose the data frame to have taxa as rows and samples as columns


sig_otu_table_log <- log10(sig_otu_table_df + 1)

sig_otu_table_log_t <- t(sig_otu_table_log)
sig_otu_table_log_matrix <- as.matrix(sig_otu_table_log_t)
sig_otu_table_log_matrix
pheatmap(sig_otu_table_log_matrix)

grouping_df = sample_data(phyloseq_filtered2)
#grouping_df$FREG5_Ethnic_Group <- factor(grouping_df$FREG5_Ethnic_Group)
#grouping_df
#annotation_data <- as.data.frame(grouping_df$FREG5_Ethnic_Group)
#rownames(annotation_data) <- grouping_df$FREG1_Barcode
#colnames(annotation_data) <- c('Group')
#annotation_data

#group_colors <- brewer.pal(n = length(levels(grouping_df$FREG5_Ethnic_Group)), "Set1")
#names(group_colors) <- levels(grouping_df$FREG5_Ethnic_Group)
```

```{r fig.width=16, fig.height=12}
#grouping_df
# Assuming you have a dataframe 'grouping_df' with 'sample_id' and 'group' columns
# Make sure sample_id matches the column names of your sig_otu_table_log_matrix

# Order the grouping_df by the group column
grouping_df <- grouping_df[order(grouping_df$FREG5_Ethnic_Group),]
sig_otu_table_log_matrix
new = as.character(grouping_df$FREG1_Barcode)
# Create an ordering vector that matches the order of the sample IDs in the grouped dataframe
order_matrix <- sig_otu_table_log_matrix[,new]

order_matrix
# Reorder the matrix columns according to the order_vector
#sig_otu_table_log_matrix_ordered <- sig_otu_table_log_matrix[, order_vector]
#sig_otu_table_log_matrix_ordered
# Now you can create the heatmap with the ordered matrix
```


```{r}

#png(file="ethinity_significance.png", width=1000, height=1000)
pheatmap(order_matrix, 
         annotation_col = annotation_data,
         cluster_cols = FALSE,
         cluster_rows = T,
         annotation_colors = list(Group = group_colors))
#dev.off()
```


```{r fig.width=16, fig.height=12}
pheatmap(order_matrix, 
         annotation_col = annotation_data,
         cluster_cols = T,
         cluster_rows = T,
         annotation_colors = list(Group = group_colors))

```





## glucose
```{r}
plot_richness(phyloseq_filtered, x = "glucose", color = "glucose", measures = c("Shannon", "Simpson", "Chao1", "Fisher")) + geom_boxplot() + theme_bw()
```
## Beta analysis
```{r}
library(vegan)
dist = phyloseq::distance(phyloseq_filtered, method="bray")
phylo.beta = ordinate(phyloseq_filtered, method="PCoA", distance=dist)

plot_ordination(phyloseq_filtered, phylo.beta, type = "samples", color = "glucose") + theme_bw() + stat_ellipse(aes(group=glucose))
```


```{r}
library(ape)
library(vegan)

# Convert glucose levels to a factor
sample_data(phyloseq_filtered)$glucose <- factor(sample_data(phyloseq_filtered)$glucose)

phyloseq_filtered <- subset_samples(phyloseq_filtered, !is.na(sample_data(phyloseq_filtered)$glucose))

# Calculate the distance matrix using the ape package
dist <- phyloseq::distance(phyloseq_filtered, method = "bray")
dist_matrix <- as.matrix(dist)

# Perform PERMANOVA using adonis2
permanova <- adonis2(dist_matrix ~ sample_data(phyloseq_filtered)$glucose)

# Print the PERMANOVA results
print(permanova)

```
#FOOD AND MICROBIOME
```{r}
food_new


food_meta = merge(x = helios_new[,c("FREG0_PID", "FREG1_Barcode")], y = food_new, by.y = "FREG0_PID", by.x = "FREG0_PID")


food_meta = food_meta[!duplicated(food_meta$FREG1_Barcode),]
food_meta

#species_data$rowname <- rownames(species_data)
species_data_row = species_data
species_data_row$rowname <- rownames(species_data_row)
```
```{r}

#rownames(food_meta) = food_meta[,2]
#food_meta
#species_data
merged_data <- merge(species_data_row, food_meta, by.x = "rowname", by.y = "FREG1_Barcode", all = FALSE)
merged_data
```

```{r}
species_data2 = subset(species_data, rownames(species_data) %in% merged_data$rowname)
species_data2
food_meta2 = subset(food_meta, food_meta$FREG1_Barcode %in% merged_data$rowname)
```

DFFQA124_coffee_sugar_g_per_day

```{r}
c("DFFQA124_coffee_sugar_g_per_day", "DFFQA124a_Coffee_condensed_milk", "DFFQA124b_Coffee_evaporated_milk", "DFFQA124c_Coffee_whole_milk", "DFFQA124d_Coffee_lowfat_milk", "DFFQA124e_Coffee_skimmed_milk", "DFFQA124f_Coffee_soya_milk", "DFFQA124g_Coffee_creamer_no_sugar", "DFFQA124h_Coffee_3in1", "DFFQA124i_Coffee_2in1", "DFFQA124j_Coffee_no_milk", "DFFQA130_bottled_coffee_serv_day")
```

```{r}
coffee_cols <- c("DFFQA124_coffee_sugar_g_per_day", "DFFQA124a_Coffee_condensed_milk",
                 "DFFQA124b_Coffee_evaporated_milk", "DFFQA124c_Coffee_whole_milk",
                 "DFFQA124d_Coffee_lowfat_milk", "DFFQA124e_Coffee_skimmed_milk",
                 "DFFQA124f_Coffee_soya_milk", "DFFQA124g_Coffee_creamer_no_sugar",
                 "DFFQA124h_Coffee_3in1", "DFFQA124i_Coffee_2in1",
                 "DFFQA124j_Coffee_no_milk", "DFFQA130_bottled_coffee_serv_day")

food_meta2$coffee_drinker <- ifelse(rowSums(food_meta2[, coffee_cols] == 0) == length(coffee_cols),
                                    "No Coffee", "Drinks Coffee")
table(food_meta2$coffee_drinker)
```

```{r}

rownames(food_meta2) = food_meta2[,2]
food_meta3 = food_meta2[, c("FREG1_Barcode", "FREG0_PID", "coffee_drinker")]
food_meta3

OTU2 = otu_table(as.matrix(species_data2), taxa_are_rows = FALSE)              
SAM2 = sample_data(food_meta3, errorIfNULL = TRUE)

phylo2 <- phyloseq(OTU2, SAM2)
phylo2
#phylo = filter_taxa(phylo, function(x) sum(x > 2) > (0.11 * length(x)), TRUE)
```
## Drop species that are present in less than 5% of the samples
## Drop species with a total abundance less than 0.01% of the total reads
```{r}
prevalence_threshold <- 0.01
otu_table_filtered3 <- filter_taxa(phylo2, 
                                  function(x) sum(x > 0) > (prevalence_threshold * nsamples(otu_table(phylo2))),
                                  prune=TRUE)


abundance_threshold <- 0.0001
phyloseq_filtered3 <- filter_taxa(otu_table_filtered3, 
                                  function(x) sum(x) > (abundance_threshold * sum(otu_table(phylo2))),
                                  prune=TRUE)

phyloseq_filtered3

```

## alpha analysis
```{r}
comparisons <- list(c("Drinks Coffee", "No Coffee"))
plot_richness(phylo2, x = "coffee_drinker", color = "coffee_drinker", measures = c("Shannon", "Simpson", "Chao1", "Fisher")) + geom_boxplot() + theme_bw() +
  scale_color_manual(values = c("Drinks Coffee" = "lightskyblue2", "No Coffee" = "indianred1")) +
  scale_fill_manual(values = c("Drinks Coffee" = "lightskyblue2", "No Coffee" = "indianred1")) +
  stat_compare_means(comparisons = comparisons, method = "t.test", label = "p.signif", symnum.args = list(cutpoints = c(0, 0.001, 0.01, 0.05, 1), symbols = c("***", "**", "*", "ns")))


```
## Beta analysis
```{r}
dist = phyloseq::distance(phylo2, method="bray")
phylo.beta = ordinate(phylo2, method="PCoA", distance=dist)

plot_ordination(phylo2, phylo.beta, type = "samples", color = "coffee_drinker") +
  theme_bw() +
  stat_ellipse(aes(group = coffee_drinker)) +
  scale_color_discrete(name = "Coffee Drinker") +
  scale_color_manual(values = c("Drinks Coffee" = "lightskyblue2", "No Coffee" = "indianred1")) +
  scale_fill_manual(values = c("Drinks Coffee" = "lightskyblue2", "No Coffee" = "indianred1"))
```

```{r}
dist_matrix <- as.matrix(dist)

# Perform PERMANOVA using adonis2
permanova <- adonis2(dist_matrix ~ sample_data(phyloseq_filtered3)$coffee_drinker)

# Print the PERMANOVA results
permanova
```

## da
```{r}

deseq_data2 <- phyloseq_to_deseq2(phyloseq_filtered3, ~ coffee_drinker)
deseq_data2 <- DESeq(deseq_data2)
res2 <- results(deseq_data2, contrast = c("coffee_drinker", "Drinks Coffee", "No Coffee"))
res2 <- res2[order(res2$padj, na.last=NA), ]
alpha <- 0.05
res2 <- res2[which(res2$padj < alpha), ]
print(res2)
rownames(res2)
```

"Streptococcus pasteurianus", "Bacteroides sp. DH3716P", "Bifidobacterium animalis", "Bacteroides luhongzhouii", "Bifidobacterium adolescentis", "Bacteroides sp. ZJ-18",
"Bacteroides ovatus", "Bacteroides sp. CBA7301", "Akkermansia sp. WON2089", "Bifidobacterium catenulatum", "Bacteroides nordii","Streptococcus infantarius","Streptococcus parasanguinis"
```{r}
library(pheatmap)
library(RColorBrewer)


sig_taxa <- rownames(res2)

sig_otu_table <- otu_table(phyloseq_filtered3)[,sig_taxa]
sig_otu_table_df <- as.data.frame(sig_otu_table)




sig_otu_table_log <- log10(sig_otu_table_df + 1)

sig_otu_table_log_t <- t(sig_otu_table_log)
sig_otu_table_log_matrix <- as.matrix(sig_otu_table_log_t)
sig_otu_table_log_matrix
pheatmap(sig_otu_table_log_matrix)

grouping_df = sample_data(phyloseq_filtered3)
grouping_df$coffee_drinker <- factor(grouping_df$coffee_drinker)
#grouping_df
annotation_data <- as.data.frame(grouping_df$coffee_drinker)
rownames(annotation_data) <- grouping_df$FREG1_Barcode
colnames(annotation_data) <- c('Group')
#annotation_data

group_colors <- brewer.pal(n = length(levels(grouping_df$coffee_drinker)), "Set1")
names(group_colors) <- levels(grouping_df$coffee_drinker)

```
```{r fig.width=16, fig.height=12}
grouping_df



grouping_df <- grouping_df[order(grouping_df$coffee_drinker),]
sig_otu_table_log_matrix
new = as.character(grouping_df$FREG1_Barcode)

order_matrix <- sig_otu_table_log_matrix[,new]


```


```{r fig.width=16, fig.height=12}

#png(file="ethinity_significance.png", width=1000, height=1000)
pheatmap(order_matrix, 
         annotation_col = annotation_data,
         cluster_cols = FALSE,
         cluster_rows = T,
         annotation_colors = list(Group = group_colors))
#dev.off()
```


```{r fig.width=16, fig.height=12}
typeof(order_matrix)
pheatmap(order_matrix, 
         annotation_col = annotation_data,
         cluster_cols = T,
         cluster_rows = T,
         annotation_colors = list(Group = group_colors))

```

"Streptococcus pasteurianus", "Bacteroides sp. DH3716P", "Bifidobacterium animalis", "Bacteroides luhongzhouii", "Bifidobacterium adolescentis", "Bacteroides sp. ZJ-18",
"Bacteroides ovatus", "Bacteroides sp. CBA7301", "Akkermansia sp. WON2089", "Bifidobacterium catenulatum", "Bacteroides nordii","Streptococcus infantarius","Streptococcus parasanguinis"
```{r}
library(pheatmap)
library(RColorBrewer)


sig_taxa <- c("Streptococcus pasteurianus", "Bacteroides sp. DH3716P", "Bifidobacterium animalis", "Bacteroides luhongzhouii", "Bifidobacterium adolescentis", "Bacteroides sp. ZJ-18", "Bacteroides ovatus", "Bacteroides sp. CBA7301", "Akkermansia sp. WON2089", "Bifidobacterium catenulatum", "Bacteroides nordii","Streptococcus infantarius","Streptococcus parasanguinis")

sig_otu_table <- otu_table(phyloseq_filtered3)[,sig_taxa]
sig_otu_table_df <- as.data.frame(sig_otu_table)




sig_otu_table_log <- log10(sig_otu_table_df + 1)

sig_otu_table_log_t <- t(sig_otu_table_log)
sig_otu_table_log_matrix <- as.matrix(sig_otu_table_log_t)
sig_otu_table_log_matrix
pheatmap(sig_otu_table_log_matrix)

grouping_df = sample_data(phyloseq_filtered3)
grouping_df$coffee_drinker <- factor(grouping_df$coffee_drinker)
#grouping_df
annotation_data <- as.data.frame(grouping_df$coffee_drinker)
rownames(annotation_data) <- grouping_df$FREG1_Barcode
colnames(annotation_data) <- c('Group')
#annotation_data

group_colors <- brewer.pal(n = length(levels(grouping_df$coffee_drinker)), "Set1")
names(group_colors) <- levels(grouping_df$coffee_drinker)

```
```{r fig.width=16, fig.height=12}
grouping_df
# Assuming you have a dataframe 'grouping_df' with 'sample_id' and 'group' columns
# Make sure sample_id matches the column names of your sig_otu_table_log_matrix

# Order the grouping_df by the group column
grouping_df <- grouping_df[order(grouping_df$coffee_drinker),]
sig_otu_table_log_matrix
new = as.character(grouping_df$FREG1_Barcode)
# Create an ordering vector that matches the order of the sample IDs in the grouped dataframe
order_matrix <- sig_otu_table_log_matrix[,new]

# Reorder the matrix columns according to the order_vector
#sig_otu_table_log_matrix_ordered <- sig_otu_table_log_matrix[, order_vector]
#sig_otu_table_log_matrix_ordered
# Now you can create the heatmap with the ordered matrix
```

```{r fig.width=20, fig.height=20}}
pheatmap(order_matrix, 
         annotation_col = annotation_data,
         cluster_cols = FALSE,
         cluster_rows = T,
         annotation_colors = list(Group = group_colors))
```

## stratification
race as confounder 
gender
economy ?
antibiotics ?
health status ?

```{r}
metadata
food_meta2
```

```{r}
coffee_race = merge(metadata, food_meta2[,c("FREG1_Barcode", "FREG0_PID", "coffee_drinker")], by = "FREG1_Barcode", all = FALSE)
```
```{r}
library(dplyr)
library(tidyr)
library(janitor)
```


```{r}
# Perform the chi-square test

chi_square_result <- coffee_race %>%
  tabyl(FREG5_Ethnic_Group, coffee_drinker) %>%
  chisq.test()

# Print the chi-square test results
print(chi_square_result)
```

```{r}
# Load the necessary library (if not already loaded)

# Create the contingency table
contingency_table <- table(coffee_race$FREG5_Ethnic_Group, coffee_race$coffee_drinker)

# Perform the chi-square test
chi_square_result <- chisq.test(contingency_table)

# Print the contingency table
print(contingency_table)

# Print the chi-square test results
print(chi_square_result)
```
p value 0.3557 > 0.05


```{r}
coffee_race
```

## DII
```{r}
dii = read_xlsx("data/Updated_HELIOS_FFQA_with_DII_v2.xlsx")
dii
```
```{r}
pid_list = helios_new$FREG0_PID
food_new_dii = filter(dii, FREG0_PID %in% pid_list)
food_new_dii
```
```{r}
#food_new_dii


food_meta_dii = merge(x = helios_new[,c("FREG0_PID", "FREG1_Barcode")], y = food_new_dii, by.y = "FREG0_PID", by.x = "FREG0_PID")


food_meta_dii = food_meta_dii[!duplicated(food_meta$FREG1_Barcode),]
food_meta_dii

#species_data$rowname <- rownames(species_data)
species_data_row = species_data
species_data_row$rowname <- rownames(species_data_row)
```
```{r}

#rownames(food_meta) = food_meta[,2]
#food_meta
#species_data
merged_data <- merge(species_data_row, food_meta_dii, by.x = "rowname", by.y = "FREG1_Barcode", all = FALSE)
merged_data
```

```{r}
species_data2 = subset(species_data, rownames(species_data) %in% merged_data$rowname)
species_data2
food_meta_dii2 = subset(food_meta_dii, food_meta_dii$FREG1_Barcode %in% merged_data$rowname)
food_meta_dii3 = subset(food_meta_dii2, food_meta_dii2$FREG1_Barcode %in% rownames(species_data))
food_meta_dii3 = food_meta_dii3[!duplicated(food_meta_dii3$FREG1_Barcode),]
food_meta_dii3
```


```{r}
#write.csv(food_meta_dii3, "food_535.csv")
```
```{r}

food_dii_core = merge(helios_new, food_meta_dii3, by = "FREG1_Barcode", all = FALSE)
food_dii_core$BMI = food_dii_core$DBI14_Weight/(food_dii_core$DBI13_Height*0.01)^2
food_dii_core$BMI

food_dii_core$BMI_groups <- cut(food_dii_core$BMI, breaks = c(0,18.5,25,30,Inf),
                  labels = c("Underweight", "Normal", "Overweight", "Obese"),
                  include.lowest = TRUE, right = FALSE)
food_dii_core
```
```{r}
nutrition_merge = merge(nutrition_new, food_meta_dii3[,c("FREG0_PID", "FREG1_Barcode")], by = "FREG0_PID", all = FALSE)
nutrition_merge2 = subset(nutrition_merge, nutrition_merge$FREG1_Barcode %in% food_meta_dii3$FREG1_Barcode)
nutrition_merge2 = nutrition_merge2[!duplicated(nutrition_merge2$FREG1_Barcode),]

rownames(nutrition_merge2) = nutrition_merge2$FREG1_Barcode
nutrition_merge2
```



```{r}
#write.csv(species_data2, "species_535.csv")
#write.csv(food_dii_core, "demo_535.csv")
#write.csv(nutrition_merge2, "nutrition_535.csv")
```


```{r}
#food_dii_core

rownames(food_dii_core) = food_dii_core[,1]
food_dii_core2 = food_dii_core[, c("FREG1_Barcode", "FREG0_PID.x", "Inflammatory_Score", "BMI", "BMI_groups", "FREG7_Gender", "FREG5_Ethnic_Group", "FREG8_Age")]
food_dii_core2

OTU3 = otu_table(as.matrix(species_data2), taxa_are_rows = FALSE)              
SAM3 = sample_data(food_dii_core2, errorIfNULL = TRUE)

phylo3 <- phyloseq(OTU3, SAM3)
phylo3
#phylo = filter_taxa(phylo, function(x) sum(x > 2) > (0.11 * length(x)), TRUE)
```
## Drop species that are present in less than 5% of the samples
## Drop species with a total abundance less than 0.01% of the total reads
```{r}
prevalence_threshold <- 0.05
otu_table_dii <- filter_taxa(phylo3, 
                                  function(x) sum(x > 0) > (prevalence_threshold * nsamples(otu_table(phylo3))),
                                  prune=TRUE)


abundance_threshold <- 0.0001
phyloseq_dii <- filter_taxa(otu_table_dii, 
                                  function(x) sum(x) > (abundance_threshold * sum(otu_table(phylo3))),
                                  prune=TRUE)

phyloseq_dii

```
```{r}

dist = phyloseq::distance(phylo3, method="bray")
#phylo.beta = ordinate(phylo3, method="PCoA", distance=dist)

#plot_ordination(phylo2, phylo.beta, type = "samples", color = "coffee_drinker") +
  #theme_bw() +
  #stat_ellipse(aes(group = coffee_drinker)) +
  #scale_color_discrete(name = "Coffee Drinker")

dist_matrix <- as.matrix(dist)

# Perform PERMANOVA using adonis2
permanova <- adonis2(dist_matrix ~ sample_data(phylo3)$Inflammatory_Score) #+ sample_data(phylo3)$FREG5_Ethnic_Group)

# Print the PERMANOVA results
permanova

```
















