---
title: "network"
author: "Ruwen Zhou"
date: "6/19/2024"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(plyr)
library(tidyverse)
library(dplyr)
library(readxl)
library(ggplot2)
library(phyloseq)
library(vegan)
library(DESeq2)
library(pheatmap)
library(RColorBrewer)
library(ggraph)
library(igraph)
library(Hmisc)
library(writexl)

library(tidygraph)

```


#demographic
```{r}
food_data = read_xlsx("data/FFQA862.xlsx")
nutrition_data = read_xlsx("data/FFQB800.xlsx")
species = read_xlsx("data/species_866.xlsx")
metadata = read_xlsx("data/ethnicity_866.xlsx")

food_data <- food_data %>%
  mutate(across(c("Mushrooms", "Papaya", "Orange", "Chapati"), as.numeric))
food_data
nutrition_data
species
metadata
```

```{r}
colnames(species)[1] = "Participant_ID"
colnames(metadata)[1] = "Participant_ID"
threshold <- nrow(species) * 0.5

species_filtered1 <- species %>% select_if(function(x) sum(x > 0) > nrow(species) * 0.2)
species_filtered1
species_filtered2 <- species %>% select_if(function(x) sum(x > 0) > threshold)

#abundance_threshold <- 0.0001
#species_filtered2 <- species_filtered %>% select_if(function(x) sum(x) > (abundance_threshold * sum(species)))
#species_filtered2
species_filtered2 <- species_filtered2[, !names(species_filtered2) %in% c("unclassified", "others", "Homo sapiens")]
#species_filtered2$`Homo sapiens`
#species_filtered2
food_data[is.na(food_data)] <- 0
#food_data <- data.frame(lapply(food_data, function(x) as.numeric(as.character(x))))
nutrition_data[is.na(nutrition_data)] <- 0
```


```{r}
food_data
nutrition_data
species_filtered2
species_filtered1
metadata
```


## knowledge graph
```{r}
# create_microbiome_knowledge_graph <- function(food_data, nutrition_data, 
#                                             species_data, metadata) {
#   # Ensure all datasets have the same participants and in the same order
#   common_participants <- Reduce(intersect, list(
#     food_data$Participant_ID,
#     nutrition_data$Participant_ID,
#     species_data$Participant_ID
#   ))
#   
#   # Filter and arrange all datasets
#   food_data <- food_data %>%
#     filter(Participant_ID %in% common_participants) %>%
#     arrange(Participant_ID)
#   
#   nutrition_data <- nutrition_data %>%
#     filter(Participant_ID %in% common_participants) %>%
#     arrange(Participant_ID)
#   
#   species_data <- species_data %>%
#     filter(Participant_ID %in% common_participants) %>%
#     arrange(Participant_ID)
#   
#   # Get categories
#   food_categories <- setdiff(colnames(food_data), "Participant_ID")
#   nutrition_categories <- setdiff(colnames(nutrition_data), "Participant_ID")
#   species_categories <- setdiff(colnames(species_data), "Participant_ID")
#   
#   # Calculate correlations with p-values using rcorr
#   calculate_correlations <- function(data1, categories1, data2, categories2) {
#     # Print dimensions for debugging
#     cat("Matrix 1 dimensions:", nrow(data1), "x", length(categories1), "\n")
#     cat("Matrix 2 dimensions:", nrow(data2), "x", length(categories2), "\n")
#     
#     # Prepare matrices
#     mat1 <- as.matrix(data1[, categories1])
#     mat2 <- as.matrix(data2[, categories2])
#     
#     # Verify dimensions
#     if(nrow(mat1) != nrow(mat2)) {
#       stop("Matrices must have same number of rows")
#     }
#     
#     # Handle NAs if present
#     complete_cases <- complete.cases(mat1, mat2)
#     if(!all(complete_cases)) {
#       cat("Removing", sum(!complete_cases), "rows with missing values\n")
#       mat1 <- mat1[complete_cases, ]
#       mat2 <- mat2[complete_cases, ]
#     }
#     
#     # Calculate correlations and p-values
#     rcorr_result <- rcorr(cbind(mat1, mat2), type = "pearson")
#     
#     # Extract relevant portions of the correlation and p-value matrices
#     n1 <- ncol(mat1)
#     n2 <- ncol(mat2)
#     cors <- rcorr_result$r[1:n1, (n1+1):(n1+n2)]
#     p_values <- rcorr_result$P[1:n1, (n1+1):(n1+n2)]
#     
#     # Set row and column names
#     rownames(cors) <- categories1
#     colnames(cors) <- categories2
#     rownames(p_values) <- categories1
#     colnames(p_values) <- categories2
#     
#     return(list(correlations = cors, p_values = p_values))
#   }
#   
#   cat("Calculating species-food correlations...\n")
#   species_food_results <- calculate_correlations(
#     species_data, species_categories,
#     food_data, food_categories
#   )
#   
#   cat("Calculating species-nutrition correlations...\n")
#   species_nutrition_results <- calculate_correlations(
#     species_data, species_categories,
#     nutrition_data, nutrition_categories
#   )
#   
#   # Create edges from correlations
#   create_correlation_edges <- function(cor_results, relationship_type) {
#     cors <- cor_results$correlations
#     p_values <- cor_results$p_values
#     edges <- which(p_values <= 0.01, arr.ind = TRUE)
#     
#     if(nrow(edges) > 0) {
#       tibble(
#         from = rownames(cors)[edges[,1]],
#         to = colnames(cors)[edges[,2]],
#         relationship = relationship_type,
#         weight = cors[cbind(edges[,1], edges[,2])],
#         p_value = p_values[cbind(edges[,1], edges[,2])]
#       )
#     } else {
#       tibble(from = character(), to = character(), 
#              relationship = character(), weight = numeric(),
#              p_value = numeric())
#     }
#   }
#   
#   # Get species that have significant connections
#   species_with_connections <- unique(c(
#     rownames(species_food_results$correlations)[rowSums(species_food_results$p_values <= 0.01) > 0],
#     rownames(species_nutrition_results$correlations)[rowSums(species_nutrition_results$p_values <= 0.01) > 0]
#   ))
#   
#   cat("\nFound", length(species_with_connections), 
#       "species with significant connections (p < 0.01)\n")
#   
#   # Create nodes
#   ethnicity_nodes <- tibble(
#     id = c("C", "I", "M"),
#     type = "ethnicity",
#     group = "ethnicity"
#   )
#   
#   food_nodes <- tibble(
#     id = food_categories,
#     type = "food",
#     group = "food"
#   )
#   
#   nutrition_nodes <- tibble(
#     id = nutrition_categories,
#     type = "nutrition",
#     group = "nutrition"
#   )
#   
#   species_nodes <- tibble(
#     id = species_with_connections,
#     type = "species",
#     group = "species"
#   )
#   
#   # Combine nodes
#   nodes <- bind_rows(
#     ethnicity_nodes,
#     food_nodes,
#     nutrition_nodes,
#     species_nodes
#   )
#   
#   # Create edges
#   cat("Creating edges...\n")
#   species_food_edges <- create_correlation_edges(species_food_results, "species_food")
#   species_nutrition_edges <- create_correlation_edges(species_nutrition_results, "species_nutrition")
#   
#   # Combine edges and add correlation direction column
#   edges <- bind_rows(
#     species_food_edges,
#     species_nutrition_edges
#   ) %>%
#     mutate(correlation_direction = ifelse(weight > 0, "positive", "negative"))
#   
#   # Create graph with absolute weights for layout
#   graph <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE) %>%
#     activate(edges) %>%
#     mutate(abs_weight = abs(weight))
#   
#   # Create visualization using absolute weights for layout but original weights for visualization
#   p <- ggraph(graph, layout = "fr", weights = abs_weight) +
#     geom_edge_link(aes(color = correlation_direction, 
#                       width = abs(weight),
#                       alpha = abs(weight))) +
#     geom_node_point(aes(color = group), size = 4) +
#     geom_node_text(aes(label = id), repel = TRUE, size = 2.5) +
#     scale_edge_width_continuous(range = c(0.2, 2)) +
#     scale_edge_color_manual(values = c("positive" = "blue", "negative" = "red")) +
#     scale_edge_alpha(range = c(0.2, 0.8)) +
#     theme_graph(base_family = "sans") +
#     scale_color_brewer(palette = "Set3") +
#     labs(title = "Microbiome Knowledge Graph",
#          subtitle = paste("Significant connections (p < 0.05) between\n",
#                          length(species_with_connections), 
#                          "species and food/nutrition variables\n",
#                          "Edge width represents correlation strength\n",
#                          "Blue = positive correlation, Red = negative correlation")) +
#     theme(
#       legend.position = "bottom",
#       legend.box = "vertical",
#       plot.margin = margin(20, 20, 20, 20)
#     )
#   
#   return(list(
#     graph = graph,
#     plot = p,
#     nodes = nodes,
#     edges = edges,
#     species_with_connections = species_with_connections,
#     results = list(
#       species_food = species_food_results,
#       species_nutrition = species_nutrition_results
#     )
#   ))
# }
```


```{r}
create_microbiome_knowledge_graph <- function(food_data, nutrition_data, 
                                            species_data, metadata,
                                            species_threshold = 0.3) {  # Add threshold parameter
  # Ensure all datasets have the same participants and in the same order
  common_participants <- Reduce(intersect, list(
    food_data$Participant_ID,
    nutrition_data$Participant_ID,
    species_data$Participant_ID
  ))
  
  # Filter and arrange all datasets
  food_data <- food_data %>%
    filter(Participant_ID %in% common_participants) %>%
    arrange(Participant_ID)
  
  nutrition_data <- nutrition_data %>%
    filter(Participant_ID %in% common_participants) %>%
    arrange(Participant_ID)
  
  species_data <- species_data %>%
    filter(Participant_ID %in% common_participants) %>%
    arrange(Participant_ID)
  
  # Get categories
  food_categories <- setdiff(colnames(food_data), "Participant_ID")
  nutrition_categories <- setdiff(colnames(nutrition_data), "Participant_ID")
  species_categories <- setdiff(colnames(species_data), "Participant_ID")
  
  # Calculate correlations with p-values using rcorr
  calculate_correlations <- function(data1, categories1, data2, categories2) {
    mat1 <- as.matrix(data1[, categories1])
    mat2 <- as.matrix(data2[, categories2])
    
    complete_cases <- complete.cases(mat1, mat2)
    if(!all(complete_cases)) {
      mat1 <- mat1[complete_cases, ]
      mat2 <- mat2[complete_cases, ]
    }
    
    rcorr_result <- rcorr(cbind(mat1, mat2), type = "pearson")
    
    n1 <- ncol(mat1)
    n2 <- ncol(mat2)
    cors <- rcorr_result$r[1:n1, (n1+1):(n1+n2)]
    p_values <- rcorr_result$P[1:n1, (n1+1):(n1+n2)]
    
    rownames(cors) <- categories1
    colnames(cors) <- categories2
    rownames(p_values) <- categories1
    colnames(p_values) <- categories2
    
    return(list(correlations = cors, p_values = p_values))
  }
  
  species_food_results <- calculate_correlations(
    species_data, species_categories,
    food_data, food_categories
  )
  
  species_nutrition_results <- calculate_correlations(
    species_data, species_categories,
    nutrition_data, nutrition_categories
  )
  
  # Create edges from correlations
  create_correlation_edges <- function(cor_results, relationship_type) {
    cors <- cor_results$correlations
    p_values <- cor_results$p_values
    edges <- which(p_values <= 0.01 & abs(cors) >= species_threshold, arr.ind = TRUE)  # Add threshold filter
    
    if(nrow(edges) > 0) {
      tibble(
        from = rownames(cors)[edges[,1]],
        to = colnames(cors)[edges[,2]],
        relationship = relationship_type,
        weight = cors[cbind(edges[,1], edges[,2])],
        p_value = p_values[cbind(edges[,1], edges[,2])]
      )
    } else {
      tibble(from = character(), to = character(), 
             relationship = character(), weight = numeric(),
             p_value = numeric())
    }
  }
  
  # Get species that have significant strong connections
  species_with_connections <- unique(c(
    rownames(species_food_results$correlations)[rowSums(species_food_results$p_values <= 0.01 & 
                                                        abs(species_food_results$correlations) >= species_threshold) > 0],
    rownames(species_nutrition_results$correlations)[rowSums(species_nutrition_results$p_values <= 0.01 & 
                                                             abs(species_nutrition_results$correlations) >= species_threshold) > 0]
  ))
  
  # Create nodes with custom colors
  ethnicity_nodes <- tibble(
    id = c("C", "I", "M"),
    type = "ethnicity",
    group = "ethnicity",
    color = "darkviolet"  # Gray
  )
  
  food_nodes <- tibble(
    id = food_categories,
    type = "food",
    group = "food",
    color = "#FF6B6B"  # Coral red
  )
  
  nutrition_nodes <- tibble(
    id = nutrition_categories,
    type = "nutrition", 
    group = "nutrition",
    color = "#4ECDC4"  # Turquoise
  )
  
  species_nodes <- tibble(
    id = species_with_connections,
    type = "species",
    group = "species",
    color = "#95A5A6"  # Light gray
  )
  
  # Combine nodes
  nodes <- bind_rows(
    ethnicity_nodes,
    food_nodes,
    nutrition_nodes,
    species_nodes
  )
  
  # Create edges
  species_food_edges <- create_correlation_edges(species_food_results, "species_food")
  species_nutrition_edges <- create_correlation_edges(species_nutrition_results, "species_nutrition")
  
  edges <- bind_rows(
    species_food_edges,
    species_nutrition_edges
  ) %>%
    mutate(correlation_direction = ifelse(weight > 0, "positive", "negative"))
  
  # Create graph
  graph <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE) %>%
    activate(edges) %>%
    mutate(abs_weight = abs(weight))
  
  # Create visualization with custom node colors
  p <- ggraph(graph, layout = "fr", weights = abs_weight) +
    geom_edge_link(aes(color = correlation_direction, 
                      width = abs(weight),
                      alpha = abs(weight))) +
    geom_node_point(aes(color = color), size = 4) +
    geom_node_text(aes(label = id, 
                   size = group),  # Map size to node group
               repel = TRUE) +
    scale_size_manual(values = c(
      "food" = 6.5,        # Larger font for food nodes
      "nutrition" = 6.5,   # Medium font for nutrition nodes
      "species" = 4,     # Smaller font for species nodes
      "ethnicity" = 6.5    # Larger font for ethnicity nodes
      )) +
    scale_edge_width_continuous(range = c(0.2, 2)) +
    scale_edge_color_manual(values = c("positive" = "salmon2", "negative" = "steelblue2")) +
    scale_edge_alpha(range = c(0.2, 0.8)) +
    scale_color_identity() +  # Use the actual colors specified in nodes
    theme_graph(base_family = "sans") +
    labs(title = "Microbiome Knowledge Graph",
         subtitle = paste("Significant connections (p < 0.05, |correlation| >=", species_threshold, ") between\n",
                         length(species_with_connections), 
                         "species and food/nutrition variables\n",
                         "Edge width represents correlation strength\n",
                         "Red = positive correlation, Blue = negative correlation")) +
    theme(
      legend.position = "bottom",
      legend.box = "vertical",
      plot.margin = margin(20, 20, 20, 20)
    )
  
  return(list(
    graph = graph,
    plot = p,
    nodes = nodes,
    edges = edges,
    species_with_connections = species_with_connections,
    results = list(
      species_food = species_food_results,
      species_nutrition = species_nutrition_results
    )
  ))
}
```


```{r fig.width=14, fig.height=14}
# result <- create_microbiome_knowledge_graph(food_data, nutrition_data, 
#                                           species_filtered1, metadata)

result = create_microbiome_knowledge_graph(food_data, nutrition_data, 
                                                species_filtered1, metadata,
                                                species_threshold = 0.15)
# View the plot
print(result$plot)

# Save the plot
ggsave("figure/microbiome_graph3.pdf", result$plot, width = 24, height = 20)

# Check correlation summary
cat("\nNumber of significant correlations:", nrow(result$edges))
cat("\nCorrelation strength summary:\n")
print(summary(abs(result$edges$weight)))
```






## put all together
```{r}

#nutrition_data <- data.frame(lapply(nutrition_data, function(x) as.numeric(as.character(x))))
full_data <- merge(food_data, species_filtered2, by = "Participant_ID")
full_data <- merge(full_data, nutrition_data, by="Participant_ID")
food_data
nutrition_data
full_data
```

```{r}
nutrition_columns <- names(nutrition_data)[-1]  
species_columns <- names(species_filtered2)[-1] 
food_columns <- names(food_data)[-1]
nutrition_columns
species_columns
food_columns
```

# all ethnicity (no need for now)
```{r}
food_species_data <- full_data[, c(food_columns, species_columns)]
nutrition_species_data <- full_data[, c(nutrition_columns, species_columns)]

calculate_significant_correlations <- function(data, food_or_nutrition_columns, species_columns) {
  cor_results <- rcorr(as.matrix(data), type = "spearman")
  adjusted_p_values <- matrix(p.adjust(cor_results$P, method = "BH"), nrow = nrow(cor_results$P))
  significant_indices <- which(adjusted_p_values < 0.01, arr.ind = TRUE)
  
  significant_correlations <- data.frame(
    from = colnames(data)[significant_indices[, "row"]],
    to = colnames(data)[significant_indices[, "col"]],
    correlation = cor_results$r[significant_indices],
    adjusted_p_value = adjusted_p_values[significant_indices]
  )
  
  significant_correlations <- significant_correlations[
    significant_correlations$from %in% food_or_nutrition_columns & significant_correlations$to %in% species_columns |
    significant_correlations$from %in% species_columns & significant_correlations$to %in% food_or_nutrition_columns, ]
  
  return(significant_correlations)
}
```


```{r}

significant_food_species_correlations <- calculate_significant_correlations(food_species_data, food_columns, species_columns)
significant_nutrition_species_correlations <- calculate_significant_correlations(nutrition_species_data, nutrition_columns, species_columns)

print(significant_food_species_correlations)
print(significant_nutrition_species_correlations)

```

```{r}
combine_t = rbind(significant_food_species_correlations, significant_nutrition_species_correlations)

combine_t  <- combine_t  %>%
  mutate(
    node1 = pmin(from, to),
    node2 = pmax(from, to)
  ) %>%
  distinct(node1, node2, .keep_all = TRUE)  # keep one of each unique pair


combine_t
```
```{r}
cor_data <- combine_t  %>%
  mutate(edge_color = ifelse(correlation > 0, "red", "cadetblue2"),
         abs_correlation = abs(correlation))

# Add a column for node shapes
# cor_data <- cor_data %>%
#   mutate(from_shape = ifelse(from %in% food_columns, "square", ifelse(from %in% nutrition_columns, "triangle", "circle")),
#          to_shape = ifelse(to %in% food_columns, "square", ifelse(to %in% nutrition_columns, "triangle", "circle")))
cor_data
```
```{r}
nodes <- unique(c(cor_data$from, cor_data$to))

node_info <- data.frame(
  name = nodes,
  type = ifelse(nodes %in% species_columns, "Species",
                ifelse(nodes %in% food_columns, "Food", "Nutrition"))
)

node_info$name <- as.character(node_info$name)
node_info
```
```{r}
edge_list <- cor_data %>% select(from, to, edge_color, abs_correlation)
edge_list
graph <- graph_from_data_frame(d = edge_list, vertices = node_info, directed = FALSE)

V(graph)$shape <- ifelse(V(graph)$type == "Species", "circle",
                         ifelse(V(graph)$type == "Food", "square", "diamond"))

V(graph)$degree <- degree(graph, mode = "all")
#plot(graph)
node_info
```


```{r fig.width=14, fig.height=14}
full_graph = ggraph(graph, layout = "fr") +  # 'fr' can be changed to another layout if needed
  geom_edge_link(aes(edge_alpha = abs_correlation, edge_width = abs_correlation, edge_color = edge_color), show.legend = FALSE) +
  geom_node_point(aes(shape = shape, size = degree , color = shape)) +
  geom_node_text(aes(label = name), repel = TRUE, fontface = "bold", color = "black", size = 3) +
  scale_size(range = c(4, 10)) + 
  scale_shape_manual(values = c(circle = 19, square = 15, diamond = 17), labels = c("Species", "Nutrition", "Food")) +
  scale_edge_color_manual(values = c("#8D3256", "#004697")) +
  scale_color_manual(values = c("slateblue", "#68B69F", "#F08E59"), labels = c("Species", "Nutrition", "Food")) +
  theme_graph() +
  theme(legend.position = "right") +
  ggtitle("Correlation Network between Species, Food, and Nutrition")
```


```{r}
png("figure/network_all.png", width = 1400, height = 1400, res = 120)
full_graph
dev.off()
```



# seperate ethnic groups

```{r}
new_full <- merge(metadata[,c("Participant_ID","FREG5_Ethnic_Group", "FREG8_Age", "FREG7_Gender", "BMI", "BMI_groups")],full_data, by="Participant_ID")
df_cm <- new_full %>% filter(FREG5_Ethnic_Group == "C" | FREG5_Ethnic_Group == "M")
df_im <- new_full %>% filter(FREG5_Ethnic_Group == "I" | FREG5_Ethnic_Group == "M")
# data_c <- new_full %>% filter(FREG5_Ethnic_Group == "C")
# data_i <- new_full %>% filter(FREG5_Ethnic_Group == "I")
# data_m <- new_full %>% filter(FREG5_Ethnic_Group == "M")
df_cm$groupM_C <- ifelse(df_cm$FREG5_Ethnic_Group == "M", 1,
                      ifelse(df_cm$FREG5_Ethnic_Group == "C", 0, NA))
df_im$groupM_I <- ifelse(df_im$FREG5_Ethnic_Group == "M", 1,
                      ifelse(df_im$FREG5_Ethnic_Group == "I", 0, NA))
# data_c <- new_full %>% filter(FREG5_Ethnic_Group == "C")
# data_i <- new_full %>% filter(FREG5_Ethnic_Group == "I")
# data_m <- new_full %>% filter(FREG5_Ethnic_Group == "M")
new_full
full_data
```

```{r}
library(MatchIt)
#library(tableone)

m.out_C <- matchit(groupM_C ~ FREG7_Gender + FREG8_Age + BMI,
                   data = df_cm, method = "nearest", ratio = 1,
                   exact = c("FREG7_Gender", "BMI_groups"),
                   caliper = 0.1)
summary(m.out_C)
matched_data_C <- match.data(m.out_C)
#matched_data_C
```


```{r}
m.out_I <- matchit(groupM_I ~ FREG7_Gender + FREG8_Age + BMI,
                   data = df_im, method = "nearest", ratio = 1,
                   exact = c("FREG7_Gender", "BMI_groups"),
                   caliper = 0.3)
summary(m.out_I)
matched_data_I <- match.data(m.out_I)
matched_data_I
```

```{r}
data_c <- matched_data_C %>% filter(FREG5_Ethnic_Group == "C")
data_i <- matched_data_I %>% filter(FREG5_Ethnic_Group == "I")
data_m <- matched_data_C %>% filter(FREG5_Ethnic_Group == "M")
```

```{r}
# data_c[is.na(data_c)] <- 0
# data_i[is.na(data_i)] <- 0
# data_m[is.na(data_m)] <- 0
# data_c <- data.frame(lapply(data_c, function(x) as.numeric(as.character(x))))
# data_i <- data.frame(lapply(data_i, function(x) as.numeric(as.character(x))))
# data_m <- data.frame(lapply(data_m, function(x) as.numeric(as.character(x))))

data_c
data_i
data_m
```


# unweighted
```{r}
calculate_significant_correlations <- function(data, food_or_nutrition_columns, species_columns) {
  cor_results <- rcorr(as.matrix(data), type = "spearman")
  adjusted_p_values <- matrix(p.adjust(cor_results$P, method = "BH"), nrow = nrow(cor_results$P))
  significant_indices <- which(adjusted_p_values < 0.02, arr.ind = TRUE)

  significant_correlations <- data.frame(
    from = colnames(data)[significant_indices[, "row"]],
    to = colnames(data)[significant_indices[, "col"]],
    correlation = cor_results$r[significant_indices],
    adjusted_p_value = adjusted_p_values[significant_indices]
  )

  significant_correlations <- significant_correlations[
    significant_correlations$from %in% food_or_nutrition_columns & significant_correlations$to %in% species_columns |
    significant_correlations$from %in% species_columns & significant_correlations$to %in% food_or_nutrition_columns, ]

  return(significant_correlations)}

```
# weighted
```{r}
# library(wCorr)  # Ensure this package is loaded
# 
# calculate_significant_correlations <- function(data, food_or_nutrition_columns, species_columns, weights) {
#   # Extract the relevant columns for the analysis
#   analysis_data <- data[, c(food_or_nutrition_columns, species_columns)]
#   
#   # Convert to numeric matrix if not already
#   analysis_data[] <- lapply(analysis_data, function(x) as.numeric(as.character(x)))
#   
#   # Prepare empty matrices for correlation and p-values
#   correlation_matrix <- matrix(0, nrow = ncol(analysis_data), ncol = ncol(analysis_data))
#   p_value_matrix <- matrix(0, nrow = ncol(analysis_data), ncol = ncol(analysis_data))
#   
#   # Calculate weighted correlations for each pair of columns
#   for (i in 1:ncol(analysis_data)) {
#     for (j in i:ncol(analysis_data)) {
#       if (i != j) {
#         # Assuming weightedCorr returns a list with 'estimate' and 'p.value'
#         corr_result <- weightedCorr(analysis_data[, i], analysis_data[, j], method = "Spearman", weights = weights)
#         correlation_matrix[i, j] <- corr_result$estimate
#         correlation_matrix[j, i] <- corr_result$estimate  # Symmetric matrix
#         p_value_matrix[i, j] <- corr_result$p.value
#         p_value_matrix[j, i] <- corr_result$p.value  # Symmetric matrix
#       } else {
#         correlation_matrix[i, j] <- 1  # Diagonal elements are perfect correlations
#         p_value_matrix[i, j] <- 0  # Diagonal elements p-value of correlation with itself
#       }
#     }
#   }
#   
#   # Adjust p-values using BH method
#   adjusted_p_values <- p.adjust(as.vector(p_value_matrix), method = "BH")
#   adjusted_p_matrix <- matrix(adjusted_p_values, nrow = ncol(analysis_data))
#   
#   # Identify significant correlations
#   significant_indices <- which(adjusted_p_matrix < 0.05, arr.ind = TRUE)
#   
#   # Create a dataframe of significant correlations
#   significant_correlations <- data.frame(
#     from = colnames(analysis_data)[significant_indices[, 1]],
#     to = colnames(analysis_data)[significant_indices[, 2]],
#     correlation = correlation_matrix[significant_indices],
#     adjusted_p_value = adjusted_p_matrix[significant_indices]
#   )
#   
#   # Filter to include only food-species interactions
#   significant_correlations <- significant_correlations[
#     (significant_correlations$from %in% food_or_nutrition_columns & significant_correlations$to %in% species_columns) |
#     (significant_correlations$from %in% species_columns & significant_correlations$to %in% food_or_nutrition_columns), ]
#   
#   return(significant_correlations)
# }

```

# chinese
```{r}
# set.seed(123) 
# 
# selected_indices <- sample(nrow(data_c), 49)
# 
# # Subset the DataFrame to get the selected rows
# data_c_49 <- data_c[selected_indices, ]
# data_c_49
data_c
```



```{r}
food_species_c <- data_c[, c(food_columns, species_columns)]

nutrition_species_c <- data_c[, c(nutrition_columns, species_columns)]

food_species_correlations_c <- calculate_significant_correlations(food_species_c, food_columns, species_columns)
nutrition_species_correlations_c <- calculate_significant_correlations(nutrition_species_c, nutrition_columns, species_columns)
```
```{r}
combine_c = rbind(food_species_correlations_c, nutrition_species_correlations_c)

combine_c  <- combine_c  %>%
  mutate(
    node1 = pmin(from, to),
    node2 = pmax(from, to)
  ) %>%
  distinct(node1, node2, .keep_all = TRUE)  # keep one of each unique pair
combine_c
```
```{r}
cor_data_c <- combine_c  %>%
  mutate(edge_color = ifelse(correlation > 0, "red", "blue"),
         abs_correlation = abs(correlation),
         log_p = -log10(adjusted_p_value))


cor_data_c
#write_xlsx(cor_data_c, "figure/network_table_c.xlsx")
```

# indian
```{r}
# set.seed(123) 
# 
# selected_indices <- sample(nrow(data_i), 49)
# 
# # Subset the DataFrame to get the selected rows
# data_i_49 <- data_c[selected_indices, ]
# data_i_49
```


```{r}
food_species_i <- data_i[, c(food_columns, species_columns)]
nutrition_species_i <- data_i[, c(nutrition_columns, species_columns)]


food_species_correlations_i <- calculate_significant_correlations(food_species_i, food_columns, species_columns)
nutrition_species_correlations_i <- calculate_significant_correlations(nutrition_species_i, nutrition_columns, species_columns)
```
```{r}
combine_i = rbind(food_species_correlations_i, nutrition_species_correlations_i)

combine_i  <- combine_i  %>%
  mutate(
    node1 = pmin(from, to),
    node2 = pmax(from, to)
  ) %>%
  distinct(node1, node2, .keep_all = TRUE)  # keep one of each unique pair
combine_i
```
```{r}
cor_data_i <- combine_i  %>%
  mutate(edge_color = ifelse(correlation > 0, "red", "blue"),
         abs_correlation = abs(correlation),
         log_p = -log10(adjusted_p_value))


cor_data_i
#write_xlsx(cor_data_i, "figure/network_table_i.xlsx")
```

# malays
```{r}
food_species_m <- data_m[, c(food_columns, species_columns)]
nutrition_species_m <- data_m[, c(nutrition_columns, species_columns)]


food_species_correlations_m <- calculate_significant_correlations(food_species_m, food_columns, species_columns)
nutrition_species_correlations_m <- calculate_significant_correlations(nutrition_species_m, nutrition_columns, species_columns)
```

```{r}
combine_m = rbind(food_species_correlations_m, nutrition_species_correlations_m)

combine_m  <- combine_m  %>%
  mutate(
    node1 = pmin(from, to),
    node2 = pmax(from, to)
  ) %>%
  distinct(node1, node2, .keep_all = TRUE)  # keep one of each unique pair
combine_m
```
```{r}
cor_data_m <- combine_m  %>%
  mutate(edge_color = ifelse(correlation > 0, "red", "blue"),
         abs_correlation = abs(correlation),
         log_p = -log10(adjusted_p_value))


cor_data_m
#write_xlsx(cor_data_m, "figure/network_table_m.xlsx")
```

## power analysis
```{r}
# library(pwr)
# r_observed <- 0.1 
# result <- pwr.r.test(n = NULL, r = r_observed, sig.level = 0.05, power = 0.80, alternative = "two.sided")
# print(result)

```

## try circuler layout
```{r}
# Assuming cor_data_c, cor_data_i, cor_data_m are data frames with columns: from, to, edge_color, abs_correlation

all_nodes <- unique(c(cor_data_c$from, cor_data_c$to, cor_data_i$from, cor_data_i$to, cor_data_m$from, cor_data_m$to))

# Create a nodes data frame
nodes_df <- data.frame(name = unique(all_nodes))
nodes_df$shape <- ifelse(nodes_df$name %in% species_columns, "circle", 
                         ifelse(nodes_df$name %in% food_columns, "square", "diamond"))
nodes_df$color <- ifelse(nodes_df$name %in% species_columns, "Species", 
                         ifelse(nodes_df$name %in% food_columns, "Food", "Nutrition"))

# Create a layout with food/nutrition in the inner circle and species in the outer circle
food_nodes <- nodes_df$name[nodes_df$name %in% food_columns]
nutrition_nodes <- nodes_df$name[nodes_df$name %in% nutrition_columns]
outer_circle_nodes <- nodes_df$name[nodes_df$name %in% species_columns]

inner_circle_positions_food <- data.frame(
  name = food_nodes,
  x = 0.5 * cos(seq(0, pi, length.out = length(food_nodes) + 1)[-length(food_nodes)-1]),
  y = 0.5 * sin(seq(0, pi, length.out = length(food_nodes) + 1)[-length(food_nodes)-1])
)

inner_circle_positions_nutrition <- data.frame(
  name = nutrition_nodes,
  x = 0.5 * cos(seq(pi, 2*pi, length.out = length(nutrition_nodes) + 1)[-length(nutrition_nodes)-1]),
  y = 0.5 * sin(seq(pi, 2*pi, length.out = length(nutrition_nodes) + 1)[-length(nutrition_nodes)-1])
)

inner_circle_positions <- rbind(inner_circle_positions_food, inner_circle_positions_nutrition)

outer_circle_positions <- data.frame(
  name = outer_circle_nodes,
  x = 1 * cos(seq(0, 2*pi, length.out = length(outer_circle_nodes) + 1)[-length(outer_circle_nodes)-1]),
  y = 1 * sin(seq(0, 2*pi, length.out = length(outer_circle_nodes) + 1)[-length(outer_circle_nodes)-1])
)

layout_positions <- rbind(inner_circle_positions, outer_circle_positions)
```


```{r fig.height=4, fig.width=6}
# Function to plot graph with fixed layout
library(scales)
library(tidygraph)
palette_colors <- brewer_pal(palette = "RdBu")(11)
red_color <- palette_colors[10]
blue_color <- palette_colors[2]

plot_graph <- function(graph, layout_positions) {
  graph <- graph %>%
    activate(nodes) %>%
    mutate(x = layout_positions$x[match(name, layout_positions$name)],
           y = layout_positions$y[match(name, layout_positions$name)],
           angle = atan2(y, x) * 180 / pi,
           hjust = ifelse(angle < -90 | angle > 90, 1, 0),
           angle = ifelse(angle < -90 | angle > 90, angle + 180, angle)) # Flip text
  
  ggraph(graph, layout = 'manual', x = x, y = y) +
    geom_edge_diagonal(aes(edge_alpha = abs_correlation, edge_color = edge_color, edge_width = abs_correlation), show.legend = FALSE) +
    geom_node_point(aes(shape = shape, color = color), size = 2) +
    geom_node_text(aes(label = name, angle = angle, hjust = hjust), size = 2) +
    scale_shape_manual(values = c(circle = 16, square = 15, diamond = 18), labels = c("Species", "Nutrition", "Food")) +  # Mapping shapes to actual ggplot2 shapes
    scale_color_manual(values = c("Species" = "#F89FA8", "Nutrition" = "#68B69F", "Food" = "#F08E59"), labels = c("Species", "Nutrition", "Food")) +
    scale_edge_color_manual(values = c(red_color, blue_color)) +
    #theme_void() +  # Remove background elements
    theme_graph() +
    theme(legend.position = "right") +
    coord_fixed(ratio = 1) +  # Ensure aspect ratio is fixed
    expand_limits(x = c(-1.7, 1.7), y = c(-1.7, 1.7))
}

# Create graph objects from edge lists and nodes
graph_c <- tbl_graph(edges = cor_data_c, nodes = nodes_df, directed = FALSE)
graph_i <- tbl_graph(edges = cor_data_i, nodes = nodes_df, directed = FALSE)
graph_m <- tbl_graph(edges = cor_data_m, nodes = nodes_df, directed = FALSE)

# Plot each graph using the same layout positions
plot_c <- plot_graph(graph_c, layout_positions)
plot_i <- plot_graph(graph_i, layout_positions)
plot_m <- plot_graph(graph_m, layout_positions)

# Print plots
print(plot_c)
print(plot_i)
print(plot_m)

```


```{r}
png("figure/network_c_circle800_2.png", width = 1600, height = 1400, res = 120)
plot_c
dev.off()
png("figure/network_i_circle800_2.png", width = 1600, height = 1400, res = 120)
plot_i
dev.off()
png("figure/network_m_circle800_2.png", width = 1600, height = 1400, res = 120)
plot_m
dev.off()
```

# permutaion test
```{r}
permutation_test_spearman <- function(data, x_cols, y_cols, n_perm = 999) {
  # Calculate observed Spearman correlations
  observed_cor <- cor(data[x_cols], data[y_cols], method = "spearman")
  
  # Perform permutation test
  perm_result <- numeric(n_perm)
  for (i in 1:n_perm) {
    perm_data <- data
    perm_data[y_cols] <- perm_data[sample(nrow(perm_data)), y_cols]
    perm_cor <- cor(perm_data[x_cols], perm_data[y_cols], method = "spearman")
    perm_result[i] <- max(abs(perm_cor))
  }
  
  # Calculate p-values
  p_values <- matrix(sapply(abs(observed_cor), function(x) mean(perm_result >= x)), 
                     nrow = nrow(observed_cor), ncol = ncol(observed_cor))
  
  return(list(observed_cor = observed_cor, p_values = p_values))
}

# Function to format and filter results
format_results <- function(test_result, x_cols, y_cols) {
  significant_results <- data.frame(
    X_Variable = rep(x_cols, each = length(y_cols)),
    Y_Variable = rep(y_cols, times = length(x_cols)),
    Spearman_Correlation = as.vector(test_result$observed_cor),
    P_value = as.vector(test_result$p_values)
  ) %>%
    filter(P_value < 0.2) %>%
    arrange(P_value)
  
  return(significant_results)
}


food_species_test <- permutation_test_spearman(new_full, food_columns, species_columns)
nutrition_species_test <- permutation_test_spearman(new_full, nutrition_columns, species_columns)

# Format and print results
cat("Significant Food-Species Spearman Correlations (p < 0.05):\n")
print(format_results(food_species_test, food_columns, species_columns))

cat("\nSignificant Nutrition-Species Spearman Correlations (p < 0.05):\n")
print(format_results(nutrition_species_test, nutrition_columns, species_columns))

# Function to plot correlation distribution
plot_correlation_distribution <- function(test_result, title) {
  correlations <- as.vector(test_result$observed_cor)
  hist(correlations, main = title, xlab = "Spearman Correlation Coefficient", 
       breaks = 20, col = "lightblue", border = "black")
  abline(v = 0, col = "red", lwd = 2, lty = 2)
}

# Plot correlation distributions
par(mfrow = c(1, 2))
plot_correlation_distribution(food_species_test, "Food-Species Correlations")
plot_correlation_distribution(nutrition_species_test, "Nutrition-Species Correlations")
```

# only permutation test significance
```{r}
food_spe = format_results(food_species_test, food_columns, species_columns)$Y_Variable
food_spe
food_perm = format_results(food_species_test, food_columns, species_columns)$X_Variable
food_perm
nutri_spe = format_results(nutrition_species_test, nutrition_columns, species_columns)$Y_Variable
nutri_spe
nutri_perm = format_results(nutrition_species_test, nutrition_columns, species_columns)$X_Variable
nutri_perm
```


# only have common food AND nutrient network
# chinese
```{r}
calculate_significant_correlations <- function(data, food_or_nutrition_columns, species_columns) {
  cor_results <- rcorr(as.matrix(data), type = "spearman")
  adjusted_p_values <- matrix(p.adjust(cor_results$P, method = "BH"), nrow = nrow(cor_results$P))
  significant_indices <- which(adjusted_p_values < 0.05, arr.ind = TRUE)

  significant_correlations <- data.frame(
    from = colnames(data)[significant_indices[, "row"]],
    to = colnames(data)[significant_indices[, "col"]],
    correlation = cor_results$r[significant_indices],
    adjusted_p_value = adjusted_p_values[significant_indices]
  )

  significant_correlations <- significant_correlations[
    significant_correlations$from %in% food_or_nutrition_columns & significant_correlations$to %in% species_columns |
    significant_correlations$from %in% species_columns & significant_correlations$to %in% food_or_nutrition_columns, ]

  return(significant_correlations)}

```

```{r}
#food_columns
#nutrition_columns
# common_food = c("Fish","Dairy", "Fruits", "Idli_Thosai", "Deep_fried","Tea")
# common_nutrition = c("Energy", "Protein", "Total_fat", "Saturated_fatty_acid", "MUFA", "PUFA", "fibre", "Iron", "Thiamin", "Riboflavin", "Potassium",  "Zinc")
# common_spec = species_columns
common_food = c("Dairy", "Cruciferous_veg", "Papaya", "Orange", "Tea", "Fish", "Legumes", "Fruits", "Wholegrain", "Deep_fried", "Poultry", 
                "Red_meat", "Processed_meat", "Noodle_dishes","Coffee")
common_nutrition = c("Protein","Total_fat", "MUFA","PUFA","CHO","starch","fibre","Calcium","Iron", "Thiamin","Riboflavin","Potassium", 
                     "sugar","Zinc", "Energy","Saturated_fatty_acid","VitaminA","VitaminC" )
common_spec = food_spe
```




```{r}

# Reduce(intersect, list(cor_data_c$to, cor_data_i$to, cor_data_m$to))
# Reduce(union, list(cor_data_c$to, cor_data_i$to, cor_data_m$to))
```

```{r}
food_species_c <- data_c[, c(common_food, common_spec)]

nutrition_species_c <- data_c[, c(common_nutrition, common_spec)]

food_species_correlations_c <- calculate_significant_correlations(food_species_c, common_food, common_spec)
nutrition_species_correlations_c <- calculate_significant_correlations(nutrition_species_c, common_nutrition, common_spec)
```
```{r}
combine_c = rbind(food_species_correlations_c, nutrition_species_correlations_c)

combine_c  <- combine_c  %>%
  mutate(
    node1 = pmin(from, to),
    node2 = pmax(from, to)
  ) %>%
  distinct(node1, node2, .keep_all = TRUE)  # keep one of each unique pair
combine_c
```
```{r}
cor_data_c <- combine_c  %>%
  mutate(edge_color = ifelse(correlation > 0, "red", "blue"),
         abs_correlation = abs(correlation),
         log_p = -log10(adjusted_p_value))


cor_data_c
#write_xlsx(cor_data_c, "figure/network_table_c.xlsx")
```


# indian


```{r}
food_species_i <- data_i[, c(common_food, common_spec)]
nutrition_species_i <- data_i[, c(common_nutrition, common_spec)]


food_species_correlations_i <- calculate_significant_correlations(food_species_i, common_food, common_spec)
nutrition_species_correlations_i <- calculate_significant_correlations(nutrition_species_i, common_nutrition, common_spec)
```
```{r}
combine_i = rbind(food_species_correlations_i, nutrition_species_correlations_i)

combine_i  <- combine_i  %>%
  mutate(
    node1 = pmin(from, to),
    node2 = pmax(from, to)
  ) %>%
  distinct(node1, node2, .keep_all = TRUE)  # keep one of each unique pair
combine_i
```
```{r}
cor_data_i <- combine_i  %>%
  mutate(edge_color = ifelse(correlation > 0, "red", "blue"),
         abs_correlation = abs(correlation),
         log_p = -log10(adjusted_p_value))


cor_data_i
#write_xlsx(cor_data_i, "figure/network_table_i.xlsx")
```

# malays
```{r}
food_species_m <- data_m[, c(common_food, common_spec)]
nutrition_species_m <- data_m[, c(common_nutrition, common_spec)]


food_species_correlations_m <- calculate_significant_correlations(food_species_m, common_food, common_spec)
nutrition_species_correlations_m <- calculate_significant_correlations(nutrition_species_m, common_nutrition, common_spec)
```

```{r}
combine_m = rbind(food_species_correlations_m, nutrition_species_correlations_m)

combine_m  <- combine_m  %>%
  mutate(
    node1 = pmin(from, to),
    node2 = pmax(from, to)
  ) %>%
  distinct(node1, node2, .keep_all = TRUE)  # keep one of each unique pair
combine_m
```
```{r}
cor_data_m <- combine_m  %>%
  mutate(edge_color = ifelse(correlation > 0, "red", "blue"),
         abs_correlation = abs(correlation),
         log_p = -log10(adjusted_p_value))


cor_data_m
#write_xlsx(cor_data_m, "figure/network_table_m.xlsx")
```


## try circuler layout
```{r}
# Assuming cor_data_c, cor_data_i, cor_data_m are data frames with columns: from, to, edge_color, abs_correlation

all_nodes <- unique(c(cor_data_c$from, cor_data_c$to, cor_data_i$from, cor_data_i$to, cor_data_m$from, cor_data_m$to))

# Create a nodes data frame
nodes_df <- data.frame(name = unique(all_nodes))
nodes_df$shape <- ifelse(nodes_df$name %in% species_columns, "circle", 
                         ifelse(nodes_df$name %in% food_columns, "square", "diamond"))
nodes_df$color <- ifelse(nodes_df$name %in% species_columns, "Species", 
                         ifelse(nodes_df$name %in% food_columns, "Food", "Nutrition"))

# Create a layout with food/nutrition in the inner circle and species in the outer circle
food_nodes <- nodes_df$name[nodes_df$name %in% food_columns]
nutrition_nodes <- nodes_df$name[nodes_df$name %in% nutrition_columns]
outer_circle_nodes <- nodes_df$name[nodes_df$name %in% species_columns]

inner_circle_positions_food <- data.frame(
  name = food_nodes,
  x = 0.5 * cos(seq(0, pi, length.out = length(food_nodes) + 1)[-length(food_nodes)-1]),
  y = 0.5 * sin(seq(0, pi, length.out = length(food_nodes) + 1)[-length(food_nodes)-1])
)

inner_circle_positions_nutrition <- data.frame(
  name = nutrition_nodes,
  x = 0.5 * cos(seq(pi, 2*pi, length.out = length(nutrition_nodes) + 1)[-length(nutrition_nodes)-1]),
  y = 0.5 * sin(seq(pi, 2*pi, length.out = length(nutrition_nodes) + 1)[-length(nutrition_nodes)-1])
)

inner_circle_positions <- rbind(inner_circle_positions_food, inner_circle_positions_nutrition)

outer_circle_positions <- data.frame(
  name = outer_circle_nodes,
  x = 1 * cos(seq(0, 2*pi, length.out = length(outer_circle_nodes) + 1)[-length(outer_circle_nodes)-1]),
  y = 1 * sin(seq(0, 2*pi, length.out = length(outer_circle_nodes) + 1)[-length(outer_circle_nodes)-1])
)

layout_positions <- rbind(inner_circle_positions, outer_circle_positions)
```


```{r fig.height=4, fig.width=6}
# Function to plot graph with fixed layout
library(scales)
library(tidygraph)
palette_colors <- brewer_pal(palette = "RdBu")(11)
red_color <- palette_colors[10]
blue_color <- palette_colors[2]

plot_graph <- function(graph, layout_positions) {
  graph <- graph %>%
    activate(nodes) %>%
    mutate(x = layout_positions$x[match(name, layout_positions$name)],
           y = layout_positions$y[match(name, layout_positions$name)],
           angle = atan2(y, x) * 180 / pi,
           hjust = ifelse(angle < -90 | angle > 90, 1, 0),
           angle = ifelse(angle < -90 | angle > 90, angle + 180, angle)) # Flip text
  
  ggraph(graph, layout = 'manual', x = x, y = y) +
    geom_edge_diagonal(aes(edge_alpha = abs_correlation, edge_color = edge_color, edge_width = abs_correlation), show.legend = FALSE) +
    geom_node_point(aes(shape = shape, color = color), size = 2) +
    geom_node_text(aes(label = name, angle = angle, hjust = hjust), size = 3) +
    scale_shape_manual(values = c(circle = 16, square = 15, diamond = 18), labels = c("Species", "Nutrition", "Food")) +  # Mapping shapes to actual ggplot2 shapes
    scale_color_manual(values = c("Species" = "#F89FA8", "Nutrition" = "#68B69F", "Food" = "#F08E59"), labels = c("Species", "Nutrition", "Food")) +
    scale_edge_color_manual(values = c(red_color, blue_color)) +
    #theme_void() +  # Remove background elements
    theme_graph() +
    theme(legend.position = "right") +
    coord_fixed(ratio = 1) +  # Ensure aspect ratio is fixed
    expand_limits(x = c(-1.7, 1.7), y = c(-1.7, 1.7))
}

# Create graph objects from edge lists and nodes
graph_c <- tbl_graph(edges = cor_data_c, nodes = nodes_df, directed = FALSE)
graph_i <- tbl_graph(edges = cor_data_i, nodes = nodes_df, directed = FALSE)
graph_m <- tbl_graph(edges = cor_data_m, nodes = nodes_df, directed = FALSE)

# Plot each graph using the same layout positions
plot_c <- plot_graph(graph_c, layout_positions)
plot_i <- plot_graph(graph_i, layout_positions)
plot_m <- plot_graph(graph_m, layout_positions)

# Print plots
print(plot_c)
print(plot_i)
print(plot_m)

```
```{r}
png("figure/permu_network_c_circle800.png", width = 1600, height = 1400, res = 120)
plot_c
dev.off()
png("figure/permu_network_i_circle800.png", width = 1600, height = 1400, res = 120)
plot_i
dev.off()
png("figure/permu_network_m_circle800.png", width = 1600, height = 1400, res = 120)
plot_m
dev.off()
```

# find out the nodes we need


```{r}
library(ggraph)
library(tidygraph)

# Assuming cor_data_c, cor_data_i, cor_data_m are data frames with columns: from, to, edge_color, log_p

all_nodes <- unique(c(cor_data_c$from, cor_data_c$to, cor_data_i$from, cor_data_i$to, cor_data_m$from, cor_data_m$to))

# Create a nodes data frame
nodes_df <- data.frame(name = unique(all_nodes))
nodes_df$shape <- ifelse(nodes_df$name %in% species_columns, "circle", 
                         ifelse(nodes_df$name %in% food_columns, "square", "diamond"))



# Create graph objects from edge lists and nodes
graph_c <- tbl_graph(edges = cor_data_c, nodes = nodes_df, directed = FALSE)
graph_i <- tbl_graph(edges = cor_data_i, nodes = nodes_df, directed = FALSE)
graph_m <- tbl_graph(edges = cor_data_m, nodes = nodes_df, directed = FALSE)

  

plot_graph <- function(graph) {
  ggraph(graph, layout = "fr") +
    geom_edge_link(aes(edge_alpha = abs_correlation, edge_color = edge_color, edge_width = abs_correlation), show.legend = FALSE) +
    geom_node_point(aes(shape = shape), size = 5) +
    scale_shape_manual(values = c(circle = 16, square = 15, diamond = 18), labels = c("Species","Nutrition", "Food")) +  # Mapping shapes to actual ggplot2 shapes
    theme_minimal() +
    scale_edge_color_manual(values = c("red", "blue"))
}

# Plot each graph
plot_c <- plot_graph(graph_c)
plot_i <- plot_graph(graph_i)
plot_m <- plot_graph(graph_m)

# Print plots
print(plot_c)
print(plot_i)
print(plot_m)

```
```{r}

all_nodes <- unique(c(cor_data_c$from, cor_data_c$to, cor_data_i$from, cor_data_i$to, cor_data_m$from, cor_data_m$to))

nodes_df <- data.frame(name = unique(all_nodes))
nodes_df$shape <- ifelse(nodes_df$name %in% species_columns, "circle", 
                         ifelse(nodes_df$name %in% food_columns, "square", "diamond"))

all_edges <- bind_rows(
  mutate(cor_data_c, graph = 'c'),
  mutate(cor_data_i, graph = 'i'),
  mutate(cor_data_m, graph = 'm')
)


combined_graph <- tbl_graph(edges = all_edges, nodes = nodes_df, directed = FALSE)

fixed_layout <- create_layout(combined_graph, layout = "fr")

nodes_df <- left_join(nodes_df, data.frame(name = nodes_df$name, x = fixed_layout$x, y = fixed_layout$y), by = "name")

graph_c <- tbl_graph(edges = cor_data_c, nodes = nodes_df, directed = FALSE)
graph_i <- tbl_graph(edges = cor_data_i, nodes = nodes_df, directed = FALSE)
graph_m <- tbl_graph(edges = cor_data_m, nodes = nodes_df, directed = FALSE)

# plot_graph_fixed <- function(graph) {
#   # Convert igraph object to tbl_graph (tidygraph object)
#   tbl_graph <- as_tbl_graph(graph)
#   
#   # Use mutate to add degree centrality
#   tbl_graph <- tbl_graph %>%
#     activate(nodes) %>%
#     mutate(degree = centrality_degree(mode = "all"))  # Use mode="all", "in", or "out" as needed
#   
#   # Create the plot
#   ggraph(tbl_graph, layout = 'manual', x = x, y = y) +
#     geom_edge_link(aes(edge_alpha = log_p, edge_color = edge_color, edge_width = log_p), show.legend = FALSE) +
#     geom_node_point(aes(shape = shape, size = degree, color = label), color = "black") +
#     scale_shape_manual(values = c(circle = 16, square = 15, diamond = 18), labels = c("Species","Nutrition", "Food")) +
#     theme_graph() +
#     scale_edge_color_manual(values = c("red", "blue")) +
#     scale_size(range = c(3, 10)) +
#     scale_color_manual(values = c("Species" = "green", "Nutrition" = "orange", "Food" = "purple"))
# }


plot_graph_fixed <- function(graph) {
  # Calculate the degree of each node
  V(graph)$degree <- degree(graph, mode = "all")
  V(graph)$shape <- ifelse(V(graph)$name %in% species_columns, "circle", 
                         ifelse(V(graph)$name %in% food_columns, "square", "diamond"))

  # Create the plot
  ggraph(graph, layout = 'manual', x = x, y = y) +
    geom_edge_link(aes(edge_alpha = abs_correlation, edge_color = edge_color, edge_width = abs_correlation), show.legend = FALSE) +
    geom_node_point(aes(shape = shape, size = degree, color = shape)) +  # Map 'size' to 'degree'
    geom_node_text(aes(label = ifelse(degree > 0, name, "")), , color = "black", repel = TRUE, fontface = "bold", size = 3) +
    scale_shape_manual(values = c(circle = 19, square = 15, diamond = 17), labels = c("Species", "Nutrition", "Food")) +
    scale_edge_color_manual(values = c("#8D3256", "#004697")) +
    scale_color_manual(values = c("slateblue", "#68B69F", "#F08E59"), labels = c("Species", "Nutrition", "Food")) +
    theme_graph() +
    scale_size(range = c(4, 7))
}
#V(graph)$shape <- ifelse(V(graph)$type == "Species", "circle",
                         #ifelse(V(graph)$type == "Food", "square", "diamond"))

plot_c_fixed <- plot_graph_fixed(graph_c)
plot_i_fixed <- plot_graph_fixed(graph_i)
plot_m_fixed <- plot_graph_fixed(graph_m)


print(plot_c_fixed)
print(plot_i_fixed)
print(plot_m_fixed)

```
```{r fig.width=10, fig.height=10}
png("figure/network_c_49.png", width = 1400, height = 1400, res = 120)
plot_c_fixed
dev.off()
png("figure/network_i_49.png", width = 1400, height = 1400, res = 120)
plot_i_fixed
dev.off()
png("figure/network_m_49.png", width = 1400, height = 1400, res = 120)
plot_m_fixed
dev.off()
```





## component
```{r}
library(igraph)
library(ggraph)
library(tidygraph)
library(dplyr)

# Combine and prepare nodes and edges
all_nodes <- unique(c(cor_data_c$from, cor_data_c$to, cor_data_i$from, cor_data_i$to, cor_data_m$from, cor_data_m$to))
nodes_df <- data.frame(name = unique(all_nodes))
nodes_df$shape <- ifelse(nodes_df$name %in% species_columns, "circle", 
                         ifelse(nodes_df$name %in% food_columns, "square", "diamond"))

all_edges <- bind_rows(
  mutate(cor_data_c, graph = 'c'),
  mutate(cor_data_i, graph = 'i'),
  mutate(cor_data_m, graph = 'm')
)

combined_graph <- tbl_graph(edges = all_edges, nodes = nodes_df, directed = FALSE)
fixed_layout <- create_layout(combined_graph, layout = "fr")
nodes_df <- left_join(nodes_df, data.frame(name = nodes_df$name, x = fixed_layout$x, y = fixed_layout$y), by = "name")

graph_c <- tbl_graph(edges = cor_data_c, nodes = nodes_df, directed = FALSE)
graph_i <- tbl_graph(edges = cor_data_i, nodes = nodes_df, directed = FALSE)
graph_m <- tbl_graph(edges = cor_data_m, nodes = nodes_df, directed = FALSE)

# Function to calculate components and add them to the nodes
add_component_membership <- function(graph) {
  components <- components(graph)
  V(graph)$component <- components$membership
  return(graph)
}

# Apply the function to each graph
graph_c <- add_component_membership(graph_c)
graph_i <- add_component_membership(graph_i)
graph_m <- add_component_membership(graph_m)

# Function to calculate and assign degree
assign_degree <- function(graph) {
  V(graph)$degree <- degree(graph, mode = "all")
  return(graph)
}

# Apply the function to each graph
graph_c <- assign_degree(graph_c)
graph_i <- assign_degree(graph_i)
graph_m <- assign_degree(graph_m)
```


```{r fig.width=10, fig.height=8}
plot_graph_fixed <- function(graph) {
  node_data <- as_tibble(as_data_frame(graph, what = "vertices"))
  ggraph(graph, layout = 'manual', x = node_data$x, y = node_data$y) +
    geom_edge_link(aes(edge_alpha = abs_correlation, edge_color = edge_color, edge_width = abs_correlation), show.legend = FALSE) +
    geom_node_point(aes(shape = shape, size = degree, color = as.factor(component)), show.legend = FALSE) +  # Map 'color' to 'component'
    geom_node_text(aes(label = ifelse(degree > 0, name, "")), color = "black", repel = TRUE, fontface = "bold", size = 3) +
    scale_shape_manual(values = c(circle = 19, square = 15, diamond = 17), labels = c("Species", "Nutrition", "Food")) +
    scale_edge_color_manual(values = c("#8D3256", "#004697")) +
    scale_color_discrete() +  # Use discrete color scale for components
    theme_graph() +
    scale_size(range = c(4, 7))
}

# Plot each graph with the fixed layout
plot_c_fixed_new <- plot_graph_fixed(graph_c)
plot_i_fixed_new <- plot_graph_fixed(graph_i)
plot_m_fixed_new <- plot_graph_fixed(graph_m)

print(plot_c_fixed_new)
print(plot_i_fixed_new)
print(plot_m_fixed_new)


```
```{r}
png("figure/network_c_new.png", width = 1400, height = 1400, res = 120)
plot_c_fixed_new
dev.off()
png("figure/network_i_new.png", width = 1400, height = 1400, res = 120)
plot_i_fixed_new
dev.off()
png("figure/network_m_new.png", width = 1400, height = 1400, res = 120)
plot_m_fixed_new
dev.off()
```






## normality test for each food
```{r}
distri_food = full_data[,-1]

distri_food = subset(distri_food, FREG5_Ethnic_Group != "O")

# Assuming 'full_data' has one row per participant and multiple food intake columns
# Reshape the data from wide to long format
long_data <- distri_food %>%
  pivot_longer(
    cols = -FREG5_Ethnic_Group,  # Exclude the group column
    names_to = "Food_Type",       # New column for food types
    values_to = "Intake"          # New column for intake values
  )
```


```{r}

food_types <- c("Fish", "Crustacean", "Mollusk", "Seafood_dishes", "Poultry", "Red_meat",
                "Meat_dishes", "Processed_meat", "Dairy", "Cruciferous_vegetables", "Legumes",
                "Mushroom", "Fruits", "Wholegrain", "Chapati", "Noodle_dishes", "Deep_fried", 
                "Coffee", "Tea")

# if (!dir.exists("figure")) {
#   dir.create("figure")
# }


for (food in food_types) {
  current_data <- long_data %>% filter(Food_Type == food)
  p <- ggplot(current_data, aes(x = Intake, fill = FREG5_Ethnic_Group)) +
    geom_density(alpha = 0.5, adjust = 1.5) +
    labs(title = paste("Density Plot of", food),
         x = "Intake per Day", y = "Density") +
    scale_fill_manual(values = c("C" = "#9FDDCA", "I" = "#9EA3F1", "M" = "#E17887")) +
    theme_minimal()+
    theme(legend.position = "none",
          plot.title = element_text(size = 20, face = "bold"))
  png_filename <- paste0("figure/", food, "_density_plot.png")
  ggsave(png_filename, plot = p, width = 10, height = 6, dpi = 300)
}

```


```{r}
library(patchwork)
# Define the vector of food types
food_types <- c("Fish", "Crustacean", "Mollusk", "Seafood_dishes", "Poultry", "Red_meat",
                "Meat_dishes", "Processed_meat", "Dairy", "Cruciferous_vegetables", "Legumes",
                "Mushroom", "Fruits", "Wholegrain", "Chapati", "Noodle_dishes", "Deep_fried", 
                "Coffee", "Tea")

# Initialize an empty list to store plots
plots_list <- list()

# Loop through each food type and create a plot
for (food in food_types) {
  current_data <- long_data %>% filter(Food_Type == food)
  
  p <- ggplot(current_data, aes(x = Intake, fill = FREG5_Ethnic_Group)) +
    geom_density(alpha = 0.5, adjust = 1.5) +
    labs(title = paste(food),
         x = "Intake per Day", y = "Density") +
    scale_fill_manual(values = c("C" = "#9FDDCA", "I" = "#9EA3F1", "M" = "#E17887")) +
    theme_minimal()+
    theme(legend.position = "none")
  
  # Add the plot to the list
  plots_list[[food]] = p
}

# Combine all plots side by side using patchwork
combined_plot <- wrap_plots(plots_list, ncol = 1) # Adjust ncol to change the number of columns
combined_plot
# Save the combined plot
#ggsave("combined_density_plots.png", combined_plot, width = 30, height = 20, dpi = 300) # Adjust size as needed

```










