---
title: "new_filter_try"
author: "Ruwen Zhou"
date: "9/11/2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(plyr)
library(tidyverse)
library(dplyr)
library(readxl)
library(ggplot2)
library(phyloseq)
library(vegan)
library(DESeq2)
library(pheatmap)
library(RColorBrewer)
```


# import 800 data species
```{r}
species_data_800 = read_excel("data/species_866.xlsx")
species_data_800
species_data_800 <- species_data_800 %>%
  dplyr::select(-c(unclassified, others, `Homo sapiens`))
ethnicity_800 = read_excel("data/ethnicity_866.xlsx")
species_data_800
ethnicity_800
```





```{r}
# Convert absolute counts -> relative abundance (row-wise proportions)
counts_to_relabund <- function(df) {
  df_num <- as.data.frame(df, stringsAsFactors = FALSE)
  # ensure numeric
  for (j in seq_len(ncol(df_num))) df_num[[j]] <- as.numeric(df_num[[j]])
  
  libsize <- rowSums(df_num, na.rm = TRUE)
  zero_rows <- which(libsize == 0)
  if (length(zero_rows) > 0) {
    warning(sprintf("Found %d samples with total count = 0; their relative abundances are set to 0.", length(zero_rows)))
  }
  rel <- sweep(df_num, 1, libsize, FUN = "/")
  rel[is.na(rel)] <- 0
  rel
}

# Filter by prevalence (≥ prev_threshold of samples have >0) 
# and mean relative abundance (≥ relabund_threshold)
filter_prev_abund <- function(rel_df, prev_threshold = 0.10, relabund_threshold = 1e-4) {
  stopifnot(prev_threshold >= 0 && prev_threshold <= 1)
  n <- nrow(rel_df)
  keep_prev  <- vapply(rel_df, function(col) sum(col > 0) >= (n * prev_threshold), logical(1))
  keep_abund <- vapply(rel_df, function(col) mean(col, na.rm = TRUE) >= relabund_threshold, logical(1))
  keep <- keep_prev & keep_abund
  rel_df[, keep, drop = FALSE]
}

# Convenience wrapper: counts -> relabund -> filter.
# Optionally return the filtered *counts* aligned to the kept taxa too.
filter_counts_by_prev_and_relabund <- function(count_df,
                                               prev_threshold = 0.10,
                                               relabund_threshold = 1e-4,
                                               return_counts_also = TRUE) {
  rel <- counts_to_relabund(count_df)
  rel_filt <- filter_prev_abund(rel, prev_threshold, relabund_threshold)
  if (return_counts_also) {
    kept_taxa <- colnames(rel_filt)
    list(
      relabund_filtered  = rel_filt,
      counts_filtered    = count_df[, kept_taxa, drop = FALSE],
      kept_taxa          = kept_taxa
    )
  } else {
    rel_filt
  }
}
```


```{r}
# ==== Example on your data ====
res <- filter_counts_by_prev_and_relabund(
  species_data_800,
  prev_threshold = 0.10,    # present in ≥10% of samples
  relabund_threshold = 0.00001 # mean relative abundance ≥0.01%
)

# Filtered relative abundance table:
relabund_filtered <- res$relabund_filtered
relabund_filtered
# (Optional) filtered counts, aligned to the same kept taxa:
species_filtered <- res$counts_filtered
species_filtered
write_xlsx(species_filtered, path = "data/filtered_species602.xlsx")
```

```{r}
# rownames(species_data) = species_data[,1]
# species_data[1] = NULL
# rownames(metadata) = metadata[,1]
# species_data
# metadata
species_filtered
```

```{r}
merged_data1 <- merge(ethnicity_800[,c("FREG1_Barcode", "FREG5_Ethnic_Group")], species_filtered, by.x = "FREG1_Barcode", by.y = "FREG1_Barcode",  all = FALSE)
merged_data1 = subset(merged_data1, FREG5_Ethnic_Group != "O")
merged_data1
ethnicity_800
```

```{r}
library(DESeq2)
library(dplyr)

# Metadata
meta <- merged_data1[, 1:2]
colnames(meta) <- c("SampleID", "Ethnicity")
rownames(meta) <- meta$SampleID
meta$Group <- ifelse(meta$Ethnicity == "C", "C", "nonC")
meta$Group <- factor(meta$Group, levels = c("nonC", "C"))
# Counts (species only)
counts <- merged_data1[, -(1:2)]
rownames(counts) <- meta$SampleID

# Transpose counts: DESeq2 wants rows = species, cols = samples
counts_t <- t(counts)

```

```{r}
dds <- DESeqDataSetFromMatrix(
  countData = counts_t,
  colData   = meta,
  design    = ~ Group
)

dds <- dds[rowSums(counts(dds)) > 10, ]  # filter low-abundance
dds <- DESeq(dds)

# Extract results: C vs nonC
res_C <- results(dds, contrast = c("Group", "C", "nonC"))
res_C <- as.data.frame(res_C) %>% mutate(Species = rownames(.))
sig_C <- res_C %>% filter(padj < 0.05) %>% arrange(padj)
sig_C

```
```{r}
# I vs non-I
meta$Group <- ifelse(meta$Ethnicity == "I", "I", "nonI")
meta$Group <- factor(meta$Group, levels = c("nonI", "I"))
dds <- DESeqDataSetFromMatrix(countData = counts_t, colData = meta, design = ~ Group)
dds <- dds[rowSums(counts(dds)) > 10, ]
dds <- DESeq(dds)
res_I <- results(dds, contrast = c("Group", "I", "nonI")) %>%
  as.data.frame() %>% mutate(Species = rownames(.))
sig_I <- res_I %>% filter(padj < 0.05) %>% arrange(padj)

# M vs non-M
meta$Group <- ifelse(meta$Ethnicity == "M", "M", "nonM")
meta$Group <- factor(meta$Group, levels = c("nonM", "M"))
dds <- DESeqDataSetFromMatrix(countData = counts_t, colData = meta, design = ~ Group)
dds <- dds[rowSums(counts(dds)) > 10, ]
dds <- DESeq(dds)
res_M <- results(dds, contrast = c("Group", "M", "nonM")) %>%
  as.data.frame() %>% mutate(Species = rownames(.))
sig_M <- res_M %>% filter(padj < 0.05) %>% arrange(padj)


```

```{r}
# Load required libraries
library(DESeq2)
library(dplyr)
library(ggplot2)

# Your existing data preparation
meta <- merged_data1[, 1:2]
colnames(meta) <- c("SampleID", "Ethnicity")
rownames(meta) <- meta$SampleID

counts <- merged_data1[, -(1:2)]
rownames(counts) <- meta$SampleID
counts_t <- t(counts)

# Function to run DESeq2 and create lollipop plot
run_comparison <- function(target_ethnicity, plot_title, filename) {
  
  # Create group variable
  meta_temp <- meta
  meta_temp$Group <- ifelse(meta_temp$Ethnicity == target_ethnicity, target_ethnicity, paste0("non", target_ethnicity))
  meta_temp$Group <- factor(meta_temp$Group, levels = c(paste0("non", target_ethnicity), target_ethnicity))
  
  # DESeq2 analysis
  dds <- DESeqDataSetFromMatrix(countData = counts_t, colData = meta_temp, design = ~ Group)
  dds <- dds[rowSums(counts(dds)) > 10, ]
  dds <- DESeq(dds)
  
  # Get results
  res <- results(dds, contrast = c("Group", target_ethnicity, paste0("non", target_ethnicity)))
  res_df <- as.data.frame(res)
  res_df$Species <- rownames(res_df)
  
  # Filter significant results
  sig_results <- res_df %>% 
    filter(!is.na(padj) & padj < 0.05) %>%
    arrange(padj)   # Top 15 most significant
  
  # Skip if no significant results
  if(nrow(sig_results) == 0) {
    return(paste("No significant species for", plot_title))
  }
  
  # Add plotting variables
  sig_results$Direction <- ifelse(sig_results$log2FoldChange > 0, "Higher", "Lower")
  sig_results$SpeciesName <- gsub("_", " ", sig_results$SpeciesName)
  sig_results$neg_log10_padj <- -log10(sig_results$padj)
  
  # Create lollipop plot
  p <- ggplot(sig_results, aes(x = reorder(SpeciesName, log2FoldChange), y = log2FoldChange)) +
    geom_segment(aes(x = reorder(SpeciesName, log2FoldChange), 
                     xend = reorder(SpeciesName, log2FoldChange),
                     y = 0, yend = log2FoldChange), 
                 linewidth = 1.2, alpha = 0.8, color = "gray60") +
    geom_point(aes(size = neg_log10_padj, color = Direction), alpha = 0.8) +
    scale_color_manual(values = c("Higher" = "#E31A1C", "Lower" = "#1F78B4")) +
    scale_size_continuous(name = "-log10(adj p-value)", range = c(3, 8)) +
    labs(title = plot_title, x = "Species", y = "log2 Fold Change") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
          legend.position = "bottom") +
    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
    coord_flip()
  
  # Save plot
  ggsave(paste0("figure/", filename), p, width = 10, height = 18, device = "pdf")
  
  return(paste("Saved", filename, "with", nrow(sig_results), "significant species"))
}



# Run all three comparisons
result1 <- run_comparison("C", "Chinese vs Non-Chinese", "chinese_vs_non.pdf")
result2 <- run_comparison("I", "Indian vs Non-Indian", "indian_vs_non.pdf")  
result3 <- run_comparison("M", "Malay vs Non-Malay", "malay_vs_non.pdf")

# Print results
cat(result1, "\n")
cat(result2, "\n")
cat(result3, "\n")
```

## bubble plot
```{r}
library(DESeq2)
library(dplyr)
library(ggplot2)
library(forcats)

# Optional: a lookup table for taxonomy if you have it
# tax_df should have columns: Species (matching rownames in counts_t), Phylum or Genus, etc.
# Example structure: tax_df <- data.frame(Species=rownames(counts_t), Phylum=..., Genus=..., stringsAsFactors=FALSE)

run_comparison_bubble <- function(target_ethnicity,
                                  plot_title,
                                  filename,
                                  top_n_per_direction = 20,
                                  facet_tax_level = NULL,      # e.g. "Phylum" or "Genus" if tax_df provided
                                  tax_df = NULL,
                                  height_per_40 = 10) {         # autoscale height: 10in per ~40 rows
  
  # Create group variable
  meta_temp <- meta
  meta_temp$Group <- ifelse(meta_temp$Ethnicity == target_ethnicity,
                            target_ethnicity, paste0("non", target_ethnicity))
  meta_temp$Group <- factor(meta_temp$Group,
                            levels = c(paste0("non", target_ethnicity), target_ethnicity))
  
  # DESeq2
  dds <- DESeqDataSetFromMatrix(countData = counts_t, colData = meta_temp, design = ~ Group)
  dds <- dds[rowSums(counts(dds)) > 10, ]
  dds <- DESeq(dds)
  
  res <- results(dds, contrast = c("Group", target_ethnicity, paste0("non", target_ethnicity)))
  res_df <- as.data.frame(res)
  res_df$Species <- rownames(res_df)

  # Keep significant rows
  sig <- res_df %>%
    filter(!is.na(padj), padj < 0.05) %>%
    mutate(Direction = ifelse(log2FoldChange > 0, "Higher", "Lower"),
           SpeciesName = gsub("_", " ", Species),
           SpeciesLabel = paste0("italic('", SpeciesName, "')"), 
           neg_log10_padj = -log10(padj))
  
  if (nrow(sig) == 0) {
    message(paste("No significant species for", plot_title))
    return(invisible(NULL))
  }

  # Attach taxonomy if provided
  if (!is.null(tax_df) && !is.null(facet_tax_level) && facet_tax_level %in% names(tax_df)) {
    sig <- sig %>%
      left_join(tax_df[, c("Species", facet_tax_level)], by = "Species") %>%
      mutate(!!facet_tax_level := ifelse(is.na(.data[[facet_tax_level]]), "Unassigned", .data[[facet_tax_level]]))
  }

  # Top-N per direction (rank by padj, then by |log2FC|)
  if (!is.null(facet_tax_level) && facet_tax_level %in% names(sig)) {
  sig_top <- sig %>%
    dplyr::group_by(Direction, .data[[facet_tax_level]]) %>%
    dplyr::arrange(padj, dplyr::desc(abs(log2FoldChange)), .by_group = TRUE) %>%
    dplyr::slice_head(n = top_n_per_direction) %>%
    dplyr::ungroup()
  } else {
    sig_top <- sig %>%
    dplyr::group_by(Direction) %>%
    dplyr::arrange(padj, dplyr::desc(abs(log2FoldChange)), .by_group = TRUE) %>%
    dplyr::slice_head(n = top_n_per_direction) %>%
    dplyr::ungroup()
  n}

  # Order species by effect size within plot for nicer layout
  sig_top <- sig_top %>%
    mutate(SpeciesName = fct_reorder(SpeciesName, log2FoldChange))

  # Bubble plot (compact)
  p <- ggplot(sig_top,
            aes(x = reorder(SpeciesLabel, log2FoldChange),
                y = log2FoldChange,
                size = neg_log10_padj,
                color = Direction)) +
  geom_point(alpha = 0.85) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.4, alpha = 0.7) +
  scale_size_continuous(name = "-log10(adj p-value)", range = c(2.5, 8)) +
  scale_color_manual(values = c("Higher" = "#E31A1C", "Lower" = "#1F78B4")) +
  labs(title = plot_title, x = "Species", y = "log2 Fold Change") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        legend.position = "bottom") +
  coord_flip() +
  scale_x_discrete(labels = function(x) parse(text = x))  # <-- render italic

  # Facet by taxonomy if requested
  if (!is.null(facet_tax_level) && facet_tax_level %in% names(sig_top)) {
    p <- p + facet_wrap(as.formula(paste("~", facet_tax_level)), scales = "free_y")
  }

  # Auto-scale figure height by number of rows shown (works for long lists)
  n_rows <- nrow(sig_top)
  fig_height <- max(6, ceiling(n_rows / 40) * height_per_40)  # ~40 rows per 10 inches (adjust if needed)
  dir.create("figure", showWarnings = FALSE, recursive = TRUE)
  ggsave(file.path("figure", filename), p, width = 10, height = fig_height, device = "pdf")

  invisible(sig_top)
}

# Example calls (mirroring your original three):
resC_bub <- run_comparison_bubble("C", "Chinese vs Non-Chinese", "chinese_vs_non_bubble.pdf",
                                  top_n_per_direction = 20)
resI_bub <- run_comparison_bubble("I", "Indian vs Non-Indian", "indian_vs_non_bubble.pdf",
                                  top_n_per_direction = 20)
resM_bub <- run_comparison_bubble("M", "Malay vs Non-Malay", "malay_vs_non_bubble.pdf",
                                  top_n_per_direction = 20)

```
```{r}
counts_t
```



```{r}
## ===== Parameters you can tweak =====
alpha <- 0.05                 # FDR cutoff (your sig_* tables likely already use this)
lfc_cutoff <- log2(2)       # require ≥ 1.5x change (set to log2(2) for 2-fold, etc.)

## ===== Helper: pick the best LFC column (shrunk if available) =====
pick_lfc <- function(df) {
  if ("log2FoldChange_shr" %in% names(df)) {
    df$log2FoldChange_shr
  } else {
    df$log2FoldChange
  }
}

## ===== Helper: annotate direction given a group label =====
annotate_direction <- function(sig_df, group_label, lfc_cut = lfc_cutoff, alpha = alpha) {
  if (is.null(sig_df) || nrow(sig_df) == 0) {
    return(sig_df)
  }
  # ensure required cols exist
  stopifnot(all(c("Species","padj") %in% names(sig_df)))
  lfc <- pick_lfc(sig_df)

  out <- sig_df
  out$LFC_used <- lfc
  out$Direction <- ifelse(out$LFC_used >=  lfc_cut, paste("Enriched in", group_label),
                   ifelse(out$LFC_used <= -lfc_cut, paste("Depleted in", group_label), "NS_by_LFC"))
  # keep only those passing the LFC cutoff as well
  out <- out[out$Direction != "NS_by_LFC" & !is.na(out$padj) & out$padj < alpha, , drop = FALSE]

  # standardize useful columns/order
  keep_cols <- intersect(c("Species","baseMean","LFC_used","lfcSE","padj","Direction",
                           "log2FoldChange","log2FoldChange_shr","stat","pvalue"),
                         names(out))
  out <- out[, keep_cols, drop = FALSE]
  names(out)[names(out) == "LFC_used"] <- "log2FC_used"

  out[order(out$padj, -abs(out$log2FC_used)), ]
}

## ===== Make sure we have the three tables (case-safe) =====
sig_C_in <- if (exists("sig_C")) sig_C else sig_c
sig_I_in <- if (exists("sig_I")) sig_I else if (exists("sig_i")) sig_i else sig_I
sig_M_in <- if (exists("sig_M")) sig_M else if (exists("sig_m")) sig_m else sig_M

## ===== Apply annotation per comparison =====
ann_C <- annotate_direction(sig_C_in, "C", lfc_cutoff, alpha)
ann_I <- annotate_direction(sig_I_in, "I", lfc_cutoff, alpha)
ann_M <- annotate_direction(sig_M_in, "M", lfc_cutoff, alpha)

# Add a Comparison column
if (!is.null(ann_C) && nrow(ann_C)>0) ann_C$Comparison <- "C_vs_nonC"
if (!is.null(ann_I) && nrow(ann_I)>0) ann_I$Comparison <- "I_vs_nonI"
if (!is.null(ann_M) && nrow(ann_M)>0) ann_M$Comparison <- "M_vs_nonM"

## ===== Combined tidy table =====
combined_dir <- dplyr::bind_rows(ann_C, ann_I, ann_M) %>%
  dplyr::select(Comparison, Species, Direction, baseMean, log2FC_used,
                dplyr::any_of(c("log2FoldChange","log2FoldChange_shr","lfcSE","stat","pvalue","padj")))

# Save if you like:
# readr::write_csv(combined_dir, "DESeq2_sig_with_direction_all.csv")

## ===== Simple species lists per group (character vectors) =====
enriched_C <- if (exists("ann_C")) ann_C$Species[grepl("^Enriched in C$", ann_C$Direction)] else character(0)
depleted_C <- if (exists("ann_C")) ann_C$Species[grepl("^Depleted in C$", ann_C$Direction)] else character(0)

enriched_I <- if (exists("ann_I")) ann_I$Species[grepl("^Enriched in I$", ann_I$Direction)] else character(0)
depleted_I <- if (exists("ann_I")) ann_I$Species[grepl("^Depleted in I$", ann_I$Direction)] else character(0)

enriched_M <- if (exists("ann_M")) ann_M$Species[grepl("^Enriched in M$", ann_M$Direction)] else character(0)
depleted_M <- if (exists("ann_M")) ann_M$Species[grepl("^Depleted in M$", ann_M$Direction)] else character(0)
```


```{r}
enriched_C
```


```{r}
depleted_C
```


```{r}
enriched_I
```


```{r}
depleted_I
```


```{r}
enriched_M
```


```{r}
depleted_M

```

## food
```{r}
food_data800 <- read_excel("data/all_food_863.xlsx")
```


```{r}
#ethnicity_800
#food_data800

merged_data2 <- merge(ethnicity_800[,c("FREG1_Barcode", "FREG5_Ethnic_Group")], food_data800, by.x = "FREG1_Barcode", by.y = "participant_id",  all = FALSE)
merged_data2 = subset(merged_data2, FREG5_Ethnic_Group != "O")
merged_data2 <- merged_data2 %>%
  dplyr::select(-c(freg0_pid))
merged_data2
```

```{r}
merged_data1
```
```{r}
## ===== Packages =====
suppressPackageStartupMessages({
  library(dplyr)
  library(purrr)
  library(readr)   # for write_csv(); swap to write.csv if you prefer
})

## ===== Inputs expected =====
## - food_data800: first col = participant_id, second col = freg0_pid (to drop), remaining = foods (numeric)
## - ethnicity_800: columns FREG1_Barcode (ID), FREG5_Ethnic_Group (C/I/M)

## ===== Parameters =====
alpha <- 0.05                         # FDR threshold
lfc_cutoff <- log2(2)               # require ≥ 1.5x change (use log2(2) for 2x)
eps <- 1e-6                           # to avoid division-by-zero in median ratios

## ===== 1) Merge & clean =====
meta <- ethnicity_800 %>%
  transmute(SampleID = FREG1_Barcode,
            Ethnicity = FREG5_Ethnic_Group) %>%
  filter(Ethnicity %in% c("C","I","M")) %>%
  mutate(Ethnicity = factor(Ethnicity, levels = c("C","I","M")))

food <- food_data800 %>%
  rename(SampleID = participant_id) %>%
  select(-freg0_pid)

df <- food %>%
  inner_join(meta, by = "SampleID") %>%
  relocate(SampleID, Ethnicity)

# Make sure all food columns are numeric
food_cols <- setdiff(names(df), c("SampleID","Ethnicity"))
df[food_cols] <- lapply(df[food_cols], function(x) suppressWarnings(as.numeric(x)))

## ===== 2) Omnibus Kruskal–Wallis across C/I/M =====
kw_results <- map_dfr(food_cols, function(var) {
  x <- df[[var]]
  d <- data.frame(x = x, Ethnicity = df$Ethnicity)
  p <- tryCatch(stats::kruskal.test(x ~ Ethnicity, data = d)$p.value, error = function(e) NA_real_)
  tibble(Food = var, p_kw = p)
}) %>%
  mutate(p_kw_adj = p.adjust(p_kw, method = "fdr"))

## ===== 3) Binary (X vs non-X) Wilcoxon + median fold-change =====
binary_compare <- function(group_label) {
  grp <- ifelse(df$Ethnicity == group_label, group_label, paste0("non", group_label))
  map_dfr(food_cols, function(var) {
    a <- df[[var]][grp == group_label]
    b <- df[[var]][grp != group_label]
    a <- a[!is.na(a)]; b <- b[!is.na(b)]
    if (length(a) < 3 || length(b) < 3) {
      return(tibble(Food = var, group = group_label, p = NA_real_,
                    median_group = NA_real_, median_non = NA_real_,
                    fc_median = NA_real_, log2FC = NA_real_))
    }
    p <- tryCatch(wilcox.test(a, b)$p.value, error = function(e) NA_real_)
    med_a <- median(a); med_b <- median(b)
    fc <- (med_a + eps) / (med_b + eps)
    tibble(Food = var, group = group_label, p = p,
           median_group = med_a, median_non = med_b,
           fc_median = fc, log2FC = log2(fc))
  }) %>%
    mutate(p_adj = p.adjust(p, method = "fdr"))
}

res_C <- binary_compare("C")
res_I <- binary_compare("I")
res_M <- binary_compare("M")

## ===== 4) Annotate direction (Enriched/Depleted) with LFC cutoff =====
annotate_dir <- function(res_tbl) {
  res_tbl %>%
    mutate(
      Direction = dplyr::case_when(
        !is.na(p_adj) & p_adj < alpha & log2FC >=  lfc_cutoff ~ paste("Enriched in", group),
        !is.na(p_adj) & p_adj < alpha & log2FC <= -lfc_cutoff ~ paste("Depleted in", group),
        TRUE ~ "NS"
      )
    ) %>%
    filter(Direction != "NS") %>%
    arrange(p_adj, desc(abs(log2FC)))
}

ann_C <- annotate_dir(res_C) %>% mutate(Comparison = "C_vs_nonC")
ann_I <- annotate_dir(res_I) %>% mutate(Comparison = "I_vs_nonI")
ann_M <- annotate_dir(res_M) %>% mutate(Comparison = "M_vs_nonM")

## ===== 5) Combined tables =====
combined_binary <- bind_rows(
  res_C %>% mutate(Comparison = "C_vs_nonC"),
  res_I %>% mutate(Comparison = "I_vs_nonI"),
  res_M %>% mutate(Comparison = "M_vs_nonM")
) %>%
  left_join(kw_results, by = "Food") %>%
  relocate(Comparison, group, Food)

combined_sig_with_dir <- bind_rows(ann_C, ann_I, ann_M) %>%
  left_join(kw_results, by = "Food") %>%
  select(Comparison, group, Food, Direction,
         median_group, median_non, fc_median, log2FC,
         p, p_adj, p_kw, p_kw_adj)
```


```{r}
## ===== 6) Simple name lists (per group) =====
enriched_C_food <- ann_C %>% filter(Direction == "Enriched in C") %>% pull(Food)
depleted_C_food <- ann_C %>% filter(Direction == "Depleted in C") %>% pull(Food)
enriched_I_food <- ann_I %>% filter(Direction == "Enriched in I") %>% pull(Food)
depleted_I_food <- ann_I %>% filter(Direction == "Depleted in I") %>% pull(Food)
enriched_M_food <- ann_M %>% filter(Direction == "Enriched in M") %>% pull(Food)
depleted_M_food <- ann_M %>% filter(Direction == "Depleted in M") %>% pull(Food)
```


```{r}
enriched_C_food
```


```{r}
depleted_C_food
```


```{r}
enriched_I_food
```


```{r}
depleted_I_food
```


```{r}
enriched_M_food
```


```{r}
depleted_M_food

```

## stratification
```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(purrr)
  library(tibble)
  library(readr)   # optional for write_csv
})

## ==== 1) Helpers ====
counts_to_relabund <- function(df_counts) {
  lib <- rowSums(df_counts, na.rm = TRUE)
  rel <- sweep(df_counts, 1, lib, "/")
  rel[is.na(rel)] <- 0
  rel
}

run_group_corr <- function(df_all, group_label, species_cols, food_cols,
                           min_nonzero_prop_species = 0.10,
                           min_nonzero_prop_food    = 0.10,
                           alpha = 0.05,
                           min_abs_rho = 0.0,
                           restrict_species = NULL,
                           restrict_foods   = NULL) {
  d <- df_all %>% filter(Ethnicity == group_label)

  # Restrict to curated lists if provided
  if (!is.null(restrict_species)) species_cols <- intersect(species_cols, restrict_species)
  if (!is.null(restrict_foods))   food_cols    <- intersect(food_cols,   restrict_foods)

  # Keep species present in ≥ X% samples (within group) and with variance > 0
  keep_species <- species_cols[vapply(species_cols, function(s)
    mean(d[[s]] > 0, na.rm = TRUE) >= min_nonzero_prop_species && stats::var(d[[s]], na.rm = TRUE) > 0,
    logical(1))]

  # Keep foods observed in ≥ X% samples (non-missing/non-constant)
  keep_foods <- food_cols[vapply(food_cols, function(f)
    mean(is.finite(d[[f]]), na.rm = TRUE) >= min_nonzero_prop_food && stats::var(d[[f]], na.rm = TRUE) > 0,
    logical(1))]

  if (length(keep_species) == 0 || length(keep_foods) == 0) {
    message(sprintf("[%s] No columns left after filtering.", group_label))
    return(tibble())
  }

  res <- map_dfr(keep_species, function(s) {
    x <- d[[s]]
    map_dfr(keep_foods, function(f) {
      y <- d[[f]]
      ok <- is.finite(x) & is.finite(y)
      if (sum(ok) < 8) return(tibble())  # need enough paired data
      ct <- suppressWarnings(cor.test(x[ok], y[ok], method = "spearman", exact = FALSE))
      tibble(Ethnicity = group_label, Species = s, Food = f,
             n = sum(ok), rho = unname(ct$estimate), p_value = ct$p.value)
    })
  })

  if (nrow(res) == 0) return(res)

  res %>%
    mutate(p_adj = p.adjust(p_value, method = "fdr")) %>%
    filter(!is.na(p_adj), p_adj < alpha, abs(rho) >= min_abs_rho) %>%
    arrange(p_adj, desc(abs(rho)))
}

## ==== 2) Build aligned data frames ====
# Inputs expected:
# - merged_data1: columns [FREG1_Barcode, FREG5_Ethnic_Group, <species counts...>]
# - food_data800: first column participant_id, second freg0_pid (drop), rest foods (numeric)
# - ethnicity_800: (if needed) FREG1_Barcode + FREG5_Ethnic_Group; but merged_data1 already has both.

# --- BEGIN replacement for species_counts build ---
stopifnot(all(c("FREG1_Barcode","FREG5_Ethnic_Group") %in% names(merged_data1)))

meta <- merged_data1[, c("FREG1_Barcode","FREG5_Ethnic_Group")]
names(meta) <- c("SampleID","Ethnicity")
meta <- subset(meta, Ethnicity %in% c("C","I","M"))
meta$Ethnicity <- factor(meta$Ethnicity, levels = c("C","I","M"))

species_counts_df <- merged_data1[, !(names(merged_data1) %in% c("FREG5_Ethnic_Group"))]
names(species_counts_df)[names(species_counts_df) == "FREG1_Barcode"] <- "SampleID"
species_counts_df <- as.data.frame(species_counts_df, stringsAsFactors = FALSE)
rownames(species_counts_df) <- NULL
rownames(species_counts_df) <- species_counts_df$SampleID
species_counts_df$SampleID <- NULL
species_counts_df[] <- lapply(species_counts_df, function(x) suppressWarnings(as.numeric(x)))
if (any(duplicated(rownames(species_counts_df)))) {
  dup_ids <- unique(rownames(species_counts_df)[duplicated(rownames(species_counts_df))])
  stop(sprintf("Duplicate SampleIDs in merged_data1: %s", paste(head(dup_ids, 10), collapse = ", ")))
}
species_counts <- as.matrix(species_counts_df)

# Relative abundance for correlations
lib <- rowSums(species_counts, na.rm = TRUE)
species_rel <- sweep(species_counts, 1, lib, "/")
species_rel[!is.finite(species_rel)] <- 0
species_rel_df <- as.data.frame(species_rel)
species_rel_df$SampleID <- rownames(species_rel_df)
# --- END replacement ---


# Convert counts -> relative abundance
species_rel <- counts_to_relabund(species_counts)
species_rel_df <- species_rel %>% as.data.frame() %>% rownames_to_column("SampleID")
species_cols <- colnames(species_rel_df)[-1]   # list of species columns

# Food table (make numeric, drop freg0_pid)
food_df <- food_data800 %>%
  rename(SampleID = participant_id) %>%
  select(-freg0_pid)
food_cols <- setdiff(colnames(food_df), "SampleID")
food_df[food_cols] <- lapply(food_df[food_cols], function(x) suppressWarnings(as.numeric(x)))

# Merge everything (inner join to ensure matched samples)
df_all <- meta %>%
  inner_join(species_rel_df, by = "SampleID") %>%
  inner_join(food_df, by = "SampleID")

## ==== 3) (Optional) restrict to your curated lists ====
# If you have DA species/foods per group, put them here. Otherwise leave as NULL to use ALL.
# Example placeholders; replace with your vectors if you want targeted correlations:
species_keep_C <- NULL   # e.g., sig_C$Species[ sig_C$padj < 0.05 & abs(sig_C$log2FoldChange) >= log2(1.5) ]
species_keep_I <- NULL
species_keep_M <- NULL

food_keep_C <- NULL      # e.g., enriched_food_C from your food DA step
food_keep_I <- NULL
food_keep_M <- NULL
```


```{r}
## ==== 4) Run within-group Spearman ====
alpha <- 0.05
min_abs_rho <- 0       # raise/lower as you like
min_prev <- 0

corr_C <- run_group_corr(df_all, "C", species_cols, food_cols,
                         min_nonzero_prop_species = min_prev,
                         min_nonzero_prop_food    = min_prev,
                         alpha = alpha, min_abs_rho = min_abs_rho,
                         restrict_species = species_keep_C,
                         restrict_foods   = food_keep_C)

corr_I <- run_group_corr(df_all, "I", species_cols, food_cols,
                         min_nonzero_prop_species = min_prev,
                         min_nonzero_prop_food    = min_prev,
                         alpha = alpha, min_abs_rho = min_abs_rho,
                         restrict_species = species_keep_I,
                         restrict_foods   = food_keep_I)

corr_M <- run_group_corr(df_all, "M", species_cols, food_cols,
                         min_nonzero_prop_species = min_prev,
                         min_nonzero_prop_food    = min_prev,
                         alpha = alpha, min_abs_rho = min_abs_rho,
                         restrict_species = species_keep_M,
                         restrict_foods   = food_keep_M)

combined_corr <- bind_rows(corr_C, corr_I, corr_M)
```


```{r}
## ==== 5) Save (optional) ====
# write_csv(corr_C, "Spearman_food_microbiome_C.csv")
# write_csv(corr_I, "Spearman_food_microbiome_I.csv")
# write_csv(corr_M, "Spearman_food_microbiome_M.csv")
# write_csv(combined_corr, "Spearman_food_microbiome_all.csv")

## Quick peek

#table(combined_corr$Ethnicity)
combined_corr

#corr_I
```


DA detects a between-group shift in means; correlation detects a within-group association (covariance of ranks). Either can be present without the other

Power/sample size
DA uses all samples at once; within-group uses only C (or I/M) samples: fewer n and more multiple testing (foods × species), so  more FDR pruning.

## food speices association
```{r}
food_group = read_xlsx("data/FFQA862.xlsx")
food_group
```

```{r}
## ===== Parameters =====
alpha <- 0.05          # FDR cutoff
min_abs_rho <- 0    # set to 0 if you don't want a rho cutoff
min_prev_species <- 0  # species present in ≥10% of samples
min_prev_food    <- 0  # food non-missing in ≥10% of samples
min_n_pair <- 8           # min paired samples per test

## ===== 0) Expect these to already exist from your DA step =====
## enriched_C, depleted_C, enriched_I, depleted_I, enriched_M, depleted_M

## ===== 1) Build the combined species list with origin labels =====
map_list <- function(vec, tag) if (exists(deparse(substitute(vec))) && length(vec)) 
  data.frame(Species = vec, Origin = tag, stringsAsFactors = FALSE) else
  data.frame(Species = character(0), Origin = character(0))

origin_tbl <- rbind(
  map_list(enriched_C, "Enriched in C"),
  map_list(depleted_C, "Depleted in C"),
  map_list(enriched_I, "Enriched in I"),
  map_list(depleted_I, "Depleted in I"),
  map_list(enriched_M, "Enriched in M"),
  map_list(depleted_M, "Depleted in M")
)
species_keep <- unique(origin_tbl$Species)

if (length(species_keep) == 0) stop("No species in combined list (species_keep).")

## ===== 2) Meta & species counts -> relative abundance (from merged_data1) =====
stopifnot(all(c("FREG1_Barcode","FREG5_Ethnic_Group") %in% names(merged_data1)))

# species counts table with SampleID rownames
sp_df <- merged_data1[, !(names(merged_data1) %in% c("FREG5_Ethnic_Group"))]
names(sp_df)[names(sp_df) == "FREG1_Barcode"] <- "SampleID"
sp_df <- as.data.frame(sp_df, stringsAsFactors = FALSE)
rownames(sp_df) <- sp_df$SampleID; sp_df$SampleID <- NULL
sp_df[] <- lapply(sp_df, function(x) suppressWarnings(as.numeric(x)))

# relative abundance
lib <- rowSums(sp_df, na.rm = TRUE)
rel <- sp_df / ifelse(lib == 0, 1, lib)
rel[!is.finite(as.matrix(rel))] <- 0

## Keep only selected species that exist in the table
species_use <- intersect(species_keep, colnames(rel))
if (length(species_use) == 0) stop("None of the selected species are present in merged_data1.")

## ===== 3) Food groups (first col = Participant_ID) =====
fd <- as.data.frame(food_group, stringsAsFactors = FALSE)
names(fd)[names(fd) == "Participant_ID"] <- "SampleID"
food_cols <- setdiff(names(fd), "SampleID")
fd[food_cols] <- lapply(fd[food_cols], function(x) suppressWarnings(as.numeric(x)))

## ===== 4) Align samples =====
ids <- intersect(rownames(rel), fd$SampleID)
rel  <- rel[ids, species_use, drop = FALSE]
fd   <- fd[match(ids, fd$SampleID), , drop = FALSE]
stopifnot(identical(rownames(rel), fd$SampleID))

## ===== 5) Filter by within-dataset prevalence/variance (pooled, no stratification) =====
keep_sp <- vapply(colnames(rel), function(s) {
  x <- rel[, s]; mean(x > 0, na.rm = TRUE) >= min_prev_species && stats::var(x, na.rm = TRUE) > 0
}, logical(1))
sp_keep <- colnames(rel)[keep_sp]

keep_food <- vapply(food_cols, function(f) {
  y <- fd[[f]]; mean(is.finite(y)) >= min_prev_food && stats::var(y, na.rm = TRUE) > 0
}, logical(1))
food_keep <- food_cols[keep_food]

if (length(sp_keep) == 0 || length(food_keep) == 0) {
  stop("After prevalence/variance filtering, no analyzable species or foods left.")
}

## ===== 6) Spearman correlations (pooled across all samples) =====
out <- vector("list", length(sp_keep))
for (i in seq_along(sp_keep)) {
  s <- sp_keep[i]; x <- rel[, s]
  rows <- lapply(food_keep, function(f) {
    y <- fd[[f]]
    ok <- is.finite(x) & is.finite(y)
    if (sum(ok) < min_n_pair) return(NULL)
    ct <- suppressWarnings(cor.test(x[ok], y[ok], method = "spearman", exact = FALSE))
    data.frame(Species = s, Food = f, n = sum(ok),
               rho = unname(ct$estimate), p_value = ct$p.value,
               stringsAsFactors = FALSE)
  })
  out[[i]] <- do.call(rbind, rows)
}
res <- do.call(rbind, out)
if (is.null(res) || nrow(res) == 0) stop("No valid species–food pairs to test.")

## ===== 7) Multiple testing & annotation =====
res$p_adj <- p.adjust(res$p_value, method = "fdr")
res <- merge(res, origin_tbl, by = "Species", all.x = TRUE)

sig_corr <- subset(res, !is.na(p_adj) & p_adj < alpha & abs(rho) >= min_abs_rho)
sig_corr <- sig_corr[order(sig_corr$p_adj, -abs(sig_corr$rho)), ]

## ===== 8) (Optional) Save & quick peek =====
# write.csv(sig_corr, "Spearman_foodGroups_selectedSpecies_allSamples.csv", row.names = FALSE)
```


```{r}
sig_corr
table(sig_corr$Origin)   # how many hits per original DA-origin tag

```
```{r}
# Load required libraries
library(dplyr)
library(corrplot)
library(ggplot2)
library(pheatmap)
library(RColorBrewer)

# Step 1: Combine all ethnic-specific species
all_ethnic_species <- c(enriched_C, depleted_C, 
                       enriched_I, depleted_I, 
                       enriched_M, depleted_M)

# Remove duplicates if any
all_ethnic_species <- unique(all_ethnic_species)

print(paste("Total number of unique ethnic-specific species:", length(all_ethnic_species)))

# Step 2: Filter species data to include only ethnic-specific species
# First check which species are actually present in the data
available_species <- intersect(all_ethnic_species, colnames(species_filtered))
missing_species <- setdiff(all_ethnic_species, colnames(species_filtered))

if (length(missing_species) > 0) {
  print(paste("Warning: The following species are not found in species_filtered:"))
  print(missing_species)
}

print(paste("Available ethnic-specific species:", length(available_species)))

species_subset <- species_filtered %>%
  select(FREG1_Barcode, all_of(available_species))

# Step 3: Merge all datasets
# First merge species with ethnicity data
merged_data <- species_subset %>%
  inner_join(ethnicity_800, by = "FREG1_Barcode")

# Then merge with food group data (matching Participant_ID with FREG1_Barcode)
final_data <- merged_data %>%
  inner_join(food_group, by = c("FREG1_Barcode" = "Participant_ID"))

print(paste("Final dataset dimensions:", nrow(final_data), "x", ncol(final_data)))

# Step 4: Prepare data for partial correlation analysis
# Get food group columns (exclude ID and metadata columns)
available_food_cols <- setdiff(colnames(food_group), "Participant_ID")
species_cols <- available_species  # Use available species instead
confounder_cols <- c("FREG7_Gender", "FREG8_Age", "BMI", "FREG5_Ethnic_Group")

# Convert gender to numeric (if not already)
final_data$FREG7_Gender_num <- ifelse(final_data$FREG7_Gender == "M", 1, 0)

# Convert ethnicity to dummy variables (handle potential issues)
tryCatch({
  ethnicity_dummies <- model.matrix(~ FREG5_Ethnic_Group - 1, data = final_data)
  # Remove the "FREG5_Ethnic_Group" prefix from column names for cleaner names
  colnames(ethnicity_dummies) <- gsub("FREG5_Ethnic_Group", "Ethnicity_", colnames(ethnicity_dummies))
}, error = function(e) {
  # If model.matrix fails, create a simple numeric encoding
  unique_ethnicities <- unique(final_data$FREG5_Ethnic_Group)
  ethnicity_dummies <- matrix(0, nrow = nrow(final_data), ncol = length(unique_ethnicities))
  colnames(ethnicity_dummies) <- paste0("Ethnicity_", unique_ethnicities)
  
  for (i in seq_along(unique_ethnicities)) {
    ethnicity_dummies[final_data$FREG5_Ethnic_Group == unique_ethnicities[i], i] <- 1
  }
})

# Create final analysis dataset with better error handling
required_cols <- c(species_cols, available_food_cols, "FREG7_Gender_num", "FREG8_Age", "BMI")
available_required_cols <- intersect(required_cols, colnames(final_data))

analysis_data <- final_data[, available_required_cols]

# Add ethnicity dummies
analysis_data <- cbind(analysis_data, ethnicity_dummies)

# Clean the data: convert all columns to numeric where possible
for (col in colnames(analysis_data)) {
  if (is.character(analysis_data[[col]]) || is.factor(analysis_data[[col]])) {
    # Try to convert to numeric, if it fails keep as is
    tryCatch({
      analysis_data[[col]] <- as.numeric(as.character(analysis_data[[col]]))
    }, warning = function(w) {
      # If conversion introduces NAs, that's okay
    })
  }
}

# Replace infinite values with NA
analysis_data[] <- lapply(analysis_data, function(x) {
  x[is.infinite(x)] <- NA
  return(x)
})

# Check for columns with too many missing values
missing_prop <- colSums(is.na(analysis_data)) / nrow(analysis_data)
cols_to_keep <- names(missing_prop[missing_prop < 0.9])  # Keep columns with <90% missing

analysis_data <- analysis_data[, cols_to_keep]

# Update species and food columns lists based on what's available
species_cols <- intersect(species_cols, colnames(analysis_data))
food_cols <- intersect(available_food_cols, colnames(analysis_data))

print(paste("Analysis dataset dimensions:", nrow(analysis_data), "x", ncol(analysis_data)))
print(paste("Species columns available:", length(species_cols)))
print(paste("Food group columns available:", length(food_cols)))

# Step 5: Function to calculate partial Spearman correlation
partial_spearman_cor <- function(x, y, confounders_matrix) {
  # Create a complete dataset
  complete_data <- data.frame(
    x = as.numeric(x),
    y = as.numeric(y),
    as.data.frame(confounders_matrix)
  )
  
  # Remove rows with any NA, NaN, or Inf values
  complete_data <- complete_data[complete.cases(complete_data), ]
  complete_data <- complete_data[apply(complete_data, 1, function(row) all(is.finite(row))), ]
  
  # Check if we have enough data points
  if (nrow(complete_data) < 15) {
    return(list(correlation = NA, p_value = NA))
  }
  
  # Check for zero variance (constant values)
  if (var(complete_data$x) == 0 || var(complete_data$y) == 0) {
    return(list(correlation = NA, p_value = NA))
  }
  
  # Check for multicollinearity in confounders
  confounder_data <- complete_data[, -(1:2)]  # exclude x and y
  if (ncol(confounder_data) > 0) {
    # Remove constant columns
    confounder_data <- confounder_data[, sapply(confounder_data, function(col) var(col) > 0)]
    
    # Check if we still have confounders after removing constant columns
    if (ncol(confounder_data) == 0) {
      # No valid confounders, just do simple correlation
      cor_test <- cor.test(complete_data$x, complete_data$y, method = "spearman")
      return(list(correlation = cor_test$estimate, p_value = cor_test$p.value))
    }
    
    # Try to fit models with error handling
    tryCatch({
      # Fit linear models to remove confounder effects
      x_model <- lm(x ~ ., data = cbind(x = complete_data$x, confounder_data))
      y_model <- lm(y ~ ., data = cbind(y = complete_data$y, confounder_data))
      
      x_residuals <- residuals(x_model)
      y_residuals <- residuals(y_model)
      
      # Calculate Spearman correlation on residuals
      cor_test <- cor.test(x_residuals, y_residuals, method = "spearman")
      return(list(correlation = cor_test$estimate, p_value = cor_test$p.value))
      
    }, error = function(e) {
      # If linear model fails, fall back to simple correlation
      cor_test <- cor.test(complete_data$x, complete_data$y, method = "spearman")
      return(list(correlation = cor_test$estimate, p_value = cor_test$p.value))
    })
    
  } else {
    # No confounders, just do simple correlation
    cor_test <- cor.test(complete_data$x, complete_data$y, method = "spearman")
    return(list(correlation = cor_test$estimate, p_value = cor_test$p.value))
  }
}

# Step 6: Calculate partial correlations
# Prepare confounder matrix (ensure all columns are numeric and finite)
confounder_cols_available <- c("FREG7_Gender_num", "FREG8_Age", "BMI", 
                               grep("Ethnicity_", colnames(analysis_data), value = TRUE))

# Only keep confounder columns that exist in the data
confounder_cols_available <- intersect(confounder_cols_available, colnames(analysis_data))

print(paste("Available confounder columns:", length(confounder_cols_available)))
print(confounder_cols_available)

confounder_matrix <- as.matrix(analysis_data[, confounder_cols_available])

# Ensure confounder matrix is numeric
confounder_matrix <- apply(confounder_matrix, 2, as.numeric)

# Initialize correlation matrix
correlation_matrix <- matrix(NA, 
                           nrow = length(species_cols), 
                           ncol = length(food_cols),
                           dimnames = list(species_cols, food_cols))

# Initialize p-value matrix
p_value_matrix <- matrix(NA, 
                        nrow = length(species_cols), 
                        ncol = length(food_cols),
                        dimnames = list(species_cols, food_cols))

# Calculate correlations
print("Calculating partial Spearman correlations...")
for (i in 1:length(species_cols)) {
  for (j in 1:length(food_cols)) {
    species_data <- analysis_data[[species_cols[i]]]
    food_data <- analysis_data[[food_cols[j]]]
    
    # Calculate partial correlation
    result <- partial_spearman_cor(species_data, food_data, confounder_matrix)
    correlation_matrix[i, j] <- result$correlation
    p_value_matrix[i, j] <- result$p_value
  }
  if (i %% 10 == 0) print(paste("Processed", i, "out of", length(species_cols), "species"))
}

# Step 7: Apply multiple testing correction
p_value_adj <- p.adjust(as.vector(p_value_matrix), method = "fdr")
p_value_adj_matrix <- matrix(p_value_adj, 
                            nrow = nrow(p_value_matrix), 
                            ncol = ncol(p_value_matrix),
                            dimnames = dimnames(p_value_matrix))

# Step 8: Create significance matrix for plotting
significance_matrix <- matrix("", 
                             nrow = nrow(correlation_matrix), 
                             ncol = ncol(correlation_matrix),
                             dimnames = dimnames(correlation_matrix))

significance_matrix[p_value_adj_matrix < 0.001] <- "***"
significance_matrix[p_value_adj_matrix >= 0.001 & p_value_adj_matrix < 0.01] <- "**"
significance_matrix[p_value_adj_matrix >= 0.01 & p_value_adj_matrix < 0.05] <- "*"

# Step 9: Create correlation heatmap
# Remove rows/columns with all NA values
valid_rows <- rowSums(!is.na(correlation_matrix)) > 0
valid_cols <- colSums(!is.na(correlation_matrix)) > 0

correlation_clean <- correlation_matrix[valid_rows, valid_cols]
significance_clean <- significance_matrix[valid_rows, valid_cols]

# Check if we have enough valid correlations for plotting
print(paste("Valid correlations for plotting:", sum(!is.na(correlation_clean))))
print(paste("Matrix dimensions:", nrow(correlation_clean), "x", ncol(correlation_clean)))

if (sum(!is.na(correlation_clean)) > 0 && nrow(correlation_clean) > 0 && ncol(correlation_clean) > 0) {
  
  # Create heatmap using pheatmap
  tryCatch({
    png("species_foodgroup_correlation_heatmap.png", width = 12, height = 10, units = "in", res = 300)
    
    pheatmap(correlation_clean,
             color = colorRampPalette(c("blue", "white", "red"))(100),
             breaks = seq(-1, 1, length.out = 101),
             display_numbers = significance_clean,
             number_color = "black",
             fontsize_number = 6,
             cluster_rows = ifelse(nrow(correlation_clean) > 2, TRUE, FALSE),
             cluster_cols = ifelse(ncol(correlation_clean) > 2, TRUE, FALSE),
             main = "Partial Spearman Correlations:\nEthnic-Specific Species vs Food Groups\n(Adjusted for Age, Gender, BMI, Ethnicity)",
             fontsize = 8,
             angle_col = 45,
             na_col = "grey90")
    
    dev.off()
    print("Pheatmap saved successfully!")
    
  }, error = function(e) {
    dev.off()
    print(paste("Pheatmap failed:", e$message))
  })
  
  # Alternative heatmap using corrplot with better error handling
  tryCatch({
    # For corrplot, we need to handle NAs and ensure proper matrix format
    correlation_for_corrplot <- correlation_clean
    
    # Replace remaining NAs with 0 for visualization purposes only
    correlation_for_corrplot[is.na(correlation_for_corrplot)] <- 0
    
    # Ensure the matrix is properly bounded between -1 and 1
    correlation_for_corrplot[correlation_for_corrplot > 1] <- 1
    correlation_for_corrplot[correlation_for_corrplot < -1] <- -1
    
    png("species_foodgroup_correlation_corrplot.png", width = 14, height = 10, units = "in", res = 300)
    
    # Use different order method if hclust fails
    plot_order <- "original"
    if (nrow(correlation_for_corrplot) > 2 && ncol(correlation_for_corrplot) > 2 && 
        sum(!is.na(correlation_clean)) > (nrow(correlation_clean) * ncol(correlation_clean) * 0.5)) {
      plot_order <- "hclust"
    }
    
    corrplot(correlation_for_corrplot,
             method = "color",
             type = "full",
             order = plot_order,
             tl.cex = 0.6,
             tl.col = "black",
             tl.srt = 45,
             col = colorRampPalette(c("blue", "white", "red"))(100),
             title = "Partial Spearman Correlations: Ethnic-Specific Species vs Food Groups\n(Adjusted for Age, Gender, BMI, Ethnicity)",
             mar = c(0, 0, 4, 0),
             na.label = "square",
             na.label.col = "grey90")
    
    dev.off()
    print("Corrplot saved successfully!")
    
  }, error = function(e) {
    dev.off()
    print(paste("Corrplot failed:", e$message))
    
    # Fallback: create a simple ggplot heatmap
    tryCatch({
      library(reshape2)
      library(ggplot2)
      
      # Convert correlation matrix to long format
      correlation_melted <- melt(correlation_clean, na.rm = FALSE)
      colnames(correlation_melted) <- c("Species", "Food_Group", "Correlation")
      
      # Create ggplot heatmap
      p <- ggplot(correlation_melted, aes(x = Food_Group, y = Species, fill = Correlation)) +
        geom_tile() +
        scale_fill_gradient2(low = "blue", mid = "white", high = "red", 
                           midpoint = 0, limit = c(-1, 1), na.value = "grey90") +
        theme_minimal() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
              axis.text.y = element_text(size = 8),
              plot.title = element_text(size = 12, hjust = 0.5)) +
        labs(title = "Partial Spearman Correlations:\nEthnic-Specific Species vs Food Groups\n(Adjusted for Age, Gender, BMI, Ethnicity)",
             x = "Food Groups", y = "Species")
      
      ggsave("species_foodgroup_correlation_ggplot.png", p, width = 12, height = 8, dpi = 300)
      print("Fallback ggplot heatmap saved successfully!")
      
    }, error = function(e2) {
      print(paste("All plotting methods failed:", e2$message))
    })
  })
  
} else {
  print("No valid correlations found for plotting. Check your data and analysis parameters.")
}

# Step 10: Summary statistics
print("\n=== SUMMARY STATISTICS ===")
print(paste("Number of species analyzed:", nrow(correlation_clean)))
print(paste("Number of food groups analyzed:", ncol(correlation_clean)))
print(paste("Total correlations calculated:", sum(!is.na(correlation_clean))))

# Significant correlations summary
sig_001 <- sum(p_value_adj_matrix < 0.001, na.rm = TRUE)
sig_01 <- sum(p_value_adj_matrix >= 0.001 & p_value_adj_matrix < 0.01, na.rm = TRUE)
sig_05 <- sum(p_value_adj_matrix >= 0.01 & p_value_adj_matrix < 0.05, na.rm = TRUE)

print(paste("Significant correlations (p < 0.001):", sig_001))
print(paste("Significant correlations (0.001 ≤ p < 0.01):", sig_01))
print(paste("Significant correlations (0.01 ≤ p < 0.05):", sig_05))

# Correlation range
print(paste("Correlation range:", round(min(correlation_clean, na.rm = TRUE), 3), "to", round(max(correlation_clean, na.rm = TRUE), 3)))

# Step 11: Save results
write.csv(correlation_matrix, "species_foodgroup_correlations.csv")
write.csv(p_value_adj_matrix, "species_foodgroup_pvalues_adjusted.csv")

# Create results summary table
results_summary <- data.frame(
  Species = rep(rownames(correlation_clean), ncol(correlation_clean)),
  Food_Group = rep(colnames(correlation_clean), each = nrow(correlation_clean)),
  Correlation = as.vector(correlation_clean),
  P_value_adjusted = as.vector(p_value_adj_matrix[valid_rows, valid_cols]),
  Significance = as.vector(significance_clean)
)

# Remove NA rows and sort by absolute correlation
results_summary <- results_summary[!is.na(results_summary$Correlation), ]
results_summary <- results_summary[order(-abs(results_summary$Correlation)), ]

write.csv(results_summary, "species_foodgroup_correlation_summary.csv", row.names = FALSE)

print("\nFiles saved:")
file_list <- c()
if (file.exists("species_foodgroup_correlation_heatmap.png")) {
  file_list <- c(file_list, "1. species_foodgroup_correlation_heatmap.png")
}
if (file.exists("species_foodgroup_correlation_corrplot.png")) {
  file_list <- c(file_list, "2. species_foodgroup_correlation_corrplot.png")
}
if (file.exists("species_foodgroup_correlation_ggplot.png")) {
  file_list <- c(file_list, "3. species_foodgroup_correlation_ggplot.png")
}
file_list <- c(file_list, "4. species_foodgroup_correlations.csv")
file_list <- c(file_list, "5. species_foodgroup_pvalues_adjusted.csv")
file_list <- c(file_list, "6. species_foodgroup_correlation_summary.csv")

for (file in file_list) {
  print(file)
}

print("\nAnalysis completed successfully!")
```


```{r}
library(openxlsx)

# Combine all ethnic-specific species
all_ethnic_species <- unique(c(enriched_C, depleted_C, 
                              enriched_I, depleted_I, 
                              enriched_M, depleted_M))

# Extract data with participant IDs and all ethnic-specific species
ethnic_species_data <- species_filtered %>%
  select(FREG1_Barcode, all_of(all_ethnic_species))
ethnic_species_data
# Export to Excel
write.xlsx(ethnic_species_data, "figure/All_Ethnic_Specific_Species602.xlsx")
```


```{r}
all_ethnic_species <- unique(c(enriched_C, depleted_C, 
                              enriched_I, depleted_I, 
                              enriched_M, depleted_M))

# Create dataframe with species names
species_df <- data.frame(species = all_ethnic_species)

# Export to CSV
write.csv(species_df, "figure/ethnic_species_names602.csv", row.names = FALSE)
```


```{r}
# ANOVA Analysis: Species Abundance vs Ethnicity (C, I, M)
# Select top species associated with ethnicity

library(dplyr)

# Assuming you have:
# - filtered_data: your species abundance data (samples x species)
# - ethnicity: vector with values "C", "I", "M" for each sample

# If you need to create the ethnicity vector from sample names or separate file:
# ethnicity <- c("C", "I", "M", "C", "I", ...)  # match your sample order
# OR read from metadata file:
# metadata <- read.csv("metadata.csv")
# ethnicity <- metadata$ethnicity

# Method 1: ANOVA for all species
# ===============================

perform_anova_analysis <- function(species_data, ethnicity_vector) {
  # Ensure ethnicity is a factor
  ethnicity_vector <- as.factor(ethnicity_vector)
  
  # Storage for results
  anova_results <- data.frame(
    species = character(),
    f_statistic = numeric(),
    p_value = numeric(),
    adj_p_value = numeric(),
    stringsAsFactors = FALSE
  )
  
  # Run ANOVA for each species
  for (i in 1:ncol(species_data)) {
    species_name <- colnames(species_data)[i]
    abundance <- species_data[, i]
    
    # Skip if all values are zero
    if (sum(abundance) == 0) next
    
    # Perform ANOVA
    anova_model <- aov(abundance ~ ethnicity_vector)
    anova_summary <- summary(anova_model)
    
    # Extract results
    f_stat <- anova_summary[[1]]["ethnicity_vector", "F value"]
    p_val <- anova_summary[[1]]["ethnicity_vector", "Pr(>F)"]
    
    # Store results
    anova_results <- rbind(anova_results, data.frame(
      species = species_name,
      f_statistic = f_stat,
      p_value = p_val,
      adj_p_value = NA
    ))
  }
  
  # Adjust p-values for multiple testing (FDR)
  anova_results$adj_p_value <- p.adjust(anova_results$p_value, method = "fdr")
  
  # Sort by significance
  anova_results <- anova_results[order(anova_results$adj_p_value), ]
  
  return(anova_results)
}

# Method 2: Quick ANOVA with top species selection
# ===============================================

select_ethnicity_species <- function(species_data, ethnicity_vector, 
                                   p_threshold = 0.05, top_n = 20) {
  
  # Run ANOVA analysis
  results <- perform_anova_analysis(species_data, ethnicity_vector)
  
  # Filter significant species
  significant_species <- results[results$adj_p_value < p_threshold, ]
  
  # Select top N species
  if (nrow(significant_species) > top_n) {
    significant_species <- significant_species[1:top_n, ]
  }
  
  print(paste("Found", nrow(significant_species), "significant species"))
  print("Top significant species:")
  print(significant_species[1:min(10, nrow(significant_species)), ])
  
  # Return filtered data with only significant species
  selected_species_data <- species_data[, significant_species$species]
  
  return(list(
    data = selected_species_data,
    results = significant_species,
    all_results = results
  ))
}

# Usage with your merged_data1:
# =============================

# Load your data


# Extract ethnicity from column 2 (FREG5_Ethnic_Group)
ethnicity <- merged_data1$FREG5_Ethnic_Group

# Extract species data (columns 3 onwards)
species_data <- merged_data1[, 3:ncol(merged_data1)]
rownames(species_data) <- merged_data1$FREG1_Barcode  # Use barcode as row names

# Convert species data to numeric and handle NAs
species_data[] <- lapply(species_data, function(x) as.numeric(as.character(x)))
species_data[is.na(species_data)] <- 0

# Check the data
print(paste("Number of samples:", nrow(species_data)))
print(paste("Number of species:", ncol(species_data)))
print("Ethnic groups:")
print(table(ethnicity))

# Run the ANOVA analysis
ethnicity_analysis <- select_ethnicity_species(species_data, ethnicity, 
                                              p_threshold = 0.05, top_n = 20)

# View results
print("ANOVA Results Summary:")
print(ethnicity_analysis$results)

# Extract the selected species data
ethnicity_associated_species <- ethnicity_analysis$data
ethnicity_analysis$results
```


```{r}
species_anova <- data.frame(species = ethnicity_analysis$results$species)

# Export to CSV
write.csv(species_anova, "data/ethnic_species_anova.csv", row.names = FALSE)
```

## common food
```{r}
# Food Intake Analysis: Most Common Foods + Ethnicity Association
# Two main outputs: Top common foods & ANOVA results for ethnicity

library(dplyr)
library(ggplot2)
library(tidyr)

# Load your data


# Extract data components
ethnicity <- merged_data2$FREG5_Ethnic_Group
food_data <- merged_data2[, 3:ncol(merged_data2)]
rownames(food_data) <- merged_data2$FREG1_Barcode

# Convert food data to numeric and handle NAs
food_data[] <- lapply(food_data, function(x) as.numeric(as.character(x)))
food_data[is.na(food_data)] <- 0

print(paste("Number of samples:", nrow(food_data)))
print(paste("Number of food items:", ncol(food_data)))
print("Ethnic groups distribution:")
print(table(ethnicity))

# ===========================================
# LIST 1: TOP MOST COMMON FOODS
# ===========================================

create_common_foods_list <- function(food_data, top_n = 20) {
  
  # Method 1: By total consumption (sum across all samples)
  total_consumption <- colSums(food_data, na.rm = TRUE)
  
  # Method 2: By frequency (how many people consume each food)
  consumption_frequency <- colSums(food_data > 0, na.rm = TRUE)
  
  # Method 3: By mean consumption (average per person)
  mean_consumption <- colMeans(food_data, na.rm = TRUE)
  
  # Create comprehensive summary
  common_foods_summary <- data.frame(
    food_item = names(total_consumption),
    total_consumption = total_consumption,
    frequency_consumers = consumption_frequency,
    percent_consumers = round((consumption_frequency / nrow(food_data)) * 100, 1),
    mean_consumption = round(mean_consumption, 3),
    stringsAsFactors = FALSE
  )
  
  # Sort by total consumption
  common_foods_by_total <- common_foods_summary[order(common_foods_summary$total_consumption, decreasing = TRUE), ]
  
  # Sort by frequency (most consumed by most people)
  common_foods_by_frequency <- common_foods_summary[order(common_foods_summary$frequency_consumers, decreasing = TRUE), ]
  
  print("=== LIST 1: TOP MOST COMMON FOODS ===")
  print("Top foods by TOTAL consumption:")
  print(head(common_foods_by_total, top_n))
  
  print("\nTop foods by FREQUENCY (most people consume):")
  print(head(common_foods_by_frequency, top_n))
  
  return(list(
    by_total = head(common_foods_by_total, top_n),
    by_frequency = head(common_foods_by_frequency, top_n),
    all_summary = common_foods_summary
  ))
}

# ===========================================
# LIST 2: ANOVA - FOODS ASSOCIATED WITH ETHNICITY
# ===========================================

perform_food_ethnicity_anova <- function(food_data, ethnicity_vector, top_n = 20) {
  
  # Ensure ethnicity is a factor
  ethnicity_vector <- as.factor(ethnicity_vector)
  
  # Storage for ANOVA results
  anova_results <- data.frame(
    food_item = character(),
    f_statistic = numeric(),
    p_value = numeric(),
    adj_p_value = numeric(),
    mean_C = numeric(),
    mean_I = numeric(),
    mean_M = numeric(),
    stringsAsFactors = FALSE
  )
  
  # Run ANOVA for each food item
  for (i in 1:ncol(food_data)) {
    food_name <- colnames(food_data)[i]
    intake <- food_data[, i]
    
    # Skip if all values are zero
    if (sum(intake, na.rm = TRUE) == 0) next
    
    # Calculate group means
    means_by_ethnicity <- tapply(intake, ethnicity_vector, mean, na.rm = TRUE)
    
    # Perform ANOVA
    anova_model <- aov(intake ~ ethnicity_vector)
    anova_summary <- summary(anova_model)
    
    # Extract F-statistic and p-value
    f_stat <- anova_summary[[1]]["ethnicity_vector", "F value"]
    p_val <- anova_summary[[1]]["ethnicity_vector", "Pr(>F)"]
    
    # Store results
    anova_results <- rbind(anova_results, data.frame(
      food_item = food_name,
      f_statistic = round(f_stat, 3),
      p_value = p_val,
      adj_p_value = NA,
      mean_C = round(means_by_ethnicity["C"], 3),
      mean_I = round(means_by_ethnicity["I"], 3),
      mean_M = round(means_by_ethnicity["M"], 3)
    ))
  }
  
  # Adjust p-values for multiple testing
  anova_results$adj_p_value <- p.adjust(anova_results$p_value, method = "fdr")
  
  # Sort by significance
  anova_results <- anova_results[order(anova_results$adj_p_value), ]
  
  # Filter significant results
  significant_foods <- anova_results[anova_results$adj_p_value < 0.05, ]
  
  print("=== LIST 2: FOODS ASSOCIATED WITH ETHNICITY (ANOVA) ===")
  print(paste("Found", nrow(significant_foods), "significantly different foods (p < 0.05)"))
  print("Top foods with ethnic differences:")
  print(head(significant_foods, top_n))
  
  return(list(
    significant = significant_foods,
    all_results = anova_results
  ))
}

# ===========================================
# RUN BOTH ANALYSES
# ===========================================

# Generate LIST 1: Most Common Foods
common_foods_analysis <- create_common_foods_list(food_data, top_n = 20)
common_foods_analysis
# Generate LIST 2: Ethnicity-Associated Foods
ethnicity_foods_analysis <- perform_food_ethnicity_anova(food_data, ethnicity, top_n = 20)
ethnicity_foods_analysis
# ===========================================
# EXPORT RESULTS
# ===========================================

# # Save LIST 1: Common Foods
# write.csv(common_foods_analysis$by_total, "LIST1_most_common_foods_by_total.csv", row.names = FALSE)
# write.csv(common_foods_analysis$by_frequency, "LIST1_most_common_foods_by_frequency.csv", row.names = FALSE)
# 
# # Save LIST 2: Ethnicity-Associated Foods
# write.csv(ethnicity_foods_analysis$significant, "LIST2_ethnicity_associated_foods.csv", row.names = FALSE)
# write.csv(ethnicity_foods_analysis$all_results, "LIST2_all_anova_results.csv", row.names = FALSE)

# ===========================================
# VISUALIZATION
# ===========================================

# Plot 1: Top common foods
# plot_common_foods <- function(common_foods, top_n = 10) {
#   top_foods <- head(common_foods$by_frequency, top_n)
#   
#   p1 <- ggplot(top_foods, aes(x = reorder(food_item, percent_consumers), y = percent_consumers)) +
#     geom_bar(stat = "identity", fill = "skyblue", alpha = 0.7) +
#     coord_flip() +
#     theme_minimal() +
#     labs(title = "Most Commonly Consumed Foods",
#          x = "Food Item", y = "% of People Who Consume") +
#     theme(axis.text.y = element_text(size = 8))
#   
#   return(p1)
# }
# 
# # Plot 2: Ethnicity differences for top significant foods
# plot_ethnicity_foods <- function(food_data, ethnicity_vector, anova_results, top_n = 6) {
#   
#   # Get top significant foods
#   top_foods <- head(anova_results$significant$food_item, top_n)
#   
#   # Prepare data
#   plot_data <- food_data[, top_foods]
#   plot_data$ethnicity <- ethnicity_vector
#   
#   # Reshape for plotting
#   plot_data_long <- pivot_longer(plot_data, 
#                                 cols = -ethnicity,
#                                 names_to = "food_item",
#                                 values_to = "intake")
#   
#   p2 <- ggplot(plot_data_long, aes(x = ethnicity, y = intake, fill = ethnicity)) +
#     geom_boxplot(alpha = 0.7) +
#     geom_jitter(width = 0.2, alpha = 0.4, size = 0.5) +
#     facet_wrap(~food_item, scales = "free_y", ncol = 3) +
#     theme_minimal() +
#     labs(title = "Foods with Significant Ethnic Differences",
#          x = "Ethnicity", y = "Food Intake") +
#     theme(axis.text.x = element_text(angle = 0),
#           strip.text = element_text(size = 8)) +
#     scale_fill_brewer(palette = "Set2")
#   
#   return(p2)
# }
# 
# # Create and save plots
# p1 <- plot_common_foods(common_foods_analysis, top_n = 15)
# ggsave("most_common_foods_plot.pdf", p1, width = 10, height = 8)
# 
# p2 <- plot_ethnicity_foods(food_data, ethnicity, ethnicity_foods_analysis, top_n = 9)
# ggsave("ethnicity_food_differences_plot.pdf", p2, width = 12, height = 8)
# 
# print("\n=== ANALYSIS COMPLETE ===")
# print("Files created:")
# print("- LIST1_most_common_foods_by_total.csv")
# print("- LIST1_most_common_foods_by_frequency.csv") 
# print("- LIST2_ethnicity_associated_foods.csv")
# print("- LIST2_all_anova_results.csv")
# print("- most_common_foods_plot.pdf")
# print("- ethnicity_food_differences_plot.pdf")
```

```{r}
common_foods_analysis
ethnicity_foods_analysis
```

common food: water, vegetable, rice, coffee, egg, wholemeal, tea, tofu, nuts butter, noodle, soup, poultry, white bread, white fish, broccoli, carrot, french_fries, white cabbage, spinach, banana, mushrooms, meat, apple
ethnic food: dhal,thosai, soup_rasam, chapati, begedil,puri, roti, idli, murtabak, pork, longtong, mutton, raita, chye_sim, seafood,  lotus_root, dumpling


1.Coffee 
2. Tea

3. Rice_Dishes (rice, fried rice, nasi lemak)

4. bakery bread (white bread, wholemeal)

5. Chapati 
6. Thosai 

7. Other_Traditional_Breads (roti, puri, murtabak)

8. Noodles_dishes
9. Leafy_Vegetables
10. mushroom
11. Root_Vegetables
12. Legumes

13. Eggs

14. Poultry
15. Red_Meat
16. Pork
17. Seafood
18. Fruits
19. Nuts_Seeds
20. Soups_Broths
21. Snacks_Fried
22. Dairy

```{r}
all_food <- read_excel("data/all_food_863.xlsx", col_names = FALSE)
ffqa <- read_excel("data/FFQA800.xlsx", col_names = FALSE)

# Find unmatched entries (in all_food but not in ffqa)
unmatched <- setdiff(ffqa, all_food)
unmatched
colnames(all_food)
ffqa
```

```{r}
library(readxl)
library(openxlsx)   # for writing back to Excel

# Read the data
all_food <- read_excel("data/all_food_863.xlsx")

# --- Define your groups ---
rice_dishes <- c(
  "plain_rice_type_mixture","plain_rice_type_unpolished","plain_rice_type_white",
  "flav_porr_type_mixture","flav_porr_type_unpolished","flav_porr_type_white",
  "plain_porr_type_mixture","plain_porr_type_unpolished","plain_porr_type_white",
  "fried_rice_with_veg","fried_rice_type_mixture","fried_rice_type_unpolished",
  "fried_rice_type_white","fried_rice_without_veg","flav_rice","flav_rice_dish",
  "glut_rice","lontong","idli","rice_roll"
)

bakery_bread <- c(
  "white_bread","wholemeal_bread","dried_fruit_bread","french_toast","pizza",
  "thosai","roti","chapati","murtabak","puri","sweet_filled_buns","savoury_filled_buns"
)

leafy_veg <- c(
  "broccoli","chye_sim","kailan","kang_kong","spinach","white_cabbage",
  "lettuce","coleslaw"
)

root_veg <- c("lotus_root","yam","potato","carrot")

eggs <- c(
  "egg_mix_pork","egg_mix_chicken","egg_mix_seafood","egg_mix_beancurd",
  "egg_mix_vege","egg_scrambled","egg_steamed"
)

# --- Add new columns with row sums ---
all_food$Rice_dishes   <- rowSums(all_food[, rice_dishes], na.rm = TRUE)
all_food$Bakery_bread  <- rowSums(all_food[, bakery_bread], na.rm = TRUE)
all_food$Leafy_veg     <- rowSums(all_food[, leafy_veg], na.rm = TRUE)
all_food$Root_veg      <- rowSums(all_food[, root_veg], na.rm = TRUE)
all_food$Eggs          <- rowSums(all_food[, eggs], na.rm = TRUE)

# --- Save back to Excel ---
write.xlsx(all_food, "data/all_food_863_new.xlsx")

```





