---
title: "pcoa_eth_food"
author: "Ruwen Zhou"
date: "9/2/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(plyr)
library(tidyverse)
library(dplyr)
library(readxl)
library(ggplot2)
library(phyloseq)
library(vegan)
library(DESeq2)
library(pheatmap)
library(RColorBrewer)
library(ggpubr)
```

```{r}
food_data = read_xlsx("data/all_food_863.xlsx")
nutrition_data = read_xlsx("data/FFQB800.xlsx")
species = read_xlsx("data/species_866.xlsx")
metadata = read_xlsx("data/ethnicity_866.xlsx")

food_data
nutrition_data
species
metadata
```

#species
```{r}
species = read_csv("data/abundance_bracken_standard_16_20221209_taxa_S_unfiltered_reads_decont2.csv")
#species$`SAM id`
helios_new_list = metadata$FREG1_Barcode
#sample_list = metadata_micro$`Original Sample name`
metadata_micro = read_csv("data/metadata_fil_01032024_redacted.csv")

micro_list <- filter(metadata_micro, metadata_micro$`Original Sample name` %in% helios_new_list)
species_new = filter(species, species$`SAM id` %in% micro_list$`Library ID`)
species_new
```

```{r}
new_merge_spe = merge(metadata_micro[,c("Library ID", "Original Sample name")], species_new, by.x = "Library ID", by.y = "SAM id")
new_merge_spe
# species_meta = merge(x = helios_new[,c("FREG1_Barcode", "FREG5_Ethnic_Group")], y = new_merge, by.y = "Original Sample name", by.x = "FREG1_Barcode")
# species_meta = species_meta %>% drop_na(FREG5_Ethnic_Group)
# species_meta
```
# pcoa of food by ethnicity
```{r}
food_fil = filter(food_data, food_data$participant_id %in% helios_new_list)
food_df = as.data.frame(food_fil)
meta_fil = filter(metadata, metadata$FREG1_Barcode %in% food_fil$participant_id)
meta_df = as.data.frame(meta_fil)

rownames(food_df) = food_df[,1]
food_df[] <- lapply(food_df, function(x) as.double(as.character(x)))
food_df[1] = NULL
food_df
rownames(meta_df) = meta_df[,1]
meta_df
```
#pcoa
```{r}
# Filter out group "O"
meta_df_filtered <- meta_df[meta_df$FREG5_Ethnic_Group != "O", ]
food_df_filtered <- food_df[rownames(meta_df_filtered), ]
library(vegan)
food_df_filtered[is.na(food_df_filtered)] <- 0
# Calculate Bray-Curtis distance matrix
food_dist <- vegdist(food_df_filtered, method = "bray")
# Identify rows with all zeroes
zero_rows <- rowSums(food_df_filtered) == 0

# Remove those rows
food_df_filtered <- food_df_filtered[!zero_rows, ]

remaining_sample_ids <- rownames(food_df_filtered)

# Filter meta_df_filtered to only include the remaining sample IDs
meta_df_filtered <- meta_df_filtered[remaining_sample_ids, ]

# Recalculate the distance matrix
food_dist <- vegdist(food_df_filtered, method = "bray")

# Perform PCoA
pcoa_result <- cmdscale(food_dist, eig = TRUE, k = 2)
# Extract the coordinates
pcoa_coords <- as.data.frame(pcoa_result$points)

colnames(pcoa_coords) <- c("PCoA1", "PCoA2")
pcoa_coords$Ethnic_Group <- meta_df_filtered$FREG5_Ethnic_Group

# Plot the PCoA
# library(ggplot2)
# ethnicity_colors = c("C" = "#9FDDCA","I"= "#9EA3F1","M" = "#E17887","O" = "#80796BFF")
# ggplot(pcoa_coords, aes(x = PCoA1, y = PCoA2, color = Ethnic_Group)) +
#   geom_point(size = 2, alpha = 0.7) +  # Adjust 'size' for dot size and 'alpha' for transparency
#   stat_ellipse(aes(group = Ethnic_Group), type = "norm", linetype = "solid", size = 1) +
#   scale_color_manual(values = ethnicity_colors) +
#   scale_fill_manual(values = ethnicity_colors) +
#   theme_minimal() +
#   labs(
#     title = "PCoA Plot Food Grouped by Ethnic Group",
#     x = "PCoA1",
#     y = "PCoA2",
#     color = "Ethnic Group"
#   ) +
#   theme(
#     text = element_text(size = 14),
#     legend.position = "right"
#   )


```
```{r}
# Perform PERMANOVA to test for differences between groups
permanova_result <- adonis2(food_dist ~ FREG5_Ethnic_Group, data = meta_df_filtered)

p_value <- permanova_result$`Pr(>F)`[1]
p = ggplot(pcoa_coords, aes(x = PCoA1, y = PCoA2, color = Ethnic_Group)) +
  geom_point(size = 2, alpha = 0.7) +
  stat_ellipse(aes(group = Ethnic_Group), type = "norm", linetype = "solid", size = 1) +
  scale_color_manual(values = ethnicity_colors) +
  scale_fill_manual(values = ethnicity_colors) +
  theme_minimal() +
  labs(
    title = "PCoA Plot Food Grouped by Ethnic Group",
    x = "PCoA1",
    y = "PCoA2",
    color = "Ethnic Group"
  ) +
  annotate("text", 
           x = Inf, y = Inf, 
           label = paste("PERMANOVA p-value:", signif(p_value, 3)), 
           hjust = 1.1, vjust = 1.5, 
           size = 3, 
           color = "black") +
  theme(
    text = element_text(size = 14),
    legend.position = "right"
  )
plot_data <- p$data
p +
  geom_point(data = subset(plot_data, Ethnic_Group == "C"), color = "#9FDDCA") +
  geom_point(data = subset(plot_data, Ethnic_Group == "I"), color = "#9EA3F1") +
  geom_point(data = subset(plot_data, Ethnic_Group == "M"), color = "#E17887")

```
# NMDS
```{r}
# Perform NMDS using the Bray-Curtis distance matrix
nmds_result <- metaMDS(food_df_filtered, distance = "bray", k = 2, trymax = 100)


nmds_coords <- as.data.frame(scores(nmds_result, display = "sites"))
nmds_coords$Ethnic_Group <- meta_df_filtered$FREG5_Ethnic_Group

# Plot NMDS
p = ggplot(nmds_coords, aes(x = NMDS1, y = NMDS2, color = Ethnic_Group)) +
  geom_point(size = 2, alpha = 0.7) +
  stat_ellipse(aes(group = Ethnic_Group), type = "norm", linetype = "solid", size = 1) +
  scale_color_manual(values = ethnicity_colors) +
  scale_fill_manual(values = ethnicity_colors) +
  theme_minimal() +
  labs(
    title = "NMDS Plot Food Grouped by Ethnic Group",
    x = "NMDS1",
    y = "NMDS2",
    color = "Ethnic Group"
  ) +
  theme(
    text = element_text(size = 14),
    legend.position = "right"
  )
plot_data <- p$data
p +
  geom_point(data = subset(plot_data, Ethnic_Group == "C"), color = "#9FDDCA") +
  geom_point(data = subset(plot_data, Ethnic_Group == "I"), color = "#9EA3F1") +
  geom_point(data = subset(plot_data, Ethnic_Group == "M"), color = "#E17887")
  #theme(text = element_text(family = "serif"))

```

```{r}

nmds_permanova_result <- adonis2(food_dist ~ FREG5_Ethnic_Group, data = meta_df_filtered)

# Display the PERMANOVA result
nmds_permanova_result

# Extract the p-value
nmds_p_value <- nmds_permanova_result$`Pr(>F)`[1]
nmds_p_value

```



# pcoa of species by ethnicity (only the significant species)
```{r}
new_merge_spe = new_merge_spe[!duplicated(new_merge_spe$`Original Sample name`),]
metadata = metadata[!duplicated(metadata$FREG1_Barcode),]
species_df = filter(new_merge_spe, new_merge_spe$`Original Sample name` %in% metadata$FREG1_Barcode)
species_df = as.data.frame(species_df)
meta_spe_df = filter(metadata, metadata$FREG1_Barcode %in% species_df$`Original Sample name`)
meta_spe_df = as.data.frame(meta_spe_df)

rownames(species_df) = species_df[,2]
species_df[1] = NULL
species_df[1] = NULL
species_df

rownames(meta_spe_df) = meta_spe_df[,1]
meta_spe_df
```

```{r}
OTU = otu_table(as.matrix(species_df), taxa_are_rows = FALSE)              
SAM = sample_data(meta_spe_df, errorIfNULL = TRUE)

phylo <- phyloseq(OTU, SAM)
phylo_no_o <- subset_samples(phylo, FREG5_Ethnic_Group != "O")
phylo_no_o
sample_data(phylo_no_o)$FREG5_Ethnic_Grou
```
```{r}
is_prevalent <- function(x, group) {
  present <- (x > 0)
  tapply(present, group, function(y) mean(y) >= 0.5)
}

# Now, let's apply this function to filter the taxa
filtered_phylo <- phyloseq::filter_taxa(phylo_no_o, function(x) {
  group <- phyloseq::sample_data(phylo_no_o)$FREG5_Ethnic_Group
  any(is_prevalent(x, group))
}, prune = TRUE)

# Print the dimensions of the filtered phyloseq object
print(filtered_phylo)
otu_table(filtered_phylo)
```

```{r}
comparisons <- list(c("C", "I"), c("C", "M"), c("I", "M"))


p <- plot_richness(filtered_phylo, x = "FREG5_Ethnic_Group", color = "FREG5_Ethnic_Group",
                  measures = c("Shannon", "Simpson", "Chao1", "Fisher")) +
    geom_boxplot() +
    theme_bw() +
    scale_color_manual(values = ethnicity_colors) +
                         
    scale_fill_manual(values = ethnicity_colors) +
                        
    stat_compare_means(comparisons = comparisons, method = "t.test", label = "p.signif", symnum.args = list(cutpoints = c(0, 0.001, 0.01, 0.05, 1), symbols = c("***", "**", "*", "ns")))
print(p)
#p + theme(text = element_text(family = "serif"))
```
## Beta analysis
```{r}
dist = phyloseq::distance(filtered_phylo, method="bray")
phylo.beta = ordinate(filtered_phylo, method="PCoA", distance=dist)
sample_data(filtered_phylo)$FREG5_Ethnic_Group <- factor(sample_data(filtered_phylo)$FREG5_Ethnic_Group,
                                                                 levels = c("C", "I", "M"))

p <- plot_ordination(filtered_phylo, phylo.beta, type = "samples", color = "FREG5_Ethnic_Group") +
  theme_bw() +
  stat_ellipse(aes(group = FREG5_Ethnic_Group)) +
  scale_color_manual(values = ethnicity_colors) +
  scale_fill_manual(values = ethnicity_colors)

# Extract the data from the plot
plot_data <- p$data
# Add the points in the desired order using geom_point()
#c("C" = "#9FDDCA","I"= "#9EA3F1","M" = "#E17887","O" = "#80796BFF")
p +
  geom_point(data = subset(plot_data, FREG5_Ethnic_Group == "C"), color = "#9FDDCA") +
  geom_point(data = subset(plot_data, FREG5_Ethnic_Group == "I"), color = "#9EA3F1") +
  geom_point(data = subset(plot_data, FREG5_Ethnic_Group == "M"), color = "#E17887")
  #theme(text = element_text(family = "serif"))

```
```{r}
permanova_result = adonis2(dist ~ FREG5_Ethnic_Group, data = as(sample_data(filtered_phylo), "data.frame"))
print(permanova_result)
```

# only the significant species
```{r}
library(DESeq2)

# Assuming filtered_phylo is your phyloseq object from previous steps

# Convert phyloseq object to DESeq2 format
dds <- phyloseq_to_deseq2(filtered_phylo, ~ FREG5_Ethnic_Group)

# Run DESeq2 analysis
dds <- DESeq(dds)

# Get results
res <- results(dds)

# Filter for significantly different species (adjust p-value threshold as needed)
sig_species <- rownames(res)[which(res$padj < 0.05)]
sig_species 
# Create a new phyloseq object with only significant species
filtered_phylo_sig <- prune_taxa(sig_species, filtered_phylo)

# Print the number of significant species
print(paste("Number of significantly different species:", ntaxa(filtered_phylo_sig)))

# Print the dimensions of the new phyloseq object
print(filtered_phylo_sig)
```

```{r}
comparisons <- list(c("C", "I"), c("C", "M"), c("I", "M"))


p <- plot_richness(filtered_phylo_sig, x = "FREG5_Ethnic_Group", color = "FREG5_Ethnic_Group",
                  measures = c("Shannon", "Simpson", "Chao1", "Fisher")) +
    geom_boxplot() +
    theme_bw() +
    scale_color_manual(values = ethnicity_colors) +
                         
    scale_fill_manual(values = ethnicity_colors) +
                        
    stat_compare_means(comparisons = comparisons, method = "t.test", label = "p.signif", symnum.args = list(cutpoints = c(0, 0.001, 0.01, 0.05, 1), symbols = c("***", "**", "*", "ns")))
print(p)
#p + theme(text = element_text(family = "serif"))
```

```{r}
dist = phyloseq::distance(filtered_phylo_sig, method="bray")
phylo.beta = ordinate(filtered_phylo_sig, method="PCoA", distance=dist)
sample_data(filtered_phylo_sig)$FREG5_Ethnic_Group <- factor(sample_data(filtered_phylo_sig)$FREG5_Ethnic_Group,
                                                                 levels = c("C", "I", "M"))

p <- plot_ordination(filtered_phylo_sig, phylo.beta, type = "samples", color = "FREG5_Ethnic_Group") +
  theme_bw() +
  stat_ellipse(aes(group = FREG5_Ethnic_Group)) +
  scale_color_manual(values = ethnicity_colors) +
  scale_fill_manual(values = ethnicity_colors)

# Extract the data from the plot
plot_data <- p$data
# Add the points in the desired order using geom_point()
#c("C" = "#9FDDCA","I"= "#9EA3F1","M" = "#E17887","O" = "#80796BFF")
p +
  geom_point(data = subset(plot_data, FREG5_Ethnic_Group == "C"), color = "#9FDDCA") +
  geom_point(data = subset(plot_data, FREG5_Ethnic_Group == "I"), color = "#9EA3F1") +
  geom_point(data = subset(plot_data, FREG5_Ethnic_Group == "M"), color = "#E17887")
  #theme(text = element_text(family = "serif"))

```
```{r}
permanova_result = adonis2(dist ~ FREG5_Ethnic_Group, data = as(sample_data(filtered_phylo_sig), "data.frame"))
print(permanova_result)
```


# pcoa within ethnic group
```{r}
prevalence_threshold <- 0.2
phylo_no_ofiltered <- filter_taxa(phylo_no_o, 
                                  function(x) sum(x > 0) > (prevalence_threshold * nsamples(otu_table(phylo))),
                                  prune=TRUE)
phylo_no_ofiltered
```

```{r}
output_dir <- "figure/"  # Replace with your desired directory

for (group in ethnic_groups) {
  phylo_subset <- subset_samples(phylo, FREG5_Ethnic_Group == group)
  otu_data <- as(otu_table(phylo_subset), "matrix")
  if (taxa_are_rows(phylo_subset)) {
    otu_data <- t(otu_data)
  }
  bray_curtis_dist <- vegdist(otu_data, method = "bray")
  
  hc <- hclust(bray_curtis_dist, method = "ward.D2")
  file_name <- paste(output_dir, paste("Hierarchical_Clustering_Group_", group, ".png", sep = ""), sep = "/")
  png(file_name, width = 800, height = 600)  # Adjust the dimensions as needed
  plot(hc, main = paste("Hierarchical Clustering - Ethnic Group", group), 
       xlab = "", sub = "", cex = 0.01)  # Adjust 'cex' to make the text smaller
  
  rect.hclust(hc, k = 3, border = "red")  # Adjust 'k' as needed to cut the tree into clusters
  dev.off()
}

```

```{r}
  phylo_subset <- subset_samples(phylo, FREG5_Ethnic_Group == "C")
  otu_data <- as(otu_table(phylo_subset), "matrix")
  if (taxa_are_rows(phylo_subset)) {
    otu_data <- t(otu_data)
  }
  bray_curtis_dist <- vegdist(otu_data, method = "bray")
  
  hc <- hclust(bray_curtis_dist, method = "ward.D2")
  file_name <- paste(output_dir, paste("Hierarchical_Clustering_Group_C.png"), sep = "/")
  #png(file_name, width = 800, height = 600)  # Adjust the dimensions as needed
  plot(hc, main = paste("Hierarchical Clustering - Ethnic Group C"), 
       xlab = "", sub = "", cex = 0.001)  # Adjust 'cex' to make the text smaller
  
  rect.hclust(hc, k = 3, border = "red")  # Adjust 'k' as needed to cut the tree into clusters
  #dev.off()
```


```{r}
ethnic_groups <- unique(sample_data(phylo)$FREG5_Ethnic_Group)

# Create an empty list to store clustering results
clustering_results <- list()

# Set the number of clusters (adjust as needed)
k <- 3

# Loop through each ethnic group
for (group in ethnic_groups) {
  # Subset the phyloseq object to the current ethnic group
  phylo_subset <- subset_samples(phylo, FREG5_Ethnic_Group == group)
  
  # Extract OTU table and convert to matrix
  otu_data <- as(otu_table(phylo_subset), "matrix")
  
  # Ensure samples are rows and taxa are columns
  if (taxa_are_rows(phylo_subset)) {
    otu_data <- t(otu_data)
  }
  
  # Calculate Bray-Curtis distance between samples
  bray_curtis_dist <- vegdist(otu_data, method = "bray")
  
  # Perform k-means clustering on the samples
  km <- kmeans(otu_data, centers = k)
  
  # Add the clustering results to the sample data
  sample_data(phylo_subset)$Cluster <- factor(km$cluster)
  
  # Store the k-means clustering results
  clustering_results[[group]] <- km
  
  # Visualize the clustering using PCoA
  pcoa <- cmdscale(bray_curtis_dist, eig = TRUE, k = 2)
  
  # Create a data frame for plotting
  pcoa_df <- data.frame(PC1 = pcoa$points[, 1], PC2 = pcoa$points[, 2], 
                        Cluster = factor(km$cluster), SampleID = rownames(otu_data))
  
  # Plot the PCoA with clusters
  p <- ggplot(pcoa_df, aes(x = PC1, y = PC2, color = Cluster)) +   #label = SampleID
    geom_point(size = 3, alpha = 0.7) +
    #geom_text(hjust = 0, nudge_x = 0.05, size = 3) +
    stat_ellipse(type = "norm", level = 0.95, linetype = 1, size = 1) +
    ggtitle(paste("PCoA with K-Means Clusters - Ethnic Group", group)) +
    theme_minimal()
  
  # Print the plot
  print(p)
}

```
#silhouette score
```{r}
library(cluster)

# Loop through each ethnic group
for (group in ethnic_groups) {
  # Subset the phyloseq object to the current ethnic group
  phylo_subset <- subset_samples(phylo, FREG5_Ethnic_Group == group)
  
  # Extract OTU table and convert to matrix
  otu_data <- as(otu_table(phylo_subset), "matrix")
  
  # Ensure samples are rows and taxa are columns
  if (taxa_are_rows(phylo_subset)) {
    otu_data <- t(otu_data)
  }
  
  # Perform k-means clustering on the samples
  km <- kmeans(otu_data, centers = 2)
  
  # Calculate silhouette scores
  silhouette_scores <- silhouette(km$cluster, dist(otu_data))
  
  # Plot silhouette analysis
  plot(silhouette_scores, main = paste("Silhouette Plot - Ethnic Group", group))
  
  # Print average silhouette width
  avg_silhouette <- mean(silhouette_scores[, "sil_width"])
  cat(paste("Average Silhouette Width for group", group, ":", avg_silhouette, "\n"))
}

```
An average silhouette width close to 1 indicates that the clusters are well-separated, while a score close to 0 suggests overlapping clusters. Negative values suggest that some samples may have been assigned to the wrong cluster.
#elbow analysis
```{r}
# Loop through each ethnic group
for (group in ethnic_groups) {
  phylo_subset <- subset_samples(phylo, FREG5_Ethnic_Group == group)
  otu_data <- as(otu_table(phylo_subset), "matrix")
  
  if (taxa_are_rows(phylo_subset)) {
    otu_data <- t(otu_data)
  }
  
  wcss <- numeric()
  
  for (i in 1:10) {
    km <- kmeans(otu_data, centers = i)
    wcss[i] <- km$tot.withinss
  }
  
  plot(1:10, wcss, type = "b", pch = 19, frame = FALSE,
       xlab = "Number of Clusters", ylab = "Total Within-Cluster Sum of Squares",
       main = paste("Elbow Method for Ethnic Group", group))
}

```
#gap statistic
```{r}
library(cluster)
library(factoextra)

# Loop through each ethnic group
for (group in ethnic_groups) {
  phylo_subset <- subset_samples(phylo, FREG5_Ethnic_Group == group)
  otu_data <- as(otu_table(phylo_subset), "matrix")
  
  if (taxa_are_rows(phylo_subset)) {
    otu_data <- t(otu_data)
  }
  
  gap_stat <- clusGap(otu_data, FUN = kmeans, nstart = 25, K.max = 10, B = 50)
  
  fviz_gap_stat(gap_stat) +
    ggtitle(paste("Gap Statistic for Ethnic Group", group))
}

```



